### 2022-08-25 paper/markdown/manuscript.Rmd
Initial commit
Analysis pipeline and R Makdown Manuscript files

---
title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19"
subtitle: "Manuscript file"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

author: 
  - Katharina Huefner:
      institute: [psych]
  - Piotr Tymoszuk:
      institute: [daas]
  - Sabina Sahanic:
      institute: [int2]
  - Anna Luger:
      institute: [radio]
  - Anna Boehm:
      institute: [int2]
  - Alex Pizzini:
      institute: [int2]
  - Christoph Schwabl:
      institute: [radio]
  - Sabine Koppelstätter:
      institute: [int2]
  - Katharina Kurz:
      institute: [int2]
  - Malte Asshoff:
      institute: [int2]
  - Birgit Mosheimer-Feistritzer:
      institute: [int2]
  - Maximilian Coen:
      institute: [int2]
  - Bernhard Pfeifer:
      institute: [umit]
  - Verena Raas:
      institute: [neuro]
  - Alexander Egger:
      institute: [zimcl]
  - Gregor Hörmann:
      institute: [zimcl]
  - Raimund Helbok:
      institute: [neuro]
  - Ewald Wöll:
      institute: [zams]
  - Günter Weiss:
      institute: [int2]
  - Gerlig Widmann:
      institute: [radio]
  - Thomas Sonnweber:
      institute: [int2]
  - Ivan Tancevski:
      institute: [int2]
  - Barbara Sperner-Unterweger:
      institute: [psych]
  - Judith Löffler-Ragg:
      institute: [int2]
      email: judith.loeffler@i-med.ac.at
      correspondence: true

institute:
  - psych: Department of Psychiatry, Psychotherapy, Psychosomatics and Medical Psychology, University Hospital for Psychiatry II, Medical University of Innsbruck, Innsbruck, Austria
  - daas: Data Analytics As a Service Tirol, Innsbruck, Austria
  - int2: Department of Internal Medicine II, Medical University of Innsbruck, Innsbruck Austria
  - radio: Department of Radiology, Medical University of Innsbruck, Innsbruck, Austria
  - umit: Division for Health Networking and Telehealth, Biomedical Informatics and Mechatronics, UMIT, Hall in Tyrol, Austria
  - neuro: Department of Neurology, Medical University of Innsbruck, Innsbruck, Austria
  - zimcl: Central Institute of Medical and Chemical Laboratory Diagnostics, University Hospital Innsbruck, Innsbruck, Austria
  - zams: Department of Internal Medicine, St. Vinzenz Hospital, Zams, Austria

keywords: "illness perception, COVID-19, post COVID-19 syndrome, persistent symptoms, fatigue"

output: 
  bookdown::word_document2:   
    reference_docx: ms_template.docx

bibliography: cov_biblio.bib

csl: frontiers-in-immunology.csl

header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)
library(flextable)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      dev = "png", dpi = 600, 
                      tab.cap.pre = 'Supplementary Table ', tab.cap.sep = " ", 
                      tab.cap.style = 'Table Heading')

set_flextable_defaults(font.family = 'Cambria', font.size = 10)

```

\newpage

# Abstract

__Background:__
Sequelae of Coronavirus disease 2019 (COVID-19) were investigated by both patient-initiated and academic initiatives. 
Patient´s subjective illness perceptions might differ from physician’s clinical assessment results. 
Herein, we explored factors influencing patient’s perception during COVID-19 recovery.

__Methods:__
Participants of the prospective observation CovILD study with persistent somatic symptoms or cardiopulmonary findings at the clinical follow-up one year after COVID-19 were analyzed (n = `r nrow(cov_data$clear_data)`). 
Explanatory variables included baseline demographic and comorbidity data, COVID-19 course and one-year follow-up data of persistent somatic symptoms, physical performance, lung function testing (LFT), chest computed tomography (CT) and trans-thoracic echocardiography (TTE). 
Factors affecting illness perception (Brief Illness Perception Questionnaire, BIPQ) were identified by penalized multi-parameter regression and unsupervised clustering.

__Results:__
In modeling, `r (blass_mod$cv_fit_stats$ipq_total$estimate[4] %>% signif(2) %>% unname) * 100`% of overall illness perception variance at one year after COVID-19 was attributed to fatigue intensity, reduced physical performance, hair loss and  baseline respiratory comorbidity. 
Overall illness perception was independent of LFT results, pulmonary lesions in CT or heart abnormality in TTE. 
As identified by clustering, persistent somatic symptom count, fatigue, diminished physical performance, dyspnea, hair loss and sleep problems at the one-year follow-up and severe acute COVID-19 were associated with the BIPQ domains of concern, emotional representation, complaints, disease timeline and consequences.

__Conclusion:__
Persistent somatic symptoms rather than clinical assessment results, revealing lung and heart abnormalities, impact on severity and quality of illness perception at one year after COVID-19 and may foster unhelpful coping mechanisms. 
Besides  COVID-19 severity, individual illness perception should be taken into account when allocating rehabilitation and psychological therapy resources.

__Study registration:__ 
ClinicalTrials.gov: NCT04416100.

\newpage

# Introduction

A sizable fraction of coronavirus disease 2019 (COVID-19) patients is affected by protracted somatic symptoms, cardiopulmonary pathology and mental health disorders [@Sonnweber2020; @Sahanic2021; @Luger2022; @Sonnweber2022; @Hufner2022; @Al-Aly2021; @Davis2021; @Soriano2022; @Evans2021a]. 
Persistent COVID-19-related symptoms have been initially described by patient and social media initiatives [@McCorkell2021] and subsequently, recognized as the 'post COVID-19 condition' by the clinical community [@Sonnweber2020; @Sahanic2021; @Davis2021; @Soriano2022; @Al-Aly2021].
Yet, the patients' and clinicians' characteristics were not always consistent, which was observed also in other conditions, such as chronic obstructive pulmonary disease [@Miravitlles2013; @Schwartz2020] or functional disorders [@Schumacher2013].
For post COVID-19 condition, the matter is further complicated by a broad range of patient-reported manifestations and widespread character of the disease [@Davis2021; @Sahanic2021].
Since it is becoming increasingly evident that many COVID-19 convalescents experience prolonged severe individual suffering [@Hufner2022; @Huang2021; @Davis2021; @Evans2021a], and pose a significant healthcare and socioeconomic challenge [@Al-Aly2021], it is critically important to characterize and understand individual illness perception following COVID-19.

Illness perceptions can be divided into cognitive or emotional components [@Broadbent2006]. 
The cognitive representations include self-perceived consequences, expected duration, personal control, expected effect of treatment, symptom perception and understanding of the disease. 
The emotional components encompass patient's concerns and emotions e.g. fear, anger or distress associated with the disease [@Broadbent2006; @Weinman1996].
According to the common-sense model of self regulation (CSM), a theoretical framework enabling understanding of how people cope with threats to their health, such illness perceptions are influenced by situational stimuli such as symptoms, health information and patient's knowledge [@Man2021; @Leventhal2016; @Diefenbach1996]. 
The individual illness perception was shown to influence disease coping, adjustment to adverse life events or chronic conditions and compliance with prevention, treatment and rehabilitation [@Hagger2010; @Figueiras2019], also in context of acute COVID-19 [@DiasNeto2021; @Lugo-Gonzalez2020; @Vollmann2021].
Finally, severe illness perception including the emotional response, concerns and consequence components was found associated with anxiety, depression and stress both in the general population and in COVID-19 patients [@Wierenga2021; @Sadeghian2022; @Hamama2022; @Burke2020].

Herein, we investigated severity of overall illness perception and the illness perception components in a cross-sectional COVID-19 convalescent collective [@Sonnweber2020; @Sahanic2021; @Luger2022; @Sonnweber2022] with incomplete somatic symptom resolution or COVID-19-related lung and heart pathology in the follow-up clinical assessment at one year after diagnosis.

\newpage

# Methods

## Study design and approval

Participants of the longitudinal observation CovILD study (ClinicalTrials.gov: NCT04416100) were recruited between April and June 2020 at three Austrian clinical centers [@Sonnweber2020; @Luger2022; @Sonnweber2022]. 
The study inclusion criteria were age $\geq$ 18 years and symptomatic PCR-confirmed SARS-CoV-2 infection. 
The follow-up visits were scheduled at two, three, six months and one year after COVID-19 diagnosis. 
A subset of participants (n = `r nrow(cov_data$clear_data)`) was analyzed who completed the Brief Illness Perception Questionnaire (BIPQ) [@Broadbent2006] and displayed (1) COVID-19-related persistent somatic symptoms or (2) any abnormality in chest computed tomography (CT) or (4) any lung function testing (LFT) deficits or (5) any heart abnormality in trans-thoracic echocardiography (TTE) at the one-year follow-up (__Figure \@ref(fig:fig-consort)__).

The study was performed in accordance with the Declaration of Helsinki and the European Data Policy. All participants gave written informed consent. 
The study protocol was approved by the ethics committee at the Medical University of Innsbruck (approval number 1103/2020).

## Procedures

For full descriptions of procedures and variables, see __Supplementary Methods__ and __Supplementary Table S1__.

Baseline clinical and acute COVID-19 data were recorded retrospectively at the two-month follow-up based on the patient's interview and electronic patient records [@Sonnweber2020]. 
Study participants were classified as ambulatory (outpatient, WHO grade 1 - 2), moderate (hospitalized, without oxygen therapy or mask/nasal prongs oxygen, WHO 3 - 4) and severe COVID-19 survivors (hospitalized with non-invasive ventilation or high-flow oxygen or mechanical ventilation, WHO 5 - 7).

Physical performance was rated with the Eastern Cooperative Oncology Group scale (ECOG). 
Dyspnea was scored with the Modified Medical British Research Council scale (mMRC).
Fatigue at one-year follow-up was rated with likert and bimodal Chalder's Fatigue Scales (CFS) [@Chalder1993; @Morriss1998]. 
Exertional capacity at the one-year follow-up was assessed by six-minute walking distance (SMWD) test and compared with the reference values [@Crapo2012].

The following COVID-19-related persistent somatic symptoms at the one-year follow-up were analyzed: 
reduced physical performance (ECOG $\geq$ 1), 
dyspnea (mMRC $\geq$ 1), 
self-reported cough (yes/no item), 
self-reported sleep problems (yes/no), 
self-reported night sweating (yes/no), 
self-reported hyposmia or anosmia (yes/no), 
self-reported dermatological symptoms (yes/no), 
self-reported gastrointestinal symptoms (yes/no), 
self-reported hair loss (yes/no), 
significant fatigue (bimodal CFS $\geq$ 4).

LFT abnormality was defined as at least one parameter < 80% (forced vital capacity [FVC], forced expiratory volume in 1 second [FEV1], total lung capacity, diffusion lung capacity for carbon monoxide) or < 70% (FEV1:FVC ratio) of the reference value [@Sonnweber2022]. 
CT images were evaluated with the Fleischner Society glossary terms [@Hansell2008] and the CT severity score [@Sonnweber2020; @Luger2022; @Sonnweber2022]. 
Blood biomarkers encompassed hemoglobin and parameters of iron turnover, inflammation and coagulation.

Illness perception was investigated with the 8-item BIPQ (Q1 - Q8) [@Broadbent2006]. 
Each item was rated with an 11-point likert scale. 
The illness perception score was defined as a sum of all BIPQ items, with the negative items Q3, Q4 and Q7 inverted.

## Analysis endpoints

The primary analysis endpoints were illness perception score values and its influencing factors among demographic, clinical and somatic symptom data at the one-year after COVID-19 in convalescents affected by persistent somatic symptoms or residual cardiopulmonary findings.
The secondary analysis endpoints were particular BIPQ component scores and their influencing factors.

## Statistical analysis

Details of statistical analysis are provided in __Supplementary Methods__. 
Data analysis was done with R version 4.2.0 (R Foundation for Statistical Computing). 
Differences in categorical variable distribution were assessed by $\chi^2$ test with Cramer V effect size statistic. 
Statistical significance for numerical variables was investigated by Mann-Whitney test with Wilcoxon r effect size statistic or Kruskal-Wallis test with $\eta^2$ effect size statistic [@Kassambara2021]. 
Correlations were investigated by Spearman's test.

Multi-parameter modeling was done with the Elastic Net (package _glmnet_) [@Zou2005; @Friedman2010], LASSO (least absolute shrinkage and selection operator; package _glmnet_) [@Tibshirani1996; @Friedman2010] and Bayesian LASSO (package _monomvn_) [@Park2012; @Gramacy2022] algorithms. 
The response, illness perception score, was square-root transformed to guarantee normality. 
Both first and second order terms of numeric explanatory variables were included in the models. 
Numeric explanatory and response variables were Z-score normalized. 
The optimal $\lambda$ for Elastic Net and LASSO were obtained by 200-repeats 10-fold cross-validation (CV). 
The 'sparsity' parameter in Bayesian LASSO was found by 10-repeats 10-fold cross-validation (package _caret_) [@Kuhn2008]. 
Explained variance ($R^2$) and model root mean squared error (RMSE) were assessed in the entire data set and 10-repeats 10-fold CV (package _caret_) [@Kuhn2008]. 
Elastic Net and LASSO coefficients were calculated for the optimal $\lambda$ values. 
Bayesian LASSO coefficients were calculated as medians over all algorithm iterations [@Park2012].
Variables with non-zero coefficients in all three models were deemed key factors for the illness perception score.

Clustering by the BIPQ items was accomplished with the PAM algorithm (partitioning around medoids; Euclidean distance) [@Schubert2019; @Drost2018].

\newpage

# Results

## Study cohort at baseline and one year after COVID-19

The CovILD cohort was recruited between April and June 2020. 
Out of 145 participants enrolled, `r nrow(cov_data$clear_data)` individuals with COVID-19-related persistent somatic symptoms, lung CT, LFT or cardiological abnormalities and complete BIPQ were included in the analysis. 
The major reasons of patient dropout were missing follow-up visits and incomplete BIPQ (__Figure \@ref(fig:fig-consort)__).
The participants were predominantly male (`r get_percent(cov_data$clear_data, "sex")['male']`%), the median age at COVID-19 diagnosis was `r get_median_iqr(cov_data$clear_data, "age")['median']` years (IQR: `r get_median_iqr(cov_data$clear_data, "age")['iqr']`), over one-third were active or ex-smokers (`r get_percent(cov_data$clear_data, "smoking_history")['yes']`%). 
Most participants suffered from comorbidities (`r get_percent(cov_data$clear_data, "comorb_present")['yes']`%), with cardiovascular disease, metabolic and respiratory disorders as the leading conditions. 
The participants were classified by COVID-19 severity as ambulatory (`r get_percent(cov_data$clear_data, "cat_WHO")[1]`%), moderate (hospitalized, no intensive care , no oxygen therapy, `r get_percent(cov_data$clear_data, "cat_WHO")[2]`%) and severe COVID-19 survivors (hospitalized, oxygen therapy or intensive care, `r get_percent(cov_data$clear_data, "cat_WHO")[3]`%). 
The ambulatory COVID-19 subset had the lowest median age, rate of smokers and comorbidity frequency (__Table \@ref(tab:tab-cohort-baseline)__).

Nearly three-quarter of participants (`r get_percent(cov_data$clear_data, "sympt_present")['yes']`%) suffered from persistent somatic symptoms at one year after COVID-19, with significant fatigue (`r get_percent(cov_data$clear_data, "Chalder_FS_bimodal")['yes']`%), reduced physical performance (`r get_percent(cov_data$clear_data, "fatigue_sympt")['yes']`%), sleep disorders (`r get_percent(cov_data$clear_data, "sleep_sympt")['yes']`%) and exertional dyspnea (`r get_percent(cov_data$clear_data, "dyspnoe_sympt")['yes']`%) being most frequent. 
The symptom frequencies and fatigue rating were comparable between the COVID-19 severity strata (__Supplementary Table S2__).
LFT abnormalities were discerned in `r get_percent(cov_data$clear_data, "lufo_red")['yes']`% of participants and tended to be most common in severe COVID-19 survivors. 
Residual lung CT lesions were found in `r get_percent(cov_data$clear_data, "ct_severity_any")['yes']`% individuals and their frequency and scoring was significantly higher in moderate and severe COVID-19 than in the ambulatory disease subset. 
The most common cardiological finding was low grade diastolic dysfunction (`r get_percent(cov_data$clear_data, "diastolic_dysf")['yes']`% of cohort), which was significantly more frequent in moderate and severe than in ambulatory COVID-19 survivors. 
Nearly 80% of severe disease survivors attended COVID-19-specific rehabilitation, the rehabilitation rates in the remaining severity strata were below 20% (__Supplementary Table S2__). 
Most laboratory parameters at one year after COVID-19 were within their normal values. Mild anemia and improper glycemia control (HbA1c) were evident solely in moderate and severe COVID-19 survivors (__Supplementary Table S3__).

## Illness perception at one year after COVID-19

As measured by Cronbach's alpha [@Cronbach1951], the BIPQ tool had an acceptable internal consistency in the study cohort ($\alpha$ = `r ipq_co$alpha_stats %>% filter(response == "ipq_total") %>% .$raw_alpha %>% signif(2)`, 95% CI: `r paste(ipq_co$alpha_stats %>% filter(response == "ipq_total") %>% .$plot_cap %>% stri_extract(regex = "\\[.*\\]$") %>% stri_replace_all(regex = "\\[|\\]", replacement = ""))`) (__Supplementary Figure S1__). 
The consequences, timeline, identity, concern and emotional representation items were strongly positively inter-correlated. 
Significant, moderate-to-strong positive association was observed between the personal control and treatment control components as well as between the coherence and treatment control items.

The median illness perception score defined as the sum of all items [@Broadbent2006] was `r get_median_iqr(cov_data$clear_data, "ipq_total")['median']` (IQR: `r get_median_iqr(cov_data$clear_data, "ipq_total")['iqr']`) in the study collective, the differences between ambulatory, moderate and severe COVID-19 were not significant.
The treatment control and coherence BIPQ items were rated the highest followed by the personal control and emotional representation components.
Significant differences between the COVID-19 severity strata were detected for the consequences, concern, emotional representation and coherence BIPQ items, which peaked in severe COVID-19 convalescents (__Table \@ref(tab:tab-psych)__).

## Key factors for overall illness perception

To identify the most important factors among `r length(globals$variables)` candidate explanatory variables (__Supplementary Table S1__) influencing the illness perception score at one year after COVID-19, three penalized multi-parameter regression algorithms: Elastic Net [@Zou2005], LASSO [@Tibshirani1996] and Bayesian LASSO [@Park2012] were employed.
The final models explained at least `r (blass_mod$cv_fit_stats$ipq_total$estimate[4] %>% signif(2) %>% unname) * 100`% and `r (blass_mod$cv_fit_stats$ipq_total$estimate[12] %>% signif(2) %>% unname) * 100`% of the illness perception score variance in the entire data set and 10-fold cross-validation, respectively (__Supplementary Figure S2__).

The number of variables with non-zero coefficients varied between `r length(an$variables$ipq_total$eln_mod)` for the Elastic Net and `r length(an$variables$ipq_total$blass_mod)` for the Bayesian LASSO regression (__Figure \@ref(fig:fig-key-factors)A__). 
The strongest positive correlates of the illness perception score in Elastic Net and LASSO modeling were pre-existing immune deficiency and respiratory disease as well as hair loss, fatigue rating and reduced physical performance at the one-year follow up. 
Fatigue scoring and respiratory comorbidity were the major positive correlates in the Bayesian LASSO model (__Supplementary Figures S2 - S5__).
Of note, neither age, sex, acute COVID-19 severity nor LFT, CT and TTE abnormalities were selected as non-zero model coefficients by any of the regression algorithms. 
Their effects on the illness perception scoring was not significant in a direct analysis either (__Supplementary Figure S6__).

The key variables associated with the illness perception rating selected by all three models were (1) reduced physical performance, (2) hair loss and (3) fatigue rating at the one-year follow-up as well as (4) baseline respiratory comorbidity (__Figure \@ref(fig:fig-key-factors)A__). Each of those parameters was found significantly associated with higher illness perception scores in a direct comparison or correlation analysis (__Figure \@ref(fig:fig-key-factors)BC__).

## Heterogeneity of illness perception

Three subsets of participants which differed qualitatively in BIPQ components, termed further 'illness perception clusters', were identified by PAM clustering (__Supplementary Figure S7__). 
Roughly half of study participants (cluster #1, `r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[1]`% of participants) had low scoring of the emotional representation, concern, identity, timeline and consequences components along with high rating of the coherence, personal control and treatment control items. 
This translated to a low median illness perception score (`r get_median_iqr(ipq_clust$analysis_tbl %>% filter(clust_id == "#1"), "ipq_total")['median']`, IQR: `r get_median_iqr(ipq_clust$analysis_tbl %>% filter(clust_id == "#1"), "ipq_total")['iqr']`). 
Another `r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[2]`% of participants assigned to the cluster #2 were characterized by low levels of self-perceived personal or treatment control of their COVID-19 sequelae but otherwise by low rating of the concern, identity, timeline, consequences and emotional representation items. 
The cluster #3 individuals (`r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[3]`% of participants) had highly elevated rating of the concern, identity, timeline, consequences and emotional representation components as compared with the clusters #1 and #2. 
As a result, the cluster #3 displayed the highest median illness perception score (`r get_median_iqr(ipq_clust$analysis_tbl %>% filter(clust_id == "#3"), "ipq_total")['median']`, IQR: `r get_median_iqr(ipq_clust$analysis_tbl %>% filter(clust_id == "#3"), "ipq_total")['iqr']`) (__Figure \@ref(fig:fig-risk-clusters)AB__).

Among baseline demographic and clinical parameters and follow-up readouts of somatic complaints and cardiopulmonary abnormalities (__Supplementary Table S1__), COVID-19-related persistent somatic symptoms demonstrated the largest significant differences between the illness perception clusters.
In particular, the cluster #3 individuals with high level of disease-related concerns suffered from multiple persistent somatic symptoms, persistent fatigue, physical performance loss, sleep problems, dyspnea and hair loss at one year after COVID-19. 
The cluster #3 comprised predominantly of individuals with higher COVID-19 severity (WHO grade), lung CT abnormalities and the peak rates of COVID-19-specific rehabilitation.
The persistent somatic symptom frequency and intensity was comparable in the clusters #1 and #2. 
Yet, in  the cluster #2 characterized by low self-perceived disease coherence and control, frequencies of smokers, metabolic comorbidity and LFT abnormalities at one year after COVID-19 tended to be higher than in the cluster #1. The rehabilitation rate in the cluster #2 was significantly lower than in the cluster #1 (__Figure \@ref(fig:fig-risk-clusters)__ and __Supplementary Table S4__).

\newpage

# Discussion

By penalized multi-parameter regression [@Park2012; @Tibshirani1996; @Zou2005] applied to `r length(globals$variables)` candidate explanatory variables including demographic, clinical, somatic symptom data, laboratory and cardiopulmonary assessment results, we could discern `r length(an$cmm_variables$ipq_total)` features explaining `r (blass_mod$cv_fit_stats$ipq_total$estimate[4] %>% signif(2) %>% unname) * 100`% of overall illness perception variance at one year after COVID-19. 
Those factors were: fatigue scoring, reduced physical performance and hair loss at the one-year follow-up and pre-existing respiratory comorbidity. 
Of note, the effects of age, sex, COVID-19 severity or residual lung lesions in CT, LFT findings and heath abnormalities in TTE on long-term illness perception were negligible. 
Furthermore, high levels of COVID-19-related concerns, consequences and emotional representation was observed primarily in the subset of participants with multiple residual somatic symptoms and moderate-to-severe acute COVID-19 course.

Literature on illness perception in COVID-19 patients is scarce [@Man2021; @Burke2020; @Hamama2022; @Sadeghian2022; @Stafford2021]. 
To our best knowledge, this is the first report assessing severity and components of illness perception in COVID-19 convalescents with persistent somatic symptoms or cardiopulmonary findings at the clinical assessment one year after diagnosis.
The link between disease severity, symptoms, therapy control and illness perception is well founded in other chronic conditions [@Broadbent2006; @Schwartz2020; @Schoormans2020]. 
In particular, we identified fatigue as a strong covariate of overall illness perception scoring after COVID-19 and as an important factor characterizing individuals with high levels of COVID-19-related concerns, emotional representation and consequences in clustering analysis. 
Similar effects of chronic fatigue were described in arthritis [@Schwartz2020] and hematological malignancy [@Schoormans2020]. 
Both acute COVID-19 and its post-acute sequelae encompass various respiratory symptoms such as cough or dyspnea [@Sahanic2021; @Sonnweber2020; @Sonnweber2022; @Davis2021]. 
In addition, COVID-19 was found to exacerbate symptoms and worsen disease control in asthma [@Philip2022]. 
Such superimposed COVID-19-dependent and independent airway manifestations may hence explain more severe illness perception in respiratory comorbidity in our study cohort.

By clustering analysis we could identify three subsets of participants differing in key illness perception components. Half of individuals displayed low severity of overall illness perception, good self-perceived coherence, personal and treatment control paralleled by low burden of persistent somatic symptoms. 
Another `r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[2]`% of participants showed a similarly low level of persistent somatic symptoms or fatigue. 
Yet their illness perception was hallmarked by poor disease understanding and disbelief in personal and treatment control. 
By contrast, the remaining minor cluster suffered from multiple somatic complaints at the one-year follow-up and was enriched in individuals with residual lung lesions in CT, severe COVID-19 course, significant fatigue, sleep problems and hair loss. 
Their illness perception was characterized primarily by intense emotional representation, concern, burden of consequences and disease identity. 
This latter subset is of particular interest and concern for psychological and psychiatric management of post COVID-19 syndrome as high scoring of emotional components of illness perception was correlated with signs of shame, guilt, stress, depression and anxiety both in the general population during the pandemic [@Wierenga2021] and in acute COVID-19 [@Burke2020; @Hamama2022; @Sadeghian2022]. 
The individuals in this cluster could potentially profit most from psychological and psychiatric interventions.

Collectively, we demonstrate an important effect of persistent somatic symptoms rather than clinical assessment results revealing lung lesions, lung function deficits or heart abnormalities on severity and quality of illness perception at one year after COVID-19. 
In the long run, this aspect may bear consequences for the patient's physical and mental health following COVID-19 and public health in general. 
Negative illness perceptions were found to accompany somatic symptom disorder, perpetuate symptoms in somatoform disorders and to predict higher future healthcare expenditure [@Frostholm2014]. 
In multiple aspects, post COVID-19 condition resembles persistent somatic manifestations of functional or somatic symptom disorders [@Ballering2021].

Our study bears limitations. 
The most important one was the low participant number and substantial participant dropout due to missing follow-up visits and the BIPQ answers, which may have resulted in a selection bias. 
Furthermore, longitudinal rating of illness perception or inclusion of a general population control [@DiasNeto2021; @Lugo-Gonzalez2020; @Wierenga2021; @Shiloh2021] would allow us to assess possible improvement or worsening and explore factors associated with illness perception at consecutive time points. 
The study collective was recruited during the pandemic onset before introduction of effective anti-viral drugs and vaccination. 
We were hence unable to assess the effect of improved treatment and prevention on illness perception [@Vollmann2021]. 
Finally, the study variable set lacks parameters which may be vital for severity and character of illness perception such as family status, education and COVID-19 knowledge [@Man2021], media consumption [@Tutzer2021] and quarantine duration [@Burke2020].

# Conclusion

Our results demonstrate, that persistent somatic symptoms rather than clinical assessment results, revealing lung lesions, lung function deficits or heart abnormalities, impact on severity and quality of illness perception at one year after COVID-19.
Hence, it is to account not only for acute COVID-19 severity but also for the interplay between persistent somatic symptoms and individual illness perceptions when allocating rehabilitation and psychological or psychiatric resources.

\newpage

# Acknowledgments

We acknowledge commitment of the study participants and the medical staff during the COVID-19 pandemic.

# Funding

The study was funded by the research fund of the State of Tyrol (Project GZ 71934, to Judith Löffler-Ragg) and an Investigator-Initiated Study grant by Boehringer Ingelheim (IIS 1199-0424 to Ivan Tancevski). 
The funders had no role in study design, data collection and interpretation, or the decision to submit the work for publication.

# Conflict of interest

Piotr Tymoszuk owns a data science company, Data Analytics as a Service Tirol. 
Other authors declare that no conflict of interest exists.

# Data and code availability

An R data (RDa) file with anonymized patient data will be made available upon request to the corresponding author. 
The study pipeline is available at https://github.com/PiotrTymoszuk/CovILD-IPQ.

\newpage

# Tables

```{r tab-cohort-baseline, tab.cap = 'Baseline demographic and clinical characteristic of the study cohort.'}

flextable::flextable(paper_tbl$baseline) %>% 
  width(1, width = 4.5, unit = 'cm') %>% 
  width(2:5, width = 2.7, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm') %>% 
  width(7, width = 2.2, unit = 'cm') %>%
  footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>% 
  footnote(4, 1, value = as_paragraph('Former or active smoker'), ref_symbols = 'b', part = 'body') %>% 
  footnote(5, 1, value = as_paragraph('Overweight: body mass index 25 - 30 kg/m\u00B2, obesity: > 30 kg/m\u00B2'), ref_symbols = 'c', part = 'body') %>% 
  theme_vanilla

```

\newpage

```{r tab-psych, tab.cap = 'Illness perception score and item values of the Brief Illness Perception Questionnaire (BIPQ, Q1 - Q8) at one year after COVID-19.'}

flextable::flextable(paper_tbl$psych) %>% 
  width(1, width = 4.3, unit = 'cm') %>% 
  width(2:5, width = 2.7, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm') %>% 
  width(7, width = 2.2, unit = 'cm') %>% 
  footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>% 
  footnote(2, 1, value = as_paragraph('Sum of BIPQ items Q1 - Q8.'), ref_symbols = 'b', part = 'body') %>% 
  theme_vanilla

```

# Figures

```{r fig-consort, fig.width = mm_inch(paper_fig$consort$w), fig.height = mm_inch(paper_fig$consort$h), fig.cap = 'Flow diagram of study enrollment and analysis inclusion process.'}

paper_fig$consort$plot

```

__Figure \@ref(fig:fig-consort). Flow diagram of study enrollment and analysis inclusion process.__

_CT: computed tomography of the chest; LFT: lung function testing, TTE: trans-thoracic echocardiography._

\newpage

```{r fig-key-factors, fig.width = mm_inch(paper_fig$key_factors$w), fig.height = mm_inch(paper_fig$key_factors$h), fig.cap = 'Key factors associated with disease perception one year after COVID-19.'}

paper_fig$key_factors$plot

```

__Figure \@ref(fig:fig-key-factors). Key factors associated with disease perception one year after COVID-19.__

_Illness perception score (sum of BIPQ items) at one year after COVID-19 was modeled as a function of `r length(globals$variables)` candidate independent variables using the Elastic Net, LASSO and Bayesian LASSO algorithms. Key factors affecting the illness perception score were identified as independent variables with non-zero coefficients in all three models._

_(A) Numbers of variables with non-zero coefficients identified by each algorithm presented in a quasi-proportional Venn diagram. The key factors are listed next to the diagram._

_(B) Relationship between the illness perception score and the categorical key factors: reduced physical performance (Eastern Cooperative Oncology Group [ECOG] score > 0) and hair loss at one year after COVID and respiratory comorbidity was investigated by Mann-Whitney test with Wilcoxon r effect size statistic. Effect size statistic and p values are indicated in the plot captions. Illness perception score values are presented in violin plots. Points represent single observations. Red diamonds and whiskers depict medians and interquartile ranges. Numbers of complete observations are displayed under the plots._

_(C) Relationship between the illness perception score and likert Chalder's fatigue score (CFS) was investigated by Spearman's correlation. The correlation coefficient ($\rho$) and p value are indicated in the plot caption. Each point represents a single observations. The blue line with gray ribbon depict the fitted second order terms and 95% confidence intervals. The number of complete observations is displayed under the plot._

\newpage

```{r fig-risk-clusters, fig.width = mm_inch(paper_fig$ipq_clusters$w), fig.height = mm_inch(paper_fig$ipq_clusters$h), fig.cap = 'Heterogeneity of illness perception and residual symptoms one year after COVID-19.'}

paper_fig$ipq_clusters$plot

```

__Figure \@ref(fig:fig-risk-clusters). Heterogeneity of illness perception and residual symptoms one year after COVID-19.__

_Three subsets of study participants (illness perception clusters) were identified by clustering in respect to the Brief Illness Perception Questionnaire items (BIPQ, Q1 - Q8) with the PAM (partitioning around medoids) clustering algorithm and Euclidean distance metric. Numbers of observations assigned to the clusters are displayed under the plots._

_(A) Mean BIPQ item scores at one year after COVID-19 in the illness perception clusters. Statistical significance was determined by Kruskal-Wallis test. P values are indicated below the variable names. Lines represent mean values, tinted ribbons depict 2$\times$SEM (standard error of the mean) intervals._

_(B) Illness perception score (sum of BIPQ items), number of symptoms and likert Chalder's fatigue score (CFS) at one year after COVID-19 in the illness perception clusters. Statistical significance was determined by Kruskal-Wallis test with $\eta^2$ effect size statistic. Effect size statistic and p values are indicated in the plot captions. Response values are presented in violin plots. Points represent single observations. Red diamonds and whiskers depict medians and interquartile ranges._

_(C) Frequencies of significant fatigue (bimodal CFS $\geq$ 4), reduced physical performance (Eastern Cooperative Oncology Group [ECOG] score > 0) and sleep problems at one year after COVID-19 in the illness perception clusters. Statistical significance was determined by $\chi^2$ test with Cramer's V effect size statistic. Effect size statistic and p values are indicated in the plot captions._

\newpage

# References

### 2022-08-25 paper/markdown/supplementary_material.Rmd
Initial commit
Analysis pipeline and R Makdown Manuscript files

---
title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19"
subtitle: "Supplementary Material"
author: "CovILD study team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  bookdown::word_document2:   
    reference_docx: ms_template.docx

bibliography: cov_biblio.bib
csl: frontiers-in-immunology.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      dev = "png", dpi = 600, 
                      tab.cap.pre = 'Supplementary Table ', tab.cap.sep = " ", 
                      tab.cap.style = 'Table Heading')

set_flextable_defaults(font.family = 'Cambria', font.size = 10)

```

\newpage

# Supplementary Methods

## Study cohort and approval

The CovILD longitudinal observation cohort study (ClinicalTrials.gov: NCT04416100) includes adult ($\geq$ 18 year) patients with a symptomatic, PCR-confirmed SARS-CoV-2 infection. 
The follow-up visits were scheduled at two, three, six months and one year after COVID-19 diagnosis. A total of 190 individuals were screened for participation.
In the current report, a subset of the original CovILD cohort was investigated, which displayed (1) persistent COVID-19-related symptoms or (2) any abnormality in chest computed tomography (CT) or (4) any deficits in lung function testing (LFT) or (5) any cardiological abnormality in trans-thoracic echocardiography (TTE) at the one-year follow-up. The final analysis inclusion criterion was availability of complete Brief Illness Perception Questionnaire (IPQ) data [@Broadbent2006].

n = 145 participants were recruited between March and June 2020 at three centers located in Tyrol, Austria: the Medical University of Innsbruck (Innsbruck), St. Vinzenz Hospital (Zams) and Karl-Landsteiner Rehabilitation Facility (Muenster). 
The reasons for screening failure were denied informed consent and self-declared logistic and temporary incompatibility with the scheduled study visits. 
Out of n = `r nrow(cov_data$data_tbl)` participants, who completed the one-year follow-up, n = `r nrow(cov_data$clear_data)` suffered from symptoms or displayed cardiopulmonary abnormalities (CT, LFT or TTE findings), filled out the BIPQ and had a complete set of modeling variables (__Supplementary Table S\@ref(tab:tab-var-dict)__). 
The recruitment process scheme is presented in __Figure 1__. 
Baseline characteristic of the study cohort is presented in __Table 1__, follow-up features are summarized in __Supplementary Tables S\@ref(tab:tab-cohort-fup)__ (symptoms, cardiopulmonary findings) and __S\@ref(tab:tab-cohort-biochemistry)__ (laboratory parameters). BIPQ item scores and the sum BIPQ score are summarized in __Table 2__.

The study was performed in accordance with the Declaration of Helsinki and the European Data Policy. All participants gave written informed consent. The study protocol was approved by the ethics committee at the Medical University of Innsbruck (approval number 1103/2020).

## Procedures and variables

The protocol of each follow-up visit included a survey of COVID-19-related symptoms, blood laboratory parameter assessment, LFT, TTE and chest CT as described in more detail before [@Sonnweber2020; @Sonnweber2022; @Luger2022]. 
Comorbidities and data on acute COVID-19 course (symptoms, laboratory parameters, functional tests and imaging) were recorded retrospectively at the two-month follow-up visit based on information provided by the participant and electronic patient records.

### Baseline demographic and clinical status, acute COVID-19 severity

The baseline demographic variables included sex (male/female), age at COVID-19 diagnosis, smoking history, body weight and height. 
The following pre-existing comorbidities were surveyed: obesity, metabolic disease, cardiovascular disease, hypertension, type II diabetes, hypercholesterolemia, gastrointestinal disease, malignancy, chronic kidney disease, immune deficiency, pulmonary and respiratory conditions. 
Among the surveyed comorbidities, asthma, COPD (chronic obstructive pulmonary disease), interstitial lung disease and other chronic pulmonary disorders were subsumed under the 'respiratory disease' variable included in the independent modeling variable set.
Study participants were stratified according to acute COVID-19 severity as ambulatory (outpatient, WHO grade 1 - 2), moderate (hospitalized at normal infection ward, no oxygen therapy, WHO grade 3 - 4) and severe COVID-19 survivors (hospitalized with oxygen therapy or mechanical ventilation or intensive care, WHO grade 5 - 7).

### Cardiopulonary assessment

LFT abnormality was defined as FVC (forced vital capacity) < 80% or FEV1 (forced expiratory volume in 1 second) < 80% or TLC (total lung capacity) < 80% or DLCO (diffusion lung capacity for carbon monoxide) < 80% of the predicted reference value or FEV1:FVC ratio < 70% of the predicted reference value [@Sonnweber2022]. 
CT images were screened for ground glass opacities (GGO), consolidations, bronchiectasis and reticulations according to the Fleischner Society glossary [@Hansell2008]. 
Chest CT abnormalities were scored separately for each lobe with the CT severity score [@Sonnweber2020; @Sonnweber2022; @Luger2022]: 0 - no abnormality, 1 - minimal (subtle GGO), 2 - mild (several GGO, subtle reticulation), 3 - moderate (multiple GGO, reticulation, small consolidation), 4 - severe (extensive GGO, consolidation, reticulation with distortion), 5 - massive (multiple findings, parenchymal destruction). 
The sum for all five lobes was used in the analysis. 
Any chest CT abnormality is defined as CT severity score $\geq$ 1. 
The most frequent cardiological abnormality in TTE was low grade diastolic dysfunction (__Supplementary Table S\@ref(tab:tab-cohort-fup)__). 
No reduced left ventricular ejection fraction was observed in the participants included in the current analysis.

### COVID-19-related persistent somatic symptoms and exertional capacity

The following features were surveyed as COVID-19-related persistent somatic symptoms:

- __reduced physical performance__ assessed with the ECOG scale (Eastern Cooperative Oncology Group). Reduced physical performance was defined as ECOG $\geq$ 1
- __dyspnea__ rated with the mMRC scale (Modified Medical British Research Council). Dyspnea was defined as mMRC $\geq$ 1
- __cough__, self-reported, surveyed as a yes/no item
- __sleep problems__, self-reported, surveyed as a yes/no item
- __night sweating__, self-reported, surveyed as a yes/no item
- __hyposmia or anosmia__, self-reported, surveyed as a yes/no item
- __dermatological symptoms__, self-reported, surveyed as a yes/no item
- __gastrointestinal symptoms__, self-reported, surveyed as a yes/no item
- __hair loss__, self-reported, surveyed as a yes/no item
- __significant fatigue__, measured with the bimodal Chalder's Fatigue Scale (CFS). Significant fatigue was defined as bimodal CFS $\geq$ 4 [@Chalder1993; @Morriss1998]

The individual counts of persistent somatic symptoms were subsequently calculated.
Since only few participants displayed values of ECOG $\geq$ 2 or mMRC $\geq$ 2, the dichotomous 'reduced performance (ECOG $\geq$ 1)' and 'dyspnea (mMRC $\geq$ 1)' independent variables were used in modeling instead of the full numeric scales.
Fatigue intensity was rated with the likert Chalder's Fatigue Scale (CFS) [@Chalder1993; @Morriss1998]. 
Exertional capacity was assessed by six-minute walking distance (SMWD) test performed according to the ATS guidelines [@Crapo2012]. 
The reference SMWD values were obtained as described by Crapo and colleagues [@Crapo2012] and differences between the actual and reference values were calculated.

### Laboratory parameters

Laboratory parameters encompassing hemoglobin (total and glycated hemoglobin [HbA1c]), biomarkers of iron turnover (ferritin [FT], transferrin saturation [TF-Sat], hepcidin and soluble transferrin receptor [sTFR]), inflammation and coagulation (C-reactive protein [CRP] and D-dimer) and cardiovascular pathology (N-terminal pro-brain natriuretic peptide [NT-proBNP]) were determined routinely at the certified laboratory of the Central Institute for Medical and Chemical Diagnostic at the Federal University Hospital in Innsbruck. 
Other markers of systemic inflammation, interleukin 6 (IL6) and procalcitonin, were within the normal ranges in the analysis collective.

### Illness perception

Illness perception was scored with the Brief Illness Perception Questionnaire (IPQ) developed by Broadbent and colleagues [@Broadbent2006]. 
Each of the 8 BIPQ items (questions Q1 - Q8) was scored in with an 11-point likert scale:

- __Q1 consequences__ (0: no consequences, 10: very severe consequences)
- __Q2 timeline__  (0: very short, 10: forever)
- __Q3 personal control__ (0: no control, 10: extreme control)
- __Q4 treatment control__ (0: no control, 10: extreme control)
- __Q5 identity__ (0: no complaints, 10: multiple severe complaints)
- __Q6 concerns__ (0: no concerns at all, 10: extreme concerns)
- __Q7 coherence__ (0: not at all, 10: I understand my illness very good)
- __Q8 emotional representation__ (0: no emotions at all, 10: extreme emotions)

The illness perception score was calculated as an arithmetic sum of items Q1 - Q8, with the negative items Q3, Q4 and Q7 inverted. 
The BIPQ tool demonstrated acceptable consistency as measured by Cronbach's $\alpha$ = `r ipq_co$alpha_stats %>% filter(response == "ipq_total") %>% .$raw_alpha %>% signif(2)` (bootstrap percentile 95% CI: `r paste(ipq_co$alpha_stats %>% filter(response == "ipq_total") %>% .$plot_cap %>% stri_extract(regex = "\\[.*\\]$"))`) (__Supplementary Figure S\@ref(fig:fig-coherence)__) [@Cronbach1951].

Variables used in the exploratory data analysis and multi-parameter modeling are listed in __Supplementary Table S\@ref(tab:tab-var-dict)__.

## Data import, transformation and visualization methods

Data import, transformation, statistical analysis and result visualization was done with R version 4.2.0 (R Foundation for Statistical Computing). 
For import of data tables and variable transformation, the _tidyverse_ package bundle [@Wickham2019] and _rlang_ package [@Henry2022] was used. 
The _ggplot2_ [@Wickham2016] and _cowplot_ [@Wilke2019] packages and the development package _figur_ (https://github.com/PiotrTymoszuk/figur) were used for data visualization. 
Quasi-proportional Venn diagrams were generated with the _nVennR_ package [@Perez-Silva2018].
The manuscript and supplementary material files were created with the _rmarkdown_ [@Allaire2022], _knitr_ [@Xie2022], _bookdown_ [@Xie2016] packages, _Affiliations_ Pandoc filter set (https://github.com/drdanholmes/Affiliations) and the development package _figur_ (https://github.com/PiotrTymoszuk/figur).

## Hypothesis testing

For statistical hypothesis testing, base R functions, the package _rstatix_ [@Kassambara2021] and development package _exda_ (https://github.com/PiotrTymoszuk/ExDA) were used. 
Statistical significance of differences in frequency of categorical variable levels was assessed by $\chi^2$ test with Cramer V effect size statistic. 
Since multiple study variables were non-normally distributed as investigated by Shapiro-Wilk test and quantile-quantile plots, differences in numeric variables were assessed by non-parametric Mann-Whitney (effect size: Wilcoxon r) or Kruskal-Wallis test (effect size: $\eta^2$), as appropriate. 
Correlations were analyzed with Spearman test (function _cor.test()_, package _stats_) and visualized by fitting a second order trend (function _geom_smooth()_, package _ggplot_, formula: $y \sim x + I(x^2)$, development package _exda_).

## BIPQ tool consistency

Spearman's pair-wise correlations between the BIPQ items were calculated (function _cor.test()_, package _stats_) and visualized in a bubble plot (__Supplementary Figure S\@ref(fig:fig-coherence)__). 
Cronbach's $\alpha$ was calculated with package _psych_ [@Revelle2022; @Cronbach1951]. 95% confidence intervals for $\alpha$ were obtained by the bootstrap technique (n = 1000).

## Multi-parameter modeling

In multi-parameter modeling, a set of `r length(globals$variables)` independent variables was used. 
The response, illness perception score defined as a sum of all BIPQ items, was square-root transformed to guarantee normal distribution. 
For numeric variables, both first- and second-order terms were included to account for non-linear relationship between the independent variable and response. 
During pre-processing, numeric variables (independent and responses) were normalized with mean centering (Z-scores). 
The full independent variable list is provides in __Supplementary Table S\@ref(tab:tab-var-dict)__.

Three L1 penalized regression algorithms were utilized to model illness perception score at one year after COVID-19: Elastic Net [@Zou2005], LASSO (least absolute shrinkage and selection operator) [@Tibshirani1996] and Bayesian LASSO [@Park2012]. 
The choice of penalized regression over canonical backward elimination or forward selection was motivated by highly multi-dimensional nature of the data set.

The Elastic Net and LASSO algorithms are implemented in R by the _glmnet_ package [@Friedman2010]. 
The optimal $\lambda$ parameters for Elastic Net and LASSO regression were found as values associated with the minimum model deviance in 200-repeats 10-fold cross-validation (CV, function _cv.glmnet()_, Elastic Net: $\alpha$ = 0.5, LASSO: $\alpha$ = 1, $\lambda$ parameter 'lambda.min' was extracted). 
The optimal $\lambda$ values were `r filter(eln_mod$opt_lambda, response == "ipq_total")$lambda %>% signif(2)` for the Elastic Net and `r filter(lasso_mod$opt_lambda, response == "ipq_total")$lambda %>% signif(2)` for the LASSO model. 
Finally, Elastic Net and LASSO models with the optimal $\lambda$ values were constructed with _caret_ package (function _train()_) [@Kuhn2008] and their performance in the training data set and 10-repeats 10-fold CV was tested (__Supplementary Figure D\@ref(fig:fig-performance)__). 
Estimates of non-zero final model coefficients (__Supplementary Figures S\@ref(fig:fig-eln-est)__ and __S\@ref(fig:fig-lasso-est)__) were extracted with the _coef_ method (s parameter set to the optimal $\lambda$ value).

The Bayesian LASSO algorithm is implemented in R by the _monomvn_ package [@Gramacy2022]. 
The Bayesian Lasso model was constructed with the _caret_ package (function _train()_, method: 'blasso', iteration number: T = 1000). 
The optimal value of the sparsity parameter, which controls the fraction of non-zero coefficients in the final model based on the posterior distributions, was found by 10-repeats 10-fold CV with the RMSE (root mean squared error) statistic as a selection criterion (__Supplementary Figure S\@ref(fig:fig-performance)__). 
The optimal sparsity value selected from the 0.1 - 0.7 range was `r blass_mod$models$ipq_total$bestTune[1, "sparsity"] %>% signif(2)`. 
The model coefficient matrix X was extracted from the 'finalModel' slot of the caret model. 
The coefficient values (__Supplementary Figure S\@ref(fig:fig-blasso-est)__) were calculated as matrix column medians over all algorithm iterations [@Park2012].

Extraction of caret model performance measures: RMSE and $R^2$ defined as $1 - MSE/Var(y)$ (__Supplementary Figure S\@ref(fig:fig-performance)__) was accomplished by the in-house-developed package _caretExtra_ (https://github.com/PiotrTymoszuk/caretExtra). 
Model assumption testing (normality of residuals, Shapiro-Wilk test, method _residuals()_) and visual quality control based on plots of residuals versus fitted and residuals quantile-quantile plots was done with the _caretExtra_ package (method _plot()_).

The key factors impacting on the illness perception score at one year after COVID-19 (`r an$cmm_variables$ipq_total %>% translate_var(out_value = "label_long") %>% de_capitalize %>% collapse_and`) were defined as variables with non-zero coefficients in all multi-parameter models.

## Clustering analysis

Study participants were clustered in respect to BIPQ items (`r colnames(model.frame(ipq_clust$clust_obj)) %>% translate_var(out_value = "label_long") %>% stri_replace(fixed = "Illness perception, ", replacement = "") %>% collapse_and`) by the PAM algorithm (partition around medoids, package _cluster_) [@Schubert2019] with the Euclidean distance measure between the observations (package _philentropy_) [@Drost2018]. 
The BIPQ item scores were not pre-processed, since they were measured with the same likert scale. 
The PAM/Euclidean distance procedure was chosen based on the high explained clustering variance (ratio of the between-cluster sum of squares to the total sum of squares) and stability (cluster assignment accuracy in 10-fold CV, cluster assignment by 7-nearest neighbors label propagation algorithm) [@Lange2004; @Leng2014] as compared with several other clustering algorithms (__Supplementary Figure S\@ref(fig:fig-clust-dev)A__). 
The choice of cluster number (k = `r ipq_clust$clust_obj$clust_assignment$clust_id %>% levels %>% length`) was based on the bend of the curve of within-cluster sum of squares (generated with package _factoextra_, __Supplementary Figure S\@ref(fig:fig-clust-dev)B__) [@Kassambara2020] and visual analysis of the distance heat map (__Supplementary Figure S\@ref(fig:fig-clust-dev)C__). 
Clustering object development, clustering variance calculation, cross-validation and visualization were performed with the in-house developed package _clustTools_ (https://github.com/PiotrTymoszuk/clustTools).

## Data and code availability

An R data (RDa) file with anonymized patient data will be made available upon request to the corresponding author. The study pipeline is available at https://github.com/PiotrTymoszuk/CovILD-IPQ.

\newpage

# Supplementary Tables

```{r tab-var-dict, tab.cap = 'Study variables'}

flextable::flextable(suppl_tbl$var_lexicon) %>% 
  width(1, width = 4.5, unit = 'cm') %>% 
  width(2, width = 5.6, unit = 'cm') %>% 
  width(3, width = 1.5, unit = 'cm') %>% 
  width(4, width = 6, unit = 'cm') %>% 
  theme_vanilla

```

\newpage

```{r tab-cohort-fup, tab.cap = 'COVID-19 symptoms, performance, fatigue, exertional capacity, cardiopulmonary abnormalities and illness perception at the one-year follow-up.'}

flextable::flextable(suppl_tbl$fup) %>% 
  width(1, width = 4.4, unit = 'cm') %>% 
  width(2:5, width = 2.8, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm') %>% 
  width(7, width = 2.1, unit = 'cm') %>%
  footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>% 
  footnote(4, 1, value = as_paragraph('Eastern Cooperative Oncology Group performance score'), ref_symbols = 'b', part = 'body') %>% 
  footnote(5, 1, value = as_paragraph('Modified Medical Research Council dyspnea score'), ref_symbols = 'c', part = 'body') %>% 
  footnote(13:14, 1, value = as_paragraph("Chalder's Fatigure Score"), ref_symbols = 'd', part = 'body') %>% 
  footnote(15, 1, value = as_paragraph('Lung function testing abnormality: FVC < 80% or FEV1 < 80% or TLC < 80% or DLCO < 80% predicted or FEV1:FVC < 70% predicted reference value'), ref_symbols = 'e', part = 'body') %>% 
  footnote(16, 1, value = as_paragraph('Any abnormality in chest computed tomography (CT), CT severity score ≥ 1'), ref_symbols = 'f', part = 'body') %>% 
  footnote(17, 1, value = as_paragraph('Chest computer tomography severity score'), ref_symbols = 'g', part = 'body') %>% 
  footnote(19, 1, value = as_paragraph('Six minute walking distance'), ref_symbols = 'h', part = 'body') %>% 
  footnote(20, 1, value = as_paragraph('Six-minute walking distance, difference between the actual and reference value'), ref_symbols = 'i', part = 'body') %>% 
  theme_vanilla

```

\newpage

```{r tab-cohort-biochemistry, tab.cap = 'Laboratory parameters at the one-year follow-up.'}

flextable::flextable(suppl_tbl$biochemistry_fup) %>%
  width(1, width = 2.8, unit = 'cm') %>% 
  width(2:5, width = 3.3, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm') %>% 
  width(7, width = 2.1, unit = 'cm') %>%
  footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>% 
  footnote(2, 1, value = as_paragraph('Females: hemoglobin (Hb) < 120 g/dL, males: Hb < 140 g/dL'), ref_symbols = 'b', part = 'body') %>% 
  footnote(3, 1, as_paragraph('Blood hemoglobin'), ref_symbols = 'c', part = 'body') %>% 
  footnote(4, 1, as_paragraph('Serum ferritin'), ref_symbols = 'd', part = 'body') %>%
  footnote(5, 1, as_paragraph('Serum transferrin saturation'), ref_symbols = 'e', part = 'body') %>% 
  footnote(6, 1, as_paragraph('Serum soluble transferrin receptor'), ref_symbols = 'f', part = 'body') %>% 
  footnote(8, 1, as_paragraph('Serum N-terminal pro-brain natriuretic peptide'), ref_symbols = 'g', part = 'body') %>% 
  footnote(10, 1, as_paragraph('Serum C-reactive protein'), ref_symbols = 'h', part = 'body') %>% 
  footnote(11, 1, as_paragraph('Blood glycated hamoglobin'), ref_symbols = 'i', part = 'body') %>% 
  theme_vanilla

```

\newpage

```{r tab-ipq-clusters, tab.cap = 'Significant (p < 0.05) and near-significant (p < 0.1) differences between the illness perception clusters.'}

flextable::flextable(suppl_tbl$ipq_clusters) %>% 
  width(1, width = 4.8, unit = 'cm') %>% 
  width(2:4, width = 2.6, unit = 'cm') %>% 
  width(5, width = 2.5, unit = 'cm') %>% 
  width(6, width = 2.1, unit = 'cm') %>%
  footnote(1, 5:6, value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>% 
  footnote(8, 1, value = as_paragraph('Eastern Cooperative Oncology Group performance score'), ref_symbols = 'b', part = 'body') %>% 
  footnote(9, 1, value = as_paragraph('Modified Medical Research Council dyspnea score'), ref_symbols = 'c', part = 'body') %>% 
  footnote(13:14, 1, value = as_paragraph("Chalder's Fatigure Score"), ref_symbols = 'd', part = 'body') %>% 
  footnote(15, 1, value = as_paragraph('Lung function testing abnormality: FVC < 80% or FEV1 < 80% or TLC < 80% or DLCO < 80% predicted or FEV1:FVC < 70% predicted reference value'), ref_symbols = 'e', part = 'body') %>% 
  footnote(16, 1, value = as_paragraph('Any abnormality in chest computed tomography (CT), CT severity score ≥ 1'), ref_symbols = 'f', part = 'body') %>% 
  theme_vanilla

```

\newpage

# Supplementary Figures

```{r fig-coherence, fig.width = mm_inch(suppl_fig$coherence$w), fig.height = mm_inch(suppl_fig$coherence$h), fig.cap = 'Coherence of the illness perception score.'}

suppl_fig$coherence$plot

```

__Supplementary Figure S\@ref(fig:fig-coherence). Coherence of the illness perception score.__

_Pair-wise correlation between particular items of the Brief Illness Perception Questionnaire (IPQ) was assessed by Spearman test and the results presented in the point plot. Point size corresponds to absolute values of the correlation coefficient $\rho$. Point color codes for the $\rho$ value. The points are labeled with $\rho$ values, significant effects are highlighted in bold. Internal consistency of the BIPQ tool was measured by Cronbach's $\alpha$. The $\alpha$ statistic with 95% confidence intervals and the number of complete observations are displayed in the plot caption._

\newpage

```{r fig-performance, fig.width = mm_inch(suppl_fig$perfomance$w), fig.height = mm_inch(suppl_fig$perfomance$h), fig.cap = 'Multi-parameter model performance.'}

suppl_fig$perfomance$plot

```

__Supplementary Figure S\@ref(fig:fig-performance). Multi-parameter model performance.__

_(A) Fraction of illness perception score variance explained by the Elastic Net, LASSO and Bayesian LASSO model in the training data set and 10-repeats 10-fold cross-validation (CV) was estimated by the $R^2$ statistic._

_(B) Model error expressed as root mean squared error (RMSE) in the training and CV data sets._

\newpage

```{r fig-eln-est, fig.width = mm_inch(suppl_fig$eln_est$w), fig.height = mm_inch(suppl_fig$eln_est$h), fig.cap = 'Non-zero coefficient estimates of the Elastic Net multi-parameter model.'}

suppl_fig$eln_est$plot

```

__Supplementary Figure S\@ref(fig:fig-eln-est). Non-zero coefficient estimates of the Elastic Net multi-parameter model.__

_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the Elastic Net model are presented in the plot as points. Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$. Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption. The data points are labeled with their $\beta$ values._

\newpage

```{r fig-lasso-est, fig.width = mm_inch(suppl_fig$lasso_est$w), fig.height = mm_inch(suppl_fig$lasso_est$h), fig.cap = 'Non-zero coefficient estimates of the LASSO multi-parameter model.'}

suppl_fig$lasso_est$plot

```

__Supplementary Figure S\@ref(fig:fig-lasso-est). Non-zero coefficient estimates of the LASSO multi-parameter model.__

_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the LASSO (least absolute shrinkage and selection operator) model are presented in the plot as points. Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$. Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption. The data points are labeled with their $\beta$ values._

\newpage

```{r fig-blasso-est, fig.width = mm_inch(suppl_fig$blasso_est$w), fig.height = mm_inch(suppl_fig$blasso_est$h), fig.cap = 'Non-zero coefficient estimates of the Bayesian LASSO multi-parameter model.'}

suppl_fig$blasso_est$plot

```

__Supplementary Figure S\@ref(fig:fig-blasso-est). Non-zero coefficient estimates of the Bayesian LASSO multi-parameter model.__

_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the Bayesian LASSO model are presented in the plot as points. Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$. Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption. The data points are labeled with their $\beta$ values._

\newpage

```{r fig-cardiopulmo, fig.width = mm_inch(suppl_fig$cardiopulmo$w), fig.height = mm_inch(suppl_fig$cardiopulmo$h), fig.cap = 'Age, sex and cardiopulmonary abnormalities at the one-year follow-up and illness perception.'}

suppl_fig$cardiopulmo$plot

```

__Supplementary Figure S\@ref(fig:fig-cardiopulmo). Age, sex and cardiopulmonary abnormalities at the one-year follow-up and illness perception..__

_(A - D) Differences in the illness perception score between males and female (A), the study participants with and without lung function (LFT) testing abnormality (B), chest computed tomography (CT) abnormality (C) and heart diastolic dysfunction (D) were assessed by Mann-Whitney test with Wilcoxon r effect size statistic. Effect size statistic and p values are indicated in the plot captions. Illness perception score values are presented i violin plots. Each point represents a single observation. Red diamonds and whiskers represent medians with interquartile ranges. Numbers of individuals with and without cardiopulmonary abnormalities are displayed under the plots._

_(F). Correlation of age at COVID-19 diagnosis and the illness perception score at one year after COVID-19 diagnosis was investigated by Spearman's test. Correlation coefficients ($\rho$) and p values are indicated in the plot captions. Each point represents a single observations. Blue lines with gray ribbons depict the fitted second order terms and 95% confidence intervals. Numbers of complete observations are displayed under the plots._

\newpage

```{r fig-clust-dev, fig.width = mm_inch(suppl_fig$clust_dev$w), fig.height = mm_inch(suppl_fig$clust_dev$h), fig.cap = 'Choice of the optimal clustering algorithm and the cluster number.'}

suppl_fig$clust_dev$plot

```

__Supplementary Figure S\@ref(fig:fig-clust-dev). Choice of the optimal clustering algorithm and the cluster number.__

_Study participants were clustered in respect to the Brief Illness Perception Questionnaire items (Q1 - Q8) using the PAM (partitioning around medoids) clustering algorithm and Euclidean distance between the observations. Numbers of complete observations and clustering variables are indicated in (A)_

_(A) Comparison of clustering algorithm performance (PAM: partitioning around medoids, HCL: Ward D2 hierarchical clustering, KMEANS). Explained clustering variance was defined as a ratio of the between-cluster sum of squares to the total sum of squares, cluster assignment accuracy was investigated in 10-fold cross-validation (CV). Note the superior CV accuracy of the PAM/Euclidean distance._

_(B, C) Choice of the optimal cluster number for the hierarchical clustering/Euclidean distance algorithm. The optimal cluster number was selected based on the bend of the curve of within-cluster sum of squares (B, dashed red line: the selected cluster number) and visual inspection of the distance heat map (C)._

\newpage

# References

### 2022-08-25 analysis scripts/clust_characteristic.R

@@ -1,0 +1,175 @@
+# Characteristic of the clusters, using the variables specified
+# for the explorative analysis
+
+  insert_head()
+  
+# container list -------
+  
+  clust_chara <- list()
+
+# globals: analysis table -----
+  
+  insert_msg('Analysis table')
+
+  clust_chara$analysis_tbl <- left_join(ipq_clust$clust_obj$clust_assignment %>% 
+                                          set_names(c('ID', 'clust_id')), 
+                                        cov_data$clear_data, 
+                                        by = 'ID')
+  
+  plan('multisession')
+  
+# descriptive stats ------
+  
+  insert_msg('Descriptive stats')
+  
+  clust_chara$desc_stats <- expl$variables %>% 
+    future_map(~explore(clust_chara$analysis_tbl, 
+                        split_factor = 'clust_id', 
+                        variables = .x, 
+                        what = 'table', 
+                        pub_styled = TRUE), 
+               .options = furrr_options(seed = TRUE)) %>% 
+    map(reduce, 
+        left_join, 
+        by = 'variable') %>% 
+    map(set_names, 
+        c('variable', levels(clust_chara$analysis_tbl$clust_id)))
+    
+# serial testing ------
+  
+  insert_msg('Serial testing')
+  
+  clust_chara$test_results <- list(variables = expl$variables, 
+                                   types = expl$test_type) %>% 
+    future_pmap(compare_variables,
+                clust_chara$analysis_tbl, 
+                split_factor = 'clust_id', 
+                what = 'test', 
+                ci = FALSE, 
+                pub_styled = TRUE, 
+                adj_method = 'none', 
+                .options = furrr_options(seed = TRUE)) %>% 
+    map(mutate, 
+        plot_cap = paste(eff_size, significance, sep = ', '))
+  
+# Plots -------
+  
+  insert_msg('Plots')
+  
+  clust_chara[c('baseline_plots', 
+                'fup_plots', 
+                'ipq_plots', 
+                'psych_plots', 
+                'lab_plots')] <- c('baseline', 
+                                   'fup', 
+                                   'ipq', 
+                                   'psych', 
+                                   'lab') %>% 
+    map(function(resp_type) list(variable = expl$variables[[resp_type]], 
+                                 type = expl$plot_type[[resp_type]], 
[…]

### 2022-08-25 analysis scripts/ipq_clust_development.R

@@ -1,0 +1,126 @@
+# Clustering by the IPQ items
+
+  insert_head()
+
+# container -----
+
+  ipq_clust_dev <- list()
+
+# globals ------
+  
+  insert_msg('Globals setup')
+  
+  ## analysis table
+  
+  ipq_clust_dev$analysis_tbl <- 
+    cov_data$clear_data[c('ID', paste0('ipq_q', 1:8))] %>% 
+    column_to_rownames('ID')
+  
+  ## distances
+  
+  ipq_clust_dev$dist <- c('euclidean', 'manhattan')
+  
+# clustering algorithms ------
+  
+  insert_msg('Clustering algorithms')
+  
+  ipq_clust_dev$algos[paste0('hcl_', ipq_clust_dev$dist)] <- 
+    ipq_clust_dev$dist %>% 
+    map(~hcluster(ipq_clust_dev$analysis_tbl, 
+                  distance_method = .x, 
+                  k = 3, 
+                  hc_method = 'ward.D2', 
+                  seed = 1234))
+  
+  ipq_clust_dev$algos[paste0('kmeans_', ipq_clust_dev$dist)] <- 
+    ipq_clust_dev$dist %>% 
+    map(~kcluster(ipq_clust_dev$analysis_tbl, 
+                  distance_method = .x, 
+                  k = 3, 
+                  clust_fun = 'kmeans', 
+                  seed = 1234))
+  
+  ipq_clust_dev$algos[paste0('pam_', ipq_clust_dev$dist)] <- 
+    ipq_clust_dev$dist %>% 
+    map(~kcluster(ipq_clust_dev$analysis_tbl, 
+                  distance_method = .x, 
+                  k = 3, 
+                  clust_fun = 'pam', 
+                  seed = 1234))
+  
+# comparing the clustering variances ------
+  
+  insert_msg('Clustering variances')
+  
+  ipq_clust_dev$variance <- ipq_clust_dev$algos %>% 
+    map(clustTools::var) %>% 
+    map_dbl(~.x$frac_var) %>% 
+    tibble(method = names(.), 
+           frac_var = .)
+  
+# cross-validation -----
+  
+  insert_msg('Cross-validation')
+  
+  ipq_clust_dev$cv <- ipq_clust_dev$algos %>% 
+    map(cv, 
+        nfolds = 10, 
+        kNN = 7, 
+        resolve_ties = TRUE) %>% 
+    map2_dfr(., names(.), 
[…]

### 2022-08-25 analysis scripts/ipq_clustering.R

@@ -1,0 +1,237 @@
+# Clustering of the study participants in respect to the IPQ-influencing factors
+# the algorithm: chosen by comparing variances and CV errors
+
+  insert_head()
+  
+# container -----
+  
+  ipq_clust <- list()
+  
+# globals: variables, analysis table and the best algorithm -----
+  
+  insert_msg('Globals setup')
+  
+  ## clustering object, renaming
+  
+  ipq_clust$clust_obj <- ipq_clust_dev$algos$pam_euclidean
+  
+  ipq_clust$clust_obj$clust_assignment <- 
+    ipq_clust$clust_obj$clust_assignment %>% 
+    mutate(clust_id = car::recode(clust_id, "'1' = '#2'; '2' = '#1'; '3' = '#3'"), 
+           clust_id = factor(clust_id))
+
+  ## clustering variables
+  
+  ipq_clust$clust_variables <- paste0('ipq_q', 1:8)
+
+  ipq_clust$clust_lexicon <- 
+    tibble(variable = c(ipq_clust$clust_variables, 
+                        'ipq_total', 'ipq_sub1', 'ipq_sub2')) %>% 
+    mutate(numeric = map_lgl(variable,  
+                             ~is.numeric(cov_data$clear_data[[.x]])), 
+           test_type = ifelse(numeric, 'kruskal_test', 'chisq_test'), 
+           plot_type = ifelse(numeric, 'violin', 'stack'))
+  
+  ipq_clust[c('fct_vars', 'num_vars')] <- dlply(ipq_clust$clust_lexicon, 
+                                                'numeric', 
+                                                function(x) x$variable)
+  
+  ## analysis table with the clustering variables, the total IPQ
+  ## and sub-scores
+  
+  ipq_clust$analysis_tbl <- 
+    cov_data$clear_data[c('ID', ipq_clust$clust_variables)] %>% 
+    right_join(ipq_clust$clust_obj$clust_assignment %>% 
+                set_names(c('ID', 'clust_id')), 
+              by = 'ID') %>% 
+    left_join(cov_data$clear_data[c('ID', 'ipq_total', 'ipq_sub1', 'ipq_sub2')], 
+              by = 'ID') %>% 
+    as_tibble
+  
+  ## plotting order for heat maps and radial plots: by p value
+  
+  ipq_clust$plot_order <- c('ipq_q3', 
+                            'ipq_q4', 
+                            'ipq_q7', 
+                            'ipq_q1', 
+                            'ipq_q2', 
+                            'ipq_q5', 
[…]

### 2022-08-25 analysis scripts/ipq_factors.R

@@ -1,0 +1,117 @@
+# Relationship between common influential factors and sub1 value
+
+  insert_head()
+
+# container -------
+
+  ipq_factors <- list()
+  
+# globals -----
+
+  insert_msg('Globals')
+  
+  ipq_factors$variables <- an$cmm_lexicon %>% 
+    filter(variable %in% unique(c(an$cmm_variables$ipq_total, 
+                                  an$conf_variables))) %>% 
+    dlply(.(what))
+  
+# Serial comparisons and correlations -----
+  
+  insert_msg('Serial comparisons')
+  
+  ## descriptive stats
+  
+  ipq_factors$desc_stats <- ipq_factors$variables$eff_size$variable %>% 
+    map(~explore(cov_data$clear_data %>% 
+                   filter(!is.na(.data[[.x]])), 
+                 split_factor = .x, 
+                 variables = 'ipq_total', 
+                 what = 'table', 
+                 pub_styled = TRUE)) %>% 
+    map(reduce, left_join, by = 'variable') %>% 
+    map2(., ipq_factors$variables$eff_size$levels, 
+         ~set_names(.x, c('variable', .y))) %>% 
+    set_names(ipq_factors$variables$eff_size$variable)
+  
+  ## comparisons
+  
+  ipq_factors$factor_test <- 
+    list(split_factor = ipq_factors$variables$eff_size$variable, 
+         types = ipq_factors$variables$eff_size$eff_type) %>% 
+    pmap_dfr(compare_variables, 
+             cov_data$clear_data, 
+             variables = 'ipq_total', 
+             what = 'eff_size', 
+             ci = FALSE, 
+             pub_styled = TRUE) %>% 
+    mutate(split_factor = ipq_factors$variables$eff_size$variable, 
+           plot_cap = paste(eff_size, significance, sep = ', '))
+  
+  ## Spearman's correlations 
+  
+  ipq_factors$corr_test <- ipq_factors$variables$correlation$variable %>% 
+    map_dfr(~correlate_variables(cov_data$clear_data[c('ipq_total', .x)] %>% 
+                                   filter(complete.cases(.)), 
+                                 variables = c('ipq_total', .x), 
+                                 what = 'correlation', 
+                                 type = 'spearman', 
+                                 ci = FALSE, 
+                                 pub_styled = TRUE)) %>% 
+    safely(mutate)(plot_cap = paste(eff_size, significance, sep = ', '), 
+                   plot_cap = stri_replace(plot_cap, 
+                                           fixed = 'rho', 
[…]

### 2022-08-25 analysis scripts/ipq_sub1_factors.R

@@ -1,0 +1,117 @@
+# Relationship between common influential factors and sub1 value
+
+  insert_head()
+
+# container -------
+
+  sub1_factors <- list()
+  
+# globals -----
+
+  insert_msg('Globals')
+  
+  sub1_factors$variables <- an$cmm_lexicon %>% 
+    filter(variable %in% unique(c(an$cmm_variables$ipq_sub1, 
+                                  an$conf_variables))) %>% 
+    dlply(.(what))
+  
+# Serial comparisons and correlations -----
+  
+  insert_msg('Serial comparisons')
+  
+  ## descriptive stats
+  
+  sub1_factors$desc_stats <- sub1_factors$variables$eff_size$variable %>% 
+    map(~explore(cov_data$clear_data %>% 
+                   filter(!is.na(.data[[.x]])), 
+                 split_factor = .x, 
+                 variables = 'ipq_sub1', 
+                 what = 'table', 
+                 pub_styled = TRUE)) %>% 
+    map(reduce, left_join, by = 'variable') %>% 
+    map2(., sub1_factors$variables$eff_size$levels, 
+         ~set_names(.x, c('variable', .y))) %>% 
+    set_names(sub1_factors$variables$eff_size$variable)
+  
+  ## comparisons
+  
+  sub1_factors$factor_test <- 
+    list(split_factor = sub1_factors$variables$eff_size$variable, 
+         types = sub1_factors$variables$eff_size$eff_type) %>% 
+    pmap_dfr(compare_variables, 
+             cov_data$clear_data, 
+             variables = 'ipq_sub1', 
+             what = 'eff_size', 
+             ci = FALSE, 
+             pub_styled = TRUE) %>% 
+    mutate(split_factor = sub1_factors$variables$eff_size$variable, 
+           plot_cap = paste(eff_size, significance, sep = ', '))
+  
+  ## Spearman's correlations 
+  
+  sub1_factors$corr_test <- sub1_factors$variables$correlation$variable %>% 
+    map_dfr(~correlate_variables(cov_data$clear_data[c('ipq_sub1', .x)] %>% 
+                                   filter(complete.cases(.)), 
+                                 variables = c('ipq_sub1', .x), 
+                                 what = 'correlation', 
+                                 type = 'spearman', 
+                                 ci = FALSE, 
+                                 pub_styled = TRUE)) %>% 
+    safely(mutate)(plot_cap = paste(eff_size, significance, sep = ', '), 
+                   plot_cap = stri_replace(plot_cap, 
[…]

### 2022-08-25 analysis scripts/ipq_sub2_factors.R

@@ -1,0 +1,117 @@
+# Relationship between common influential factors and sub2 value
+
+  insert_head()
+
+# container -------
+
+  sub2_factors <- list()
+  
+# globals -----
+
+  insert_msg('Globals')
+  
+  sub2_factors$variables <- an$cmm_lexicon %>% 
+    filter(variable %in% unique(c(an$cmm_variables$ipq_sub2, 
+                                  an$conf_variables))) %>% 
+    dlply(.(what))
+  
+# Serial comparisons and correlations -----
+  
+  insert_msg('Serial comparisons')
+  
+  ## descriptive stats
+  
+  sub2_factors$desc_stats <- sub2_factors$variables$eff_size$variable %>% 
+    map(~explore(cov_data$clear_data %>% 
+                   filter(!is.na(.data[[.x]])), 
+                 split_factor = .x, 
+                 variables = 'ipq_sub2', 
+                 what = 'table', 
+                 pub_styled = TRUE)) %>% 
+    map(reduce, left_join, by = 'variable') %>% 
+    map2(., sub2_factors$variables$eff_size$levels, 
+         ~set_names(.x, c('variable', .y))) %>% 
+    set_names(sub2_factors$variables$eff_size$variable)
+  
+  ## comparisons
+  
+  sub2_factors$factor_test <- 
+    list(split_factor = sub2_factors$variables$eff_size$variable, 
+         types = sub2_factors$variables$eff_size$eff_type) %>% 
+    pmap_dfr(compare_variables, 
+             cov_data$clear_data, 
+             variables = 'ipq_sub2', 
+             what = 'eff_size', 
+             ci = FALSE, 
+             pub_styled = TRUE) %>% 
+    mutate(split_factor = sub2_factors$variables$eff_size$variable, 
+           plot_cap = paste(eff_size, significance, sep = ', '))
+  
+  ## Spearman's correlations 
+  
+  sub2_factors$corr_test <- sub2_factors$variables$correlation$variable %>% 
+    map_dfr(~correlate_variables(cov_data$clear_data[c('ipq_sub2', .x)] %>% 
+                                   filter(complete.cases(.)), 
+                                 variables = c('ipq_sub2', .x), 
+                                 what = 'correlation', 
+                                 type = 'spearman', 
+                                 ci = FALSE, 
+                                 pub_styled = TRUE)) %>% 
+    safely(mutate)(plot_cap = paste(eff_size, significance, sep = ', '), 
+                   plot_cap = stri_replace(plot_cap, 
[…]

### 2022-08-25 analysis scripts/plots.R

@@ -1,0 +1,63 @@
+# Venn plots with the common influnetial factors and plots with the 
+# response-wise performance measures
+
+  insert_head()
+  
+# container -----
+  
+  mod_plots <- list()
+  
+# Performance measures: R-squared and RMSE ------
+  
+  insert_msg('Performance measures')
+  
+  ## data
+  
+  mod_plots$perf_data <- list(eln_mod = eln_mod, 
+                              lasso_mod = lasso_mod, 
+                              blass_mod = blass_mod) %>% 
+    map(~.x$cv_fit_stats) %>% 
+    map(reduce, rbind) %>% 
+    map(filter, statistic %in% c('RMSE', 'rsq')) %>% 
+    map2_dfr(., names(.), 
+             ~mutate(.x, 
+                     method = factor(.y, c(names(an$methods), 'eli_mod')), 
+                     response = factor(response, an$responses)))
+  
+  ## plotting 
+  
+  mod_plots$perf_plots <- list(cv_stats =  mod_plots$perf_data %>% 
+                                 dlply(.(response)), 
+                               plot_subtitle = an$responses %>% 
+                                 translate_var %>% 
+                                 paste('1-year follow-up', sep = ', ')) %>% 
+    pmap(plot_rsq_rmse,
+         y_factor = 'method') %>% 
+    map(~map(.x, 
+             ~.x + scale_y_discrete(labels = c(an$methods, 
+                                               c(eli_mod = 'Backward elimination')))))
+  
+# Numbers of selected factors: Venn plots -----
+  
+  insert_msg('Numbers of selected factors')
+  
+  mod_plots$venn_plots <- 
+    list(data = map(an$variables, 
+                    ~set_names(.x, an$methods[names(.x)])), 
+         plot_title = an$responses %>% 
+           translate_var %>% 
+           paste('1-year follow-up', sep = ', ')) %>% 
+    pmap(plot_n_venn, 
+         plot_subtitle = 'variables with non-zero coefficients', 
+         plot_tag = paste0('variables: n = ', 
+                           length(mod$variables), 
+                           ', observations: n = ', 
+                           nrow(mod$multi_tbl)), 
+         fill_color = set_names(an$method_colors, 
+                                an$methods), 
+         fontScale = 3, 
+         panel = FALSE)
+
+# END -----
+  
+  insert_tail()

### 2022-08-25 analysis.R

@@ -1,0 +1,103 @@
+# Detailed analysis of SES, SOCL-9 and KW IPQ
+# as a function of common influential paramaters identified by 
+# backwards elimination, Elastic Net and Bayesian Lasso modeling.
+
+# tools ------
+
+  library(plyr)
+  library(tidyverse)
+  library(exda)
+  library(stringi)
+  library(soucer)
+  library(clustTools)
+  library(furrr)
+
+  insert_head()
+  
+  source_all('./tools/project_tools.R')
+
+# analysis globals -----
+
+  insert_msg('Analysis globals')
+
+  an <- list()
+
+  ## variables selected by each algorithm
+  ## second and first term subsumed under one variable
+  
+  an$variables <- list(eln_mod = eln_mod, 
+                       lasso_mod = lasso_mod, 
+                       blass_mod = blass_mod) %>% 
+    map(~.x$inference) %>% 
+    transpose %>% 
+    map(compact) %>% 
+    map(~map(.x, filter, parameter != '(Intercept)')) %>% 
+    map(~map(.x, ~.x$variable) %>% 
+          map(stri_replace, fixed = '_sq', replacement = '') %>% 
+          map(stri_replace, fixed = 'ctss', replacement = 'ct_severity_score') %>% 
+          map(unique))
+  
+  ## common variables selected by at least two penalized modeling methods
+  ## determining the variable type and number of levels
+  
+  an$cmm_variables <- an$variables %>% 
+    map(reduce, intersect)
+  
+  an$conf_variables <- c('age', 
+                         'sex', 
+                         'sympt_present', 
+                         'lufo_red', 
+                         'ct_severity_any', 
+                         'diastolic_dysf', 
+                         'cat_WHO')
+  
+  an$cmm_lexicon <- 
+    tibble(variable = c(reduce(an$cmm_variables, union), 
+                        an$conf_variables)) %>% 
+    filter(!duplicated(variable)) %>% 
+    mutate(numeric = map_lgl(variable, 
+                             ~is.numeric(cov_data$clear_data[[.x]])), 
+           plot_type = ifelse(numeric, 'correlation', 'violin'), 
+           levels = map(variable, 
+                        ~levels(cov_data$clear_data[[.x]])), 
+           level_no = map_dbl(variable, 
+                              ~length(levels(cov_data$clear_data[[.x]]))), 
+           eff_type = ifelse(level_no == 2, 
[…]

### 2022-08-25 exec.R

@@ -1,0 +1,13 @@
+# Launches the entire pipeline
+
+  library(soucer)
+
+  print(source_all(c('import.R', 
+                     'exploration.R', 
+                     'modeling.R', 
+                     'analysis.R', 
+                     'paper.R'), 
+                   crash = FALSE, 
+                   message = TRUE))
+  
+  save.image()

### 2022-08-25 exploration scripts/cohort_characteristic.R

@@ -1,0 +1,83 @@
+# Baseline clinical characteristic of the study cohort and recovery status 
+# at the one-year follow-up
+
+  insert_head()
+  
+# container -----
+  
+  cohort <- list()
+  
+# globals -----
+  
+  insert_msg('Globals setup')
+  
+  cohort$analysis_tbl <- rbind(cov_data$clear_data %>% 
+                                 mutate(subset_var = 'cohort'), 
+                               cov_data$clear_data %>% 
+                                 mutate(subset_var = cat_WHO)) %>% 
+    mutate(subset_var = factor(subset_var, c('cohort', 'A', 'HM', 'HS')))
+  
+  plan('multisession')
+  
+# descriptive stats ------
+  
+  insert_msg('Descriptive stats')
+  
+  cohort$desc_stats <- expl$variables %>% 
+    future_map(~explore(cohort$analysis_tbl, 
+                 split_factor = 'subset_var', 
+                 variables = .x, 
+                 what = 'table', 
+                 pub_styled = TRUE), 
+               .options = furrr_options(seed = TRUE)) %>% 
+    map(reduce, 
+        left_join, 
+        by = 'variable') %>% 
+    map(set_names, 
+        c('variable', levels(cohort$analysis_tbl$subset_var)))
+  
+# Testing for differences between the severity strata groups -------
+  
+  insert_msg('Testing')
+  
+  cohort$test_results <- list(variables = expl$variables, 
+                              types = expl$test_type) %>% 
+    future_pmap(compare_variables,
+                cov_data$clear_data, 
+                split_factor = 'cat_WHO', 
+                what = 'test', 
+                ci = FALSE, 
+                pub_styled = TRUE, 
+                adj_method = 'none', 
+                .options = furrr_options(seed = TRUE)) %>% 
+    map(mutate, 
+        plot_cap = paste(eff_size, significance, sep = ', '))
+  
+# Plots of the psych variables in the severity groups ------
+  
+  insert_msg('Plots of the psych variable values in the severity groups')
+  
+  cohort$psy_plots <- list(variable = expl$variables$psych, 
+                           plot_title = expl$variables$psych %>% 
+                             translate_var(out_value = 'label_long'), 
+                           plot_subtitle = cohort$test_results$psych$plot_cap, 
+                           y_lab = expl$variables$psych %>% 
[…]

### 2022-08-25 exploration scripts/distribution.R

@@ -1,0 +1,186 @@
+# Distribution tests and plots for numeric variables ------
+
+  insert_head()
+
+# container ------
+
+  distr_tests <- list()
+  
+# globals ------
+  
+  insert_msg('Globals setup')
+  
+  distr_tests$num_vars <- map2(expl$variables, 
+                               expl$var_numeric, 
+                              ~.x[.y]) %>% 
+    reduce(c)
+  
+  plan('multisession')
+  
+# normality testing: identity transformation -------
+  
+  insert_msg('Normality testing, identity transformation')
+  
+  ## Shapiro - Wilk tests
+  
+  distr_tests$norm_test <- explore(cov_data$clear_data, 
+                                   variables = distr_tests$num_vars, 
+                                   what = 'normality', 
+                                   pub_styled = TRUE)
+  
+  ## QQ plots
+  
+  distr_tests$norm_plots <- explore(cov_data$clear_data, 
+                                    variables = distr_tests$num_vars, 
+                                    what = 'plots', 
+                                    type = 'qq', 
+                                    cust_theme = globals$common_theme) %>% 
+    map2(., translate_var(distr_tests$num_vars), 
+         ~.x + 
+           labs(title = .y, 
+                subtitle = .x$labels$tag %>% 
+                  stri_replace_all(fixed = '\n', replacement = ', ') %>% 
+                  stri_replace(fixed = 'Complete: ', replacement = '')) + 
+           theme(plot.tag = element_blank()))
+
+# normality testing: log transformation ------
+  
+  insert_msg('Normality testing, log transformation')
+  
+  ## Shapiro - Wilk tests
+  
+  distr_tests$norm_test_log <- 
+    explore(cov_data$clear_data %>% 
+              map_dfc(function(x) if(is.numeric(x)) log(x + 1) else x), 
+            variables = distr_tests$num_vars, 
+            what = 'normality', 
+            pub_styled = TRUE)
+  
+  ## QQ plots
+  
+  distr_tests$norm_plots_log <- 
+    explore(cov_data$clear_data %>% 
+              map_dfc(function(x) if(is.numeric(x)) log(x + 1) else x), 
+            variables = distr_tests$num_vars, 
+            what = 'plots', 
+            type = 'qq', 
[…]

### 2022-08-25 exploration scripts/missingness.R

@@ -1,0 +1,43 @@
+# Missingness fraction. Removal of high missing answer records (> 25%)
+
+  insert_head()
+  
+# container ------
+  
+  mis <- list()
+
+# calculating percents missing for variables -------
+  
+  insert_msg('Percentages of missing values for variables')
+  
+  mis$var_frac_mis <- cov_data$clear_data[globals$variables] %>% 
+    map_dbl(~sum(is.na(.x))) %>% 
+    tibble(variable = names(.), 
+           n_missing = .) %>% 
+    mutate(frac_missing = n_missing/nrow(cov_data$clear_data), 
+           perc_missing = frac_missing * 100) %>% 
+    mutate(var_lab = translate_var(variable))
+  
+# Plotting variable missingness -------
+  
+  insert_msg('Plotting variable missngness')
+  
+  mis$var_mis_plot <- mis$var_frac_mis %>%
+    filter(perc_missing > 0) %>%
+    ggplot(aes(x = perc_missing, 
+               y = reorder(var_lab, perc_missing))) + 
+    geom_bar(stat = 'identity', 
+             color = 'black', 
+             fill = 'steelblue') + 
+    geom_text(aes(label = signif(perc_missing, 2)), 
+              size = 2.75, 
+              hjust = -0.5) + 
+    globals$common_theme + 
+    theme(axis.title.y = element_blank()) + 
+    labs(title = 'Variable missingness', 
+         subtitle = paste('total participants: n =', nrow(cov_data$clear_data)), 
+         x = '% missing answers')
+  
+# END -----
+  
+  insert_tail()

### 2022-08-25 exploration scripts/psy_correlation.R

@@ -1,0 +1,87 @@
+# Correlation of the IPQ items, sum and subscores with other psy variables
+
+  insert_head()
+  
+# container -----
+  
+  psy_cor <- list()
+  
+# globals ------
+  
+  insert_msg('Globals: variables')
+  
+  psy_cor$var1 <- 
+    expl$variables$ipq[!expl$variables$ipq %in% c('ipq_sub1', 'ipq_sub2')]
+  
+  psy_cor$var2 <- globals$psy_variables
+  
+  psy_cor$pairs <- psy_cor$var1 %>% 
+    map(function(var1) map(psy_cor$var2, ~c(var1, .x))) %>% 
+    unlist(recursive = FALSE)
+  
+  psy_cor$pair_names <- psy_cor$pairs %>% 
+    map_chr(paste, collapse = '_')
+  
+  psy_cor$pairs <- set_names(psy_cor$pairs, 
+                             psy_cor$pair_names)
+  
+# serial correlation, Spearman test -----
+  
+  insert_msg('Spearman correlation')
+  
+  psy_cor$corr_results <- psy_cor$pairs %>% 
+    map_dfr(~correlate_variables(cov_data$clear_data[.x] %>% 
+                                   filter(complete.cases(.)), 
+                                 variables = .x, 
+                                 what = 'correlation', 
+                                 type = 'spearman', 
+                                 ci = TRUE)) %>% 
+    mutate(highlight = ifelse(p_value < 0.05, 'bold', 'plain'), 
+           significant = ifelse(p_value < 0.05, 'p < 0.05', 'ns'), 
+           significant = factor(significant, c('p < 0.05', 'ns')))
+  
+# Plotting the correlation matrix -------
+  
+  insert_msg('Plotting of the correlation matrix')
+  
+  psy_cor$corr_mtx_plot <- psy_cor$corr_results %>% 
+    mutate(variable1 = ifelse(variable1 %in% names(globals$ipq_sublabs), 
+                              globals$ipq_sublabs[variable1], 
+                              translate_var(variable1))) %>% 
+    ggplot(aes(x = variable1, 
+               y = variable2, 
+               size = abs(estimate), 
+               fill = estimate)) + 
+    geom_point(shape = 21, 
+               color = 'black') + 
+    geom_text(aes(fontface = highlight, 
+                  label = signif(estimate, 2), 
+                  color = significant), 
+              size = 2.75, 
+              vjust = -1.5) + 
[…]

### 2022-08-25 exploration scripts/score_coherence.R

@@ -1,0 +1,168 @@
+# IPQ score coherence with Cronbach's alpha
+
+  insert_head()
+  
+# container ------
+  
+  ipq_co <- list()
+  
+# globals ------
+  
+  insert_msg('Analysis tables and variable pairs')
+  
+  ipq_co$analysis_tbl <- list(ipq_total = paste0('ipq_q', 1:8), 
+                              ipq_sub1 = paste0('ipq_q', c(1, 2, 5, 6, 8)), 
+                              ipq_sub2 = paste0('ipq_q', c(3, 4, 7))) %>% 
+    map(~select(cov_data$clear_data, 
+                ID, 
+                all_of(.x))) %>%
+    map(~filter(.x, complete.cases(.x))) %>% 
+    map(select, - ID)
+  
+  ipq_co$pairs <- ipq_co$analysis_tbl %>%
+    map(names) %>% 
+    map(~combn(.x, 2, simplify = FALSE))
+  
+# pair-wise correlations of the sub-scores, Spearman test ------
+  
+  insert_msg('Pair-wise correlation coefs')
+  
+  ipq_co$corr_results <- 
+    list(data = ipq_co$analysis_tbl, 
+         var_lst = ipq_co$pairs) %>% 
+    pmap(function(data, var_lst) map_dfr(var_lst, 
+                                         ~correlate_variables(data, 
+                                                              variables = .x, 
+                                                              type = 'spearman', 
+                                                              ci = TRUE, 
+                                                              pub_styled = TRUE))) %>% 
+    map(mutate, 
+        eff_size = stri_replace(eff_size, fixed = 'rho', replacement = '\u03C1'), 
+        plot_cap = paste(eff_size, significance, sep = ', '))
+  
+# single correlation plots ------
+  
+  insert_msg('Single correlation plots')
+  
+  ipq_co$corr_plots <- 
+    list(data = ipq_co$analysis_tbl, 
+         var_lst = ipq_co$pairs, 
+         pt_colors = globals$ipq_colors, 
+         stats = ipq_co$corr_results) %>% 
+    pmap(function(data, var_lst, pt_colors, stats) map2(set_names(var_lst, 
+                                                                  map(var_lst, paste, collapse = '_')), 
+                                                        stats$plot_cap,  
+                                                        ~plot_correlation(data = data, 
+                                                                          variables = .x, 
+                                                                          type = 'correlation', 
+                                                                          point_color = pt_colors, 
+                                                                          cust_theme = globals$common_theme, 
[…]

### 2022-08-25 exploration scripts/score_pca.R

@@ -1,0 +1,101 @@
+# Principal component analysis of the IPQ sub-scores. Median centering
+
+  insert_head()
+  
+# container -----
+  
+  ipq_pca <- list()
+  
+# globals -----
+  
+  insert_msg('Globals: analysis table')
+  
+  ipq_pca$analysis_tbl <- cov_data$clear_data %>% 
+    select(ID, starts_with('ipq_q')) %>% 
+    column_to_rownames('ID') %>% 
+    center_data(type = 'median', 
+                complete_cases = TRUE) %>% 
+    set_names(globals$ipq_sublabs[names(.)])
+  
+# PCA: 4 dimensions ------
+  
+  insert_msg('4D PCA')
+  
+  ## PCA object
+  
+  ipq_pca$pca_obj <- reduce_data(ipq_pca$analysis_tbl, 
+                                 kdim = 4, 
+                                 red_fun = 'pca')
+  
+# PCA plots ------
+  
+  insert_msg('Base PCA plots')
+
+  ## scree plot
+  
+  ipq_pca$scree_plot <- plot(ipq_pca$pca_obj, 
+                             type = 'scree', 
+                             cust_theme = globals$common_theme) + 
+    expand_limits(y = 0) + 
+    labs(title = 'PCA variance per component', 
+         subtitle = '4D PCA, median centering', 
+         tag = paste('\nn =', nrow(ipq_pca$analysis_tbl)))
+  
+  ## PCA scores
+  
+  ipq_pca$score_plot <- plot(ipq_pca$pca_obj, 
+                             type = 'scores', 
+                             cust_theme = globals$common_theme) + 
+    geom_hline(yintercept = 0, 
+               linetype = 'dashed') + 
+    geom_vline(xintercept = 0, 
+               linetype = 'dashed') + 
+    labs(title = 'PCA scores', 
+         subtitle = '4D PCA, median centering', 
+         tag = paste('\nn =', nrow(ipq_pca$analysis_tbl)))
+  
+  ## loadings
+  
+  ipq_pca$loadings_plot <- plot(ipq_pca$pca_obj, 
+                                type = 'loadings', 
+                                cust_theme = globals$common_theme) + 
+    geom_hline(yintercept = 0, 
+               linetype = 'dashed') + 
+    geom_vline(xintercept = 0, 
+               linetype = 'dashed') + 
+    labs(title = 'PCA loadings', 
+         subtitle = '4D PCA, median centering', 
[…]

### 2022-08-25 exploration.R

@@ -1,0 +1,113 @@
+# Exploratory data analysis
+# Remova of participants with > 25% missing answers if any detected
+
+# tools -------
+
+  library(plyr)
+  library(tidyverse)
+  library(stringi)
+  library(exda)
+  library(soucer)
+  library(furrr)
+  library(psych)
+  library(clustTools)
+
+  insert_head()
+  
+  source_all('./tools/project_tools.R')
+  
+# common variables and parameters ------
+  
+  insert_msg('Exploration globals')
+  
+  expl <- list()
+  
+  ## variables
+  
+  expl$variables <- list(baseline = c('sex', 
+                                      'age', 
+                                      'smoking_history', 
+                                      'weight_class', 
+                                      'comorb_present', 
+                                      'no_comorb', 
+                                      'cardiovascular_comorb', 
+                                      'hypertension_comorb', 
+                                      'endometabolic_comorb', 
+                                      'hyperchol_comorb', 
+                                      'diabetes_comorb', 
+                                      'gastro_comorb', 
+                                      'malingancy_comorb', 
+                                      'respi_comorb', 
+                                      'ckd_comorb', 
+                                      'immdef_comorb', 
+                                      'WHO', 
+                                      'cat_WHO'), 
+                         fup = c('sympt_present', 
+                                 'sympt_number', 
+                                 'fatigue_sympt', 
+                                 'dyspnoe_sympt', 
+                                 'cough_sympt', 
+                                 'sleep_sympt', 
+                                 'night_sweat_sympt', 
+                                 'anosmia_sympt', 
+                                 'derma_sympt', 
+                                 'gastro_sympt', 
+                                 'hair_loss_sympt', 
+                                 'Chalder_FS', 
+                                 'Chalder_FS_bimodal', 
+                                 'lufo_red', 
+                                 'ct_severity_any', 
+                                 'ct_severity_score', 
+                                 'diastolic_dysf', 
+                                 'smwd', 
+                                 'smwd_dref', 
+                                 'rehabilitation'), 
+                         ipq = c('ipq_total',
+                                   'ipq_sub1',
+                                   'ipq_sub2', 
+                                   'ipq_q1', 
+                                   'ipq_q2', 
+                                   'ipq_q3', 
[…]

### 2022-08-25 import.R

@@ -1,0 +1,416 @@
+# Import of the patient's data cleared before during the COVILD-12 month project
+# See: https://github.com/PiotrTymoszuk/CovILD-Plus for the import script code
+# A subset of participants is selected with at least 
+# one COVID-19-related symptom
+
+# tools -------
+
+  library(plyr)
+  library(tidyverse)
+  library(readxl)
+  library(stringi)
+  library(soucer)
+  library(rlang)
+
+  insert_head()
+  
+  source_all('./tools/project_tools.R')
+  
+# containers -----
+  
+  cov_data <- list()
+  globals <- list()
+  
+# Setup of globals -------
+  
+  insert_msg('Graphics globals setup')
+  
+  ## modeling variables
+  
+  globals$variables <- c('sex', 
+                         'age', 
+                         'age_sq', 
+                         'smoking_history', 
+                         'weight_class', 
+                         'comorb_present', 
+                         'no_comorb', 
+                         'no_comorb_sq', 
+                         'cardiovascular_comorb', 
+                         'hypertension_comorb', 
+                         'endometabolic_comorb', 
+                         'hyperchol_comorb', 
+                         'diabetes_comorb', 
+                         'respi_comorb', 
+                         'ckd_comorb', 
+                         'gastro_comorb', 
+                         'malingancy_comorb', 
+                         'immdef_comorb', 
+                         'cat_WHO', 
+                         'WHO', 
+                         'WHO_sq', 
+                         'sympt_present', 
+                         'sympt_number', 
+                         'sympt_number_sq', 
+                         'fatigue_sympt', 
+                         'dyspnoe_sympt', 
+                         'cough_sympt', 
+                         'sleep_sympt', 
+                         'night_sweat_sympt', 
+                         'anosmia_sympt', 
+                         'derma_sympt', 
+                         'gastro_sympt', 
+                         'hair_loss_sympt', 
+                         'Chalder_FS', 
+                         'Chalder_FS_sq', 
+                         'Chalder_FS_bimodal', 
+                         'smwd', 
+                         'smwd_sq', 
+                         'smwd_dref', 
+                         'smwd_dref_sq', 
+                         'lufo_red', 
+                         'ct_severity_any', 
[…]

### 2022-08-25 modeling scripts/bayes_lasso.R

@@ -1,0 +1,146 @@
+# Modeling with Bayesian Lasso
+
+  insert_head()
+  
+# container -----
+  
+  blass_mod <- list()
+  
+  registerDoParallel(7)
+  
+# model construction -------
+  
+  insert_msg('Model construction')
+  
+  ## tune grid
+  
+  blass_mod$tune_grid <- data.frame(sparsity = seq(0.3, 0.7, by = 0.1))
+  
+  ## models
+  
+  set.seed(1234)
+  
+  blass_mod$models <- mod$multi_form %>% 
+    map(train, 
+        data = mod$multi_tbl, 
+        method = 'blasso', 
+        metric = 'RMSE', 
+        trControl = mod$tr_control, 
+        tuneGrid = blass_mod$tune_grid) %>% 
+    map(as_caretx)
+  
+# Residuals -----
+  
+  insert_msg('Model residuals')
+  
+  blass_mod$resid_tbl <- blass_mod$models %>% 
+    map(resid)
+  
+  ## normality testing
+  
+  blass_mod$resid_norm <- blass_mod$resid_tbl %>% 
+    map(~map(.x, 
+             explore, 
+             what = 'normality', 
+             variables = '.resid') %>% 
+          map2_dfr(., names(.), ~mutate(.x, mod_type = .y))) %>% 
+    map2_dfr(., names(.), ~mutate(.x, response = .y))
+  
+  ## residual plots
+  
+  blass_mod$resid_plots <- list(x = blass_mod$models, 
+                                plot_title = names(blass_mod$models) %>% 
+                                  translate_var %>% 
+                                  map(paste, c('training', 'CV'), sep = ', ')) %>% 
+    pmap(plot, 
+         type = 'diagnostic', 
+         cust_theme = globals$common_theme) %>% 
+    transpose
+  
+# Fit stats ------
+  
+  insert_msg('Fit stats')
+  
+  ## stat table
+  
+  blass_mod$cv_fit_stats <- blass_mod$models %>% 
+    map(summary, plain = TRUE) %>% 
+    map2(., names(.), 
+             ~mutate(.x, response = .y)) %>% 
+    map(mutate, prediction = factor(prediction, c('train', 'cv')))
+  
+  blass_mod$plot_cap <- blass_mod$cv_fit_stats %>% 
+    map(as.data.frame) %>% 
+    map(~paste0('R\u00B2 = ', signif(.x[4, 2], 2), 
+                ', RMSE = ', signif(.x[3, 2], 2))) %>% 
[…]

### 2022-08-25 modeling scripts/elanet.R

@@ -1,0 +1,194 @@
+# Elastic net/LASSO modeling 
+
+  insert_head()
+  
+# container -----
+  
+  eln_mod <- list()
+  
+# globals ------
+  
+  insert_msg('Globals setup')
+  
+  ## response vectors and explanatory variable matrices
+  
+  eln_mod$response <- mod$responses %>% 
+    map(~mod$multi_tbl[[.x]])
+
+  eln_mod$variables <- model.matrix(~., data = mod$multi_tbl[globals$variables])
+  
+  ## alpha and n repetitions
+  
+  eln_mod$alpha <- 0.5
+  eln_mod$n_rep <- 200
+  
+  ## folds for CV
+  
+  set.seed(1234)
+  
+  eln_mod$folds <- 1:eln_mod$n_rep %>% 
+    map(~createFolds(y = eln_mod$response[[1]], 
+                     k = 10, 
+                     list = FALSE, 
+                     returnTrain = TRUE)) %>% 
+    set_names(paste0('rep_', 1:eln_mod$n_rep))
+
+  ## parallel backends
+  
+  plan('multisession')
+  
+# lambda finding -------
+  
+  insert_msg('Lambda finding')
+  
+  ## model construction
+  
+  eln_mod$lambda_tune <- eln_mod$response %>% 
+    map(function(resp) eln_mod$folds %>% 
+          future_map(~cv.glmnet(x = eln_mod$variables, 
+                                y = resp, 
+                                foldid = .x, 
+                                type.measure = 'default', 
+                                family = 'gaussian', 
+                                alpha = eln_mod$alpha), 
+                     .options = furrr_options(seed = TRUE)))
+
+  ## extracting the optimal lambdas
+  
+  eln_mod$lambda_tbl <- eln_mod$lambda_tune %>% 
+    map(~map(.x, 
+             ~as_tibble(.x[c('lambda', 'cvm', 'cvup', 'cvlo', 'lambda.min')]))) %>% 
+    map(~map_dfr(.x, filter, lambda == lambda.min))
+  
+  eln_mod$opt_lambda <- eln_mod$lambda_tbl %>% 
+    map(filter, cvm == min(cvm)) %>% 
+    map2_dfr(., names(.), ~mutate(.x, response = .y))
+  
+  eln_mod$tune_grids <- eln_mod$opt_lambda$lambda %>% 
+    map(~data.frame(alpha = eln_mod$alpha, 
+                    lambda = .x)) %>% 
[…]

### 2022-08-25 modeling scripts/lasso.R

@@ -1,0 +1,194 @@
+# LASSO modeling 
+
+  insert_head()
+  
+# container -----
+  
+  lasso_mod <- list()
+  
+# globals ------
+  
+  insert_msg('Globals setup')
+  
+  ## response vectors and explanatory variable matrices
+  
+  lasso_mod$response <- mod$responses %>% 
+    map(~mod$multi_tbl[[.x]])
+
+  lasso_mod$variables <- model.matrix(~., data = mod$multi_tbl[globals$variables])
+  
+  ## alpha and n repetitions
+  
+  lasso_mod$alpha <- 1
+  lasso_mod$n_rep <- 200
+  
+  ## folds for CV
+  
+  set.seed(1234)
+  
+  lasso_mod$folds <- 1:lasso_mod$n_rep %>% 
+    map(~createFolds(y = lasso_mod$response[[1]], 
+                     k = 10, 
+                     list = FALSE, 
+                     returnTrain = TRUE)) %>% 
+    set_names(paste0('rep_', 1:lasso_mod$n_rep))
+
+  ## parallel backends
+  
+  plan('multisession')
+  
+# lambda finding -------
+  
+  insert_msg('Lambda finding')
+  
+  ## model construction
+  
+  lasso_mod$lambda_tune <- lasso_mod$response %>% 
+    map(function(resp) lasso_mod$folds %>% 
+          future_map(~cv.glmnet(x = lasso_mod$variables, 
+                                y = resp, 
+                                foldid = .x, 
+                                type.measure = 'default', 
+                                family = 'gaussian', 
+                                alpha = lasso_mod$alpha), 
+                     .options = furrr_options(seed = TRUE)))
+
+  ## extracting the optimal lambdas
+  
+  lasso_mod$lambda_tbl <- lasso_mod$lambda_tune %>% 
+    map(~map(.x, 
+             ~as_tibble(.x[c('lambda', 'cvm', 'cvup', 'cvlo', 'lambda.min')]))) %>% 
+    map(~map_dfr(.x, filter, lambda == lambda.min))
+  
+  lasso_mod$opt_lambda <- lasso_mod$lambda_tbl %>% 
+    map(filter, cvm == min(cvm)) %>% 
+    map2_dfr(., names(.), ~mutate(.x, response = .y))
+  
+  lasso_mod$tune_grids <- lasso_mod$opt_lambda$lambda %>% 
+    map(~data.frame(alpha = lasso_mod$alpha, 
+                    lambda = .x)) %>% 
[…]

### 2022-08-25 modeling scripts/uni_variate.R

@@ -1,0 +1,148 @@
+# Uni-variate modeling
+
+  insert_head()
+  
+# container ------
+  
+  uni_mod <- list()
+  
+# model construction: normalized and non-normalized models -------
+  
+  insert_msg('Model construction')
+  
+  ## without normalization
+  
+  uni_mod$models <- mod$responses  %>% 
+    map(function(resp) set_names(globals$variables, 
+                                 globals$variables) %>% 
+          map(~make_lm(data = mod$multi_tbl, 
+                       response = resp, 
+                       indep_variable = .x, 
+                       mod_fun = lm, 
+                       family = NULL)))
+  
+# model assumptions ------
+  
+  insert_msg('Model assumptions')
+  
+  uni_mod$assumptions <- uni_mod$models %>% 
+    map(~map(.x, 
+             summary, 
+             type = 'assumptions', 
+             type.predict = 'response') %>% 
+          map2_dfr(., names(.), ~mutate(.x, variable = .y))) %>% 
+    map2(., names(.), ~mutate(.x, response = .y))
+
+# model fit statistics -----
+  
+  insert_msg('Model fit stats')
+  
+  uni_mod$fit_stats <- uni_mod$models %>% 
+    map(~map(.x, 
+             summary, 
+             type = 'fit') %>% 
+          map2_dfr(., names(.), ~mutate(.x, variable = .y))) %>% 
+    map(mutate, 
+        rsq_eff_size = cut(raw_rsq, 
+                           c(-Inf, 0.01, 0.09, 0.25, Inf), 
+                           c('none', 'weak', 'moderate', 'large')))
+  
+# model inference for normalized and non-normalized models ------
+  
+  insert_msg('Model inference')
+  
+  uni_mod$inference <- uni_mod$models %>% 
+    map(~map(.x, 
+             summary, 
+             type = 'inference'))
+
+  ## filling up with the baseline levels
+  
+  uni_mod$inference <- uni_mod$inference %>% 
+    map(~map2_dfr(., names(.), 
+                  ~mutate(.x, 
+                          level_mod = levels(cov_data$clear_data[[.y]])))) %>% 
+    map(mutate, 
+        level_mod = ifelse(is.na(level_mod), 
+                           0, level_mod))
+
+# Plots of R squares ------
+  
+  insert_msg('R square value plots')
+  
+  uni_mod$rsq_plots <- uni_mod$fit_stats %>% 
+    map2(., translate_var(names(.)), 
+         ~ggplot(.x, 
[…]

### 2022-08-25 modeling.R

@@ -1,0 +1,96 @@
+# Analysis of differences in mental health status as a function
+# of baseline factors, symptoms and cardiopulmonary features
+# responses: the concern, identity, time (ipq_sub1) and 
+# the control, coherence (ipq_sub2) components of the IPQ questionnaire
+# and the total IPQ item sum.
+# Square roots of them used because of improved normality
+# methods: 
+# (1) uni-variate Gaussian-family modeling, 
+# (2) multi-variate backward elimination Gaussian-family modeling
+# (3) multi-variate elastic net modeling
+# multi-variate models are validated by 10-fold CV
+
+# tools -----
+
+  library(plyr)
+  library(tidyverse)
+  library(stringi)
+  library(lmqc)
+  library(caret)
+  library(caretExtra)
+  library(soucer)
+  library(glmnet)
+  library(doParallel)
+
+  insert_head()
+  
+  source_all('./tools/project_tools.R')
+  
+# common modeling globals ------
+  
+  insert_msg('Common modeling globals')
+  
+  mod <- list()
+
+  mod$responses <- c(ipq_sub1 = 'sqrt_ipq_sub1', 
+                     ipq_sub2 = 'sqrt_ipq_sub2', 
+                     ipq_total = 'sqrt_ipq_total')
+  
+  ## tables for multi-variate modeling with complete cases only
+  ## normalization with mean centering
+  
+  mod$multi_tbl <- cov_data$clear_data %>% 
+    select(ID, 
+           all_of(unname(mod$responses)), 
+           all_of(globals$variables)) %>% 
+    filter(complete.cases(.)) %>% 
+    map_dfc(function(x) if(is.numeric(x)) scale(x)[, 1] else x) %>% 
+    column_to_rownames('ID')
+
+  ## full model formulas
+  
+  mod$multi_form <- mod$responses %>% 
+    map(make_lm, 
+        data = mod$multi_tbl, 
+        indep_variable = globals$variables, 
+        mod_fun = lm, 
+        family = NULL) %>% 
+    map(formula)
+  
+  ## training control object for caret models
+  
+  mod$tr_control <- trainControl(method = 'repeatedcv', 
+                                 number = 10, 
+                                 repeats = 10, 
+                                 savePredictions = 'final')
+  
+  ## level n numbers
+  
+  mod$level_n <- set_names(globals$variables, 
[…]

### 2022-08-25 paper scripts/figures.R

@@ -1,0 +1,309 @@
+# Paper and supplement figures
+
+  insert_head()
+  
+# containers -----
+  
+  paper_fig <- list()
+  suppl_fig <- list()
+  
+# Figure 1: study CONSORT diagram ------
+  
+  insert_msg('Figure 1: study consort')
+  
+  paper_fig$consort <- 
+    plot_grid(ggdraw() + 
+                draw_image('./study consort/consort_diagram.png')) %>% 
+    as_figure(label = 'figure_1_consort', 
+              ref_name = 'consort', 
+              caption = 'Flow diagram of study enrollment and analysis inclusion process.', 
+              w = 90, 
+              h = 90 * 2766/1728)
+  
+# Figure 2: Key factors impacting on disease perception ------
+  
+  insert_msg('Figure 2: key factors impacting on disease perception')
+  
+  ## top panel: Venn plot
+  
+  paper_fig$key_factors$upper_panel <- 
+    plot_grid(mod_plots$venn_plots$ipq_total$venn, 
+              plot_grid(mod_plots$venn_plots$ipq_total$legend_grob, 
+                        ggdraw() + 
+                          draw_text(paste(an$cmm_variables$ipq_total %>% 
+                                            translate_var(out_value = 'label_long'), 
+                                          collapse = '\n'), 
+                                    hjust = 0, 
+                                    x = 0.1, 
+                                    size = 8), 
+                        ncol = 2, 
+                        rel_widths = c(0.4, 0.6)), 
+              ncol = 2, 
+              rel_widths = c(0.5, 0.5))
+  
+  ## middle panel: violin plots
+  
+  paper_fig$key_factors$middle_panel <- 
+    ipq_factors$violin_plots[c('fatigue_sympt', 
+                               'respi_comorb', 
+                               'hair_loss_sympt')] %>% 
+    map(~.x + theme(legend.position = 'none')) %>% 
+    plot_grid(plotlist = ., 
+              ncol = 3, 
+              align = 'hv')
+  
+  ## bottom panel: correlations
+  
+  paper_fig$key_factors$bottom_panel <- 
+    plot_grid(ipq_factors$corr_plots$Chalder_FS, 
+              ncol = 3, 
+              align = 'hv')
+
+  ## entire figure
+  
+  paper_fig$key_factors <- plot_grid(paper_fig$key_factors$upper_panel, 
+                                     ggdraw(), 
+                                     paper_fig$key_factors$middle_panel, 
[…]

### 2022-08-25 paper scripts/render.R

@@ -1,0 +1,37 @@
+# Renders the manuscript and the supplementary material
+
+  insert_head()
+  
+# supplementary material ------
+  
+  insert_msg('Rendering the supplements')
+  
+  render('./paper/markdown/supplementary_material.Rmd', 
+         output_format = word_document2(number_sections = FALSE, 
+                                        reference_docx = 'ms_template.docx'), 
+         output_dir = './paper')
+  
+# paper -----
+  
+  insert_msg('Rendering the paper')
+  
+  my_word <- function(...) {
+    
+    form <- word_document2(number_sections = FALSE, 
+                           reference_docx = 'ms_template.docx')
+    
+    form$pandoc$lua_filters <- c(form$pandoc$lua_filters, 
+                                 'scholarly-metadata.lua', 
+                                 'author-info-blocks.lua')
+    
+    form
+    
+  }
+  
+  render('./paper/markdown/manuscript.Rmd', 
+         output_format = my_word(), 
+         output_dir = './paper')
+  
+# END -----
+  
+  insert_tail()

### 2022-08-25 paper scripts/tables.R

@@ -1,0 +1,228 @@
+# Paper and supplement tables
+
+  insert_head()
+  
+# containers -----
+  
+  paper_tbl <- list()
+  suppl_tbl <- list()
+
+# Table 1: characteristic of the study cohort -----
+  
+  insert_msg('Table 1: baseline cohort characteristic')
+  
+  paper_tbl$baseline <- left_join(cohort$desc_stats$baseline, 
+                                  cohort$test_results$baseline[c('variable', 
+                                                                 'significance', 
+                                                                 'eff_size')], 
+                                  by = 'variable') %>% 
+    filter(!variable %in% c('cat_WHO', 'no_comorb')) %>% 
+    format_summ_tbl(rm_n = TRUE) %>% 
+    rbind(tibble(variable = 'n, participants', 
+                 cohort = pap$n_baseline$n[1], 
+                 A = pap$n_baseline$n[2], 
+                 HM = pap$n_baseline$n[3], 
+                 HS = pap$n_baseline$n[4], 
+                 significance = NA, 
+                 eff_size = NA), 
+          .) %>% 
+    mutate(variable = stri_replace(variable, 
+                                   fixed = ', %', 
+                                   replacement = ''), 
+           variable = ifelse(variable == 'Sex', 'Male sex', variable)) %>%
+    map_dfc(stri_replace, regex = '^.*\\nmale:\\s{1}', replacement = '') %>% 
+    set_names(c('Variable', 
+                'Cohort', 
+                'Ambulatory COVID-19', 
+                'Moderate COVID-19', 
+                'Severe COVID-19', 
+                'Significance', 
+                'Effect size'))
+  
+  paper_tbl$baseline <- 
+    as_mdtable(paper_tbl$baseline , 
+               label = 'cohort_baseline', 
+               ref_name = 'cohort_baseline', 
+               caption = 'Baseline demographic and clinical characteristic of the study cohort.')
+  
+# Table 2: IPQ components and sum in the cohort and severity groups -----
+  
+  insert_msg('IPQ in the severity groups and the cohort')
+  
+  paper_tbl$psych <- left_join(cohort$desc_stats$ipq, 
+                               cohort$test_results$ipq[c('variable', 
+                                                         'significance', 
+                                                         'eff_size')], 
+                               by = 'variable') %>% 
+    filter(variable %in% c(paste0('ipq_q', 1:8), 'ipq_total')) %>% 
[…]

### 2022-08-25 paper scripts/values.R

@@ -1,0 +1,438 @@
+# Values used in the manuscript text
+
+  insert_head()
+  
+# container -----
+  
+  paper_val <- list()
+  
+# Participant n numbers ------
+  
+  insert_msg('N numbers of participants')
+  
+  paper_val$n_fup_missing <- 
+    mdexpr(145 - nrow(cov_data$data_tbl), 
+           ref_name = 'n_fup_missing', 
+           caption = 'Participnts with the missing one-year follow-up.')
+  
+  paper_val$n_fup_complete <- 
+    mdexpr(nrow(cov_data$data_tbl), 
+           ref_name = 'n_fup_complete', 
+           caption = 'Participants having complete the one-year follow-up.')
+  
+  paper_val$n_no_recovery <- 
+    mdexpr(length(cov_data$cov_residuals), 
+           ref_name = 'n_no_recovery', 
+           caption = 'Participants with incomplete recovery.')
+  
+  paper_val$n_ipq_complete <- 
+    mdexpr(length(cov_data$ipq_complete), 
+           ref_name = 'n_ipq_complete', 
+           caption = 'Participants with complete IPQ answers.')
+  
+  paper_val$n_analysis <- 
+    mdexpr(nrow(cov_data$clear_data), 
+           ref_name = 'n_analysis', 
+           caption = 'Participants in the analysis.')
+  
+# Numbers of factors included in clustering analysis -------
+  
+  insert_msg('N numbers and names of clustering factors')
+  
+  paper_val$n_clust_factors <- 
+    mdexpr(clustTools::nobs(ipq_clust$clust_obj)$variables, 
+           ref_name = 'n_clust_factors', 
+           caption = 'Number of clustering factors.')
+  
+  paper_val$clust_factors <- 
+    mdexpr(colnames(model.frame(ipq_clust$clust_obj)) %>% 
+             translate_var(out_value = 'label_long') %>% 
+             stri_replace(fixed = 'Illness perception, ', 
+                          replacement = '') %>% 
+             paste(collapse = ', '), 
+           ref_name = 'clust_factors', 
+           caption = 'Names of clustering factors')
+  
+  paper_val$clust_factors_and <- 
+    mdexpr(colnames(model.frame(ipq_clust$clust_obj)) %>% 
+             translate_var(out_value = 'label_long') %>% 
+             stri_replace(fixed = 'Illness perception, ', 
[…]

### 2022-08-25 paper.R

@@ -1,0 +1,59 @@
+# Paper-rendering scripts
+
+# tools -----
+
+  library(plyr)
+  library(tidyverse)
+  library(stringi)
+  library(soucer)
+  library(cowplot)
+  library(figur)
+  library(rmarkdown)
+  library(knitr)
+  library(bookdown)
+  library(flextable)
+  library(writexl)
+  library(clustTools)
+
+  insert_head()
+  
+  source_all('./tools/project_tools.R')
+  
+# globals -----
+  
+  insert_msg('Paper globals')
+  
+  pap <- list()
+  
+  ## n numbers
+  
+  pap$n_baseline <- cohort$analysis_tbl %>% 
+    count(subset_var)
+  
+  pap$n_clust <- ipq_clust$clust_obj$clust_assignment %>% 
+    count(clust_id)
+  
+  ## common scales for the estimate plots
+  
+  pap$est_limits <- c(-0.35, 0.49)
+  
+  pap$est_fills <- scale_fill_gradient2(low = 'steelblue', 
+                                        high = 'firebrick', 
+                                        mid = 'white', 
+                                        midpoint = 0, 
+                                        limits = pap$est_limits, 
+                                        name = expression(beta))
+  
+# paper scripts -------
+
+  insert_msg()
+  
+  c('./paper scripts/tables.R', 
+    './paper scripts/figures.R', 
+    './paper scripts/values.R', 
+    './paper scripts/render.R') %>% 
+    source_all(message = TRUE, crash = TRUE)
+  
+# END ------
+  
+  insert_tail()

### 2022-08-25 tools/project_tools.R

@@ -1,0 +1,550 @@
+# A medley of functional project tools
+
+  library(plyr)
+  library(tidyverse)
+  library(rlang)
+  library(cowplot)
+  library(glue)
+  library(ggtext)
+  library(nVennR)
+  library(cowplot)
+  library(magick)
+  
+# data import and transformation -----
+
+  outer_rbind <- function(tbl1, tbl2) {
+    
+    ## binds two data frames by rows, missing variables are filled with NA
+    
+    ## missing variables
+    
+    miss1 <- names(tbl2)[!names(tbl2) %in% names(tbl1)]
+    miss2 <- names(tbl1)[!names(tbl1) %in% names(tbl2)]
+    
+    ## filling the tables
+    
+    for(i in miss1){
+      
+      tbl1 <- tbl1 %>% 
+        mutate(!!sym(i) := NA)
+      
+    }
+    
+    for(i in miss2){
+      
+      tbl2 <- tbl2 %>% 
+        mutate(!!sym(i) := NA)
+      
+    }
+    
+    return(rbind(tbl1, tbl2))
+    
+  }
+  
+  fill_forward <- function(data_chunk, variable) {
+    
+    ## if complete recovery was observed at the 6 month visit and the one-year
+    ## follow-up is missing, the last non-missing is filled forward
+    
+    data_chunk <- data_chunk %>% 
+      arrange(time)
+    
+    
+    last_rec <- data_chunk[nrow(data_chunk), variable]
+    
+    if(!is.na(last_rec)) {
+      
+      return(data_chunk)
+      
+    }
+    
+    for(i in nrow(data_chunk):1) {
+      
+      last_full <- data_chunk[nrow(data_chunk) - i, variable]
+      
+      if(!is.na(last_rec)) {
+        
+        break
+        
+      }
+      
+    }
+
+    data_chunk[nrow(data_chunk), variable] <- last_full
+    
+    data_chunk
+    
+  }
+
+# variable:label translation, color setup -----
+  
+  translate_var <- function(variable, 
+                            key = 'variable', 
+                            out_value = 'label', 
+                            dict = globals$var_lexicon, 
+                            unit = FALSE) {
+    
+    naming_vec <- dict[[out_value]]
+
+    if(unit) {
+      
+      naming_vec <- ifelse(is.na(dict[['unit']]), 
+                           naming_vec, 
+                           paste(naming_vec, dict[['unit']], sep = ', '))
+      
+    }
+    
[…]

### 2023-01-09 paper/markdown/manuscript.Rmd
Revision

@@
 ---
-title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19"
-subtitle: "Manuscript file"
+title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19 in a cross-sectional prospective cohort study"
+subtitle: "Illness perception one year after COVID-19"
 date: "`r format(Sys.time(), '%Y-%m-%d')`"
@@
 author: 
-  - Katharina Huefner:
+  - Katharina Hüfner, MD:
       institute: [psych]
-  - Piotr Tymoszuk:
+  - Piotr Tymoszuk, PhD:
       institute: [daas]
-  - Sabina Sahanic:
+  - Sabina Sahanic, MD:
       institute: [int2]
-  - Anna Luger:
+  - Anna Luger, MD:
       institute: [radio]
-  - Anna Boehm:
+  - Anna Boehm, MD:
       institute: [int2]
-  - Alex Pizzini:
+  - Alex Pizzini, MD:
       institute: [int2]
-  - Christoph Schwabl:
+  - Christoph Schwabl, MD:
       institute: [radio]
-  - Sabine Koppelstätter:
+  - Sabine Koppelstätter, MD:
       institute: [int2]
-  - Katharina Kurz:
+  - Katharina Kurz, MD:
       institute: [int2]
-  - Malte Asshoff:
+  - Malte Asshoff, MD:
       institute: [int2]
-  - Birgit Mosheimer-Feistritzer:
+  - Birgit Mosheimer-Feistritzer, MD:
       institute: [int2]
-  - Maximilian Coen:
+  - Bernhard Pfeifer, PhD:
+      institute: [umit]
+  - Verena Rass, MD PhD:
+      institute: [neuro]
+  - Andrea Schroll, MD PhD:
       institute: [int2]
-  - Bernhard Pfeifer:
-      institute: [umit]
-  - Verena Raas:
+  - Sarah Iglseder, MD:
       institute: [neuro]
-  - Alexander Egger:
+  - Alexander Egger, MD:
       institute: [zimcl]
-  - Gregor Hörmann:
-      institute: [zimcl]
-  - Raimund Helbok:
+  - Ewald Wöll, MD:
+      institute: [zams]
+  - Günter Weiss, MD:
+      institute: [int2]
+  - Raimund Helbok, MD:
       institute: [neuro]
-  - Ewald Wöll:
-      institute: [zams]
-  - Günter Weiss:
+  - Gerlig Widmann, MD:
+      institute: [radio]
+  - Thomas Sonnweber, MD:
       institute: [int2]
-  - Gerlig Widmann:
-      institute: [radio]
-  - Thomas Sonnweber:
+  - Ivan Tancevski, MD:
       institute: [int2]
-  - Ivan Tancevski:
-      institute: [int2]
-  - Barbara Sperner-Unterweger:
+  - Barbara Sperner-Unterweger, MD:
       institute: [psych]
-  - Judith Löffler-Ragg:
+  - Judith Löffler-Ragg, MD:
       institute: [int2]
       email: judith.loeffler@i-med.ac.at
+      phone: +43 (0)512/504-23251
+      fax: +43 (0)512/504-23317
       correspondence: true
@@
   - daas: Data Analytics As a Service Tirol, Innsbruck, Austria
-  - int2: Department of Internal Medicine II, Medical University of Innsbruck, Innsbruck Austria
+  - int2: Department of Internal Medicine II, Medical University of Innsbruck, Innsbruck, Austria
   - radio: Department of Radiology, Medical University of Innsbruck, Innsbruck, Austria
@@
 
-csl: frontiers-in-immunology.csl
+csl: jpr.csl
 
@@
 
-__Background:__
-Sequelae of Coronavirus disease 2019 (COVID-19) were investigated by both patient-initiated and academic initiatives. 
-Patient´s subjective illness perceptions might differ from physician’s clinical assessment results. 
-Herein, we explored factors influencing patient’s perception during COVID-19 recovery.
+__Objective:__
+Subjective illness perception (IP) can differ from physician’s clinical assessment results. 
+Herein, we explored patient’s IP during <span custom-style = 'revision'>coronavirus disease 2019 (COVID-19)</span> recovery.
 
 __Methods:__
-Participants of the prospective observation CovILD study with persistent somatic symptoms or cardiopulmonary findings at the clinical follow-up one year after COVID-19 were analyzed (n = `r nrow(cov_data$clear_data)`).
-Explanatory variables included baseline demographic and comorbidity data, COVID-19 course and one-year follow-up data of persistent somatic symptoms, physical performance, lung function testing (LFT), chest computed tomography (CT) and trans-thoracic echocardiography (TTE).
-Factors affecting illness perception (Brief Illness Perception Questionnaire, BIPQ) were identified by penalized multi-parameter regression and unsupervised clustering.
+Participants of the prospective observation CovILD study <span custom-style = 'revision'>(ClinicalTrials.gov: NCT04416100)</span> with persistent somatic symptoms or cardiopulmonary findings one year after COVID-19 were analyzed (n = `r nrow(cov_data$clear_data)`).
+Explanatory variables included demographic and comorbidity, COVID-19 course and one-year follow-up data of persistent somatic symptoms, physical performance, lung function testing, chest computed tomography and trans-thoracic echocardiography.
+Factors affecting IP (Brief Illness Perception Questionnaire) <span custom-style = 'revision'>one year after COVID-19</span> were identified by regularized modeling and unsupervised clustering.
 
 __Results:__
-In modeling, `r (blass_mod$cv_fit_stats$ipq_total$estimate[4] %>% signif(2) %>% unname) * 100`% of overall illness perception variance at one year after COVID-19 was attributed to fatigue intensity, reduced physical performance, hair loss and  baseline respiratory comorbidity.
-Overall illness perception was independent of LFT results, pulmonary lesions in CT or heart abnormality in TTE. 
-As identified by clustering, persistent somatic symptom count, fatigue, diminished physical performance, dyspnea, hair loss and sleep problems at the one-year follow-up and severe acute COVID-19 were associated with the BIPQ domains of concern, emotional representation, complaints, disease timeline and consequences.
+In modeling, 
+<span custom-style = 'revision'>`r 100 * (lasso_mod$cv_fit_stats$ipq_total$estimate[4] %>% signif(2) %>% unname)`%</span> 
+of overall IP variance <span custom-style = 'revision'>(R^2^)</span> was attributed to fatigue intensity, reduced physical performance <span custom-style = 'revision'>and persistent somatic symptom count.</span>
+Overall IP was <span custom-style = 'revision'>largely</span> independent of <span custom-style = 'revision'>lung and heart findings revealed by imaging and function testing.</span> 
+<span custom-style = 'revision'>In<span> clustering, 
+persistent somatic symptom count <span custom-style = 'revision'>(Kruskal-Wallis test: `r clust_chara$test_results$fup %>% filter(variable == 'sympt_number') %>% .$plot_cap`)</span>, 
+fatigue <span custom-style>(`r clust_chara$test_results$fup %>% filter(variable == 'Chalder_FS') %>% .$plot_cap`)</span>, 
+diminished physical performance <span custom-style = 'revision'>($\chi^2$ test, Cramer V effect size statistic: `r clust_chara$test_results$fup %>% filter(variable == 'fatigue_sympt') %>% .$plot_cap`)<span>,
+dyspnea <span custom-style = 'revision'>(`r clust_chara$test_results$fup %>% filter(variable == 'dyspnoe_sympt') %>% .$plot_cap`)</span>, 
+hair loss <span custom-style = 'revision'>(`r clust_chara$test_results$fup %>% filter(variable == 'hair_loss_sympt') %>% .$plot_cap`)</span>
+and sleep problems <span custom-style = 'revision'>(`r clust_chara$test_results$fup %>% filter(variable == 'sleep_sympt') %>% .$plot_cap`)</span> 
+were <span custom-style = 'revision'>strongly</span> associated with the concern, emotional representation, complaints, disease timeline and consequences <span custom-style = 'revision'>IP dimensions.</span>
 
 __Conclusion:__
-Persistent somatic symptoms rather than clinical assessment results, revealing lung and heart abnormalities, impact on severity and quality of illness perception at one year after COVID-19 and may foster unhelpful coping mechanisms.
-Besides  COVID-19 severity, individual illness perception should be taken into account when allocating rehabilitation and psychological therapy resources.
+Persistent somatic symptoms rather than <span custom-style = 'revision'>abnormalities in cardiopulmonary testing influence IP</span> one year after COVID-19. 
+<span custom-style = 'revision'>Modifying IP represents a promising innovative approach to treatment of post-COVID-19 condition.
+Besides COVID-19 severity, individual IP should guide rehabilitation and psychological therapy decisions.</span>
 
-__Study registration:__ 
-ClinicalTrials.gov: NCT04416100.
+\newpage
+
+# Keywords
+
+COVID-19, clustering, illness perception, persistent somatic symptoms, post-COVID-19 condition, regularized regression
 
@@
 
-A sizable fraction of coronavirus disease 2019 (COVID-19) patients is affected by protracted somatic symptoms, cardiopulmonary pathology and mental health disorders [@Sonnweber2020; @Sahanic2021; @Luger2022; @Sonnweber2022; @Hufner2022; @Al-Aly2021; @Davis2021; @Soriano2022; @Evans2021a].
-Persistent COVID-19-related symptoms have been initially described by patient and social media initiatives [@McCorkell2021] and subsequently, recognized as the 'post COVID-19 condition' by the clinical community [@Sonnweber2020; @Sahanic2021; @Davis2021; @Soriano2022; @Al-Aly2021].
-Yet, the patients' and clinicians' characteristics were not always consistent, which was observed also in other conditions, such as chronic obstructive pulmonary disease [@Miravitlles2013; @Schwartz2020] or functional disorders [@Schumacher2013].
-For post COVID-19 condition, the matter is further complicated by a broad range of patient-reported manifestations and widespread character of the disease [@Davis2021; @Sahanic2021].
-Since it is becoming increasingly evident that many COVID-19 convalescents experience prolonged severe individual suffering [@Hufner2022; @Huang2021; @Davis2021; @Evans2021a], and pose a significant healthcare and socioeconomic challenge [@Al-Aly2021], it is critically important to characterize and understand individual illness perception following COVID-19.
+A subset of coronavirus disease 2019 (COVID-19) patients is affected by protracted somatic symptoms, cardiopulmonary pathology and mental health disorders [@Sonnweber2020; @Sahanic2021; @Luger2022; @Sonnweber2022; @Hufner2022; @Al-Aly2021; @Davis2021; @Soriano2022; @Evans2021a; @Sudre2021; @Chen2022].
+Persistent COVID-19-related symptoms have been initially described by patient initiatives [@McCorkell2021] and subsequently recognized as 'post-COVID-19 condition' by the clinical community [@Sonnweber2020; @Sahanic2021; @Davis2021; @Soriano2022; @Al-Aly2021; @Sudre2021].
+<span custom-style = 'revision'>Most patients recover completely from acute COVID-19 symptoms within four weeks. 
+The clearance rate is particularly fast for non-specific infection complaints [@Sudre2021; @Sahanic2021; @Rass2022]. 
+In a recent metaanalysis of 1.7 million patients, the pooled prevalence estimate of post-COVID-19 condition, i.e. symptoms present for 3 months or longer, was 43% [@Soriano2022; @Chen2022]. 
+Yet, in a rigorously controlled study, approximately 12.7% of SARS-CoV-2-positive individuals suffered from persistent somatic symptoms directly associated with COVID-19 [@Ballering2022].
+Fatigue, concentration and memory problems, sleep disorders, physical performance deficits, smell/taste disorders and hair loss, dyspnea and pain are the most common, frequently overlapping and relapsing manifestations [@Chen2022; @Sonnweber2022; @Sahanic2021; @Davis2021; @Rass2022; @Evans2021a; @Huang2021; @Ballering2022].</span>
 
-Illness perceptions can be divided into cognitive or emotional components [@Broadbent2006]. 
-The cognitive representations include self-perceived consequences, expected duration, personal control, expected effect of treatment, symptom perception and understanding of the disease. 
-The emotional components encompass patient's concerns and emotions e.g. fear, anger or distress associated with the disease [@Broadbent2006; @Weinman1996].
-According to the common-sense model of self regulation (CSM), a theoretical framework enabling understanding of how people cope with threats to their health, such illness perceptions are influenced by situational stimuli such as symptoms, health information and patient's knowledge [@Man2021; @Leventhal2016; @Diefenbach1996].
-The individual illness perception was shown to influence disease coping, adjustment to adverse life events or chronic conditions and compliance with prevention, treatment and rehabilitation [@Hagger2010; @Figueiras2019], also in context of acute COVID-19 [@DiasNeto2021; @Lugo-Gonzalez2020; @Vollmann2021].
-Finally, severe illness perception including the emotional response, concerns and consequence components was found associated with anxiety, depression and stress both in the general population and in COVID-19 patients [@Wierenga2021; @Sadeghian2022; @Hamama2022; @Burke2020].
+Yet, patients' and clinicians' perceptions 
+<span custom-style = 'revision'>of post-COVID-19 condition <span custom-style = 'revision'>and its impact on patients´ lives</span> are not always consistent, as observed also in chronic obstructive pulmonary disease [@Miravitlles2013; @Schwartz2020] or functional disorders [@Schumacher2013].
+More specifically, persistent symptoms in many COVID-19 patients were not accompanied by lung, heart, biochemical, inflammatory and immunity abnormalities [@Sonnweber2022; @Sneller2022]. 
+<span custom-style = 'revision'>Persistent somatic symptoms (PSS) are defined as any subjectively distressing somatic complaints, including medically explained, unexplained and functional symptoms [@Lehmann2021].</span>
+<span custom-style = 'revision'>Manifestations of post-COVID-19 condition</span> like fatigue or neuro-cognitive disorders were also associated with depression, anxiety, stress and other signs of mental health impairment [@Al-Aly2021; @Hufner2022; @Sneller2022; @Bierbauer2022; @Wang2022; @Kachaner2022].
+Intriguingly, somatic complaints comparable to post-COVID-19 condition symptoms were also observed in COVID-19-free individuals [@Wisk2022; @Behnood2022; @Matta2022] suggestive of effects of socioeconomic background or altered social behavior during the SARS-CoV-2 pandemic [@Man2021; @Tutzer2021].
+Fatigue, sleep problems and dyspnea were also widespread in the general pre-pandemic population [@Hinz2017].</span>
 
-Herein, we investigated severity of overall illness perception and the illness perception components in a cross-sectional COVID-19 convalescent collective [@Sonnweber2020; @Sahanic2021; @Luger2022; @Sonnweber2022] with incomplete somatic symptom resolution or COVID-19-related lung and heart pathology in the follow-up clinical assessment at one year after diagnosis.
+Illness perception (IP) consists of cognitive and emotional dimensions [@Broadbent2006]. 
+The cognitive dimensions comprise of self-perceived consequences, expected duration, personal control, expected treatment effect, symptom perception and understanding of the disease. 
+The emotional dimensions encompass disease-associated concerns and emotions e.g. fear, anger or distress [@Broadbent2006; @Weinman1996].
+According to the common-sense model of self regulation, aiming to explain coping with health threats, IP is influenced by situational stimuli such as symptoms, health information and patient's knowledge [@Man2021; @Leventhal2016; @Diefenbach1996].
+IP was shown to influence adjustment to adverse life events or chronic conditions and compliance with prevention, treatment and rehabilitation [@Hagger2010; @Figueiras2019], also in COVID-19 [@DiasNeto2021; @Lugo-Gonzalez2020; @Vollmann2021].
+Emotional response, concerns and consequence IP dimensions correlate with anxiety, depression and stress both in the general population and acute COVID-19 patients [@Wierenga2021; @Sadeghian2022; @Hamama2022; @Burke2020; @Chen2022].
+<span custom-style = 'revision'>Similar observations were made recently in individuals with post-COVID-19 condition recruited from self-help and social media platforms [@Bierbauer2022]. 
+However, characterization of IP and its influencing factors in a defined observational COVID-19 cohort is still missing. 
+In particular, it is unclear how PSS, lung and heart pathology [@Sonnweber2020; @Sonnweber2022; @Luger2022] shape IP during long-term COVID-19 recovery.</span>
+
+<span custom-style = 'revision'>We aimed to describe IP and IP components, identify demographic and clinical variables affecting IP and explore IP patterns in convalescents suffering from persistent somatic symptoms (PSS), lung or heart abnormalities one year after COVID-19.
+To this end, we gauged IP in a long-term observatory cross-sectional COVID-19 collective [@Sonnweber2020; @Sahanic2021; @Luger2022; @Sonnweber2022] with the Brief Illness Perception Questionnaire (BIPQ) [@Broadbent2006] and investigated components, influential factors and patterns of IP by factor analysis, regularized multi-parameter modeling and unsupervised PAM clustering.</span>
 
@@
 
-Participants of the longitudinal observation CovILD study (ClinicalTrials.gov: NCT04416100) were recruited between April and June 2020 at three Austrian clinical centers [@Sonnweber2020; @Luger2022; @Sonnweber2022].
-The study inclusion criteria were age $\geq$ 18 years and symptomatic PCR-confirmed SARS-CoV-2 infection. 
-The follow-up visits were scheduled at two, three, six months and one year after COVID-19 diagnosis. 
-A subset of participants (n = `r nrow(cov_data$clear_data)`) was analyzed who completed the Brief Illness Perception Questionnaire (BIPQ) [@Broadbent2006] and displayed (1) COVID-19-related persistent somatic symptoms or (2) any abnormality in chest computed tomography (CT) or (4) any lung function testing (LFT) deficits or (5) any heart abnormality in trans-thoracic echocardiography (TTE) at the one-year follow-up (__Figure \@ref(fig:fig-consort)__).
+Participants of the longitudinal observation CovILD study (ClinicalTrials.gov: NCT04416100) were recruited <span custom-style = 'revision'>during the first wild-type SARS-CoV-2 outbreak</span> between April and June 2020 at <span custom-style = 'revision'>the University Hospital of Innsbruck, the St.
+Vinzenz Hospital in Zams and the Karl-Landsteiner Rehabilitation Facility in Münster (all in Austria) [@Sonnweber2020; @Luger2022; @Sonnweber2022].</span>
+The study inclusion criteria were age $\geq$ 18 years and symptomatic, PCR-confirmed SARS-CoV-2 infection. 
+The follow-ups were scheduled at two, three, six months and one year after COVID-19 diagnosis.
 
-The study was performed in accordance with the Declaration of Helsinki and the European Data Policy. All participants gave written informed consent. 
+<span custom-style = 'revision'>In total, 190 individuals were screened for participation. 
+The screened subset size was limited by the number of COVID-19 cases diagnosed at the study centers, their follow-up availability (residency in the study region) and available resources for the follow-up assessments.
+No statistical sample size estimation was performed, as the kinetic of COVID-19 recovery was unknown at the pandemic onset.
+Out of 145 enrolled individuals, `r nrow(cov_data$raw_data)` completed the one-year follow-up.</span>
+A subset of participants (n = `r nrow(cov_data$clear_data)`) was analyzed who displayed (1) PSS or (2) any abnormality in chest computed tomography (CT) or (3) any lung function testing (LFT) deficits or (4) any heart abnormality in trans-thoracic echocardiography (TTE) at the one-year follow-up <span custom-style = 'revision'>and had a complete set of BIPQ, demographic and clinical analysis variables (__Figure \@ref(fig:fig-consort)__, __Supplementary Tables S1 - S2__).</span>
+<span custom-style = 'revision'>We decided to restrict the analysis to incomplete COVID-19 recovery, because BIPQ was initially designed for and validated in somatic conditions [@Broadbent2006].</span>
+
+The study was performed in accordance with the Declaration of Helsinki and European Data Policy. 
+All participants gave written informed consent. 
 The study protocol was approved by the ethics committee at the Medical University of Innsbruck (approval number 1103/2020).
@@
 
-For full descriptions of procedures and variables, see __Supplementary Methods__ and __Supplementary Table S1__.
+For full descriptions of procedures and variables, see __Supplementary Methods__ and __Supplementary Tables S1__ and __S2__.
 
-Baseline clinical and acute COVID-19 data were recorded retrospectively at the two-month follow-up based on the patient's interview and electronic patient records [@Sonnweber2020]. 
-Study participants were classified as ambulatory (outpatient, WHO grade 1 - 2), moderate (hospitalized, without oxygen therapy or mask/nasal prongs oxygen, WHO 3 - 4) and severe COVID-19 survivors (hospitalized with non-invasive ventilation or high-flow oxygen or mechanical ventilation, WHO 5 - 7).
+Baseline clinical and acute COVID-19 data were recorded retrospectively at the two-month follow-up [@Sonnweber2020]. 
+Study participants were classified as ambulatory (outpatient, WHO <span custom-style = 'revision'>ordinal scale for clinical improvement</span> 1 - 2), moderate (hospitalized without oxygen therapy, WHO 3 - 4) and severe COVID-19 survivors (hospitalized with non-invasive ventilation or high-flow oxygen or mechanical ventilation, WHO 5 - 7).
 
@@
 Dyspnea was scored with the Modified Medical British Research Council scale (mMRC).
-Fatigue at one-year follow-up was rated with likert and bimodal Chalder's Fatigue Scales (CFS) [@Chalder1993; @Morriss1998]. 
-Exertional capacity at the one-year follow-up was assessed by six-minute walking distance (SMWD) test and compared with the reference values [@Crapo2012].
+Fatigue was rated with likert and bimodal Chalder's Fatigue Scales (CFS) [@Chalder1993; @Morriss1998]. 
+Exertional capacity was assessed by six-minute walking distance (SMWD) [@Crapo2012].
 
-The following COVID-19-related persistent somatic symptoms at the one-year follow-up were analyzed: 
+The following, <span custom-style = 'revision'>PSS,</span> were analyzed at the one-year follow-up: 
 reduced physical performance (ECOG $\geq$ 1), 
@@
 significant fatigue (bimodal CFS $\geq$ 4). 
+<span custom-style = 'revision'>Records of reduced physical performance, dyspnea, cough, sleep problems, night sweating, hyposmia/anosmia and gastrointestinal symptoms were available for acute COVID-19 and all follow-ups [@Sonnweber2022].</span>
 
@@
 
-Illness perception was investigated with the 8-item BIPQ (Q1 - Q8) [@Broadbent2006]. 
-Each item was rated with an 11-point likert scale. 
-The illness perception score was defined as a sum of all BIPQ items, with the negative items Q3, Q4 and Q7 inverted.
+IP was investigated with <span custom-style = 'revision'>a German translation of BIPQ [@Broadbent2006] with the following 8 items:</span> 
+<span custom-style = 'revision'>Q1 consequences (0: no consequences, 10: very severe consequences), 
+Q2 timeline (0: very short, 10: forever), 
+Q3 lacking personal control (negative item, 0: full control, 10: no control at all), 
+Q4 lacking treatment control (negative item, 0: full control, 10: no control at all), 
+Q5 identity (0: no complaints, 10: multiple severe complaints), 
+Q6 concerns (0: not at all, 10: extreme concerns), 
+Q7 lacking coherence (negative item, 0: very good understanding, 10: no understanding at all) and 
+Q8 emotional representation (0: no emotions at all, 10: extreme emotions).</span> 
+Items were rated with 11-point likert scales, <span custom-style = 'revision'>with 0 corresponding to neutral or positive and 10 representing severely distorted or negative IP dimension. 
+The Q3, Q4 and Q7 conceptualized as negative items were re-coded prior to analysis.</span>
+<span custom-style = 'revision'>The total IP score was defined as all BIPQ item sum as a measure of overall negative IP alterations (0: neutral - 80: extremely severe) [@Broadbent2006].</span> 
+<span custom-style = 'revision'>The emotion/concern/consequences score was defined as the Q1, Q2, Q5, Q6 and Q8 item sum (0: neutral - 50: extremely severe) and the lacking control/coherence score was defined as the Q3, Q4 and Q7 item sum (range: 0: good control/understanding - 30: no control/understanding at all) to measure intensity of the two main IP components identified by factor analysis.</span>
 
@@
 
-The primary analysis endpoints were illness perception score values and its influencing factors among demographic, clinical and somatic symptom data at the one-year after COVID-19 in convalescents affected by persistent somatic symptoms or residual cardiopulmonary findings.
-The secondary analysis endpoints were particular BIPQ component scores and their influencing factors.
+<span custom-style = 'revision'>The primary analysis endpoints were demographic and clinical factors affecting total IP and its emotion/concern/consequences and lacking control/coherence components in COVID-19 convalescents with PSS or cardiopulmonary abnormalities at the one-year follow-up.</span>
+<span custom-style = 'revision'>The secondary analysis endpoints were components of IP, patterns of IP dimensions explored by clustering as well as demographic and clinical factors differentiating between the IP clusters.</span>
 
@@
 
-Details of statistical analysis are provided in __Supplementary Methods__. 
-Data analysis was done with R version 4.2.0 (R Foundation for Statistical Computing). 
+Statistical analysis details are provided in __Supplementary Methods__. 
+Analysis was done with R version 4.2.0. 
+<span custom-style = 'revision'>Numeric variables were presented as medians with interquartile ranges. 
+Categorical variables were shown as percentages and counts.</span> 
 Differences in categorical variable distribution were assessed by $\chi^2$ test with Cramer V effect size statistic. 
-Statistical significance for numerical variables was investigated by Mann-Whitney test with Wilcoxon r effect size statistic or Kruskal-Wallis test with $\eta^2$ effect size statistic [@Kassambara2021].
+Statistical significance for numerical variables was investigated by Mann-Whitney test with r effect size statistic or Kruskal-Wallis test with $\eta^2$ effect size statistic. 
 Correlations were investigated by Spearman's test.
+<span custom-style = 'revision'>Differences in symptom frequency over time were investigated by Cochran Q test. 
+Interaction effects were analyzed by two-way ANOVA with $\eta^2$ effect size statistic.</span>
+<span custom-style = 'revision'>BIPQ dimensionality and consistency was investigated by factor analysis [@Bartlett1937] and McDonald's $\omega$ [@McDonald1999].</span>
 
-Multi-parameter modeling was done with the Elastic Net (package _glmnet_) [@Zou2005; @Friedman2010], LASSO (least absolute shrinkage and selection operator; package _glmnet_) [@Tibshirani1996; @Friedman2010] and Bayesian LASSO (package _monomvn_) [@Park2012; @Gramacy2022] algorithms.
-The response, illness perception score, was square-root transformed to guarantee normality. 
+Multi-parameter modeling was done with <span custom-style = 'revision'>regularized regression algorithms:</span> Elastic Net [@Zou2005; @Friedman2010], LASSO (least absolute shrinkage and selection operator) [@Tibshirani1996; @Friedman2010] and Bayesian LASSO [@Park2012; @Gramacy2022].
+<span custom-style = 'revision'>The choice of regularized linear regression was motivated by multi-dimensional character of the study data set, i.e.
+comparable number of explanatory variables and observations [@Kirpich2018; @Morozova2015].
+Ordinary least square models with backward term elimination developed in the study data set had poor reproducibility and over-parameterization as demonstrated by fit error expansion in cross-validation.</span>
+
+<span custom-style = 'revision'>Modeling responses: the total IP score, emotion/concern/consequence and lacking control/coherence scores, were square root-transformed to guarantee normality.</span> 
 Both first and second order terms of numeric explanatory variables were included in the models. 
 Numeric explanatory and response variables were Z-score normalized. 
-The optimal $\lambda$ for Elastic Net and LASSO were obtained by 200-repeats 10-fold cross-validation (CV). 
-The 'sparsity' parameter in Bayesian LASSO was found by 10-repeats 10-fold cross-validation (package _caret_) [@Kuhn2008]. 
-Explained variance ($R^2$) and model root mean squared error (RMSE) were assessed in the entire data set and 10-repeats 10-fold CV (package _caret_) [@Kuhn2008]. 
-Elastic Net and LASSO coefficients were calculated for the optimal $\lambda$ values. 
+The optimal $\lambda$ for Elastic Net and LASSO were obtained by 200-repeats 10-fold cross-validation. 
+The 'sparsity' parameter in Bayesian LASSO was found by 10-repeats 10-fold cross-validation [@Kuhn2008]. 
+Explained variance ($R^2$) and model root mean squared error were assessed in the entire data set and 10-repeats 10-fold cross-validation [@Kuhn2008]. 
+Elastic Net and LASSO model coefficients were calculated for the optimal $\lambda$ values. 
 Bayesian LASSO coefficients were calculated as medians over all algorithm iterations [@Park2012]. 
-Variables with non-zero coefficients in all three models were deemed key factors for the illness perception score.
+<span custom-style = 'revision'>The number of non-zero model coefficients was found to differ substantially between the models as reported for data sets of similar size [@Morozova2015].</span>
+<span custom-style = 'revision'>Hence,</span> variables with non-zero coefficients in <span custom-style = 'revision'>models constructed with all three algorithms</span> were deemed the key factors for
+<span custom-style = 'revision'>the total IP score, emotion/concern/consequences and lacking control/coherence scores.</span>
 
-Clustering by the BIPQ items was accomplished with the PAM algorithm (partitioning around medoids; Euclidean distance) [@Schubert2019; @Drost2018].
+Clustering by the BIPQ items was accomplished by PAM (partitioning around medoids) <span custom-style = 'revision'>with Euclidean distance [@Schubert2019; @Drost2018].</span> 
+<span custom-style = 'revision'>The clustering algorithm choice was motivated by its optimal stability in 10-fold cross-validation [@Lange2004] and good explanatory performance measured by the fraction of clustering variance as compared with hierarchical and KMEANS clustering.</span>
 
@@
 
-The CovILD cohort was recruited between April and June 2020. 
-Out of 145 participants enrolled, `r nrow(cov_data$clear_data)` individuals with COVID-19-related persistent somatic symptoms, lung CT, LFT or cardiological abnormalities and complete BIPQ were included in the analysis.
-The major reasons of patient dropout were missing follow-up visits and incomplete BIPQ (__Figure \@ref(fig:fig-consort)__).
-The participants were predominantly male (`r get_percent(cov_data$clear_data, "sex")['male']`%), the median age at COVID-19 diagnosis was `r get_median_iqr(cov_data$clear_data, "age")['median']` years (IQR: `r get_median_iqr(cov_data$clear_data, "age")['iqr']`), over one-third were active or ex-smokers (`r get_percent(cov_data$clear_data, "smoking_history")['yes']`%).
-Most participants suffered from comorbidities (`r get_percent(cov_data$clear_data, "comorb_present")['yes']`%), with cardiovascular disease, metabolic and respiratory disorders as the leading conditions.
-The participants were classified by COVID-19 severity as ambulatory (`r get_percent(cov_data$clear_data, "cat_WHO")[1]`%), moderate (hospitalized, no intensive care , no oxygen therapy, `r get_percent(cov_data$clear_data, "cat_WHO")[2]`%) and severe COVID-19 survivors (hospitalized, oxygen therapy or intensive care, `r get_percent(cov_data$clear_data, "cat_WHO")[3]`%).
-The ambulatory COVID-19 subset had the lowest median age, rate of smokers and comorbidity frequency (__Table \@ref(tab:tab-cohort-baseline)__).
+The CovILD cohort was recruited <span custom-style = 'revision'>during the first outbreak of SARS-CoV-2</span> between April and June 2020 [@Sonnweber2020]. 
+Out of 145 participants enrolled, `r nrow(cov_data$clear_data)` individuals with PSS, lung CT, LFT or cardiological abnormalities, <span custom-style = 'revision'>complete BIPQ and a complete set of study variables set at the one-year follow-up</span> were analyzed.
+The major reasons of patient dropout were missing follow-ups and incomplete BIPQ (__Figure \@ref(fig:fig-consort)__).
+<span custom-style = 'revision'>The excluded participants tended towards less severe acute COVID-19, more dermatological and gastrointestinal symptoms at the one-year follow-up, and lower blood hemoglobin and ferritin as compared with the analysis collective; those effects were, however, not significant.
+The excluded participants had also significantly lower SMWD (__Supplementary Figure S3__).</span>
 
-Nearly three-quarter of participants (`r get_percent(cov_data$clear_data, "sympt_present")['yes']`%) suffered from persistent somatic symptoms at one year after COVID-19, with significant fatigue (`r get_percent(cov_data$clear_data, "Chalder_FS_bimodal")['yes']`%), reduced physical performance (`r get_percent(cov_data$clear_data, "fatigue_sympt")['yes']`%), sleep disorders (`r get_percent(cov_data$clear_data, "sleep_sympt")['yes']`%) and exertional dyspnea (`r get_percent(cov_data$clear_data, "dyspnoe_sympt")['yes']`%) being most frequent.
-The symptom frequencies and fatigue rating were comparable between the COVID-19 severity strata (__Supplementary Table S2__).
-LFT abnormalities were discerned in `r get_percent(cov_data$clear_data, "lufo_red")['yes']`% of participants and tended to be most common in severe COVID-19 survivors. 
-Residual lung CT lesions were found in `r get_percent(cov_data$clear_data, "ct_severity_any")['yes']`% individuals and their frequency and scoring was significantly higher in moderate and severe COVID-19 than in the ambulatory disease subset.
-The most common cardiological finding was low grade diastolic dysfunction (`r get_percent(cov_data$clear_data, "diastolic_dysf")['yes']`% of cohort), which was significantly more frequent in moderate and severe than in ambulatory COVID-19 survivors.
-Nearly 80% of severe disease survivors attended COVID-19-specific rehabilitation, the rehabilitation rates in the remaining severity strata were below 20% (__Supplementary Table S2__). 
-Most laboratory parameters at one year after COVID-19 were within their normal values.
-Mild anemia and improper glycemia control (HbA1c) were evident solely in moderate and severe COVID-19 survivors (__Supplementary Table S3__).
+The <span custom-style = 'revision'>analyzed</span> participants were predominantly male (`r get_percent(cov_data$clear_data, "sex")['male']`%), the median age at COVID-19 diagnosis was `r get_median_iqr(cov_data$clear_data, "age")['median']` years (IQR: `r get_median_iqr(cov_data$clear_data, "age")['iqr']`), over one-third were active or ex-smokers (`r get_percent(cov_data$clear_data, "smoking_history")['yes']`%).
+Most participants suffered from comorbidities (`r get_percent(cov_data$clear_data, "comorb_present")['yes']`%), with cardiovascular disease, metabolic and respiratory disorders as leading conditions. 
+The participants were classified by COVID-19 severity as ambulatory (`r get_percent(cov_data$clear_data, "cat_WHO")[1]`%), moderate (hospitalized, no intensive care, no oxygen therapy, `r get_percent(cov_data$clear_data, "cat_WHO")[2]`%) and severe COVID-19 survivors (hospitalized, oxygen therapy or intensive care, `r get_percent(cov_data$clear_data, "cat_WHO")[3]`%).
+The ambulatory COVID-19 subset had the lowest median age, smoking and comorbidity rates (__Table \@ref(tab:tab-cohort-baseline)__).
 
-## Illness perception at one year after COVID-19
+Nearly three-quarter of participants (`r get_percent(cov_data$clear_data, "sympt_present")['yes']`%) suffered from PSS one year after COVID-19. 
+Significant fatigue (`r get_percent(cov_data$clear_data, "Chalder_FS_bimodal")['yes']`%), reduced physical performance (`r get_percent(cov_data$clear_data, "fatigue_sympt")['yes']`%), sleep disorders (`r get_percent(cov_data$clear_data, "sleep_sympt")['yes']`%) and exertional dyspnea (`r get_percent(cov_data$clear_data, "dyspnoe_sympt")['yes']`%) were the most frequent PSS.
+<span custom-style = 'revision'>Recovery pace for the longitudinally recorded PSS was the fastest within the first 3 months after COVID-19. 
+There was little improvement during later convalescence and percentages of performance deficits, sleep problems and smell disorders were substantially higher at the one-year than at the 6 month follow-up suggestive of relapse.
+Except for self-reported anosmia, which was absent in severe COVID-19 survivors, </span>PSS frequencies and fatigue rating were comparable between the COVID-19 severity strata (__Supplementary Figure S1__, __Supplementary Table S4__).
 
-As measured by Cronbach's alpha [@Cronbach1951], the BIPQ tool had an acceptable internal consistency in the study cohort ($\alpha$ = `r ipq_co$alpha_stats %>% filter(response == "ipq_total") %>% .$raw_alpha %>% signif(2)`, 95% CI: `r paste(ipq_co$alpha_stats %>% filter(response == "ipq_total") %>% .$plot_cap %>% stri_extract(regex = "\\[.*\\]$") %>% stri_replace_all(regex = "\\[|\\]", replacement = ""))`) (__Supplementary Figure S1__).
-The consequences, timeline, identity, concern and emotional representation items were strongly positively inter-correlated. 
-Significant, moderate-to-strong positive association was observed between the personal control and treatment control components as well as between the coherence and treatment control items.
+LFT abnormalities affected `r get_percent(cov_data$clear_data, "lufo_red")['yes']`% of participants and were substantially enriched in severe COVID-19 survivors. 
+Lung CT abnormalities were found in `r get_percent(cov_data$clear_data, "ct_severity_any")['yes']`% individuals and their frequency and scoring was significantly higher in moderate-to-severe COVID-19 than in the ambulatory subset.
+The leading cardiological finding was low grade diastolic dysfunction (`r get_percent(cov_data$clear_data, "diastolic_dysf")['yes']`%), which was significantly enriched in moderate and severe COVID-19 survivors.
+<span custom-style = 'revision'>Concomitant PSS, abnormal lung CT, LFT and TTE one year after COVID-19 were observed in `r signif(fup_overlap$all_intersection/nrow(cov_data$clear_data), 3) * 100`% of participants.
+In `r signif(14/nrow(cov_data$clear_data), 3) * 100`% of participants, PSS were not accompanied by cardiopulmonary abnormalities (__Supplementary Figure S2__).</span>
+Nearly 80% of severe disease survivors attended COVID-19-specific rehabilitation, the rehabilitation rates in the remaining severity strata were below 20% (__Supplementary Table S4__). 
+Most laboratory parameters at the one-year follow-up were within their normal values. 
+Mild anemia and improper glycemia control (HbA1c) were evident primarily in moderate and severe COVID-19 survivors (__Supplementary Table S5__).
 
-The median illness perception score defined as the sum of all items [@Broadbent2006] was `r get_median_iqr(cov_data$clear_data, "ipq_total")['median']` (IQR: `r get_median_iqr(cov_data$clear_data, "ipq_total")['iqr']`) in the study collective, the differences between ambulatory, moderate and severe COVID-19 were not significant.
-The treatment control and coherence BIPQ items were rated the highest followed by the personal control and emotional representation components.
-Significant differences between the COVID-19 severity strata were detected for the consequences, concern, emotional representation and coherence BIPQ items, which peaked in severe COVID-19 convalescents (__Table \@ref(tab:tab-psych)__).
+<span custom-style = 'revision'>Females displayed significantly less severe COVID-19 and less chest CT abnormalities but higher PSS rates and counts than males at the one-year follow-up. 
+Differences between the genders in frequencies and scoring of specific PSS were not significant (__Supplementary Table S6__ and not shown).</span>
 
-## Key factors for overall illness perception
+## Illness perception one year after COVID-19
 
-To identify the most important factors among `r length(globals$variables)` candidate explanatory variables (__Supplementary Table S1__) influencing the illness perception score at one year after COVID-19, three penalized multi-parameter regression algorithms: Elastic Net [@Zou2005], LASSO [@Tibshirani1996] and Bayesian LASSO [@Park2012] were employed.
-The final models explained at least `r (blass_mod$cv_fit_stats$ipq_total$estimate[4] %>% signif(2) %>% unname) * 100`% and `r (blass_mod$cv_fit_stats$ipq_total$estimate[12] %>% signif(2) %>% unname) * 100`% of the illness perception score variance in the entire data set and 10-fold cross-validation, respectively (__Supplementary Figure S2__).
+<span custom-style = 'revision'>Factor analysis of BIPQ [@Broadbent2006; @Bartlett1937] revealed two IP components one year after COVID-19: (1) an emotional representation, concern, identity, consequences and timeline component (Q1, Q2, Q5, Q6 and Q8), and (2) a lacking treatment/personal control and lacking coherence component (Q3, Q4 and Q7) (__Supplementary Table S3__).
+The entire BIPQ ($\omega$ = `r ipq_co$omega_obj$omega.group$total[1] %>% signif(2)`) and the emotion/concern/consequence component ($\omega$ = `r ipq_co$omega_obj$omega.group$total[2] %>% signif(2)`) had good internal consistencies as determined by McDonald's omega [@McDonald1999].
+The lacking control/coherence component consistency was markedly lower ($\omega$ = `r ipq_co$omega_obj$omega.group$total[3] %>% signif(2)`), presumably due to weak correlation of the lacking personal control (Q3) with the remaining items (__Supplementary Figure S4__).</span>
 
-The number of variables with non-zero coefficients varied between `r length(an$variables$ipq_total$eln_mod)` for the Elastic Net and `r length(an$variables$ipq_total$blass_mod)` for the Bayesian LASSO regression (__Figure \@ref(fig:fig-key-factors)A__).
-The strongest positive correlates of the illness perception score in Elastic Net and LASSO modeling were pre-existing immune deficiency and respiratory disease as well as hair loss, fatigue rating and reduced physical performance at the one-year follow up.
-Fatigue scoring and respiratory comorbidity were the major positive correlates in the Bayesian LASSO model (__Supplementary Figures S2 - S5__).
-Of note, neither age, sex, acute COVID-19 severity nor LFT, CT and TTE abnormalities were selected as non-zero model coefficients by any of the regression algorithms. 
-Their effects on the illness perception scoring was not significant in a direct analysis either (__Supplementary Figure S6__).
+The median <span custom-style = 'revision'>total</span> IP score representing the sum of all BIPQ items was `r get_median_iqr(cov_data$clear_data, "ipq_total")['median']` (IQR: `r get_median_iqr(cov_data$clear_data, "ipq_total")['iqr']`) in the study collective, the differences between ambulatory, moderate and severe COVID-19 were not significant.
+<span custom-style = 'revision'>The emotion/concern/consequences IP component rated by the Q1, Q2, Q5, Q6 and Q8 item sum was the lowest in moderate and the highest in severe COVID-19. 
+Moderate COVID-19 convalescents had in turn the highest intensity of the lacking control/coherence IP component scored as the Q3, Q4 and Q7 item sum as compared with the remaining severity subsets; this difference was not significant (__Table \@ref(tab:tab-psych)__).</span>
 
-The key variables associated with the illness perception rating selected by all three models were (1) reduced physical performance, (2) hair loss and (3) fatigue rating at the one-year follow-up as well as (4) baseline respiratory comorbidity (__Figure \@ref(fig:fig-key-factors)A__).
-Each of those parameters was found significantly associated with higher illness perception scores in a direct comparison or correlation analysis (__Figure \@ref(fig:fig-key-factors)BC__).
+Significant differences between the COVID-19 severity strata were detected for the BIPQ items of consequences, concern, emotional representation, which peaked in severe COVID-19 convalescents. 
+<span custom-style = 'revision'>The maximum scores of the lacking coherence BIPQ item were detected in mild COVID-19 individuals (__Table \@ref(tab:tab-psych)__). 
+Differences in IP between the genders were not significant (__Supplementary Table S6__).</span>
 
-## Heterogeneity of illness perception
+## Key factors influencing illness perception
 
-Three subsets of participants which differed qualitatively in BIPQ components, termed further 'illness perception clusters', were identified by PAM clustering (__Supplementary Figure S7__). 
-Roughly half of study participants (cluster #1, `r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[1]`% of participants) had low scoring of the emotional representation, concern, identity, timeline and consequences components along with high rating of the coherence, personal control and treatment control items.
-This translated to a low median illness perception score (`r get_median_iqr(ipq_clust$analysis_tbl %>% filter(clust_id == "#1"), "ipq_total")['median']`, IQR: `r get_median_iqr(ipq_clust$analysis_tbl %>% filter(clust_id == "#1"), "ipq_total")['iqr']`).
-Another `r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[2]`% of participants assigned to the cluster #2 were characterized by low levels of self-perceived personal or treatment control of their COVID-19 sequelae but otherwise by low rating of the concern, identity, timeline, consequences and emotional representation items.
-The cluster #3 individuals (`r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[3]`% of participants) had highly elevated rating of the concern, identity, timeline, consequences and emotional representation components as compared with the clusters #1 and #2.
-As a result, the cluster #3 displayed the highest median illness perception score (`r get_median_iqr(ipq_clust$analysis_tbl %>% filter(clust_id == "#3"), "ipq_total")['median']`, IQR: `r get_median_iqr(ipq_clust$analysis_tbl %>% filter(clust_id == "#3"), "ipq_total")['iqr']`) (__Figure \@ref(fig:fig-risk-clusters)AB__).
+The most important factors among `r length(mod$variables)` candidate explanatory variables (__Supplementary Tables S2__) influencing <span custom-style = 'revision'>total IP and the emotion/concern/consequences and lacking control/coherence IP components one year after COVID-19, </span>were identified with three <span custom-style = 'revision'>regularized</span> multi-parameter regression algorithms: Elastic Net [@Zou2005], LASSO [@Tibshirani1996] and Bayesian LASSO [@Park2012].
+<span custom-style = 'revision'>The choice of regularized regression instead of canonical linear regression was motivated by multi-dimensional character of the study data, i.e.
+comparable number of observations and explanatory variables [@Kirpich2018; @Morozova2015].</span>
 
-Among baseline demographic and clinical parameters and follow-up readouts of somatic complaints and cardiopulmonary abnormalities (__Supplementary Table S1__), COVID-19-related persistent somatic symptoms demonstrated the largest significant differences between the illness perception clusters.
-In particular, the cluster #3 individuals with high level of disease-related concerns suffered from multiple persistent somatic symptoms, persistent fatigue, physical performance loss, sleep problems, dyspnea and hair loss at one year after COVID-19.
-The cluster #3 comprised predominantly of individuals with higher COVID-19 severity (WHO grade), lung CT abnormalities and the peak rates of COVID-19-specific rehabilitation.
-The persistent somatic symptom frequency and intensity was comparable in the clusters #1 and #2. 
-Yet, in  the cluster #2 characterized by low self-perceived disease coherence and control, frequencies of smokers, metabolic comorbidity and LFT abnormalities at one year after COVID-19 tended to be higher than in the cluster #1.
-The rehabilitation rate in the cluster #2 was significantly lower than in the cluster #1 (__Figure \@ref(fig:fig-risk-clusters)__ and __Supplementary Table S4__).
+<span custom-style = 'revision'>The final regularized models explained at least `r (lasso_mod$cv_fit_stats$ipq_total$estimate[4] %>% signif(2) %>% unname) * 100`% and `r (lasso_mod$cv_fit_stats$ipq_total$estimate[12] %>% signif(2) %>% unname) * 100`% of the total IP score variance in the entire data set and cross-validation, respectively.
+Explained variances of the emotion/concern/consequences IP component score were substantially higher with at least `r (eln_mod$cv_fit_stats$ipq_sub1$estimate[4] %>% signif(2) %>% unname) * 100`% in the entire collective and `r (eln_mod$cv_fit_stats$ipq_sub1$estimate[12] %>% signif(2) %>% unname) * 100`% in cross-validation.
+For the lacking control/coherence component, no meaningful models could be established with either of the algorithms (__Supplementary Figure S5__).
+The number of explanatory factors with non-zero coefficients in the total IP models varied between `r length(an$variables$ipq_total$lasso_mod)` (LASSO) and `r length(an$variables$ipq_total$eln_mod)` (Elastic Net) (__Figure \@ref(fig:fig-key-factors)A__, __Supplementary Figures S6 - S8__).
+Between `r length(an$variables$ipq_sub1$blass_mod)` (Bayesian LASSO) and `r length(an$variables$ipq_sub1$lasso_mod)` (LASSO) explanatory variables were linked to the emotion/concern/consequences rating (__Figure \@ref(fig:fig-key-factors)B__, __Supplementary Figures S9 - S11__).</span>
+<span custom-style = 'revision'>The key variables affecting total IP identified by all three models were PSS count, reduced physical performance and fatigue scoring at the one-year follow-up. 
+In uni-variable analysis, each of them was significantly associated with total IP (__Figure \@ref(fig:fig-total-bipq)__). 
+The key explanatory factors for the emotion/concern/consequences IP component identified by all three modeling algorithms were 
+PSS count, reduced physical performance, fatigue scoring, significant fatigue, hair loss, cough, diastolic dysfunction, elevated C-reactive protein (CRP) at the one-year follow-up as well as rehabilitation during COVID-19 recovery, respiratory comorbidity and body weight class.
+In uni-variable analysis, PSS number, fatigue scoring and significant fatigue, reduced performance, hair loss, rehabilitation and respiratory comorbidity were significantly associated with increased emotion/concern/consequences scores (__Figure \@ref(fig:fig-sub1-bipq)__).</span>
+
+<span custom-style = 'revision'>Age, gender, CT, LFT or TTE abnormalities alone had no significant effects on total IP or its components in uni-variable analysis (__Supplementary Figures S12 - S13__).
+However, lung CT abnormalities at the one-year follow-up amplified the effects of diastolic dysfunction, fatigue and fatigue rating on total IP and the emotion/concern/consequences component, as revealed by significant interactions in two-way ANOVA.
+In particular, individuals with concomitant diastolic dysfunction plus CT abnormalities or with high fatigue scoring plus CT abnormalities displayed significantly higher emotion/concern/consequences scores than participants without lung CT findings.
+Similar interaction effects of LFT deficits, fatigue and symptom numbers and of diastolic dysfunction and fatigue were discerned as well (__Supplementary Figures S14__ - __S16__).
+Participants with respiratory comorbidities had significantly elevated fatigue scores, LFT abnormality rates, total IP and emotion/concern/consequences ratings as compared with the remaining collective.
+This may suggest additional interaction effects of respiratory conditions, LFT deficits and persistent fatigue on IP (__Supplementary Table S7__).</span>
+
+## Illness perception clusters
+
+<span custom-style = 'revision'>IP patterns were investigated by unsupervised clustering of the study participants in respect to the BIPQ items. 
+Among several algorithms tested, PAM with Euclidean distance [@Drost2018; @Schubert2019] demonstrated the best reproducibility [@Lange2004] and good explanatory performance and was hence used for definition of three IP clusters (__Supplementary Figure S17__).</span>
+Roughly half of participants <span custom-style = 'revision'>assigned to the cluster #1 displayed low scores for each BIPQ item.</span> 
+This translated to low total IP severity, <span custom-style = 'revision'>, and low scoring of the emotion/concern/consequences and lacking control/coherence IP components.</span>
+Another `r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[2]`% of participants assigned to the cluster #2 were characterized by low self-perceived personal or treatment control but also by low rating of the concern, identity, timeline, consequences and emotional representation items.
+<span custom-theme = 'revision'>As a result, the maximum scores of the lacking control/consequence components and low emotion/concern/consequences IP component scores were observed in the cluster #2.</span>
+Cluster #3 individuals (`r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[3]`% of participants) had highly elevated rating of the concern, identity, timeline, consequences and emotional representation items as compared with the clusters #1 and #2.
+As a result, the cluster #3 displayed the highest total IP severity, <span custom-style = 'revision'>and the highest emotion/concern/consequence ratings. 
+In turn, the scoring  of the lacking control/coherence IP component was low, suggestive of good understanding of the disease and trust in therapy or personal control of COVID-19 sequelae (__Figure \@ref(fig:fig-ipq-clusters)__).
+Effect sizes of differences in total IP and its components between the IP clusters were large (Kruskal-Wallis test, total IP: `r clust_chara$test_results$ipq %>% filter(variable == 'ipq_total') %>% .$plot_cap`, emotion/concern/consequences: `r clust_chara$test_results$ipq %>% filter(variable == 'ipq_total') %>% .$plot_cap`, lacking control/coherence: `r clust_chara$test_results$ipq %>% filter(variable == 'ipq_sub2') %>% .$plot_cap`).</span>
+
+<span custom-style = 'revision'>The cluster #1 individuals tended towards the lowest rates of smoking, metabolic and respiratory comorbidities (__Supplementary Table S8__). 
+In the cluster #2 the lowest acute COVID-19 severity, rates of rehabilitation and lung CT abnormality grades were observed. 
+The cluster #2 tended also towards the lowest frequency of diastolic dysfunction.
+By contrast, in the cluster #3 affected by disease-related concerns, the frequency of severe COVID-19, COVID-19-specific rehabilitation rate, chest CT finding severity and PSS frequency at the one-year follow-up were the highest (__Figure \@ref(fig:fig-clust-cov)__).
+In particular, cluster #3 individuals suffered from protracted fatigue, reduced physical performance, dyspnea and sleep problems to the highest extend, and hair loss and gastrointestinal complaints were observed solely in the cluster #3.
+Consequently, the PSS number and fatigue scoring peaked in the cluster #3. 
+PSS frequencies and intensity were comparable in the clusters #1 and #2 (__Figure \@ref(fig:fig-clust-symptoms)__). </span>
+
+<span custom-style = 'revision'>Generally, differences between the IP clusters in PSS readouts like 
+PSS number (Kruskal-Wallis test: `r clust_chara$test_results$fup %>% filter(variable == 'sympt_number') %>% .$plot_cap`), 
+fatigue rating (`r clust_chara$test_results$fup %>% filter(variable == 'Chalder_FS') %>% .$plot_cap`), 
+significant fatigue ($\chi^2$ test: `r clust_chara$test_results$fup %>% filter(variable == 'Chalder_FS-bimodal') %>% .$plot_cap`), 
+sleep problems (`r clust_chara$test_results$fup %>% filter(variable == 'sleep_sympt') %>% .$plot_cap`) or 
+hair loss (`r clust_chara$test_results$fup %>% filter(variable == 'hair_loss_sympt') %>% .$plot_cap`) and differences in rehabilitation rates (`r clust_chara$test_results$fup %>% filter(variable == 'rehabilitation') %>% .$plot_cap`) were of moderate-to-large effect size.
+Effects of differences in 
+lung CT abnormality (`r clust_chara$test_results$fup %>% filter(variable == 'ct_severity_any') %>% .$plot_cap`), 
+chest CT abnormality scoring (`r clust_chara$test_results$fup %>% filter(variable == 'ct_severity_score') %>% .$plot_cap`), 
+LFT deficits (`r clust_chara$test_results$fup %>% filter(variable == 'lufo_red') %>% .$plot_cap`) or 
+diastolic dysfunction (`r clust_chara$test_results$fup %>% filter(variable == 'diastolic_dysf') %>% .$plot_cap`) were substantially weaker (__Supplementary Table S8__).</span>
 
@@
 
-By penalized multi-parameter regression [@Park2012; @Tibshirani1996; @Zou2005] applied to `r length(globals$variables)` candidate explanatory variables including demographic, clinical, somatic symptom data, laboratory and cardiopulmonary assessment results, we could discern `r length(an$cmm_variables$ipq_total)` features explaining `r (blass_mod$cv_fit_stats$ipq_total$estimate[4] %>% signif(2) %>% unname) * 100`% of overall illness perception variance at one year after COVID-19.
-Those factors were: fatigue scoring, reduced physical performance and hair loss at the one-year follow-up and pre-existing respiratory comorbidity. 
-Of note, the effects of age, sex, COVID-19 severity or residual lung lesions in CT, LFT findings and heath abnormalities in TTE on long-term illness perception were negligible. 
-Furthermore, high levels of COVID-19-related concerns, consequences and emotional representation was observed primarily in the subset of participants with multiple residual somatic symptoms and moderate-to-severe acute COVID-19 course.
+<span custom-style = 'revision'>In our observational collective affected by PSS or abnormalities in cardiopulmonary tests one year after COVID, two IP facets were identified: an emotion, concern and consequences component, and a lacking control and coherence component.
+By multi-parameter modeling, we established signatures of total IP and the emotion/concern/consequences IP component. 
+They included reduced physical performance, fatigue intensity and PSS number. 
+The  emotions, concerns and consequences IP component correlated additionally with significant fatigue, hair loss, COVID-19-specific rehabilitation, respiratory comorbidity, and, to a lesser degree, with diastolic dysfunction, elevated levels of the inflammatory marker CRP and obesity.
+Three IP clusters of study participants were identified based on the BIPQ items. 
+Differences between the IP clusters in PPS readouts such as PSS number, fatigue rating or sleep problems were much more pronounced than differing rates of llung and heart abnormalities in CT, LFT and TTE.</span>
 
-Literature on illness perception in COVID-19 patients is scarce [@Man2021; @Burke2020; @Hamama2022; @Sadeghian2022; @Stafford2021]. 
-To our best knowledge, this is the first report assessing severity and components of illness perception in COVID-19 convalescents with persistent somatic symptoms or cardiopulmonary findings at the clinical assessment one year after diagnosis.
-The link between disease severity, symptoms, therapy control and illness perception is well founded in other chronic conditions [@Broadbent2006; @Schwartz2020; @Schoormans2020]. 
-In particular, we identified fatigue as a strong covariate of overall illness perception scoring after COVID-19 and as an important factor characterizing individuals with high levels of COVID-19-related concerns, emotional representation and consequences in clustering analysis.
-Similar effects of chronic fatigue were described in arthritis [@Schwartz2020] and hematological malignancy [@Schoormans2020]. 
-Both acute COVID-19 and its post-acute sequelae encompass various respiratory symptoms such as cough or dyspnea [@Sahanic2021; @Sonnweber2020; @Sonnweber2022; @Davis2021]. 
+We identified <span custom-style = 'revision'>fatigue, a leading persistent COVID-19 manifestation [@Chen2022; @Sudre2021; @Davis2021; @Sahanic2021; @Bierbauer2022]</span> as a strong covariate <span custom-style = 'revision'>of total IP severity and the emotion/concern/consequences IP component.</span>.
+Similar effects were described in arthritis [@Schwartz2020], hematological malignancy [@Schoormans2020], <span custom-style = 'revision'>and recently in post-COVID-19 condition [@Bierbauer2022].</span>
+<span custom-style = 'revision'>Physical performance was found to be another key factor of IP and its emotion/concern/consequences component, in line with findings in hospitalized COVID-19 survivors, where reduced mobility was paralleled by anxiety and depression signs [@Evans2021].
+Objective readouts of good physical fitness in turn correlated with shorter symptom duration, lower PSS count, less fatigue and improved mental health in ambulatory COVID-19 patients [@Jimeno-Almazan2022].</span>
+<span custom-style = 'revision'>Hair loss belongs to the most frequent post-COVID-19 condition manifestations [@Chen2022; @Sahanic2021; @Jimeno-Almazan2022] and may result from psychological stress and/or inflammation [@Sattur2021].
+Hair loss, along with CRP, was linked by us with the emotion/concern/consequences IP component and may reflect an interplay between protracted inflammation and deteriorated mental health.</span>
+Post-COVID-19 condition encompass various respiratory symptoms [@Sahanic2021; @Sonnweber2020; @Sonnweber2022; @Davis2021; @Sudre2021; @Chen2022; @Bierbauer2022]. 
 In addition, COVID-19 was found to exacerbate symptoms and worsen disease control in asthma [@Philip2022]. 
-Such superimposed COVID-19-dependent and independent airway manifestations may hence explain more severe illness perception in respiratory comorbidity in our study cohort.
+<span custom-style = 'revision'>In our collective, respiratory comorbidity was associated with the emotion/concern/consequences IP component, higher fatigue rating and LFT abnormalities. 
+Such superimposed subjective fatigue feeling and objective LFT deficits may explain more severe IP in individuals with respiratory comorbidity. 
+We proposed PSS count as a strong covariate of IP severity and the emotion/concern/consequences IP component. 
+In line with literature evidence [@Sudre2021; @Sahanic2021; @Sneller2022; @Chen2022], significantly more PSS at the one-year follow-up were discerned in females than male participants. 
+However, IP rating was comparable between the genders. 
+This may be explained by the less severe COVID-19 and less lung CT abnormalities in females. 
+Interestingly, we observed significant super-additive effects of fatigue, reduced physical performance plus lung CT abnormalities on IP. 
+Since pulmonary CT findings affected nearly 75% of male participants, these interactions may influence IP in a male-specific manner and compensate for the higher PSS burden in females.</span>
 
-By clustering analysis we could identify three subsets of participants differing in key illness perception components.
-Half of individuals displayed low severity of overall illness perception, good self-perceived coherence, personal and treatment control paralleled by low burden of persistent somatic symptoms.
-Another `r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[2]`% of participants showed a similarly low level of persistent somatic symptoms or fatigue. 
-Yet their illness perception was hallmarked by poor disease understanding and disbelief in personal and treatment control. 
-By contrast, the remaining minor cluster suffered from multiple somatic complaints at the one-year follow-up and was enriched in individuals with residual lung lesions in CT, severe COVID-19 course, significant fatigue, sleep problems and hair loss.
-Their illness perception was characterized primarily by intense emotional representation, concern, burden of consequences and disease identity. 
-This latter subset is of particular interest and concern for psychological and psychiatric management of post COVID-19 syndrome as high scoring of emotional components of illness perception was correlated with signs of shame, guilt, stress, depression and anxiety both in the general population during the pandemic [@Wierenga2021] and in acute COVID-19 [@Burke2020; @Hamama2022; @Sadeghian2022].
-The individuals in this cluster could potentially profit most from psychological and psychiatric interventions.
+<span custom-style = 'revision'>No explanatory factors affecting the lacking control/coherence IP component were identified by any of the modeling algorithms. 
+This may reflect two non-exclusive phenomena. 
+First, the internal consistency of the lacking control/coherence component was remarkably lower than consistency of the BIPQ tool or the emotion/concern/consequences component. 
+Second, the lacking control/coherence component covering cognitive IP aspects [@Broadbent2006] may be influenced primarily by COVID-19-independent factors such as socioeconomic background, education level or follow-up care.
+Low socioeconomic status was proposed as a risk factor of post-COVID-19 condition [@Dumont2022] and can affect IP measured by BIPQ [@Mafla2018].</span>
 
-Collectively, we demonstrate an important effect of persistent somatic symptoms rather than clinical assessment results revealing lung lesions, lung function deficits or heart abnormalities on severity and quality of illness perception at one year after COVID-19.
-In the long run, this aspect may bear consequences for the patient's physical and mental health following COVID-19 and public health in general. 
-Negative illness perceptions were found to accompany somatic symptom disorder, perpetuate symptoms in somatoform disorders and to predict higher future healthcare expenditure [@Frostholm2014]. 
-In multiple aspects, post COVID-19 condition resembles persistent somatic manifestations of functional or somatic symptom disorders [@Ballering2021].
+We identified three clusters of participants differing in key IP components. 
+<span custom-style = 'revision'>The cluster #1 encompassing roughly</span> half of the participants displayed low total IP severity, good self-perceived coherence, personal and treatment control paralleled by low PSS rates.
+Another `r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[2]`% of participants <span custom-style = 'revision'>assigned to the cluster #2 enriched in moderate COVID-19 survivors with the lowest grade of lung CT abnormalities</span> showed a similarly low levels of PSS or fatigue.
+Yet their IP was hallmarked by poor disease understanding and disbelief in personal or treatment control. 
+By contrast, the cluster #3 individuals suffered from multiple PSS such as significant fatigue, sleep problems, physical performance deficits or hair loss, <span custom-style = 'revision'>had the highest rehabilitation rate</span> and tended to have more residual lung lesions and severe acute COVID-19 course.
+Their IP was characterized by high severity of the emotion/concern/consequences IP component <span custom-style = 'revision'>and, paradoxically, good coherence and belief in disease control.</span> 
+Emotional IP dimensions were correlated with shame, guilt, stress, depression and anxiety both in the general population during the pandemic [@Wierenga2021] and COVID-19 [@Burke2020; @Hamama2022; @Sadeghian2022; @Bierbauer2022].
+Cluster #3 individuals <span custom-style = 'revision'>may be at particular risk of COVID-19-related mental disorders and profit most from psychological and psychiatric interventions, i.e.
+aiming at IP improvement [@Broadbent2009].</span>
+<span custom-style = 'revision'>Of note, differences in PSS and rehabilitation frequencies between the IP clusters were way stronger than differences in cardiopulmonary abnormalities, which corroborates our modeling results.</span>
+<span custom-style = 'revision'>Rehabilitation associated by us with high emotion/concern/consequences IP scoring may reflect physical and social disability following COVID-19. 
+However, rehabilitation might have helped cluster #3 individuals to understand the disease and foster trust in treatment and personal control. 
+As inferred from high rates of smoking, metabolic and respiratory comorbidities, the clusters #2 and #3 may include patients with lower socioeconomic status, which likely contributed to the more severe IP in these two subsets [@Dumont2022; @Mafla2018].</span>
+
+Collectively, we demonstrate an important effect of 
+<span custom-style = 'revision'>PSS on severity and quality of IP at the one-year follow-up which was found to outweigh the effects of objective lung and heart abnormalities. 
+Yet, our data suggests that particular PSS such as fatigue or physical performance loss may be amplified by cardiopulmonary sequelae at shaping negative IP.</span> 
+<span custom-style = 'revision'>Mounting evidence suggests that post-COVID-19 condition occurs independently of COVID-19 severity, cardiopulmonary findings, biochemical and immunity markers [@Sneller2022; @Sonnweber2022], and proposes demographic, acute disease-related [@Chen2022; @Sahanic2021; @Sudre2021; @Sneller2022], mental health and mental stress [@Sneller2022; @Hufner2022; @Wang2022; @Al-Aly2021] as explanatory factors.
+Hence, a positive feedback between IP, mental health and PSS may contribute to post-COVID-19 syndrome [@Sneller2022; @Bierbauer2022; @Hufner2022] in analogy to somatic symptom disorders [@Frostholm2014; @Ballering2021; @Kachaner2022].
+Additionally, IP may be affected by comorbidities, SARS-CoV-2-independent somatic symptoms, social and economic consequences of the pandemic [@Wisk2022; @Behnood2022; @Tutzer2021; @Matta2022; @Hinz2017].
+The symptom kinetics in our cohort [@Sonnweber2022; @Rass2022] showed an early rapid recovery followed by a plateau and/or relapse at 3 - 12 months after COVID-19. 
+This indicates acute SARS-CoV-2 infection as a likely trigger and reflects pathogen clearance in early convalescence. 
+Yet, it is likely that COVID-19-independent factors contribute to the stalled PSS recovery and recurrence during later convalescence.</span>
 
 Our study bears limitations. 
-The most important one was the low participant number and substantial participant dropout due to missing follow-up visits and the BIPQ answers, which may have resulted in a selection bias. 
-Furthermore, longitudinal rating of illness perception or inclusion of a general population control [@DiasNeto2021; @Lugo-Gonzalez2020; @Wierenga2021; @Shiloh2021] would allow us to assess possible improvement or worsening and explore factors associated with illness perception at consecutive time points.
-The study collective was recruited during the pandemic onset before introduction of effective anti-viral drugs and vaccination. 
-We were hence unable to assess the effect of improved treatment and prevention on illness perception [@Vollmann2021]. 
-Finally, the study variable set lacks parameters which may be vital for severity and character of illness perception such as family status, education and COVID-19 knowledge [@Man2021], media consumption [@Tutzer2021] and quarantine duration [@Burke2020].
+The most important one was the low participant number and substantial dropout due to missing follow-up visits and the BIPQ answers. 
+<span custom-style = 'revision'>This limitation was partly addressed by the regularized regression modeling methodology described as robust in low power, high dimension datasets [@Kirpich2018; @Morozova2015].
+Yet the IP clusters require validation in a larger collective.</span>
+<span custom-style = 'revision'>As inferred from the substantially lower severity COVID-19 severity in excluded individuals, the analysis inclusion process may have resulted in a selection bias towards moderate-to-severe COVID-19.</span>
+Furthermore, longitudinal IP rating or a general population control [@DiasNeto2021; @Lugo-Gonzalez2020; @Wierenga2021; @Shiloh2021; @Sneller2022] would allow us to assess temporal IP changes during recovery, IP-influencing factors at consecutive time points <span custom-style = 'revision'>and disease-independent effects [@Wisk2022; @Behnood2022; @Hinz2017].</span>
+The study collective was recruited during the <span custom-style = 'revision'>first outbreak of wild-type SARS-CoV-2</span>. 
+We were hence unable to assess the effects of improved treatment, prevention and vaccination [@Vollmann2021], <span custom-style = 'revision'>and of virus variants differing in pathogenicity and symptom spectrum.</span>
+The study variable set lacks potentially vital parameters for IP such as family status, education and COVID-19 knowledge [@Man2021], media consumption [@Tutzer2021], quarantine duration [@Burke2020] <span custom-style = 'revision'>and socioeconomic status [@Dumont2022].</span>
+<span custom-style = 'revision'>At design of our study initially conceptualized as a cardiopulmonary observation cohort [@Sonnweber2020], the multi-organ spectrum of COVID-19 symptoms was unknown. 
+For this reason, multiple PSS of relevance for mental health and IP like neuro-cognitive symptoms [@Hufner2022; @Chen2022; @Bierbauer2022] were not surveyed.</span>
 
@@
 
-Our results demonstrate, that persistent somatic symptoms rather than clinical assessment results, revealing lung lesions, lung function deficits or heart abnormalities, impact on severity and quality of illness perception at one year after COVID-19.
-Hence, it is to account not only for acute COVID-19 severity but also for the interplay between persistent somatic symptoms and individual illness perceptions when allocating rehabilitation and psychological or psychiatric resources.
+<span custom-style = 'revision'>One year after COVID, persistent somatic symptoms such as fatigue or physical performance loss impact on severity and quality of IP to a higher degree than lung and heart abnormalities revealed by a systematic clinical assessment.
+Hence, besides  COVID-19 severity, individual IP should be addressed when allocating rehabilitation, psychological or psychiatric resources. 
+Interventions addressing specifically unhelpful IP are a promising approach in treatment of post-COVID-19 condition.</span>
 
@@
 We acknowledge commitment of the study participants and the medical staff during the COVID-19 pandemic. 
-
-# Funding
-
 The study was funded by the research fund of the State of Tyrol (Project GZ 71934, to Judith Löffler-Ragg) and an Investigator-Initiated Study grant by Boehringer Ingelheim (IIS 1199-0424 to Ivan Tancevski).
@@
 
+# Author's contribution
+
+KH, SS, RH, BP, Günter Weiss, Gerlig Widmann, TS, IT, BSU and JLR designed the study; KH, SS, AL, AB, AP, CS, SK, KK, MA, BMF, BP, VR, AS, SI, AE, EW collected the study data; KH, PT, SS and JLR analyzed the data and wrote the manuscript, IT and JLR acquired the funding.
+All authors reviewed and approved the final version of the manuscript.
+
 # Conflict of interest
 
-Piotr Tymoszuk owns a data science company, Data Analytics as a Service Tirol. 
+All authors have completed the ICMJE uniform disclosure form at www.icmje.org/coi_disclosure.pdf and declare:
+Katharina Hüfner has received research grants from Austria Wirtchaftsservice (AWS) and the State of Tyrol as well as lecturer's honoraria from Forum Medizinische Fortbildung (FOMF), the Anton Proksch Institute and the Hospital of Schwaz.
+Piotr Tymoszuk owns a data science company, Data Analytics as a Service Tirol, and receives payments from statistical data analysis, bioinformatic and scientific writing services.
 Other authors declare that no conflict of interest exists.
@@
 An R data (RDa) file with anonymized patient data will be made available upon request to the corresponding author. 
-The study pipeline is available at https://github.com/PiotrTymoszuk/CovILD-IPQ.
+The study analysis pipeline is available at https://github.com/PiotrTymoszuk/CovILD-IPQ.
 
@@
 
-```{r tab-cohort-baseline, tab.cap = 'Baseline demographic and clinical characteristic of the study cohort.'}
+```{r tab-cohort-baseline, tab.cap = 'Baseline demographic and clinical characteristic of the study cohort.
+Numeric variables are presented as medians with interquartile ranges (IQR).
+Categorical variables are shown as percentages and counts within the strata.'}
 
@@
   width(7, width = 2.2, unit = 'cm') %>%
-  footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>%
-  footnote(4, 1, value = as_paragraph('Former or active smoker'), ref_symbols = 'b', part = 'body') %>% 
-  footnote(5, 1, value = as_paragraph('Overweight: body mass index 25 - 30 kg/m\u00B2, obesity: > 30 kg/m\u00B2'), ref_symbols = 'c', part = 'body') %>% 
+  footnote(1, 6:7, 
+           value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'),
+           ref_symbols = 'a', 
+           part = 'header') %>% 
+  footnote(4, 1, 
+           value = as_paragraph('Former or active smoker'), 
+           ref_symbols = 'b', 
+           part = 'body') %>% 
+  footnote(5, 1, 
+           value = as_paragraph('Overweight: body mass index 25 - 30 kg/m\u00B2, obesity: > 30 kg/m\u00B2'), 
+           ref_symbols = 'c', 
+           part = 'body') %>% 
+  footnote(17, 1, 
+           value = as_paragraph('WHO ordinal scale for clinical improvement'), 
+           ref_symbols = 'c', 
+           part = 'body') %>% 
   theme_vanilla
@@
 
-```{r tab-psych, tab.cap = 'Illness perception score and item values of the Brief Illness Perception Questionnaire (BIPQ, Q1 - Q8) at one year after COVID-19.'}
+```{r tab-psych, tab.cap = 'The total illness perception (IP) score (sum of all BIPQ items), emotion/concern/consequences IP component score (sum of items 1, 2, 5, 6 and 8), lacking control/coherence IP component score (sum of items 3, 4 and 7) and values of single Brief Illness Perception Questionnaire items (BIPQ, Q1 - Q8) at one year after COVID-19.
+Values are presented as medians with interquartile ranges (IQR).'}
 
@@
   footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>% 
-  footnote(2, 1, value = as_paragraph('Sum of BIPQ items Q1 - Q8.'), ref_symbols = 'b', part = 'body') %>% 
   theme_vanilla
@@
 ```
+
+\newpage
 
@@
 
-```{r fig-key-factors, fig.width = mm_inch(paper_fig$key_factors$w), fig.height = mm_inch(paper_fig$key_factors$h), fig.cap = 'Key factors associated with disease perception one year after COVID-19.'}
+```{r fig-key-factors, fig.width = mm_inch(paper_fig$key_factors$w), fig.height = mm_inch(paper_fig$key_factors$h), fig.cap = 'Key factors associated with illness perception one year after COVID-19.'}
 
@@
 
-__Figure \@ref(fig:fig-key-factors). Key factors associated with disease perception one year after COVID-19.__
+__Figure \@ref(fig:fig-key-factors). Key factors associated with illness perception one year after COVID-19.__
 
-_Illness perception score (sum of BIPQ items) at one year after COVID-19 was modeled as a function of `r length(globals$variables)` candidate independent variables using the Elastic Net, LASSO and Bayesian LASSO algorithms.
-Key factors affecting the illness perception score were identified as independent variables with non-zero coefficients in all three models._
-
-_(A) Numbers of variables with non-zero coefficients identified by each algorithm presented in a quasi-proportional Venn diagram. The key factors are listed next to the diagram._
-
-_(B) Relationship between the illness perception score and the categorical key factors: reduced physical performance (Eastern Cooperative Oncology Group [ECOG] score > 0) and hair loss at one year after COVID and respiratory comorbidity was investigated by Mann-Whitney test with Wilcoxon r effect size statistic.
-Effect size statistic and p values are indicated in the plot captions.
-Illness perception score values are presented in violin plots.
-Points represent single observations.
-Red diamonds and whiskers depict medians and interquartile ranges.
-Numbers of complete observations are displayed under the plots._
-
-_(C) Relationship between the illness perception score and likert Chalder's fatigue score (CFS) was investigated by Spearman's correlation.
-The correlation coefficient ($\rho$) and p value are indicated in the plot caption.
-Each point represents a single observations.
-The blue line with gray ribbon depict the fitted second order terms and 95% confidence intervals.
-The number of complete observations is displayed under the plot._
+_The total illness perception (IP) score (A) and the emotion/concern/consequences IP component score (B) at the one-year follow-up were modeled as a function of `r length(mod$variables)` candidate explanatory variables by Elastic Net, LASSO and Bayesian LASSO.
+Key factors for the total IP score and emotion/concern/consequences component score were identified as variables with non-zero coefficients in all three models.
+Key factor numbers identified by each algorithm are presented as quasi-proportional Venn diagrams.
+The key factors are listed next to the diagrams.
+CFS: Chalder's fatigue score; reduced performance: Eastern Cooperative Oncology Group (ECOG) $\geq$ 1; # PSS: number of persistent somatic symptoms at the one-year follow-up; CRP: C-reactive protein._
 
@@
 
-```{r fig-risk-clusters, fig.width = mm_inch(paper_fig$ipq_clusters$w), fig.height = mm_inch(paper_fig$ipq_clusters$h), fig.cap = 'Heterogeneity of illness perception and residual symptoms one year after COVID-19.'}
+```{r fig-total-bipq, fig.width = mm_inch(paper_fig$total_bipq$w), fig.height = mm_inch(paper_fig$total_bipq$h), fig.cap = 'Persistent somatic symptom number, reduced physical performance, fatigue and the total illness perception scoring.'}
+
+paper_fig$total_bipq$plot
+
+```
+
+__Figure \@ref(fig:fig-total-bipq). Persistent somatic symptom number, reduced physical performance, fatigue and the total illness perception scoring.__
+
+_Association of the total illness perception (IP) score with persistent somatic symptom number, reduced physical performance and fatigue scoring at the one-year follow-up.
+Statistical significance was determined by Kruskal-Wallis test with $\eta^2$ effect size statistic (symptom number), Mann-Whitney test with r effect size statistic (reduced performance) and Spearman's correlation (fatigue scoring).
+The total IP scores in the persistent somatic symptom number or physical performance strata are presented in violin plots; single observations are depicted as points, red diamonds with whiskers represent medians with interquartile ranges.
+The correlation with fatigue scoring is presented in a point plot; the blue line represents the fitted second-order trend and the gray ribbon depicts the 95% confidence intervals.
+Effect size statistic and p values are indicated in the plot captions.
+Numbers of complete observations are displayed in the plot captions or in the plot X axes.
+# PSS: number of persistent somatic symptoms at the one-year follow-up; CFS: Chalder's fatigue score; ECOG: Eastern Cooperative Oncology Group._
+
+\newpage
+
+```{r fig-sub1-bipq, fig.width = mm_inch(paper_fig$sub1_bipq$w), fig.height = mm_inch(paper_fig$sub1_bipq$h), fig.cap = 'Key factors influencing the emotional representation, concern and consequences component of illness perception.'}
+
+paper_fig$sub1_bipq$plot
+
+```
+
+__Figure \@ref(fig:fig-sub1-bipq). Key factors influencing the emotional representation, concern and consequences component of illness perception.__
+
+_Association of the total emotion/concern/consequences IP component score at the one-year follow-up with persistent somatic symptoms (A), rehabilitation status, cardiological and inflammatory abnormalities (B), and respiratory comorbidity and body weight (C).
+Statistical significance was assessed by Kruskal-Wallis test with $\eta^2$ effect size statistic (number of symptoms), Spearman's correlation (fatigue scoring) and Mann-Whitney test with r effect size statistic (remaining independent variables).The correlation with fatigue scoring is presented in a point plot; the blue line represents the fitted second-order trend and the gray ribbon depicts the 95% confidence intervals.
+For the remaining independent variables, the score values are presented in violin plots with single observations depicted as points, and red diamonds and whiskers representing medians with interquartile ranges.
+Effect size statistic and p values are indicated in the plot captions.
+Numbers of complete observations are displayed in the plot captions or in the plot X axes.
+# PSS: number of persistent somatic symptoms at the one-year follow-up, CFS: Chalder's fatigue score; ECOG: Eastern Cooperative Oncology Group; CRP: serum C-reactive protein._
+
+\newpage
+
+```{r fig-ipq-clusters, fig.width = mm_inch(paper_fig$ipq_clusters$w), fig.height = mm_inch(paper_fig$ipq_clusters$h), fig.cap = 'Clusters of illness perception'}
 
@@
 
-__Figure \@ref(fig:fig-risk-clusters). Heterogeneity of illness perception and residual symptoms one year after COVID-19.__
+__Figure \@ref(fig:fig-ipq-clusters). Clusters of illness perception__
 
-_Three subsets of study participants (illness perception clusters) were identified by clustering in respect to the Brief Illness Perception Questionnaire items (BIPQ, Q1 - Q8) with the PAM (partitioning around medoids) clustering algorithm and Euclidean distance metric.
-Numbers of observations assigned to the clusters are displayed under the plots._
+_Three subsets of study participants (illness perception [IP] clusters) were identified by clustering in respect to the Brief Illness Perception Questionnaire items (BIPQ, Q1 - Q8) with PAM (partitioning around medoids) and Euclidean distance metric.
+Numbers of observations assigned to the clusters are displayed next to the plots or in the plot X axes._
 
-_(A) Mean BIPQ item scores at one year after COVID-19 in the illness perception clusters.
+_(A) Mean BIPQ item scores at one year after COVID-19 in the IP clusters.
 Statistical significance was determined by Kruskal-Wallis test.
-P values are indicated below the variable names.
-Lines represent mean values, tinted ribbons depict 2$\times$SEM (standard error of the mean) intervals._
+P values are indicated below the item names.
+Lines represent mean values, tinted ribbons depict 2 $\times$ SEM (standard error of the mean) intervals._
 
-_(B) Illness perception score (sum of BIPQ items), number of symptoms and likert Chalder's fatigue score (CFS) at one year after COVID-19 in the illness perception clusters.
+_(B) The total IP score, emotion/concern/consequences and the lacking control/coherence IP component scores in the IP clusters.
 Statistical significance was determined by Kruskal-Wallis test with $\eta^2$ effect size statistic.
-Effect size statistic and p values are indicated in the plot captions.
 Response values are presented in violin plots.
 Points represent single observations.
-Red diamonds and whiskers depict medians and interquartile ranges._
-
-_(C) Frequencies of significant fatigue (bimodal CFS $\geq$ 4), reduced physical performance (Eastern Cooperative Oncology Group [ECOG] score > 0) and sleep problems at one year after COVID-19 in the illness perception clusters.
-Statistical significance was determined by $\chi^2$ test with Cramer's V effect size statistic.
+Red diamonds and whiskers depict medians and interquartile ranges.
 Effect size statistic and p values are indicated in the plot captions._
@@
 
+```{r fig-clust-cov, fig.width = mm_inch(paper_fig$clust_cov$w), fig.height = mm_inch(paper_fig$clust_cov$h), fig.cap = 'Course of COVID-19, rehabilitation and one-year follow-up sequelae of COVID-19 in the illness perception clusters.'}
+
+paper_fig$clust_cov$plot
+
+```
+
+__Figure \@ref(fig:fig-clust-cov). Course of COVID-19, rehabilitation and one-year follow-up sequelae of COVID-19 in the illness perception clusters.__
+
+_COVID-19 severity and post-COVID-19 rehabilitation status (A), persistent somatic symptom and cardiopulmonary abnormality rates (B) and chest computed tomography abnormality severity (C) in the illness perception (IP) clusters.
+Statistical significance was assessed by Kruskal-Wallis test with $\eta^2$ effect size statistic (numeric variables) or by $\chi^2$ test with Cramer V effect size statistic (categorical variables).
+Numeric variable values are presented in violin plots with single observations depicted as points, and red diamonds and whiskers representing medians with interquartile ranges.
+Categorical variable strata frequencies expressed as percentages within the IP cluster are presented in stack plots.
+Effect size statistic and p values are indicated in the plot captions or plot facets.
+Numbers of observations assigned to the clusters are displayed below the plots or in the plot X axes.
+A, HM, HS: ambulatory, hospitalized moderate and hospitalized severe acute COVID-19, PSS: persistent somatic symptoms, WHO COVID-19 severity: WHO ordinal scale for clinical improvement; LFT: lung function testing; CT: chest computed tomography._
+
+\newpage
+
+```{r fig-clust-symptoms, fig.width = mm_inch(paper_fig$clust_symptoms$w), fig.height = mm_inch(paper_fig$clust_symptoms$h), fig.cap = 'Persistent symptoms at the one-year follow-up in the illness perception clusters.'}
+
+paper_fig$clust_symptoms$plot
+
+```
+
+__Figure \@ref(fig:fig-clust-symptoms). Persistent somatic symptoms at the one-year follow-up in the illness perception clusters.__
+
+_(A) Frequencies of persistent somatic symptoms (PSS) in the illness perception (IP) clusters.
+Statistical significance was determined by $\chi^2$ test.
+Percentages of individuals with and without the specific PSS within the IP cluster are presented in a stack plot.
+Significance is indicated next in the plot facets.
+Numbers of complete observations are displayed next to the plot.
+bi.
+CFS: bimodal Chalder's fatigue score; reduced performance: ECOG (Eastern Cooperative Oncology Group) $\geq$ 1._
+
+_(B) Numbers of PSS and fatigue scoring in the IP clusters.
+Statistical significance was determined by Kruskal-Wallis test with $\eta^2$ effect size statistic.
+Response values are presented in violin plots.
+Points represent single observations.
+Red diamonds and whiskers depict medians and interquartile ranges.
+Effect size statistic and p values are indicated in the plot captions.
+CFS: Chalder's fatigue score; # PSS: number of persistent somatic symptoms._
+
+\newpage
+
 # References

### 2023-01-09 paper/markdown/reviewer_material.Rmd
Revision

--- 2022-08-25 paper/markdown/supplementary_material.Rmd
@@
 ---
-title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19"
-subtitle: "Supplementary Material"
+title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19 in a cross-sectional prospective cohort study"
+subtitle: "Figures for Reviewers"
 author: "CovILD study team"
@@
 bibliography: cov_biblio.bib
-csl: frontiers-in-immunology.csl
+csl: jpr.csl
+
 header-includes:
@@
 
-# Supplementary Methods
+# Figures for Reviewers
 
-## Study cohort and approval
+```{r fig-eli-performance, fig.width = mm_inch(rev_fig$eli_performance$w), fig.height = mm_inch(rev_fig$eli_performance$h), fig.cap = 'Performance and reproducibility of backward elimination multi-parameter models.'}
 
-The CovILD longitudinal observation cohort study (ClinicalTrials.gov: NCT04416100) includes adult ($\geq$ 18 year) patients with a symptomatic, PCR-confirmed SARS-CoV-2 infection. 
-The follow-up visits were scheduled at two, three, six months and one year after COVID-19 diagnosis. A total of 190 individuals were screened for participation.
-In the current report, a subset of the original CovILD cohort was investigated, which displayed (1) persistent COVID-19-related symptoms or (2) any abnormality in chest computed tomography (CT) or (4) any deficits in lung function testing (LFT) or (5) any cardiological abnormality in trans-thoracic echocardiography (TTE) at the one-year follow-up.
-The final analysis inclusion criterion was availability of complete Brief Illness Perception Questionnaire (IPQ) data [@Broadbent2006].
-
-n = 145 participants were recruited between March and June 2020 at three centers located in Tyrol, Austria: the Medical University of Innsbruck (Innsbruck), St.
-Vinzenz Hospital (Zams) and Karl-Landsteiner Rehabilitation Facility (Muenster).
-The reasons for screening failure were denied informed consent and self-declared logistic and temporary incompatibility with the scheduled study visits. 
-Out of n = `r nrow(cov_data$data_tbl)` participants, who completed the one-year follow-up, n = `r nrow(cov_data$clear_data)` suffered from symptoms or displayed cardiopulmonary abnormalities (CT, LFT or TTE findings), filled out the BIPQ and had a complete set of modeling variables (__Supplementary Table S\@ref(tab:tab-var-dict)__).
-The recruitment process scheme is presented in __Figure 1__. 
-Baseline characteristic of the study cohort is presented in __Table 1__, follow-up features are summarized in __Supplementary Tables S\@ref(tab:tab-cohort-fup)__ (symptoms, cardiopulmonary findings) and __S\@ref(tab:tab-cohort-biochemistry)__ (laboratory parameters).
-BIPQ item scores and the sum BIPQ score are summarized in __Table 2__.
-
-The study was performed in accordance with the Declaration of Helsinki and the European Data Policy.
-All participants gave written informed consent.
-The study protocol was approved by the ethics committee at the Medical University of Innsbruck (approval number 1103/2020).
-
-## Procedures and variables
-
-The protocol of each follow-up visit included a survey of COVID-19-related symptoms, blood laboratory parameter assessment, LFT, TTE and chest CT as described in more detail before [@Sonnweber2020; @Sonnweber2022; @Luger2022].
-Comorbidities and data on acute COVID-19 course (symptoms, laboratory parameters, functional tests and imaging) were recorded retrospectively at the two-month follow-up visit based on information provided by the participant and electronic patient records.
-
-### Baseline demographic and clinical status, acute COVID-19 severity
-
-The baseline demographic variables included sex (male/female), age at COVID-19 diagnosis, smoking history, body weight and height. 
-The following pre-existing comorbidities were surveyed: obesity, metabolic disease, cardiovascular disease, hypertension, type II diabetes, hypercholesterolemia, gastrointestinal disease, malignancy, chronic kidney disease, immune deficiency, pulmonary and respiratory conditions.
-Among the surveyed comorbidities, asthma, COPD (chronic obstructive pulmonary disease), interstitial lung disease and other chronic pulmonary disorders were subsumed under the 'respiratory disease' variable included in the independent modeling variable set.
-Study participants were stratified according to acute COVID-19 severity as ambulatory (outpatient, WHO grade 1 - 2), moderate (hospitalized at normal infection ward, no oxygen therapy, WHO grade 3 - 4) and severe COVID-19 survivors (hospitalized with oxygen therapy or mechanical ventilation or intensive care, WHO grade 5 - 7).
-
-### Cardiopulonary assessment
-
-LFT abnormality was defined as FVC (forced vital capacity) < 80% or FEV1 (forced expiratory volume in 1 second) < 80% or TLC (total lung capacity) < 80% or DLCO (diffusion lung capacity for carbon monoxide) < 80% of the predicted reference value or FEV1:FVC ratio < 70% of the predicted reference value [@Sonnweber2022].
-CT images were screened for ground glass opacities (GGO), consolidations, bronchiectasis and reticulations according to the Fleischner Society glossary [@Hansell2008]. 
-Chest CT abnormalities were scored separately for each lobe with the CT severity score [@Sonnweber2020; @Sonnweber2022; @Luger2022]: 0 - no abnormality, 1 - minimal (subtle GGO), 2 - mild (several GGO, subtle reticulation), 3 - moderate (multiple GGO, reticulation, small consolidation), 4 - severe (extensive GGO, consolidation, reticulation with distortion), 5 - massive (multiple findings, parenchymal destruction).
-The sum for all five lobes was used in the analysis. 
-Any chest CT abnormality is defined as CT severity score $\geq$ 1. 
-The most frequent cardiological abnormality in TTE was low grade diastolic dysfunction (__Supplementary Table S\@ref(tab:tab-cohort-fup)__). 
-No reduced left ventricular ejection fraction was observed in the participants included in the current analysis.
-
-### COVID-19-related persistent somatic symptoms and exertional capacity
-
-The following features were surveyed as COVID-19-related persistent somatic symptoms:
-
-- __reduced physical performance__ assessed with the ECOG scale (Eastern Cooperative Oncology Group). Reduced physical performance was defined as ECOG $\geq$ 1
-- __dyspnea__ rated with the mMRC scale (Modified Medical British Research Council). Dyspnea was defined as mMRC $\geq$ 1
-- __cough__, self-reported, surveyed as a yes/no item
-- __sleep problems__, self-reported, surveyed as a yes/no item
-- __night sweating__, self-reported, surveyed as a yes/no item
-- __hyposmia or anosmia__, self-reported, surveyed as a yes/no item
-- __dermatological symptoms__, self-reported, surveyed as a yes/no item
-- __gastrointestinal symptoms__, self-reported, surveyed as a yes/no item
-- __hair loss__, self-reported, surveyed as a yes/no item
-- __significant fatigue__, measured with the bimodal Chalder's Fatigue Scale (CFS). Significant fatigue was defined as bimodal CFS $\geq$ 4 [@Chalder1993; @Morriss1998]
-
-The individual counts of persistent somatic symptoms were subsequently calculated.
-Since only few participants displayed values of ECOG $\geq$ 2 or mMRC $\geq$ 2, the dichotomous 'reduced performance (ECOG $\geq$ 1)' and 'dyspnea (mMRC $\geq$ 1)' independent variables were used in modeling instead of the full numeric scales.
-Fatigue intensity was rated with the likert Chalder's Fatigue Scale (CFS) [@Chalder1993; @Morriss1998]. 
-Exertional capacity was assessed by six-minute walking distance (SMWD) test performed according to the ATS guidelines [@Crapo2012]. 
-The reference SMWD values were obtained as described by Crapo and colleagues [@Crapo2012] and differences between the actual and reference values were calculated.
-
-### Laboratory parameters
-
-Laboratory parameters encompassing hemoglobin (total and glycated hemoglobin [HbA1c]), biomarkers of iron turnover (ferritin [FT], transferrin saturation [TF-Sat], hepcidin and soluble transferrin receptor [sTFR]), inflammation and coagulation (C-reactive protein [CRP] and D-dimer) and cardiovascular pathology (N-terminal pro-brain natriuretic peptide [NT-proBNP]) were determined routinely at the certified laboratory of the Central Institute for Medical and Chemical Diagnostic at the Federal University Hospital in Innsbruck.
-Other markers of systemic inflammation, interleukin 6 (IL6) and procalcitonin, were within the normal ranges in the analysis collective.
-
-### Illness perception
-
-Illness perception was scored with the Brief Illness Perception Questionnaire (IPQ) developed by Broadbent and colleagues [@Broadbent2006]. 
-Each of the 8 BIPQ items (questions Q1 - Q8) was scored in with an 11-point likert scale:
-
-- __Q1 consequences__ (0: no consequences, 10: very severe consequences)
-- __Q2 timeline__  (0: very short, 10: forever)
-- __Q3 personal control__ (0: no control, 10: extreme control)
-- __Q4 treatment control__ (0: no control, 10: extreme control)
-- __Q5 identity__ (0: no complaints, 10: multiple severe complaints)
-- __Q6 concerns__ (0: no concerns at all, 10: extreme concerns)
-- __Q7 coherence__ (0: not at all, 10: I understand my illness very good)
-- __Q8 emotional representation__ (0: no emotions at all, 10: extreme emotions)
-
-The illness perception score was calculated as an arithmetic sum of items Q1 - Q8, with the negative items Q3, Q4 and Q7 inverted. 
-The BIPQ tool demonstrated acceptable consistency as measured by Cronbach's $\alpha$ = `r ipq_co$alpha_stats %>% filter(response == "ipq_total") %>% .$raw_alpha %>% signif(2)` (bootstrap percentile 95% CI: `r paste(ipq_co$alpha_stats %>% filter(response == "ipq_total") %>% .$plot_cap %>% stri_extract(regex = "\\[.*\\]$"))`) (__Supplementary Figure S\@ref(fig:fig-coherence)__) [@Cronbach1951].
-
-Variables used in the exploratory data analysis and multi-parameter modeling are listed in __Supplementary Table S\@ref(tab:tab-var-dict)__.
-
-## Data import, transformation and visualization methods
-
-Data import, transformation, statistical analysis and result visualization was done with R version 4.2.0 (R Foundation for Statistical Computing). 
-For import of data tables and variable transformation, the _tidyverse_ package bundle [@Wickham2019] and _rlang_ package [@Henry2022] was used. 
-The _ggplot2_ [@Wickham2016] and _cowplot_ [@Wilke2019] packages and the development package _figur_ (https://github.com/PiotrTymoszuk/figur) were used for data visualization. 
-Quasi-proportional Venn diagrams were generated with the _nVennR_ package [@Perez-Silva2018].
-The manuscript and supplementary material files were created with the _rmarkdown_ [@Allaire2022], _knitr_ [@Xie2022], _bookdown_ [@Xie2016] packages, _Affiliations_ Pandoc filter set (https://github.com/drdanholmes/Affiliations) and the development package _figur_ (https://github.com/PiotrTymoszuk/figur).
-
-## Hypothesis testing
-
-For statistical hypothesis testing, base R functions, the package _rstatix_ [@Kassambara2021] and development package _exda_ (https://github.com/PiotrTymoszuk/ExDA) were used. 
-Statistical significance of differences in frequency of categorical variable levels was assessed by $\chi^2$ test with Cramer V effect size statistic. 
-Since multiple study variables were non-normally distributed as investigated by Shapiro-Wilk test and quantile-quantile plots, differences in numeric variables were assessed by non-parametric Mann-Whitney (effect size: Wilcoxon r) or Kruskal-Wallis test (effect size: $\eta^2$), as appropriate.
-Correlations were analyzed with Spearman test (function _cor.test()_, package _stats_) and visualized by fitting a second order trend (function _geom_smooth()_, package _ggplot_, formula: $y \sim x + I(x^2)$, development package _exda_).
-
-## BIPQ tool consistency
-
-Spearman's pair-wise correlations between the BIPQ items were calculated (function _cor.test()_, package _stats_) and visualized in a bubble plot (__Supplementary Figure S\@ref(fig:fig-coherence)__). 
-Cronbach's $\alpha$ was calculated with package _psych_ [@Revelle2022; @Cronbach1951]. 95% confidence intervals for $\alpha$ were obtained by the bootstrap technique (n = 1000).
-
-## Multi-parameter modeling
-
-In multi-parameter modeling, a set of `r length(globals$variables)` independent variables was used. 
-The response, illness perception score defined as a sum of all BIPQ items, was square-root transformed to guarantee normal distribution. 
-For numeric variables, both first- and second-order terms were included to account for non-linear relationship between the independent variable and response. 
-During pre-processing, numeric variables (independent and responses) were normalized with mean centering (Z-scores). 
-The full independent variable list is provides in __Supplementary Table S\@ref(tab:tab-var-dict)__.
-
-Three L1 penalized regression algorithms were utilized to model illness perception score at one year after COVID-19: Elastic Net [@Zou2005], LASSO (least absolute shrinkage and selection operator) [@Tibshirani1996] and Bayesian LASSO [@Park2012].
-The choice of penalized regression over canonical backward elimination or forward selection was motivated by highly multi-dimensional nature of the data set.
-
-The Elastic Net and LASSO algorithms are implemented in R by the _glmnet_ package [@Friedman2010]. 
-The optimal $\lambda$ parameters for Elastic Net and LASSO regression were found as values associated with the minimum model deviance in 200-repeats 10-fold cross-validation (CV, function _cv.glmnet()_, Elastic Net: $\alpha$ = 0.5, LASSO: $\alpha$ = 1, $\lambda$ parameter 'lambda.min' was extracted).
-The optimal $\lambda$ values were `r filter(eln_mod$opt_lambda, response == "ipq_total")$lambda %>% signif(2)` for the Elastic Net and `r filter(lasso_mod$opt_lambda, response == "ipq_total")$lambda %>% signif(2)` for the LASSO model.
-Finally, Elastic Net and LASSO models with the optimal $\lambda$ values were constructed with _caret_ package (function _train()_) [@Kuhn2008] and their performance in the training data set and 10-repeats 10-fold CV was tested (__Supplementary Figure D\@ref(fig:fig-performance)__).
-Estimates of non-zero final model coefficients (__Supplementary Figures S\@ref(fig:fig-eln-est)__ and __S\@ref(fig:fig-lasso-est)__) were extracted with the _coef_ method (s parameter set to the optimal $\lambda$ value).
-
-The Bayesian LASSO algorithm is implemented in R by the _monomvn_ package [@Gramacy2022]. 
-The Bayesian Lasso model was constructed with the _caret_ package (function _train()_, method: 'blasso', iteration number: T = 1000). 
-The optimal value of the sparsity parameter, which controls the fraction of non-zero coefficients in the final model based on the posterior distributions, was found by 10-repeats 10-fold CV with the RMSE (root mean squared error) statistic as a selection criterion (__Supplementary Figure S\@ref(fig:fig-performance)__).
-The optimal sparsity value selected from the 0.1 - 0.7 range was `r blass_mod$models$ipq_total$bestTune[1, "sparsity"] %>% signif(2)`. 
-The model coefficient matrix X was extracted from the 'finalModel' slot of the caret model. 
-The coefficient values (__Supplementary Figure S\@ref(fig:fig-blasso-est)__) were calculated as matrix column medians over all algorithm iterations [@Park2012].
-
-Extraction of caret model performance measures: RMSE and $R^2$ defined as $1 - MSE/Var(y)$ (__Supplementary Figure S\@ref(fig:fig-performance)__) was accomplished by the in-house-developed package _caretExtra_ (https://github.com/PiotrTymoszuk/caretExtra).
-Model assumption testing (normality of residuals, Shapiro-Wilk test, method _residuals()_) and visual quality control based on plots of residuals versus fitted and residuals quantile-quantile plots was done with the _caretExtra_ package (method _plot()_).
-
-The key factors impacting on the illness perception score at one year after COVID-19 (`r an$cmm_variables$ipq_total %>% translate_var(out_value = "label_long") %>% de_capitalize %>% collapse_and`) were defined as variables with non-zero coefficients in all multi-parameter models.
-
-## Clustering analysis
-
-Study participants were clustered in respect to BIPQ items (`r colnames(model.frame(ipq_clust$clust_obj)) %>% translate_var(out_value = "label_long") %>% stri_replace(fixed = "Illness perception, ", replacement = "") %>% collapse_and`) by the PAM algorithm (partition around medoids, package _cluster_) [@Schubert2019] with the Euclidean distance measure between the observations (package _philentropy_) [@Drost2018].
-The BIPQ item scores were not pre-processed, since they were measured with the same likert scale. 
-The PAM/Euclidean distance procedure was chosen based on the high explained clustering variance (ratio of the between-cluster sum of squares to the total sum of squares) and stability (cluster assignment accuracy in 10-fold CV, cluster assignment by 7-nearest neighbors label propagation algorithm) [@Lange2004; @Leng2014] as compared with several other clustering algorithms (__Supplementary Figure S\@ref(fig:fig-clust-dev)A__).
-The choice of cluster number (k = `r ipq_clust$clust_obj$clust_assignment$clust_id %>% levels %>% length`) was based on the bend of the curve of within-cluster sum of squares (generated with package _factoextra_, __Supplementary Figure S\@ref(fig:fig-clust-dev)B__) [@Kassambara2020] and visual analysis of the distance heat map (__Supplementary Figure S\@ref(fig:fig-clust-dev)C__).
-Clustering object development, clustering variance calculation, cross-validation and visualization were performed with the in-house developed package _clustTools_ (https://github.com/PiotrTymoszuk/clustTools).
-
-## Data and code availability
-
-An R data (RDa) file with anonymized patient data will be made available upon request to the corresponding author. The study pipeline is available at https://github.com/PiotrTymoszuk/CovILD-IPQ.
-
-\newpage
-
-# Supplementary Tables
-
-```{r tab-var-dict, tab.cap = 'Study variables'}
-
-flextable::flextable(suppl_tbl$var_lexicon) %>% 
-  width(1, width = 4.5, unit = 'cm') %>% 
-  width(2, width = 5.6, unit = 'cm') %>% 
-  width(3, width = 1.5, unit = 'cm') %>% 
-  width(4, width = 6, unit = 'cm') %>% 
-  theme_vanilla
+rev_fig$eli_performance$plot
 
@@
 
-\newpage
+__Supplementary Figure S\@ref(fig:fig-eli-performance). Performance and reproducibility of backward elimination multi-parameter models.__
 
-```{r tab-cohort-fup, tab.cap = 'COVID-19 symptoms, performance, fatigue, exertional capacity, cardiopulmonary abnormalities and illness perception at the one-year follow-up.'}
+_The total illness perception (IP) score, the emotion/concern/consequences and the lacking control/coherence sub-scores at the one-year follow-up were modeled by linear regression with Akaike information criterion-driven backwards elimination of model terms._
 
-flextable::flextable(suppl_tbl$fup) %>% 
-  width(1, width = 4.4, unit = 'cm') %>% 
-  width(2:5, width = 2.8, unit = 'cm') %>% 
-  width(6, width = 2.5, unit = 'cm') %>% 
-  width(7, width = 2.1, unit = 'cm') %>%
-  footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>%
-  footnote(4, 1, value = as_paragraph('Eastern Cooperative Oncology Group performance score'), ref_symbols = 'b', part = 'body') %>% 
-  footnote(5, 1, value = as_paragraph('Modified Medical Research Council dyspnea score'), ref_symbols = 'c', part = 'body') %>% 
-  footnote(13:14, 1, value = as_paragraph("Chalder's Fatigure Score"), ref_symbols = 'd', part = 'body') %>% 
-  footnote(15, 1, value = as_paragraph('Lung function testing abnormality: FVC < 80% or FEV1 < 80% or TLC < 80% or DLCO < 80% predicted or FEV1:FVC < 70% predicted reference value'), ref_symbols = 'e', part = 'body') %>%
-  footnote(16, 1, value = as_paragraph('Any abnormality in chest computed tomography (CT), CT severity score ≥ 1'), ref_symbols = 'f', part = 'body') %>% 
-  footnote(17, 1, value = as_paragraph('Chest computer tomography severity score'), ref_symbols = 'g', part = 'body') %>% 
-  footnote(19, 1, value = as_paragraph('Six minute walking distance'), ref_symbols = 'h', part = 'body') %>% 
-  footnote(20, 1, value = as_paragraph('Six-minute walking distance, difference between the actual and reference value'), ref_symbols = 'i', part = 'body') %>% 
-  theme_vanilla
-
-```
-
-\newpage
-
-```{r tab-cohort-biochemistry, tab.cap = 'Laboratory parameters at the one-year follow-up.'}
-
-flextable::flextable(suppl_tbl$biochemistry_fup) %>%
-  width(1, width = 2.8, unit = 'cm') %>% 
-  width(2:5, width = 3.3, unit = 'cm') %>% 
-  width(6, width = 2.5, unit = 'cm') %>% 
-  width(7, width = 2.1, unit = 'cm') %>%
-  footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>%
-  footnote(2, 1, value = as_paragraph('Females: hemoglobin (Hb) < 120 g/dL, males: Hb < 140 g/dL'), ref_symbols = 'b', part = 'body') %>% 
-  footnote(3, 1, as_paragraph('Blood hemoglobin'), ref_symbols = 'c', part = 'body') %>% 
-  footnote(4, 1, as_paragraph('Serum ferritin'), ref_symbols = 'd', part = 'body') %>%
-  footnote(5, 1, as_paragraph('Serum transferrin saturation'), ref_symbols = 'e', part = 'body') %>% 
-  footnote(6, 1, as_paragraph('Serum soluble transferrin receptor'), ref_symbols = 'f', part = 'body') %>% 
-  footnote(8, 1, as_paragraph('Serum N-terminal pro-brain natriuretic peptide'), ref_symbols = 'g', part = 'body') %>% 
-  footnote(10, 1, as_paragraph('Serum C-reactive protein'), ref_symbols = 'h', part = 'body') %>% 
-  footnote(11, 1, as_paragraph('Blood glycated hamoglobin'), ref_symbols = 'i', part = 'body') %>% 
-  theme_vanilla
-
-```
-
-\newpage
-
-```{r tab-ipq-clusters, tab.cap = 'Significant (p < 0.05) and near-significant (p < 0.1) differences between the illness perception clusters.'}
-
-flextable::flextable(suppl_tbl$ipq_clusters) %>% 
-  width(1, width = 4.8, unit = 'cm') %>% 
-  width(2:4, width = 2.6, unit = 'cm') %>% 
-  width(5, width = 2.5, unit = 'cm') %>% 
-  width(6, width = 2.1, unit = 'cm') %>%
-  footnote(1, 5:6, value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>%
-  footnote(8, 1, value = as_paragraph('Eastern Cooperative Oncology Group performance score'), ref_symbols = 'b', part = 'body') %>% 
-  footnote(9, 1, value = as_paragraph('Modified Medical Research Council dyspnea score'), ref_symbols = 'c', part = 'body') %>% 
-  footnote(13:14, 1, value = as_paragraph("Chalder's Fatigure Score"), ref_symbols = 'd', part = 'body') %>% 
-  footnote(15, 1, value = as_paragraph('Lung function testing abnormality: FVC < 80% or FEV1 < 80% or TLC < 80% or DLCO < 80% predicted or FEV1:FVC < 70% predicted reference value'), ref_symbols = 'e', part = 'body') %>%
-  footnote(16, 1, value = as_paragraph('Any abnormality in chest computed tomography (CT), CT severity score ≥ 1'), ref_symbols = 'f', part = 'body') %>% 
-  theme_vanilla
-
-```
-
-\newpage
-
-# Supplementary Figures
-
-```{r fig-coherence, fig.width = mm_inch(suppl_fig$coherence$w), fig.height = mm_inch(suppl_fig$coherence$h), fig.cap = 'Coherence of the illness perception score.'}
-
-suppl_fig$coherence$plot
-
-```
-
-__Supplementary Figure S\@ref(fig:fig-coherence). Coherence of the illness perception score.__
-
-_Pair-wise correlation between particular items of the Brief Illness Perception Questionnaire (IPQ) was assessed by Spearman test and the results presented in the point plot.
-Point size corresponds to absolute values of the correlation coefficient $\rho$.
-Point color codes for the $\rho$ value.
-The points are labeled with $\rho$ values, significant effects are highlighted in bold.
-Internal consistency of the BIPQ tool was measured by Cronbach's $\alpha$.
-The $\alpha$ statistic with 95% confidence intervals and the number of complete observations are displayed in the plot caption._
-
-\newpage
-
-```{r fig-performance, fig.width = mm_inch(suppl_fig$perfomance$w), fig.height = mm_inch(suppl_fig$perfomance$h), fig.cap = 'Multi-parameter model performance.'}
-
-suppl_fig$perfomance$plot
-
-```
-
-__Supplementary Figure S\@ref(fig:fig-performance). Multi-parameter model performance.__
-
-_(A) Fraction of illness perception score variance explained by the Elastic Net, LASSO and Bayesian LASSO model in the training data set and 10-repeats 10-fold cross-validation (CV) was estimated by the $R^2$ statistic._
+_(A) Fraction of variance explained by the model in the training data set and 10-repeats 10-fold cross-validation (CV) was estimated by the $R^2$ statistic._
 
 _(B) Model error expressed as root mean squared error (RMSE) in the training and CV data sets._
-
-\newpage
-
-```{r fig-eln-est, fig.width = mm_inch(suppl_fig$eln_est$w), fig.height = mm_inch(suppl_fig$eln_est$h), fig.cap = 'Non-zero coefficient estimates of the Elastic Net multi-parameter model.'}
-
-suppl_fig$eln_est$plot
-
-```
-
-__Supplementary Figure S\@ref(fig:fig-eln-est). Non-zero coefficient estimates of the Elastic Net multi-parameter model.__
-
-_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the Elastic Net model are presented in the plot as points.
-Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$.
-Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption.
-The data points are labeled with their $\beta$ values._
-
-\newpage
-
-```{r fig-lasso-est, fig.width = mm_inch(suppl_fig$lasso_est$w), fig.height = mm_inch(suppl_fig$lasso_est$h), fig.cap = 'Non-zero coefficient estimates of the LASSO multi-parameter model.'}
-
-suppl_fig$lasso_est$plot
-
-```
-
-__Supplementary Figure S\@ref(fig:fig-lasso-est). Non-zero coefficient estimates of the LASSO multi-parameter model.__
-
-_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the LASSO (least absolute shrinkage and selection operator) model are presented in the plot as points.
-Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$.
-Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption.
-The data points are labeled with their $\beta$ values._
-
-\newpage
-
-```{r fig-blasso-est, fig.width = mm_inch(suppl_fig$blasso_est$w), fig.height = mm_inch(suppl_fig$blasso_est$h), fig.cap = 'Non-zero coefficient estimates of the Bayesian LASSO multi-parameter model.'}
-
-suppl_fig$blasso_est$plot
-
-```
-
-__Supplementary Figure S\@ref(fig:fig-blasso-est). Non-zero coefficient estimates of the Bayesian LASSO multi-parameter model.__
-
-_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the Bayesian LASSO model are presented in the plot as points.
-Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$.
-Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption.
-The data points are labeled with their $\beta$ values._
-
-\newpage
-
-```{r fig-cardiopulmo, fig.width = mm_inch(suppl_fig$cardiopulmo$w), fig.height = mm_inch(suppl_fig$cardiopulmo$h), fig.cap = 'Age, sex and cardiopulmonary abnormalities at the one-year follow-up and illness perception.'}
-
-suppl_fig$cardiopulmo$plot
-
-```
-
-__Supplementary Figure S\@ref(fig:fig-cardiopulmo). Age, sex and cardiopulmonary abnormalities at the one-year follow-up and illness perception..__
-
-_(A - D) Differences in the illness perception score between males and female (A), the study participants with and without lung function (LFT) testing abnormality (B), chest computed tomography (CT) abnormality (C) and heart diastolic dysfunction (D) were assessed by Mann-Whitney test with Wilcoxon r effect size statistic.
-Effect size statistic and p values are indicated in the plot captions.
-Illness perception score values are presented i violin plots.
-Each point represents a single observation.
-Red diamonds and whiskers represent medians with interquartile ranges.
-Numbers of individuals with and without cardiopulmonary abnormalities are displayed under the plots._
-
-_(F).
-Correlation of age at COVID-19 diagnosis and the illness perception score at one year after COVID-19 diagnosis was investigated by Spearman's test.
-Correlation coefficients ($\rho$) and p values are indicated in the plot captions.
-Each point represents a single observations.
-Blue lines with gray ribbons depict the fitted second order terms and 95% confidence intervals.
-Numbers of complete observations are displayed under the plots._
-
-\newpage
-
-```{r fig-clust-dev, fig.width = mm_inch(suppl_fig$clust_dev$w), fig.height = mm_inch(suppl_fig$clust_dev$h), fig.cap = 'Choice of the optimal clustering algorithm and the cluster number.'}
-
-suppl_fig$clust_dev$plot
-
-```
-
-__Supplementary Figure S\@ref(fig:fig-clust-dev). Choice of the optimal clustering algorithm and the cluster number.__
-
-_Study participants were clustered in respect to the Brief Illness Perception Questionnaire items (Q1 - Q8) using the PAM (partitioning around medoids) clustering algorithm and Euclidean distance between the observations.
-Numbers of complete observations and clustering variables are indicated in (A)_
-
-_(A) Comparison of clustering algorithm performance (PAM: partitioning around medoids, HCL: Ward D2 hierarchical clustering, KMEANS).
-Explained clustering variance was defined as a ratio of the between-cluster sum of squares to the total sum of squares, cluster assignment accuracy was investigated in 10-fold cross-validation (CV).
-Note the superior CV accuracy of the PAM/Euclidean distance._
-
-_(B, C) Choice of the optimal cluster number for the hierarchical clustering/Euclidean distance algorithm.
-The optimal cluster number was selected based on the bend of the curve of within-cluster sum of squares (B, dashed red line: the selected cluster number) and visual inspection of the distance heat map (C)._
-
-\newpage
-
-# References

### 2023-01-09 paper/markdown/supplementary_material.Rmd
Revision

@@
 ---
-title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19"
+title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19 in a cross-sectional prospective cohort study"
 subtitle: "Supplementary Material"
@@
 
-The CovILD longitudinal observation cohort study (ClinicalTrials.gov: NCT04416100) includes adult ($\geq$ 18 year) patients with a symptomatic, PCR-confirmed SARS-CoV-2 infection. 
-The follow-up visits were scheduled at two, three, six months and one year after COVID-19 diagnosis. A total of 190 individuals were screened for participation.
-In the current report, a subset of the original CovILD cohort was investigated, which displayed (1) persistent COVID-19-related symptoms or (2) any abnormality in chest computed tomography (CT) or (4) any deficits in lung function testing (LFT) or (5) any cardiological abnormality in trans-thoracic echocardiography (TTE) at the one-year follow-up.
-The final analysis inclusion criterion was availability of complete Brief Illness Perception Questionnaire (IPQ) data [@Broadbent2006].
+The CovILD longitudinal observation cohort (ClinicalTrials.gov: NCT04416100) includes adult ($\geq$ 18 year) patients with a symptomatic, PCR-confirmed SARS-CoV-2 infection. 
+The follow-up visits were scheduled at two, three, six months and one year after COVID-19 diagnosis. 
+In the current report, a subset of the original CovILD cohort was investigated, which displayed (1) persistent somatic symptoms (PSS) or (2) any abnormality in chest computed tomography (CT) or (4) any deficits in lung function testing (LFT) or (5) any cardiological abnormality in trans-thoracic echocardiography (TTE) at the one-year follow-up.
+The final analysis inclusion criterion was availability of complete Brief Illness Perception Questionnaire (IPQ) [@Broadbent2006] and the complete set of variables used in explorative data analysis, modeling and clustering (__Supplementary Table S\@ref(tab:tab-var-dict)__ and __S\@ref(tab:tab-modeling-vars)__).
 
-n = 145 participants were recruited between March and June 2020 at three centers located in Tyrol, Austria: the Medical University of Innsbruck (Innsbruck), St.
+The decision to analyze the study participant subset with COVID-19 sequlae was motivated by the design of the illness perception-measuring tool used in our report, the Brief Illness Perception Questionnaire (BIPQ), which was initially developed for and validated in somatic conditions [@Broadbent2006].
+
+A total of 190 individuals were screened for participation. 
+The screened subset size was limited by the number of individuals diagnosed with COVID-19 at the study centers, their potential availability for follow-up visits (residency in the study region) and available resources for the follow-up assessments.
+No statistical estimation of the sample size was performed, as the kinetic of COVID-19 recovery was entirely unknown at the pandemic onset. 
+Out of the positively screened individuals, 145 participants were recruited between March and June 2020 at three centers located in Tyrol, Austria: the Medical University of Innsbruck (Innsbruck), St.
 Vinzenz Hospital (Zams) and Karl-Landsteiner Rehabilitation Facility (Muenster).
 The reasons for screening failure were denied informed consent and self-declared logistic and temporary incompatibility with the scheduled study visits. 
-Out of n = `r nrow(cov_data$data_tbl)` participants, who completed the one-year follow-up, n = `r nrow(cov_data$clear_data)` suffered from symptoms or displayed cardiopulmonary abnormalities (CT, LFT or TTE findings), filled out the BIPQ and had a complete set of modeling variables (__Supplementary Table S\@ref(tab:tab-var-dict)__).
+Out of `r nrow(cov_data$raw_data)` participants, who completed the one-year follow-up, `r nrow(cov_data$clear_data)` met the inclusion criteria and were analyzed here. 
 The recruitment process scheme is presented in __Figure 1__. 
-Baseline characteristic of the study cohort is presented in __Table 1__, follow-up features are summarized in __Supplementary Tables S\@ref(tab:tab-cohort-fup)__ (symptoms, cardiopulmonary findings) and __S\@ref(tab:tab-cohort-biochemistry)__ (laboratory parameters).
-BIPQ item scores and the sum BIPQ score are summarized in __Table 2__.
+Significant and near significant differences between the CovILD participants included in the analysis and excluded due to lacking COVID-19 sequelae at the one-year follow-up or missing BIPQ data are listed in __Supplementary Table S\@ref(tab:tab-excluded)__
+Baseline characteristic of the study participants included in the analysis is presented in __Table 1__, follow-up features are summarized in __Supplementary Tables S\@ref(tab:tab-cohort-fup)__ (symptoms, cardiopulmonary findings) and __S\@ref(tab:tab-cohort-biochemistry)__ (laboratory parameters).
+BIPQ item scores, the sum BIPQ score, and the emotion/concern/consequences and lacking control/coherence IP component scores are summarized in __Table 2__.
 
@@
 
-The protocol of each follow-up visit included a survey of COVID-19-related symptoms, blood laboratory parameter assessment, LFT, TTE and chest CT as described in more detail before [@Sonnweber2020; @Sonnweber2022; @Luger2022].
+The protocol of each follow-up visit included a survey of PSS, blood laboratory parameter assessment, LFT, TTE and chest CT as described in more detail before [@Sonnweber2020; @Sonnweber2022; @Luger2022].
 Comorbidities and data on acute COVID-19 course (symptoms, laboratory parameters, functional tests and imaging) were recorded retrospectively at the two-month follow-up visit based on information provided by the participant and electronic patient records.
+
+Variables used in the exploratory data analysis and multi-parameter modeling are listed in __Supplementary Table S\@ref(tab:tab-var-dict)__ and __Supplementary Table S\@ref(tab:tab-modeling-vars)__.
 
@@
 Among the surveyed comorbidities, asthma, COPD (chronic obstructive pulmonary disease), interstitial lung disease and other chronic pulmonary disorders were subsumed under the 'respiratory disease' variable included in the independent modeling variable set.
-Study participants were stratified according to acute COVID-19 severity as ambulatory (outpatient, WHO grade 1 - 2), moderate (hospitalized at normal infection ward, no oxygen therapy, WHO grade 3 - 4) and severe COVID-19 survivors (hospitalized with oxygen therapy or mechanical ventilation or intensive care, WHO grade 5 - 7).
+Study participants were stratified according to the acute COVID-19 severity as ambulatory (outpatient, WHO ordinal scale for clinical improvement 1 - 2), moderate (hospitalized at normal infection ward, no oxygen therapy, WHO 3 - 4) and severe COVID-19 survivors (hospitalized with oxygen therapy or mechanical ventilation or intensive care, WHO 5 - 7).
 
-### Cardiopulonary assessment
+### Cardiopulmonary assessment
 
@@
 CT images were screened for ground glass opacities (GGO), consolidations, bronchiectasis and reticulations according to the Fleischner Society glossary [@Hansell2008]. 
-Chest CT abnormalities were scored separately for each lobe with the CT severity score [@Sonnweber2020; @Sonnweber2022; @Luger2022]: 0 - no abnormality, 1 - minimal (subtle GGO), 2 - mild (several GGO, subtle reticulation), 3 - moderate (multiple GGO, reticulation, small consolidation), 4 - severe (extensive GGO, consolidation, reticulation with distortion), 5 - massive (multiple findings, parenchymal destruction).
-The sum for all five lobes was used in the analysis. 
-Any chest CT abnormality is defined as CT severity score $\geq$ 1. 
-The most frequent cardiological abnormality in TTE was low grade diastolic dysfunction (__Supplementary Table S\@ref(tab:tab-cohort-fup)__). 
+Chest CT abnormalities were scored separately for each lobe with the CT severity score (CTSS) [@Sonnweber2020; @Sonnweber2022; @Luger2022]: 0 - no abnormality, 1 - minimal (subtle GGO), 2 - mild (several GGO, subtle reticulation), 3 - moderate (multiple GGO, reticulation, small consolidation), 4 - severe (extensive GGO, consolidation, reticulation with distortion), 5 - massive (multiple findings, parenchymal destruction).
+The sum CTSS for all five lobes was used in the analysis. 
+Any chest CT abnormality is defined as CTSS $\geq$ 1. 
+The most frequent cardiological abnormality in TTE was low grade diastolic dysfunction. 
 No reduced left ventricular ejection fraction was observed in the participants included in the current analysis. 
+Frequencies of cardiopulmonary findings at the one-year follow-up are presented in __Supplementary Table S\@ref(tab:tab-cohort-fup)__ and __Supplementary Figure S\@ref(fig:fig-overlap)__.
 
-### COVID-19-related persistent somatic symptoms and exertional capacity
+### Persistent somatic symptoms and exertional capacity
 
-The following features were surveyed as COVID-19-related persistent somatic symptoms:
+The following features were surveyed as PSS at the one-year follow-up:
 
@@
 
-The individual counts of persistent somatic symptoms were subsequently calculated.
+Records of reduced physical performance, dyspnea, cough, sleep problems, night sweating, hyposmia/anosmia and gastrointestinal symptoms were available for acute disease and all scheduled study visits [@Sonnweber2022].
+
 Since only few participants displayed values of ECOG $\geq$ 2 or mMRC $\geq$ 2, the dichotomous 'reduced performance (ECOG $\geq$ 1)' and 'dyspnea (mMRC $\geq$ 1)' independent variables were used in modeling instead of the full numeric scales.
 Fatigue intensity was rated with the likert Chalder's Fatigue Scale (CFS) [@Chalder1993; @Morriss1998]. 
-Exertional capacity was assessed by six-minute walking distance (SMWD) test performed according to the ATS guidelines [@Crapo2012]. 
+Exertional capacity was assessed by six-minute walking distance (SMWD) test performed according to the American Thoracic Society guidelines [@Crapo2012]. 
 The reference SMWD values were obtained as described by Crapo and colleagues [@Crapo2012] and differences between the actual and reference values were calculated.
+Frequencies of symptoms, ratings of fatigue and exertional capacity at the one-year follow-up are shown in __Supplementary Table S\@ref(tab:tab-cohort-fup)__, __Supplementary Figures S\@ref(fig:fig-sev-symptoms)__ and  __S\@ref(fig:fig-overlap)__.
 
@@
 
-Laboratory parameters encompassing hemoglobin (total and glycated hemoglobin [HbA1c]), biomarkers of iron turnover (ferritin [FT], transferrin saturation [TF-Sat], hepcidin and soluble transferrin receptor [sTFR]), inflammation and coagulation (C-reactive protein [CRP] and D-dimer) and cardiovascular pathology (N-terminal pro-brain natriuretic peptide [NT-proBNP]) were determined routinely at the certified laboratory of the Central Institute for Medical and Chemical Diagnostic at the Federal University Hospital in Innsbruck.
-Other markers of systemic inflammation, interleukin 6 (IL6) and procalcitonin, were within the normal ranges in the analysis collective.
+Laboratory parameters encompassing hemoglobin (total and glycated hemoglobin [HbA1c]), biomarkers of iron turnover (ferritin [FT], transferrin saturation [TF-Sat], hepcidin and soluble transferrin receptor [sTFR]), inflammation and coagulation (C-reactive protein [CRP] and D-dimer) and cardiovascular pathology (N-terminal pro-brain natriuretic peptide [NT-proBNP]) were determined routinely at the certified laboratory of the Central Institute for Medical and Chemical Diagnostic at the University Hospital in Innsbruck.
+Other markers of systemic inflammation, interleukin 6 (IL6) and procalcitonin (PCT), were within the normal ranges in the analysis collective. 
+Values of laboratory parameters at the one-year follow-up are listed in __Supplementary Table S\@ref(tab:tab-cohort-biochemistry)__.
 
@@
 
-Illness perception was scored with the Brief Illness Perception Questionnaire (IPQ) developed by Broadbent and colleagues [@Broadbent2006]. 
-Each of the 8 BIPQ items (questions Q1 - Q8) was scored in with an 11-point likert scale:
+Illness perception (IP) was scored with a German translation of the Brief Illness Perception Questionnaire (BIPQ) [@Broadbent2006]. 
+Each of the 8 BIPQ items (questions Q1 - Q8) was scored in with an 11-point likert scale, the items Q3, Q4 and Q7 were conceptualized as negative items are were inverted prior to the analysis:
 
 - __Q1 consequences__ (0: no consequences, 10: very severe consequences)
-- __Q2 timeline__  (0: very short, 10: forever)
-- __Q3 personal control__ (0: no control, 10: extreme control)
-- __Q4 treatment control__ (0: no control, 10: extreme control)
+
+- __Q2 timeline__ (0: very short disease duration, 10: forever)
+
+- __Q3 lacking personal control__ (negative item, 0: full control, 10: no control at all)
+
+- __Q4 lacking treatment control__ (negative item, 0: full control, 10: no control at all)
+
 - __Q5 identity__ (0: no complaints, 10: multiple severe complaints)
+
 - __Q6 concerns__ (0: no concerns at all, 10: extreme concerns)
-- __Q7 coherence__ (0: not at all, 10: I understand my illness very good)
+
+- __Q7 lacking coherence__ (negative item, 0: very good disease understanding, 10: no understanding at all)
+
 - __Q8 emotional representation__ (0: no emotions at all, 10: extreme emotions)
 
-The illness perception score was calculated as an arithmetic sum of items Q1 - Q8, with the negative items Q3, Q4 and Q7 inverted. 
-The BIPQ tool demonstrated acceptable consistency as measured by Cronbach's $\alpha$ = `r ipq_co$alpha_stats %>% filter(response == "ipq_total") %>% .$raw_alpha %>% signif(2)` (bootstrap percentile 95% CI: `r paste(ipq_co$alpha_stats %>% filter(response == "ipq_total") %>% .$plot_cap %>% stri_extract(regex = "\\[.*\\]$"))`) (__Supplementary Figure S\@ref(fig:fig-coherence)__) [@Cronbach1951].
+The total IP score was calculated as an arithmetic sum of items Q1 - Q8. 
+As shown by the results of factor analysis [@Bartlett1937] of the BIPQ tool items (__Supplementary Figure S\@ref(fig:fig-factor-analysis)__), two separate factors could be identified: the emotion/concern/consequences (items Q1, Q2, Q5, Q6, Q8) and the lacking control/coherence component (items Q3, Q4, Q7).
+For this reason, two IP component scores were calculated and analysed as well: the emotion/concern/consequences score defined as the Q1, Q2, Q5, Q6, Q8 item sum and the lacking control/coherence score defined as the Q3, Q4, Q7 item sum.
+Values of the total IP score, the emotion/concern/consequences and lacking control/coherence IP component scores as well as values of the BIPQ items recorded at the one-year follow-up are presented in __Table 2__.
 
-Variables used in the exploratory data analysis and multi-parameter modeling are listed in __Supplementary Table S\@ref(tab:tab-var-dict)__.
+As measured by McDonald's total $\omega$ values [@McDonald1999], the full BIPQ tool ($\omega$ = `r ipq_co$omega_obj$omega.group$total[1] %>% signif(2)`), the emotion, concern, consequence ($\omega$ = `r ipq_co$omega_obj$omega.group$total[2] %>% signif(2)`) and the lacking control/coherence ($\omega$ = `r ipq_co$omega_obj$omega.group$total[3] %>% signif(2)`) demonstrated acceptable-to-good internal consistency (__Supplementary Figure S\@ref(fig:fig-coherence)__).
 
-## Data import, transformation and visualization methods
+## Software
 
 Data import, transformation, statistical analysis and result visualization was done with R version 4.2.0 (R Foundation for Statistical Computing). 
-For import of data tables and variable transformation, the _tidyverse_ package bundle [@Wickham2019] and _rlang_ package [@Henry2022] was used. 
-The _ggplot2_ [@Wickham2016] and _cowplot_ [@Wilke2019] packages and the development package _figur_ (https://github.com/PiotrTymoszuk/figur) were used for data visualization. 
+For import, handling and transformation of tabular data, the _tidyverse_ package bundle [@Wickham2019] and _rlang_ package [@Henry2022] was used. 
+Handling of text variables (search, replacement) was done with _stringi_ [@Gagolewski2021].
+
+For descriptive statistic, correlation and statistical testing for differences between analysis groups the package _rstatix_ [@Kassambara2021] and the development package _ExDA_ were used (https://github.com/PiotrTymoszuk/ExDA) were employed.
+Factor analysis was accomplished with the _stats_ package of base R and the development package _clustTools_ (https://github.com/PiotrTymoszuk/clustTools). 
+The BIPQ score consistency was investigated with the _psych_ package [@Revelle2022].
+
+For modeling and ANOVA, the development package _lmqc_ (https://github.com/PiotrTymoszuk/lmqc, backward elimination and two-way ANOVA) and the packages _glmnet_ (Elastic Net and LASSO) [@Friedman2010], _monomvn_ [@Gramacy2022] (Bayesian LASSO) and _caret_ [@Kuhn2008] were employed.
+Model quality control, cross-validation and fit statistics were computed with _caret_ and the development package _caretExtra_ (https://github.com/PiotrTymoszuk/caretExtra).
+
+Clustering was accomplished with tools provided by the packages _cluster_ [@Schubert2019], _philentropy_ [@Drost2018], _factoextra_ [@Kassambara2020] and the development package _clustTools_ (https://github.com/PiotrTymoszuk/clustTools).
+
+The _ggplot2_ [@Wickham2016] package were used for data visualization. 
+Bar and violin plots were generated with _ExDA_. 
 Quasi-proportional Venn diagrams were generated with the _nVennR_ package [@Perez-Silva2018]. 
-The manuscript and supplementary material files were created with the _rmarkdown_ [@Allaire2022], _knitr_ [@Xie2022], _bookdown_ [@Xie2016] packages, _Affiliations_ Pandoc filter set (https://github.com/drdanholmes/Affiliations) and the development package _figur_ (https://github.com/PiotrTymoszuk/figur).
+Figures were constructed with _cowplot_ [@Wilke2019]. 
+Tables were generated with _flextable_ [@Gohel2022]. 
+The manuscript and supplementary material files were created with the _rmarkdown_ [@Allaire2022], _knitr_ [@Xie2022], _bookdown_ [@Xie2016] packages and _Affiliations_ Pandoc filter set (https://github.com/drdanholmes/Affiliations).
+For management of figures, tables and R code in the RMarkdown documents, the development package _figur_ (https://github.com/PiotrTymoszuk/figur) was used.
 
-## Hypothesis testing
+## Descriptive statistic, distribution testing and hypothesis testing
 
-For statistical hypothesis testing, base R functions, the package _rstatix_ [@Kassambara2021] and development package _exda_ (https://github.com/PiotrTymoszuk/ExDA) were used. 
-Statistical significance of differences in frequency of categorical variable levels was assessed by $\chi^2$ test with Cramer V effect size statistic. 
-Since multiple study variables were non-normally distributed as investigated by Shapiro-Wilk test and quantile-quantile plots, differences in numeric variables were assessed by non-parametric Mann-Whitney (effect size: Wilcoxon r) or Kruskal-Wallis test (effect size: $\eta^2$), as appropriate.
-Correlations were analyzed with Spearman test (function _cor.test()_, package _stats_) and visualized by fitting a second order trend (function _geom_smooth()_, package _ggplot_, formula: $y \sim x + I(x^2)$, development package _exda_).
+Descriptive statistic: median with interquartile ranges for numeric variables, and percentages and counts for categorical variables were calculated with the `explore()` function (_ExDA_ package). 
+Statistical significance of differences in frequency of categorical variable levels was assessed by $\chi^2$ test with Cramer V effect size statistic (`compare_variables()`, the _ExDA_ package). 
+Normality was tested with Shapiro-Wilk test and quantile-quantile plots (`explore()`, _ExDA_ package).
 
-## BIPQ tool consistency
+Since multiple study variables were non-normally distributed, differences in numeric variables were assessed by non-parametric Mann-Whitney (effect size: r) or Kruskal-Wallis test (effect size: $\eta^2$), as appropriate.
+Group comparisons were done with the `compare_variables()` function (package _ExDA_).
+Correlations were analyzed with Spearman test (`correlate_variables()` package _ExDA_) and visualized by fitting a second order trend (`geom_smooth()`, package _ggplot_, formula: $y \sim x + I(x^2)$).
+Statistical significance for changes in frequency of PSS in time was investigated with Cochran Q test (function `cochran_qtest()`, package _rstatix_) [@Kassambara2021].
 
-Spearman's pair-wise correlations between the BIPQ items were calculated (function _cor.test()_, package _stats_) and visualized in a bubble plot (__Supplementary Figure S\@ref(fig:fig-coherence)__). 
-Cronbach's $\alpha$ was calculated with package _psych_ [@Revelle2022; @Cronbach1951]. 95% confidence intervals for $\alpha$ were obtained by the bootstrap technique (n = 1000).
+## BIPQ tool factor analysis and consistency assessment
+
+BIPQ dimensionality was assessed by maximum  likelihood factor analysis [@Bartlett1937] (`reduce_data(red_fun = 'fa')`, package _clustTools_, a wrapper around `factanal()` from _stats_) and visual investigation of the BIPQ item loadings (method `plot(type = 'loadings')` called for the factor analysis object).
+The factor analysis revealed presence of two components of the BIPQ tool: the emotion/concern/consequences and the lacking control/coherence factor (__Figure \@ref(fig:fig-factor-analysis)__).
+Spearman's pair-wise correlations between the BIPQ items were calculated (function 'correlate_variables()', package _ExDA_) and visualized in a bubble plot. 
+Due to multi-dimensionality of the BIPQ tool McDonald's $\omega$ was calculated with the function `omega()` from the package _psych_ [@Revelle2022; @McDonald1999] was used as a reliability measure for the global BIPQ and the emotion/concern/consequences and lacking control/coherence factors (__Supplementary Figure S\@ref(fig:fig-coherence)__).
 
@@
 
-In multi-parameter modeling, a set of `r length(globals$variables)` independent variables was used. 
-The response, illness perception score defined as a sum of all BIPQ items, was square-root transformed to guarantee normal distribution. 
-For numeric variables, both first- and second-order terms were included to account for non-linear relationship between the independent variable and response. 
+In multi-parameter modeling, a set of `r length(mod$variables)` independent variables was used (__Supplementary Table S\@ref(tab:tab-modeling-vars)__). 
+To improve normality, FT and sTFR were transformed with the logarithm function, and TF-Sat and hepcidin were transformed with square root. 
+Strongly non-normally or bimodally distributed numeric explanatory variables: 
+number of comorbidities (`r levels(cov_data$clear_data$no_comorb_class) %>% paste(collapse = ', ')`), 
+number of symptoms (`r levels(cov_data$clear_data$sympt_number_class) %>% paste(collapse = ', ')`), 
+CTSS ('none': CTSS 0, 'mild': 1 - 5, 'moderate': 5 - 10, 'severe': 11+) [@Luger2022], 
+NT-proBNP (`r levels(cov_data$clear_data$NTproBNP_class) %>% paste(collapse = ', ')`), 
+D-Dimer (`r levels(cov_data$clear_data$DDimer_class) %>% paste(collapse = ', ')`), 
+CRP (`r levels(cov_data$clear_data$DDimer_class) %>% paste(collapse = ', ')`) and 
+HbA1c (`r levels(cov_data$clear_data$DDimer_class) %>% paste(collapse = ', ')`) were stratified with clinically applicable cutoffs or cutoffs determined by visual analysis of the plots of modeling variable and the explanatory factor with fitted LOESS trends.
+For numeric explanatory variables, both first- and second-order terms were included in the model to account for possible non-linear relationship between the independent variable and response. 
 During pre-processing, numeric variables (independent and responses) were normalized with mean centering (Z-scores).
-The full independent variable list is provides in __Supplementary Table S\@ref(tab:tab-var-dict)__.
 
-Three L1 penalized regression algorithms were utilized to model illness perception score at one year after COVID-19: Elastic Net [@Zou2005], LASSO (least absolute shrinkage and selection operator) [@Tibshirani1996] and Bayesian LASSO [@Park2012].
-The choice of penalized regression over canonical backward elimination or forward selection was motivated by highly multi-dimensional nature of the data set.
+The modeling responses: (1) the total IP score (all BIPQ item sum), (2) the emotion/concern/consequences (Q1, Q2, Q5, Q6 and Q8 BIPQ item sum) and (3) the lacking control/coherence (Q3, Q4 and Q7 BIPQ item sum) IP component scores were square root-transformed to guarantee normal distribution.
+
+Three L1 regularized regression algorithms were utilized to model the total IP score and the IP component scores at one year after COVID-19: Elastic Net [@Zou2005], LASSO (least absolute shrinkage and selection operator) [@Tibshirani1996] and Bayesian LASSO [@Park2012].
+The choice of penalized regression over canonical backward elimination or forward selection was motivated by highly multi-dimensional nature of the data set [@Kirpich2018]. 
+As expected, modeling of the IP responses with canonical linear regression with Akaike information criterion-driven backward elimination (function `stepAIC()`, package MASS) [@Ripley2022] led to highly over-parameterized models as evident from huge expansion of the fit errors in cross-validation (CV) (not shown).
 
 The Elastic Net and LASSO algorithms are implemented in R by the _glmnet_ package [@Friedman2010]. 
-The optimal $\lambda$ parameters for Elastic Net and LASSO regression were found as values associated with the minimum model deviance in 200-repeats 10-fold cross-validation (CV, function _cv.glmnet()_, Elastic Net: $\alpha$ = 0.5, LASSO: $\alpha$ = 1, $\lambda$ parameter 'lambda.min' was extracted).
-The optimal $\lambda$ values were `r filter(eln_mod$opt_lambda, response == "ipq_total")$lambda %>% signif(2)` for the Elastic Net and `r filter(lasso_mod$opt_lambda, response == "ipq_total")$lambda %>% signif(2)` for the LASSO model.
-Finally, Elastic Net and LASSO models with the optimal $\lambda$ values were constructed with _caret_ package (function _train()_) [@Kuhn2008] and their performance in the training data set and 10-repeats 10-fold CV was tested (__Supplementary Figure D\@ref(fig:fig-performance)__).
-Estimates of non-zero final model coefficients (__Supplementary Figures S\@ref(fig:fig-eln-est)__ and __S\@ref(fig:fig-lasso-est)__) were extracted with the _coef_ method (s parameter set to the optimal $\lambda$ value).
+The optimal $\lambda$ parameters for Elastic Net and LASSO regression were found as values associated with the minimum model deviance of the optimally regularized model in 200-repeats 10-fold CV (function `cv.glmnet()`, Elastic Net: $\alpha$ = 0.5, LASSO: $\alpha$ = 1, $\lambda$ parameter 'lambda.1se' corresponding to the optimally regularized model).
+The optimal $\lambda$ values in Elastic Net were 
+`r filter(eln_mod$opt_lambda, response == "ipq_total")$lambda %>% signif(2)` for the total IP score, 
+`r filter(eln_mod$opt_lambda, response == "ipq_sub1")$lambda %>% signif(2)` for the emotion/concern/consequences and 
+`r filter(eln_mod$opt_lambda, response == "ipq_sub2")$lambda %>% signif(2)` for the lacking control/coherence IP component scores.
+The optimal $\lambda$ values in LASSO were 
+`r filter(lasso_mod$opt_lambda, response == "ipq_total")$lambda %>% signif(2)` for the total IP score, 
+`r filter(lasso_mod$opt_lambda, response == "ipq_sub1")$lambda %>% signif(2)` for the emotion/concern/consequences and 
+`r filter(lasso_mod$opt_lambda, response == "ipq_sub2")$lambda %>% signif(2)` for the lacking control/coherence IP component score. 
+Elastic Net and LASSO models with the optimal $\lambda$ values were constructed with _caret_ package (function `train(method = 'glmnet')`) [@Kuhn2008] and their performance in the training data set and 10-repeats 10-fold CV was tested.
+As shown in __Supplementary Figure S\@ref(fig:fig-performance)__, for the total IP score and the emotion/concern/consequences IP component score, meaningful, reproducible and acceptably parameterized models could be built with either of the algorithms as suggested by comparison of the model performance in the training and CV data sets.
+In turn, poorly performing, estimate-only models could be obtained for the lacking control/coherence IP component. 
+Estimates of non-zero coefficients were extracted from the final models with the _coef_ method (the `s` parameter set to the optimal $\lambda$ value) (__Supplementary Figures S\@ref(fig:fig-eln-est-total)__ - __S\@ref(fig:fig-blasso-est-sub1)__).
 
 The Bayesian LASSO algorithm is implemented in R by the _monomvn_ package [@Gramacy2022]. 
-The Bayesian Lasso model was constructed with the _caret_ package (function _train()_, method: 'blasso', iteration number: T = 1000). 
-The optimal value of the sparsity parameter, which controls the fraction of non-zero coefficients in the final model based on the posterior distributions, was found by 10-repeats 10-fold CV with the RMSE (root mean squared error) statistic as a selection criterion (__Supplementary Figure S\@ref(fig:fig-performance)__).
-The optimal sparsity value selected from the 0.1 - 0.7 range was `r blass_mod$models$ipq_total$bestTune[1, "sparsity"] %>% signif(2)`. 
+The Bayesian LASSO model was constructed with the _caret_ package (function `train(method = 'blasso')`, iteration number: T = 1000). 
+The optimal value of the sparsity parameter, which controls the fraction of non-zero coefficients in the final model based on the posterior distributions, was found by 10-repeats 10-fold CV with the RMSE (root mean squared error) statistic as a selection criterion.
+The optimal sparsity values selected from the 0.1 - 0.7 range were `r blass_mod$models$ipq_total$bestTune[1, "sparsity"] %>% signif(2)`, `r blass_mod$models$ipq_sub1$bestTune[1, "sparsity"] %>% signif(2)` and `r blass_mod$models$ipq_sub2$bestTune[1, "sparsity"] %>% signif(2)` for the total IP score, the emotion/concern/consequences and the lacking control/coherence IP component scores, respectively.
+Similar to the Elastic Net and LASSO algorithms, meaningful models could be obtained for the total IP score and the emotion/concern/consequences IP component score but not for the lacking control/coherence IP component score (__Supplementary Figure S\@ref(fig:fig-performance)__).
 The model coefficient matrix X was extracted from the 'finalModel' slot of the caret model. 
-The coefficient values (__Supplementary Figure S\@ref(fig:fig-blasso-est)__) were calculated as matrix column medians over all algorithm iterations [@Park2012].
+The coefficient values (__Supplementary Figures S\@ref(fig:fig-blasso-est-total)__ and __S\@ref(fig:fig-blasso-est-sub1)__) were calculated as matrix column medians over all algorithm iterations [@Park2012].
 
-Extraction of caret model performance measures: RMSE and $R^2$ defined as $1 - MSE/Var(y)$ (__Supplementary Figure S\@ref(fig:fig-performance)__) was accomplished by the in-house-developed package _caretExtra_ (https://github.com/PiotrTymoszuk/caretExtra).
-Model assumption testing (normality of residuals, Shapiro-Wilk test, method _residuals()_) and visual quality control based on plots of residuals versus fitted and residuals quantile-quantile plots was done with the _caretExtra_ package (method _plot()_).
+Extraction of caret model performance measures: RMSE and $R^2$ defined as $1 - MSE/Var(y)$ (__Supplementary Figure S\@ref(fig:fig-performance)__) was accomplished by the in-house-developed package _caretExtra_ (modeling objects created with `as_caretex()`).
+Model assumption testing (normality of residuals, Shapiro-Wilk test, method `residuals()`) and visual quality control based on plots of residuals versus fitted and residuals quantile-quantile plots was done with the _caretExtra_ package (method `plot()` called for the `caretex` object).
 
-The key factors impacting on the illness perception score at one year after COVID-19 (`r an$cmm_variables$ipq_total %>% translate_var(out_value = "label_long") %>% de_capitalize %>% collapse_and`) were defined as variables with non-zero coefficients in all multi-parameter models.
+The number of non-zero model coefficients was found to differ substantially between the regularized regression models (__Supplementary Figures S\@ref(fig:fig-eln-est-total)__ - __S\@ref(fig:fig-blasso-est-sub1)__), which was reported for data sets of similar size [@Morozova2015].
+Hence, the key factors impacting on the total IP score and the emotion/concern/consequences IP component at one year after COVID-19 (`r an$cmm_variables$ipq_total %>% translate_var(out_value = "label_long") %>% de_capitalize %>% collapse_and`) were defined as variables with non-zero coefficients in all multi-parameter models.
+Since no meaningful models for the lacking control/coherence could be obtained, no influential factors for the lacking control/coherence component could be identified (__Figure 2__).
+
+## Analysis of interaction with two-way ANOVA
+
+Interaction between the key factors affecting the total IP score and the emotion/concern/consequences IP component score and presence of cardiopulmonary abnormalities (CT, LFT findings or diastolic dysfunction in TTE) was analyzed by two-way ANOVA.
+Briefly, the interaction ANOVA models were constructed with the function `make_lm()` (employing `lm()` from base R, package _lmqc_). 
+The dependent variables were square root-transformed to improve normality. 
+The assumptions of normality and homogeneity of distribution of the model residuals were checked by Shapiro-Wilk and Levene tests, respectively (method `summary(type = 'assumptions')`) and standard diagnostic plots of model residuals (method `plot()`).
+ANOVA results with the $\eta^2$ effect size statistic (fraction of variance explained by the specific ANOVA term) were extracted from the models with the `anova()` method (package _lmqc_). 
+The significant and near significant (p < 0.1) interactions are presented in __Supplementary Figures S\@ref(fig:fig-interaction-total)__ - __S\@ref(fig:fig-interaction-sub1-lufo-dia)__.
 
@@
 
-Study participants were clustered in respect to BIPQ items (`r colnames(model.frame(ipq_clust$clust_obj)) %>% translate_var(out_value = "label_long") %>% stri_replace(fixed = "Illness perception, ", replacement = "") %>% collapse_and`) by the PAM algorithm (partition around medoids, package _cluster_) [@Schubert2019] with the Euclidean distance measure between the observations (package _philentropy_) [@Drost2018].
+Study participants were clustered in respect to the BIPQ items by PAM (partition around medoids, originally implemented by the package _cluster_) [@Schubert2019] with Euclidean distance between the observations (originally implemented by the package _philentropy_) [@Drost2018].
 The BIPQ item scores were not pre-processed, since they were measured with the same likert scale. 
-The PAM/Euclidean distance procedure was chosen based on the high explained clustering variance (ratio of the between-cluster sum of squares to the total sum of squares) and stability (cluster assignment accuracy in 10-fold CV, cluster assignment by 7-nearest neighbors label propagation algorithm) [@Lange2004; @Leng2014] as compared with several other clustering algorithms (__Supplementary Figure S\@ref(fig:fig-clust-dev)A__).
-The choice of cluster number (k = `r ipq_clust$clust_obj$clust_assignment$clust_id %>% levels %>% length`) was based on the bend of the curve of within-cluster sum of squares (generated with package _factoextra_, __Supplementary Figure S\@ref(fig:fig-clust-dev)B__) [@Kassambara2020] and visual analysis of the distance heat map (__Supplementary Figure S\@ref(fig:fig-clust-dev)C__).
-Clustering object development, clustering variance calculation, cross-validation and visualization were performed with the in-house developed package _clustTools_ (https://github.com/PiotrTymoszuk/clustTools).
+Technically, the PAM clustering object was constructed with the `kcluster(distance_method = 'euclidean', clust_fun = 'pam')` function from the development package _clustTools_.
+
+The PAM/Euclidean distance procedure was chosen based on the high explained clustering variance (ratio of the between-cluster sum of squares to the total sum of squares, method `var()`, package _clustTools_) and stability (cluster assignment accuracy in 10-fold CV, cluster assignment by 7-nearest neighbors classifier, method `cv()`, package _clustTools_) [@Lange2004; @Leng2014] as compared with several other clustering algorithms (hierarchical clustering and KMEANS) (__Supplementary Figure S\@ref(fig:fig-clust-dev)A__).
+The choice of cluster number (k = `r ipq_clust$clust_obj$clust_assignment$clust_id %>% levels %>% length`) was based on the bend of the curve of within-cluster sum of squares (method `plot()`, package _clustTools_, a wrapper around `fviz_nbclust()`, package _factoextra_) [@Kassambara2020] and visual analysis of the distance heat map (__Supplementary Figure S\@ref(fig:fig-clust-dev)BC__).
 
@@
 
-An R data (RDa) file with anonymized patient data will be made available upon request to the corresponding author. The study pipeline is available at https://github.com/PiotrTymoszuk/CovILD-IPQ.
+An R data (RDa) file with anonymized patient data will be made available upon request to the corresponding author. 
+The study analysis pipeline is available at https://github.com/PiotrTymoszuk/CovILD-IPQ.
 
@@
 
-```{r tab-var-dict, tab.cap = 'Study variables'}
+```{r tab-var-dict, tab.cap = 'Study variables. The table is available as a supplementary Excel file.'}
 
-flextable::flextable(suppl_tbl$var_lexicon) %>% 
-  width(1, width = 4.5, unit = 'cm') %>% 
-  width(2, width = 5.6, unit = 'cm') %>% 
-  width(3, width = 1.5, unit = 'cm') %>% 
-  width(4, width = 6, unit = 'cm') %>% 
+suppl_tbl$var_lexicon[1, 1] %>%
+  mutate(`R name` = '') %>% 
+  set_names(' ') %>% 
+  flextable
+
+```
+
+\newpage
+
+```{r tab-modeling-vars, tab.cap = 'Initial set of candidate explanatory variables in modeling of illness perception.'}
+
+flextable::flextable(suppl_tbl$modeling_vars) %>% 
+  width(1, width = 14, unit = 'cm') %>% 
   theme_vanilla
@@
 
-```{r tab-cohort-fup, tab.cap = 'COVID-19 symptoms, performance, fatigue, exertional capacity, cardiopulmonary abnormalities and illness perception at the one-year follow-up.'}
+```{r tab-excluded, tab.cap = 'Significant and near-significant (p < 0.1) factors differentiating between the CovILD cohort participants included in the analysis and excluded due to variable missingness or lacking long-term COVID-19 sequelae.
+Numeric variables are presented as medians with interquartile ranges (IQR).
+Categorical variables are shown as percentages and counts within the strata.'}
+
+flextable::flextable(suppl_tbl$excluded) %>% 
+  width(1, width = 4.8, unit = 'cm') %>% 
+  width(2:3, width = 3.5, unit = 'cm') %>% 
+  width(4, width = 2.5, unit = 'cm') %>% 
+  width(5, width = 2.1, unit = 'cm') %>% 
+  footnote(1, 4:5, 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Mann-Whitney test with r effect size statistic'), 
+           ref_symbols = 'a', 
+           part = 'header') %>% 
+  footnote(2, 1, 
+           value = as_paragraph('WHO ordinal scale for clinical improvement'), 
+           ref_symbols = 'b', 
+           part = 'body') %>% 
+  footnote(5, 1, 
+           value = as_paragraph('Six minute walking distance'), 
+           ref_symbols = 'b', 
+           part = 'body') %>% 
+  footnote(6, 1, 
+           value = as_paragraph('Blood hemoglobin'), 
+           ref_symbols = 'b', 
+           part = 'body') %>% 
+  footnote(7, 1, 
+           value = as_paragraph('Serum ferritin'), 
+           ref_symbols = 'b', 
+           part = 'body') %>% 
+  theme_vanilla()
+
+```
+
+\newpage
+
+```{r tab-cohort-fup, tab.cap = 'COVID-19 symptoms, performance, fatigue, exertional capacity, cardiopulmonary abnormalities and rehabilitation status at the one-year follow-up.
+Numeric variables are presented as medians with interquartile ranges (IQR).
+Categorical variables are shown as percentages and counts within the strata.'}
 
 flextable::flextable(suppl_tbl$fup) %>% 
-  width(1, width = 4.4, unit = 'cm') %>% 
-  width(2:5, width = 2.8, unit = 'cm') %>% 
-  width(6, width = 2.5, unit = 'cm') %>% 
+  width(1, width = 3.6, unit = 'cm') %>% 
+  width(2:5, width = 3.2, unit = 'cm') %>% 
+  width(6, width = 2.4, unit = 'cm') %>% 
   width(7, width = 2.1, unit = 'cm') %>%
-  footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>%
-  footnote(4, 1, value = as_paragraph('Eastern Cooperative Oncology Group performance score'), ref_symbols = 'b', part = 'body') %>% 
-  footnote(5, 1, value = as_paragraph('Modified Medical Research Council dyspnea score'), ref_symbols = 'c', part = 'body') %>% 
-  footnote(13:14, 1, value = as_paragraph("Chalder's Fatigure Score"), ref_symbols = 'd', part = 'body') %>% 
-  footnote(15, 1, value = as_paragraph('Lung function testing abnormality: FVC < 80% or FEV1 < 80% or TLC < 80% or DLCO < 80% predicted or FEV1:FVC < 70% predicted reference value'), ref_symbols = 'e', part = 'body') %>%
-  footnote(16, 1, value = as_paragraph('Any abnormality in chest computed tomography (CT), CT severity score ≥ 1'), ref_symbols = 'f', part = 'body') %>% 
-  footnote(17, 1, value = as_paragraph('Chest computer tomography severity score'), ref_symbols = 'g', part = 'body') %>% 
-  footnote(19, 1, value = as_paragraph('Six minute walking distance'), ref_symbols = 'h', part = 'body') %>% 
-  footnote(20, 1, value = as_paragraph('Six-minute walking distance, difference between the actual and reference value'), ref_symbols = 'i', part = 'body') %>% 
+  footnote(1, 6:7, 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), 
+           ref_symbols = 'a', 
+           part = 'header') %>% 
+  footnote(2:3, 1, 
+           value = as_paragraph('PSS: persistent somatic symptoms'), 
+           ref_symbols = 'c', 
+           part = 'body') %>% 
+  footnote(4, 1, 
+           value = as_paragraph('ECOG: Eastern Cooperative Oncology Group performance score'), 
+           ref_symbols = 'c', 
+           part = 'body') %>% 
+  footnote(5, 1, 
+           value = as_paragraph('mMRC: Modified Medical Research Council dyspnea score'), 
+           ref_symbols = 'd', 
+           part = 'body') %>% 
+  footnote(13:14, 1, 
+           value = as_paragraph("CFS: Chalder's Fatigure Score"), 
+           ref_symbols = 'e', 
+           part = 'body') %>% 
+  footnote(15, 1, 
+           value = as_paragraph('Lung function testing abnormality: FVC < 80% or FEV1 < 80% or TLC < 80% or DLCO < 80% predicted or FEV1:FVC < 70% predicted reference value'), 
+           ref_symbols = 'f', 
+           part = 'body') %>% 
+  footnote(17, 1, 
+           value = as_paragraph('Chest computer tomography severity score'), 
+           ref_symbols = 'g', 
+           part = 'body') %>% 
+  footnote(19, 1, 
+           value = as_paragraph('SMWD: Six minute walking distance'), 
+           ref_symbols = 'h', 
+           part = 'body') %>% 
+  footnote(20, 1, 
+           value = as_paragraph('SMWD, difference between the actual and reference value'), 
+           ref_symbols = 'i', 
+           part = 'body') %>% 
   theme_vanilla
@@
 
-```{r tab-cohort-biochemistry, tab.cap = 'Laboratory parameters at the one-year follow-up.'}
+```{r tab-cohort-biochemistry, tab.cap = 'Laboratory parameters at the one-year follow-up. Values are presented as medians with interquartile ranges (IQR).'}
 
 flextable::flextable(suppl_tbl$biochemistry_fup) %>% 
-  width(1, width = 2.8, unit = 'cm') %>% 
-  width(2:5, width = 3.3, unit = 'cm') %>% 
-  width(6, width = 2.5, unit = 'cm') %>% 
+  width(1, width = 2.5, unit = 'cm') %>% 
+  width(2:5, width = 3.5, unit = 'cm') %>% 
+  width(6, width = 2.4, unit = 'cm') %>% 
   width(7, width = 2.1, unit = 'cm') %>%
-  footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>%
-  footnote(2, 1, value = as_paragraph('Females: hemoglobin (Hb) < 120 g/dL, males: Hb < 140 g/dL'), ref_symbols = 'b', part = 'body') %>% 
-  footnote(3, 1, as_paragraph('Blood hemoglobin'), ref_symbols = 'c', part = 'body') %>% 
-  footnote(4, 1, as_paragraph('Serum ferritin'), ref_symbols = 'd', part = 'body') %>%
-  footnote(5, 1, as_paragraph('Serum transferrin saturation'), ref_symbols = 'e', part = 'body') %>% 
-  footnote(6, 1, as_paragraph('Serum soluble transferrin receptor'), ref_symbols = 'f', part = 'body') %>% 
-  footnote(8, 1, as_paragraph('Serum N-terminal pro-brain natriuretic peptide'), ref_symbols = 'g', part = 'body') %>% 
-  footnote(10, 1, as_paragraph('Serum C-reactive protein'), ref_symbols = 'h', part = 'body') %>% 
-  footnote(11, 1, as_paragraph('Blood glycated hamoglobin'), ref_symbols = 'i', part = 'body') %>% 
+  footnote(1, 6:7, 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), 
+           ref_symbols = 'a', 
+           part = 'header') %>% 
+  footnote(2, 1, 
+           value = as_paragraph('Females: hemoglobin (Hb) < 120 g/dL, males: Hb < 140 g/dL'), 
+           ref_symbols = 'b', 
+           part = 'body') %>% 
+  footnote(3, 1, 
+           as_paragraph('Blood hemoglobin'), 
+           ref_symbols = 'c', 
+           part = 'body') %>% 
+  footnote(4, 1, 
+           as_paragraph('Serum ferritin'), 
+           ref_symbols = 'd', 
+           part = 'body') %>%
+  footnote(5, 1, 
+           as_paragraph('Serum transferrin saturation'), 
+           ref_symbols = 'e', 
+           part = 'body') %>% 
+  footnote(6, 1, 
+           as_paragraph('Serum soluble transferrin receptor'), 
+           ref_symbols = 'f', 
+           part = 'body') %>% 
+  footnote(8, 1, 
+           as_paragraph('Serum N-terminal pro-brain natriuretic peptide'), 
+           ref_symbols = 'g', part = 'body') %>% 
+  footnote(10, 1, as_paragraph('Serum C-reactive protein'), 
+           ref_symbols = 'h', 
+           part = 'body') %>% 
+  footnote(11, 1, 
+           as_paragraph('Blood glycated hamoglobin'), 
+           ref_symbols = 'i', 
+           part = 'body') %>% 
   theme_vanilla
@@
 
-```{r tab-ipq-clusters, tab.cap = 'Significant (p < 0.05) and near-significant (p < 0.1) differences between the illness perception clusters.'}
+```{r tab-gender, tab.cap = 'Significant and near-significant (p < 0.1) factors differentiating between female and male participants and illness perception in the genders (Brief Illness Perception Questionnaire, BIPQ).
+Numeric variables are presented as medians with interquartile ranges (IQR).
+Categorical variables are shown as percentages and counts within the strata.'}
+
+flextable::flextable(suppl_tbl$gender) %>% 
+  width(1, width = 5.2, unit = 'cm') %>% 
+  width(2:3, width = 3.5, unit = 'cm') %>% 
+  width(4, width = 2.5, unit = 'cm') %>% 
+  width(5, width = 2.1, unit = 'cm') %>% 
+  footnote(1, 4:5, 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Mann-Whitney test with r effect size statistic'), 
+           ref_symbols = 'a', 
+           part = 'header') %>% 
+  footnote(2, 1, 
+           value = as_paragraph('Former or active smoker'), 
+           ref_symbols = 'b', 
+           part = 'body') %>% 
+  footnote(3, 1, 
+           value = as_paragraph('Overweight: body mass index 25 - 30 kg/m\u00B2, obesity: > 30 kg/m\u00B2'), 
+           ref_symbols = 'c', 
+           part = 'body') %>% 
+  footnote(5, 1, 
+           value = as_paragraph('WHO ordinal scale for clinical improvement'), 
+           ref_symbols = 'd', 
+           part = 'body') %>% 
+  footnote(7:8, 1, 
+           value = as_paragraph('PSS: persistent somatic symptoms'), 
+           ref_symbols = 'e', 
+           part = 'body') %>% 
+  footnote(9, 1, 
+           value = as_paragraph('mMRC: Modified Medical Research Council dyspnea score'), 
+           ref_symbols = 'f', 
+           part = 'body') %>% 
+  footnote(11, 1, 
+           value = as_paragraph('Chest computer tomography severity score'), 
+           ref_symbols = 'g', 
+           part = 'body') %>% 
+  footnote(12, 1, 
+           value = as_paragraph('Blood hemoglobin'), 
+           ref_symbols = 'h', 
+           part = 'body') %>% 
+  footnote(13, 1, 
+           value = as_paragraph('Serum ferritin'), 
+           ref_symbols = 'i', 
+           part = 'body') %>% 
+  theme_vanilla
+
+```
+
+\newpage
+
+```{r tab-respi, tab.cap = 'Significant and near-significant (p < 0.1) factors differentiating between participants with and without respiratory comorbidity.
+Numeric variables are presented as medians with interquartile ranges (IQR).
+Categorical variables are shown as percentages and counts within the strata.'}
+
+flextable::flextable(suppl_tbl$respi) %>% 
+  width(1, width = 4.8, unit = 'cm') %>% 
+  width(2:3, width = 3.3, unit = 'cm') %>% 
+  width(4, width = 2.5, unit = 'cm') %>% 
+  width(5, width = 2.1, unit = 'cm') %>% 
+  footnote(1, 4:5, 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Mann-Whitney test with r effect size statistic'), 
+           ref_symbols = 'a', 
+           part = 'header') %>% 
+  footnote(5:6, 1, 
+           value = as_paragraph("CFS: Chalder's fatigue score"), 
+           ref_symbols = 'b', 
+           part = 'body') %>% 
+  footnote(7, 1, 
+           as_paragraph('Lung function testing abnormality: FVC < 80% or FEV1 < 80% or TLC < 80% or DLCO < 80% predicted or FEV1:FVC < 70% predicted reference value'), 
+           ref_symbols = 'c', 
+           part = 'body') %>% 
+  footnote(8:12, 1, 
+           as_paragraph('BIPQ: brief illness perception questionnaire'), 
+           ref_symbols = 'd', 
+           part = 'body') %>% 
+  theme_vanilla
+
+```
+
+\newpage
+
+```{r tab-ipq-clusters, tab.cap = 'Significant and near-significant (p < 0.1) differences between the illness perception clusters.
+Numeric variables are presented as medians with interquartile ranges (IQR).
+Categorical variables are shown as percentages and counts within the strata.'}
 
 flextable::flextable(suppl_tbl$ipq_clusters) %>% 
-  width(1, width = 4.8, unit = 'cm') %>% 
-  width(2:4, width = 2.6, unit = 'cm') %>% 
+  width(1, width = 5.4, unit = 'cm') %>% 
+  width(2:4, width = 3, unit = 'cm') %>% 
   width(5, width = 2.5, unit = 'cm') %>% 
   width(6, width = 2.1, unit = 'cm') %>%
-  footnote(1, 5:6, value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>%
-  footnote(8, 1, value = as_paragraph('Eastern Cooperative Oncology Group performance score'), ref_symbols = 'b', part = 'body') %>% 
-  footnote(9, 1, value = as_paragraph('Modified Medical Research Council dyspnea score'), ref_symbols = 'c', part = 'body') %>% 
-  footnote(13:14, 1, value = as_paragraph("Chalder's Fatigure Score"), ref_symbols = 'd', part = 'body') %>% 
-  footnote(15, 1, value = as_paragraph('Lung function testing abnormality: FVC < 80% or FEV1 < 80% or TLC < 80% or DLCO < 80% predicted or FEV1:FVC < 70% predicted reference value'), ref_symbols = 'e', part = 'body') %>%
-  footnote(16, 1, value = as_paragraph('Any abnormality in chest computed tomography (CT), CT severity score ≥ 1'), ref_symbols = 'f', part = 'body') %>% 
+  footnote(1, 5:6, 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), 
+           ref_symbols = 'a', 
+           part = 'header') %>% 
+  footnote(2, 1, 
+           value = as_paragraph('Former or active smoker'), 
+           ref_symbols = 'b', 
+           part = 'body') %>% 
+  footnote(5, 1, 
+           value = as_paragraph('WHO ordinal scale for clinical improvement'), 
+           ref_symbols = 'c', 
+           part = 'body') %>% 
+  footnote(6:7, 1, 
+           value = as_paragraph('PSS: persistent somatic symptoms'), 
+           ref_symbols = 'd', 
+           part = 'body') %>% 
+  footnote(8, 1, 
+           value = as_paragraph('ECOG: Eastern Cooperative Oncology Group performance score'), 
+           ref_symbols = 'e', 
+           part = 'body') %>% 
+  footnote(9, 1, 
+           value = as_paragraph('mMRC: Modified Medical Research Council dyspnea score'), 
+           ref_symbols = 'f', 
+           part = 'body') %>% 
+  footnote(13:14, 1, value = as_paragraph("Chalder's Fatigure Score"), 
+           ref_symbols = 'g', 
+           part = 'body') %>% 
+  footnote(15, 1, 
+           value = as_paragraph('Lung function testing abnormality: FVC < 80% or FEV1 < 80% or TLC < 80% or DLCO < 80% predicted or FEV1:FVC < 70% predicted reference value'), 
+           ref_symbols = 'h', 
+           part = 'body') %>% 
   theme_vanilla
@@
 
-```{r fig-coherence, fig.width = mm_inch(suppl_fig$coherence$w), fig.height = mm_inch(suppl_fig$coherence$h), fig.cap = 'Coherence of the illness perception score.'}
+```{r fig-sev-symptoms, fig.width = mm_inch(suppl_fig$sev_symptoms$w), fig.height = mm_inch(suppl_fig$sev_symptoms$h), fig.cap = 'Longitudinal changes of symptom rates, frequency of persistent somatic symptoms in the COVID-19 severity strata at the one-year follow-up.'}
+
+suppl_fig$sev_symptoms$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-sev-symptoms).
+Longitudinal changes of persistent somatic symptom rates, frequency of persistent somatic symptoms in the COVID-19 severity strata at the one-year follow-up.__
+
+_(A) Frequencies of persistent somatic symptoms (PSS) in acute COVID-19 (time 0) and at the consecutive follow-ups after COVID-19 diagnosis expressed as percentages of the analysis collective.
+Only PSS recorded at every follow-up were analyzed.
+Statistical significance of changes in frequencies in time was investigated by Cochran Q test.
+Number of complete observations is indicated in the plot caption.
+P values for specific PSS are presented in the plot legend._
+
+_(B) Frequencies of PSS at the one-year follow up were compared between survivors of ambulatory, moderate and severe COVID-19 by $\chi^2$ test.
+Percentages of individuals with and without the specific PSS within the severity strata are presented in a stack plot.
+Significance is indicated in the plot facets.
+Numbers of complete observations are displayed next to the plot.
+bi.
+CFS: bimodal Chalder's fatigue score; reduced performance: ECOG (Eastern Cooperative Oncology Group) $\geq$ 1._
+
+\newpage
+
+```{r fig-overlap, fig.width = mm_inch(suppl_fig$overlap$w), fig.height = mm_inch(suppl_fig$overlap$h), fig.cap = 'Overlap between persistent somatic symptoms and cardiopulmonary findings at the one-year follow-up.'}
+
+suppl_fig$overlap$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-overlap). Overlap between persistent somatic symptoms and cardiopulmonary findings at the one-year follow-up.__
+
+_Numbers of the analyzed CovILD study participants (total: n = `r nrow(cov_data$clear_data)`) with at least one persistent somatic symptom (PSS), any abnormality in chest computed tomography (CT findings) or lung function testing (LFT abnormality) or diastolic dysfunction in trans-thoracic echocardiography (TTE abnormality) at the one-year follow-up presented in a quasi-proportional Venn plot._
+
+\newpage
+
+```{r fig-factor-analysis, fig.width = mm_inch(suppl_fig$factor_analysis$w), fig.height = mm_inch(suppl_fig$factor_analysis$h), fig.cap = 'Factor analysis of the BIPQ tool items.'}
+
+suppl_fig$factor_analysis$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-factor-analysis). Factor analysis of the BIPQ tool items.__
+
+_Brief illness perception questionnaire (BIPQ) items were subjected to three-dimensional factor analysis.
+Loadings of the BIPQ items for the first two factors are presented as points.
+Variances associated with the factors are shown in the plot axes.
+Numbers of complete observations are indicated in the plot caption._
+
+\newpage
+
+```{r fig-coherence, fig.width = mm_inch(suppl_fig$coherence$w), fig.height = mm_inch(suppl_fig$coherence$h), fig.cap = 'Coherence of the BIPQ tool.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-coherence). Coherence of the illness perception score.__
+__Supplementary Figure S\@ref(fig:fig-coherence). Coherence of the BIPQ tool.__
 
-_Pair-wise correlation between particular items of the Brief Illness Perception Questionnaire (IPQ) was assessed by Spearman test and the results presented in the point plot.
+_Pair-wise correlation between the Brief Illness Perception Questionnaire (BIPQ) items was assessed by Spearman test (observations: n = `r nrow(cov_data$clear_data)`) and the results presented in a bubble plot.
 Point size corresponds to absolute values of the correlation coefficient $\rho$.
@@
 The points are labeled with $\rho$ values, significant effects are highlighted in bold.
-Internal consistency of the BIPQ tool was measured by Cronbach's $\alpha$.
-The $\alpha$ statistic with 95% confidence intervals and the number of complete observations are displayed in the plot caption._
+Internal consistency of the BIPQ tool was measured by McDonald's $\omega$.
+$\omega$ values for the entire BIPQ tool, the emotion/concern/consequences (Q1/2/3/6/8) and the lacking control/coherence (Q3/Q4/Q7) factors are indicated in the plot caption._
 
@@
 
-```{r fig-performance, fig.width = mm_inch(suppl_fig$perfomance$w), fig.height = mm_inch(suppl_fig$perfomance$h), fig.cap = 'Multi-parameter model performance.'}
+```{r fig-performance, fig.width = mm_inch(suppl_fig$perfomance$w), fig.height = mm_inch(suppl_fig$perfomance$h), fig.cap = 'Performance and reproducibility of regularized multi-parameter models.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-performance). Multi-parameter model performance.__
+__Supplementary Figure S\@ref(fig:fig-performance). Performance and reproducibility of regularized multi-parameter models.__
 
-_(A) Fraction of illness perception score variance explained by the Elastic Net, LASSO and Bayesian LASSO model in the training data set and 10-repeats 10-fold cross-validation (CV) was estimated by the $R^2$ statistic._
-
-_(B) Model error expressed as root mean squared error (RMSE) in the training and CV data sets._
+_The total illness perception (IP) score (A), the emotion/concern/consequences (B) and the lacking control/coherence (C) IP component scores at the one-year follow-up were modeled by the Elastic Net, LASSO and Bayesian LASSO.
+Fraction of the modeling response variance explained by the model in the training data set and 10-repeats 10-fold cross-validation (CV) was estimated by the $R^2$ statistic.
+Model error was expressed as root mean squared error (RMSE) in the training and CV data sets._
 
@@
 
-```{r fig-eln-est, fig.width = mm_inch(suppl_fig$eln_est$w), fig.height = mm_inch(suppl_fig$eln_est$h), fig.cap = 'Non-zero coefficient estimates of the Elastic Net multi-parameter model.'}
+```{r fig-eln-est-total, fig.width = mm_inch(suppl_fig$eln_est_total$w), fig.height = mm_inch(suppl_fig$eln_est_total$h), fig.cap = 'Non-zero coefficient estimates of the Elastic Net multi-parameter model of the total illness perception score at the one-year follow-up.'}
 
-suppl_fig$eln_est$plot
+suppl_fig$eln_est_total$plot
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-eln-est). Non-zero coefficient estimates of the Elastic Net multi-parameter model.__
+__Supplementary Figure S\@ref(fig:fig-eln-est-total). Non-zero coefficient estimates of the Elastic Net multi-parameter model of the total illness perception (IP) score at the one-year follow-up.__
 
-_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the Elastic Net model are presented in the plot as points.
+_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the model are presented in the plot as points.
 Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$.
 Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption.
-The data points are labeled with their $\beta$ values._
+The data points are labeled with their $\beta$ values.
+CFS: Chalder's fatigue score; reduced performance: ECOG (Eastern Cooperative Oncology Group) $\geq$ 1; # PSS: number of persistent somatic symptoms at the one-year follow-up._
 
@@
 
-```{r fig-lasso-est, fig.width = mm_inch(suppl_fig$lasso_est$w), fig.height = mm_inch(suppl_fig$lasso_est$h), fig.cap = 'Non-zero coefficient estimates of the LASSO multi-parameter model.'}
+```{r fig-lasso-est-total, fig.width = mm_inch(suppl_fig$lasso_est_total$w), fig.height = mm_inch(suppl_fig$lasso_est_total$h), fig.cap = 'Non-zero coefficient estimates of the LASSO multi-parameter model of the total illness perception score at the one-year follow-up.'}
 
-suppl_fig$lasso_est$plot
+suppl_fig$lasso_est_total$plot
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-lasso-est). Non-zero coefficient estimates of the LASSO multi-parameter model.__
+__Supplementary Figure S\@ref(fig:fig-lasso-est-total). Non-zero coefficient estimates of the LASSO multi-parameter model of the total illness perception (IP) score at the one-year follow-up.__
 
-_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the LASSO (least absolute shrinkage and selection operator) model are presented in the plot as points.
+_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the model are presented in the plot as points.
 Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$.
 Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption.
-The data points are labeled with their $\beta$ values._
+The data points are labeled with their $\beta$ values.
+CFS: Chalder's fatigue score; reduced performance: ECOG (Eastern Cooperative Oncology Group) $\geq$ 1; # PSS: number of persistent somatic symptoms at the one-year follow-up._
 
@@
 
-```{r fig-blasso-est, fig.width = mm_inch(suppl_fig$blasso_est$w), fig.height = mm_inch(suppl_fig$blasso_est$h), fig.cap = 'Non-zero coefficient estimates of the Bayesian LASSO multi-parameter model.'}
+```{r fig-blasso-est-total, fig.width = mm_inch(suppl_fig$blasso_est_total$w), fig.height = mm_inch(suppl_fig$blasso_est_total$h), fig.cap = 'Non-zero coefficient estimates of the Bayesian LASSO multi-parameter model of the total illness perception score at the one-year follow-up.'}
 
-suppl_fig$blasso_est$plot
+suppl_fig$blasso_est_total$plot
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-blasso-est). Non-zero coefficient estimates of the Bayesian LASSO multi-parameter model.__
+__Supplementary Figure S\@ref(fig:fig-blasso-est-total).
+Non-zero coefficient estimates of the Bayesian LASSO multi-parameter model of the total illness perception (IP) score at the one-year follow-up.__
 
-_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the Bayesian LASSO model are presented in the plot as points.
+_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the model are presented in the plot as points.
 Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$.
 Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption.
-The data points are labeled with their $\beta$ values._
+The data points are labeled with their $\beta$ values.
+CFS: Chalder's fatigue score; reduced performance: ECOG (Eastern Cooperative Oncology Group) $\geq$ 1; # PSS: number of persistent somatic symptoms at the one-year follow-up._
 
@@
 
-```{r fig-cardiopulmo, fig.width = mm_inch(suppl_fig$cardiopulmo$w), fig.height = mm_inch(suppl_fig$cardiopulmo$h), fig.cap = 'Age, sex and cardiopulmonary abnormalities at the one-year follow-up and illness perception.'}
+```{r fig-eln-est-sub1, fig.width = mm_inch(suppl_fig$eln_est_sub1$w), fig.height = mm_inch(suppl_fig$eln_est_sub1$h), fig.cap = 'Non-zero coefficient estimates of the Elastic Net multi-parameter model of the emotion/concern/consequences illness perception component at the one-year follow-up.'}
 
-suppl_fig$cardiopulmo$plot
+suppl_fig$eln_est_sub1$plot
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-cardiopulmo). Age, sex and cardiopulmonary abnormalities at the one-year follow-up and illness perception..__
+__Supplementary Figure S\@ref(fig:fig-eln-est-sub1).
+Non-zero coefficient estimates of the Elastic Net multi-parameter model of the emotion/concern/consequences illness perception component the one-year follow-up.__
 
-_(A - D) Differences in the illness perception score between males and female (A), the study participants with and without lung function (LFT) testing abnormality (B), chest computed tomography (CT) abnormality (C) and heart diastolic dysfunction (D) were assessed by Mann-Whitney test with Wilcoxon r effect size statistic.
+_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the model are presented in the plot as points.
+Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$.
+Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption.
+The data points are labeled with their $\beta$ values.
+CFS: Chalder's fatigue score; reduced performance: ECOG (Eastern Cooperative Oncology Group) $\geq$ 1; # PSS: number of persistent somatic symptoms at the one-year follow-up; CT lesion severity: chest computed tomography abnormality severity at the one-year follow-up; GID: gastrointestinal disease; CRP: serum C-reactive protein; COVID-19 severity: HS: hospitalized/severe; NT-proBNP; N-terminal pro-brain natriuretic peptide._
+
+\newpage
+
+```{r fig-lasso-est-sub1, fig.width = mm_inch(suppl_fig$lasso_est_sub1$w), fig.height = mm_inch(suppl_fig$lasso_est_sub1$h), fig.cap = 'Non-zero coefficient estimates of the LASSO multi-parameter model of the emotion/concern/consequences illness perception component at the one-year follow-up.'}
+
+suppl_fig$lasso_est_sub1$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-lasso-est-sub1).
+Non-zero coefficient estimates of the LASSO multi-parameter model of the emotion/concern/consequences illness perception component at the one-year follow-up.__
+
+_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the model are presented in the plot as points.
+Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$.
+Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption.
+The data points are labeled with their $\beta$ values.
+CFS: Chalder's fatigue score; reduced performance: ECOG (Eastern Cooperative Oncology Group) $\geq$ 1; # PSS: number of persistent somatic symptoms at the one-year follow-up; CT lesion severity: chest computed tomography abnormality severity at the one-year follow-up; GID: gastrointestinal disease; CRP: serum C-reactive protein; COVID-19 severity: HS: hospitalized/severe; NT-proBNP; N-terminal pro-brain natriuretic peptide._
+
+\newpage
+
+```{r fig-blasso-est-sub1, fig.width = mm_inch(suppl_fig$blasso_est_sub1$w), fig.height = mm_inch(suppl_fig$blasso_est_sub1$h), fig.cap = 'Non-zero coefficient estimates of the Bayesian LASSO multi-parameter model of the emotion/concern/consequences illness perception illness perception component at the one-year follow-up.'}
+
+suppl_fig$blasso_est_sub1$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-blasso-est-sub1).
+Non-zero coefficient estimates of the Bayesian LASSO multi-parameter model of the emotion/concern/consequences illness perception component at the one-year follow-up.__
+
+_Z-score normalized estimates ($\beta$) of the non-zero coefficients of the model are presented in the plot as points.
+Point color codes $\beta$ value, point size corresponds to the absolute value of $\beta$.
+Model's performance measures in the training data set ($R^2$ and root mean squared error [RMSE]) and numbers of complete observations are indicated in the plot caption.
+The data points are labeled with their $\beta$ values.
+CFS: Chalder's fatigue score; reduced performance: ECOG (Eastern Cooperative Oncology Group) $\geq$ 1; # PSS: number of persistent somatic symptoms at the one-year follow-up; CRP: serum C-reactive protein._
+
+\newpage
+
+```{r fig-cardiopulmo-total, fig.width = mm_inch(suppl_fig$cardiopulmo_total$w), fig.height = mm_inch(suppl_fig$cardiopulmo_total$h), fig.cap = 'Age, sex, cardiopulmonary abnormalities and persistent somatic symptoms at the one-year follow-up and the total illness perception score.'}
+
+suppl_fig$cardiopulmo_total$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-cardiopulmo-total). Age, sex, cardiopulmonary abnormalities and persistent somatic symptoms at the one-year follow-up and the total illness perception score.__
+
+_Association of the total illness perception (IP) score with age at COVID-19 diagnosis, sex, and lung function (LFT) testing abnormality, chest computed tomography (CT) abnormality and heart diastolic dysfunction and presence of any persistent somatic symptoms (PSS) at the one-year follow-up.
+Statistical significance was assessed by Spearman's correlation (age) or Mann-Whitney test with r effect size statistic (remaining independent variables).
+The correlation with age is presented in a point plot; the blue line represents the fitted second-order trend and the gray ribbon depicts the 95% confidence intervals.
+For the remaining independent variables, the score values are presented in violin plots with single observations depicted as points, and red diamonds and whiskers representing medians with interquartile ranges.
 Effect size statistic and p values are indicated in the plot captions.
-Illness perception score values are presented i violin plots.
-Each point represents a single observation.
-Red diamonds and whiskers represent medians with interquartile ranges.
-Numbers of individuals with and without cardiopulmonary abnormalities are displayed under the plots._
+Numbers of complete observations are displayed in the plot captions or in the plot X axes._
 
-_(F).
-Correlation of age at COVID-19 diagnosis and the illness perception score at one year after COVID-19 diagnosis was investigated by Spearman's test.
-Correlation coefficients ($\rho$) and p values are indicated in the plot captions.
-Each point represents a single observations.
-Blue lines with gray ribbons depict the fitted second order terms and 95% confidence intervals.
-Numbers of complete observations are displayed under the plots._
+\newpage
+
+```{r fig-cardiopulmo-sub, fig.width = mm_inch(suppl_fig$cardiopulmo_sub$w), fig.height = mm_inch(suppl_fig$cardiopulmo_sub$h), fig.cap = 'Age, sex, cardiopulmonary abnormalities and persistent somatic symptoms at the one-year follow-up and the illness perception components.'}
+
+suppl_fig$cardiopulmo_sub$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-cardiopulmo-sub). Age, sex, cardiopulmonary abnormalities and persistent somatic symptoms at the one-year follow-up and the illness perception components.__
+
+_Association of the emotion/concern/consequences (A) and the lacking control/coherence (B) IP component scores with age at COVID-19 diagnosis, sex, and lung function (LFT) testing abnormality, chest computed tomography (CT) abnormality and heart diastolic dysfunction and presence of any persistent somatic symptoms (PSS) at the one-year follow-up.
+Statistical significance was assessed by Spearman's correlation (age) or Mann-Whitney test with r effect size statistic (remaining independent variables).
+The correlation with age is presented in a point plot; the blue line represents the fitted second-order trend and the gray ribbon depicts the 95% confidence intervals.
+For the remaining independent variables, the score values are presented in violin plots with single observations depicted as points, and red diamonds and whiskers representing medians with interquartile ranges.
+Effect size statistic and p values are indicated in the plot captions.
+Numbers of complete observations are displayed in the plot captions or in the plot X axes._
+
+\newpage
+
+```{r fig-interaction-total, fig.width = mm_inch(suppl_fig$interaction_total$w), fig.height = mm_inch(suppl_fig$interaction_total$h), fig.cap = 'Interactions of diastolic dysfunction, fatigue and fatigue intensity with cardiopulmonary findings follow-up affecting the total illness perception at the one-year follow-up.'}
+
+suppl_fig$interaction_total$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-interaction-total).
+Interactions of diastolic dysfunction, fatigue and fatigue intensity with cardiopulmonary findings follow-up affecting the total illness perception at the one-year follow-up.__
+
+_Interaction effects between the key factors affecting the total illness perception (IP) score and cardiopulmonary abnormalities (A: chest computed tomography [CT] abnormality, B: lung function testing [LFT] abnormality, C: diastolic dysfunction) at the one-year follow-up were investigated by two-way ANOVA with $\eta^2$ effect size statistic.
+The key factors with significant and near-significant interaction effects (p < 0.1) are presented.
+The associations with fatigue scoring are presented in a point plots; the black line represents the fitted linear trend and the tinted ribbon depicts the 95% confidence intervals.
+The total score values for the remaining explanatory variables are presented in violin plots with single observations depicted as points, and red diamonds and whiskers representing medians with interquartile ranges.
+Numbers of complete observations in the strata are displayed in the plot or in the plot facets.
+Effect size statistic and p values are shown under the plots.
+CFS: Chalder's fatigue score._
+
+\newpage
+
+```{r fig-interaction-sub1-ct, fig.width = mm_inch(suppl_fig$interaction_sub1_ct$w), fig.height = mm_inch(suppl_fig$interaction_sub1_ct$h), fig.cap = 'Interactions of diastolic dysfunction, fatigue and fatigue scoring with chest CT abnormality affecting the emotion/concern/consequences component of illness perception.'}
+
+suppl_fig$interaction_sub1_ct$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-interaction-sub1-ct).
+Interactions of diastolic dysfunction, fatigue and fatigue scoring with chest CT abnormality affecting the emotion/concern/consequences component of illness perception.__
+
+_Interaction effects between the key factors affecting the emotion/concern/consequences IP component score and chest computed tomography (CT) abnormalities at the one-year follow-up were investigated by two-way ANOVA with $\eta^2$ effect size statistic.
+The key factors with significant and near-significant interaction effects (p < 0.1) are presented.
+The associations with fatigue scoring are presented in a point plots; the black line represents the fitted linear trend and the tinted ribbon depicts the 95% confidence intervals.
+The total score values for the remaining explanatory variables are presented in violin plots with single observations depicted as points, and red diamonds and whiskers representing medians with interquartile ranges.
+Numbers of complete observations in the strata are displayed in the plot or in the plot facets.
+Effect size statistic and p values are shown under the plots.
+CFS: Chalder's fatigue score._
+
+\newpage
+
+```{r fig-interaction-sub1-lufo-dia, fig.width = mm_inch(suppl_fig$interaction_sub1_lufo_dia$w), fig.height = mm_inch(suppl_fig$interaction_sub1_lufo_dia$h), fig.cap = 'Interactions of symptom number, fatigue and fatigue scoring with LFT abnormality and diastolic dysfunction affecting the emotion/concern/consequences component of illness perception.'}
+
+suppl_fig$interaction_sub1_lufo_dia$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-interaction-sub1-lufo-dia).
+Interactions of persistent somatic symptom number, fatigue and fatigue scoring with LFT abnormality and diastolic dysfunction affecting the emotion/concern/consequences component of illness perception.__
+
+_Interaction effects between the key factors affecting the emotion/concern/consequences IP component score and cardiopulmonary abnormalities (A: lung function testing [LFT] abnormality, B: diastolic dysfunction) at the one-year follow-up were investigated by two-way ANOVA with $\eta^2$ effect size statistic.
+The key factors with significant and near-significant interaction effects (p < 0.1) are presented.
+The associations with fatigue scoring are presented in a point plots; the black line represents the fitted linear trend and the tinted ribbon depicts the 95% confidence intervals.
+The total score values for the remaining explanatory variables are presented in violin plots with single observations depicted as points, and red diamonds and whiskers representing medians with interquartile ranges.
+Numbers of complete observations in the strata are displayed in the plot or in the plot facets.
+Effect size statistic and p values are shown under the plots.
+CFS: Chalder's fatigue score._
 
@@
 
-_(A) Comparison of clustering algorithm performance (PAM: partitioning around medoids, HCL: Ward D2 hierarchical clustering, KMEANS).
+_(A) Comparison of clustering algorithm performance (PAM, HCL: Ward D2 hierarchical clustering, KMEANS).
 Explained clustering variance was defined as a ratio of the between-cluster sum of squares to the total sum of squares, cluster assignment accuracy was investigated in 10-fold cross-validation (CV).

### 2023-01-09 analysis scripts/clust_characteristic.R

@@ -18,6 +18,16 @@
   
   plan('multisession')
   
+  ## n numbers to be presented in the X axes
+  
+  clust_chara$clust_numbers <- expl$variables %>% 
+    map(~map(.x, 
+             ~filter(clust_chara$analysis_tbl, 
+                     !is.na(.data[[.x]]))) %>% 
+          map(count, clust_id) %>% 
+          map(~map2_chr(.x[[1]], .x[[2]], 
+                   paste, sep = '\nn = ')))
+  
 # descriptive stats ------
   
   insert_msg('Descriptive stats')
@@ -68,7 +78,7 @@
     map(function(resp_type) list(variable = expl$variables[[resp_type]], 
                                  type = expl$plot_type[[resp_type]], 
                                  plot_title = expl$variables[[resp_type]] %>% 
-                                   translate_var(out_value = 'label_long'), 
+                                   translate_var(out_value = 'label'), 
                                  plot_subtitle = clust_chara$test_results[[resp_type]]$plot_cap, 
                                  y_lab = expl$variables[[resp_type]] %>% 
                                    translate_var(out_value = 'axis_lab')) %>% 
@@ -77,12 +87,13 @@
                split_factor = 'clust_id',
                scale = 'percent', 
                x_lab = 'IP cluster', 
+               point_hjitter = 0, 
                cust_theme = globals$common_theme) %>% 
-          map(~.x + 
-                labs(tag = .x$labels$tag %>% 
-                       stri_replace_all(fixed = '\n', replacement = ', ') %>% 
-                       paste0('\n', .)) + 
-                scale_fill_manual(values = globals$clust_colors)) %>% 
+          map2(clust_chara$clust_numbers[[resp_type]], 
+               ~.x + 
+                 scale_x_discrete(labels = .y) + 
+                 scale_fill_manual(values = globals$clust_colors) + 
+                 theme(plot.tag = element_blank())) %>% 
           set_names(expl$variables[[resp_type]]))
   
 # A summary heat map of the psych variables in the clusters ------
@@ -169,6 +180,111 @@
          subtitle = 'Kruskal-Wallis test, normalized scores', 
          x = 'Patricipant',
          tag = clust_chara$psy_plot_tag )
+  
+# A summary plot of the symptom frequency in the clusters -----
+  
+  insert_msg('Symptom frequency plot')
+  
+  ## symptom variables, ordered by their effect size of the difference
+  ## between the clusters
+  
[…]

### 2023-01-09 analysis scripts/interaction.R

@@ -1,0 +1,437 @@
+# Analyzing the interaction effects between the key factors
+# affecting the overall illness perception score and the 
+# emotion/concern/consequence sub-score with cardiopulmonary abnormality
+# 
+# Done with ANOVA/linear modeling with the interaction effect
+
+
+  insert_head()
+  
+# container -----
+  
+  inter <- list()
+  
+# Analysis globals ------
+  
+  insert_msg('Analysis globals')
+  
+  ## the key factors influencing the illness perception responses
+  
+  inter$variables <- an$cmm_variables[c('ipq_total',
+                                        'ipq_sub1')] %>% 
+    reduce(union)
+  
+  inter$variables <- set_names(inter$variables)
+
+  ## the analysis table
+  ## defining the cardiopulmonary abnormality variable
+  
+  inter$analysis_tbl <- cov_data$clear_data %>% 
+    mutate(cardiopulmo = ifelse(ct_severity_any == 'yes' | 
+                                  lufo_red == 'yes' |
+                                  diastolic_dysf == 'yes', 
+                                'yes', 'no')) %>% 
+    select(sqrt_ipq_total, 
+           sqrt_ipq_sub1, 
+           ipq_total, 
+           ipq_sub1, 
+           ct_severity_any, 
+           lufo_red, 
+           diastolic_dysf, 
+           all_of(unname(inter$variables)))
+  
+  ## variable type: numeric vs factor
+  
+  inter$var_types <- inter$analysis_tbl[inter$variables] %>% 
+    map_lgl(is.numeric)
+  
+  ## model formulas with the interaction term
+  
+  inter$formulas_ct$ipq_total <- inter$variables %>% 
+    map(~paste('sqrt_ipq_total ~', .x, '* ct_severity_any')) %>% 
+    map(as.formula)
+  
+  inter$formulas_ct$ipq_sub1 <- inter$variables %>% 
+    map(~paste('sqrt_ipq_sub1 ~', .x, '* ct_severity_any')) %>% 
+    map(as.formula)
+  
+  inter$formulas_lufo$ipq_total <- inter$variables %>% 
+    map(~paste('sqrt_ipq_total ~', .x, '* lufo_red')) %>% 
+    map(as.formula)
+  
+  inter$formulas_lufo$ipq_sub1 <- inter$variables %>% 
+    map(~paste('sqrt_ipq_sub1 ~', .x, '* lufo_red')) %>% 
+    map(as.formula)
+  
+  inter$formulas_dia$ipq_total <- inter$variables %>% 
[…]

### 2023-01-09 analysis scripts/ipq_factors.R

@@ -15,6 +15,15 @@
                                   an$conf_variables))) %>% 
     dlply(.(what))
   
+  ## n numbers to be presented in the X axis
+  
+  ipq_factors$n_numbers <- ipq_factors$variables$eff_size$variable %>% 
+    map(~count(mod$multi_tbl, .data[[.x]])) %>% 
+    set_names(ipq_factors$variables$eff_size$variable)
+  
+  ipq_factors$x_axes <- ipq_factors$n_numbers %>% 
+    map(~map2_chr(.x[[1]], .x[[2]], paste, sep = '\nn = '))
+  
 # Serial comparisons and correlations -----
   
   insert_msg('Serial comparisons')
@@ -43,6 +52,7 @@
              variables = 'ipq_total', 
              what = 'eff_size', 
              ci = FALSE, 
+             exact = FALSE, 
              pub_styled = TRUE) %>% 
     mutate(split_factor = ipq_factors$variables$eff_size$variable, 
            plot_cap = paste(eff_size, significance, sep = ', '))
@@ -75,16 +85,18 @@
                         split_factor = .x, 
                         variable = 'ipq_total', 
                         type = 'violin', 
-                        plot_title = translate_var(.x, out_value = 'label_long'), 
+                        plot_title = translate_var(.x, 
+                                                   out_value = 'label_long'), 
                         plot_subtitle = .y, 
                         y_lab = translate_var('ipq_total'), 
                         x_lab = translate_var(.x), 
+                        point_hjitter = 0, 
                         cust_theme = globals$common_theme)) %>% 
-    map(~.x + 
-          scale_fill_brewer() + 
-          labs(tag = .x$labels$tag %>% 
-                 stri_replace_all(fixed = '\n', replacement = ', ') %>% 
-                 paste0('\n', .))) %>% 
+    map2(ipq_factors$x_axes, 
+         ~.x + 
+           scale_fill_brewer() + 
+           scale_x_discrete(labels = .y) + 
+           theme(plot.tag = element_blank())) %>% 
     set_names(ipq_factors$factor_test$split_factor)
   
 # Plotting the correlations -----
@@ -98,17 +110,23 @@
                              filter(complete.cases(.)), 
                            variables = c(.x, 'ipq_total'), 
                            type = 'correlation',
-                           plot_title = translate_var(.x, out_value = 'label_long'), 
+                           plot_title = translate_var(.x, 
+                                                      out_value = 'label_long'), 
                            plot_subtitle = .y, 
[…]

### 2023-01-09 analysis scripts/ipq_sub1_factors.R

@@ -1,4 +1,5 @@
-# Relationship between common influential factors and sub1 value
+# Relationship between common influential factors and sub1 value 
+# (subscore for emotion/concern/consequences)
 
   insert_head()
 
@@ -14,6 +15,15 @@
     filter(variable %in% unique(c(an$cmm_variables$ipq_sub1, 
                                   an$conf_variables))) %>% 
     dlply(.(what))
+  
+  ## n numbers to be presented in the X axis
+  
+  sub1_factors$n_numbers <- sub1_factors$variables$eff_size$variable %>% 
+    map(~count(mod$multi_tbl, .data[[.x]])) %>% 
+    set_names(sub1_factors$variables$eff_size$variable)
+  
+  sub1_factors$x_axes <- sub1_factors$n_numbers %>% 
+    map(~map2_chr(.x[[1]], .x[[2]], paste, sep = '\nn = '))
   
 # Serial comparisons and correlations -----
   
@@ -43,6 +53,7 @@
              variables = 'ipq_sub1', 
              what = 'eff_size', 
              ci = FALSE, 
+             exact = FALSE, 
              pub_styled = TRUE) %>% 
     mutate(split_factor = sub1_factors$variables$eff_size$variable, 
            plot_cap = paste(eff_size, significance, sep = ', '))
@@ -75,16 +86,18 @@
                         split_factor = .x, 
                         variable = 'ipq_sub1', 
                         type = 'violin', 
-                        plot_title = translate_var(.x, out_value = 'label_long'), 
+                        plot_title = translate_var(.x, 
+                                                   out_value = 'label_long'), 
                         plot_subtitle = .y, 
                         y_lab = translate_var('ipq_sub1'), 
                         x_lab = translate_var(.x), 
+                        point_hjitter = 0, 
                         cust_theme = globals$common_theme)) %>% 
-    map(~.x + 
+    map2(sub1_factors$x_axes, 
+         ~.x + 
+          scale_x_discrete(labels = .y) + 
           scale_fill_brewer() + 
-          labs(tag = .x$labels$tag %>% 
-                 stri_replace_all(fixed = '\n', replacement = ', ') %>% 
-                 paste0('\n', .))) %>% 
+          theme(plot.tag = element_blank())) %>% 
     set_names(sub1_factors$factor_test$split_factor)
   
 # Plotting the correlations -----
@@ -103,12 +116,16 @@
                            x_lab = translate_var(.x, out_value = 'axis_lab'), 
[…]

### 2023-01-09 analysis scripts/ipq_sub2_factors.R

@@ -1,4 +1,7 @@
-# Relationship between common influential factors and sub2 value
+# Relationship between common influential factors and sub2 value 
+# (control/coherence)
+#
+# Note: there were no factor identified by modeling!
 
   insert_head()
 
@@ -14,6 +17,15 @@
     filter(variable %in% unique(c(an$cmm_variables$ipq_sub2, 
                                   an$conf_variables))) %>% 
     dlply(.(what))
+  
+  ## n numbers to be presented in the X axis
+  
+  sub2_factors$n_numbers <- sub2_factors$variables$eff_size$variable %>% 
+    map(~count(mod$multi_tbl, .data[[.x]])) %>% 
+    set_names(sub2_factors$variables$eff_size$variable)
+  
+  sub2_factors$x_axes <- sub2_factors$n_numbers %>% 
+    map(~map2_chr(.x[[1]], .x[[2]], paste, sep = '\nn = '))
   
 # Serial comparisons and correlations -----
   
@@ -43,6 +55,7 @@
              variables = 'ipq_sub2', 
              what = 'eff_size', 
              ci = FALSE, 
+             exact = FALSE, 
              pub_styled = TRUE) %>% 
     mutate(split_factor = sub2_factors$variables$eff_size$variable, 
            plot_cap = paste(eff_size, significance, sep = ', '))
@@ -79,12 +92,13 @@
                         plot_subtitle = .y, 
                         y_lab = translate_var('ipq_sub2'), 
                         x_lab = translate_var(.x), 
+                        point_hjitter = 0, 
                         cust_theme = globals$common_theme)) %>% 
-    map(~.x + 
+    map2(sub2_factors$x_axes, 
+        ~.x + 
+          scale_x_discrete(labels = .y) + 
           scale_fill_brewer() + 
-          labs(tag = .x$labels$tag %>% 
-                 stri_replace_all(fixed = '\n', replacement = ', ') %>% 
-                 paste0('\n', .))) %>% 
+          theme(plot.tag = element_blank())) %>% 
     set_names(sub2_factors$factor_test$split_factor)
   
 # Plotting the correlations -----
@@ -103,15 +117,18 @@
                            x_lab = translate_var(.x, out_value = 'axis_lab'), 
                            y_lab = translate_var('ipq_sub2', out_value = 'axis_lab'), 
                            show_trend = FALSE, 
+                           point_hjitter = 0, 
+                           point_wjitter = 0, 
[…]

### 2023-01-09 analysis scripts/plots.R

@@ -1,4 +1,4 @@
-# Venn plots with the common influnetial factors and plots with the 
+# Venn plots with the common influential factors and plots with the 
 # response-wise performance measures
 
   insert_head()
@@ -41,22 +41,25 @@
   
   insert_msg('Numbers of selected factors')
   
+  ## working with safely, since there are likely no factors found
+  ## for the control/coherence subscore
+  
   mod_plots$venn_plots <- 
     list(data = map(an$variables, 
                     ~set_names(.x, an$methods[names(.x)])), 
          plot_title = an$responses %>% 
-           translate_var %>% 
-           paste('1-year follow-up', sep = ', ')) %>% 
-    pmap(plot_n_venn, 
-         plot_subtitle = 'variables with non-zero coefficients', 
-         plot_tag = paste0('variables: n = ', 
-                           length(mod$variables), 
-                           ', observations: n = ', 
-                           nrow(mod$multi_tbl)), 
+           translate_var) %>% 
+    pmap(safely(plot_n_venn),  
          fill_color = set_names(an$method_colors, 
                                 an$methods), 
-         fontScale = 3, 
-         panel = FALSE)
+         subset_names = an$methods[names(an$variables$ipq_total)], 
+         fontScale = 4, 
+         panel = FALSE, 
+         legend_position = 'bottom', 
+         comp_heights = c(0.08, 0, 0.92, 0))
+  
+  mod_plots$venn_plots <- mod_plots$venn_plots %>% 
+    map(~.x$result)
 
 # END -----
   

### 2023-01-09 analysis.R

@@ -1,6 +1,18 @@
-# Detailed analysis of SES, SOCL-9 and KW IPQ
-# as a function of common influential paramaters identified by 
-# backwards elimination, Elastic Net and Bayesian Lasso modeling.
+# Detailed analysis of factor influencing illness perception
+# identified by Elastic Net, LASSO and Bayesian Lasso modeling.
+#
+# 1) Fit stats (training and CV) are compared for each of the algorithms
+# and factors influencing the total illness perceptions score and its sub-scores
+# are identified as overlapping non-zero coefficients computed by all 
+# three modeling algorithms. Note: no factor influencing the control/coherence
+# component could be discerned (COVID-19-independent effects?)
+#
+# 2) Detailed analysis of the influential factors by correlation and statistical 
+# hypothesis testing and analysis of their interaction with cardiopulmonary
+# abnormalities (two-way ANOVA)
+#
+# 3) Unsupervised clustering of the study participants in respect to 
+
 
 # tools ------
 
@@ -34,7 +46,7 @@
     map(~map(.x, filter, parameter != '(Intercept)')) %>% 
     map(~map(.x, ~.x$variable) %>% 
           map(stri_replace, fixed = '_sq', replacement = '') %>% 
-          map(stri_replace, fixed = 'ctss', replacement = 'ct_severity_score') %>% 
+          map(stri_replace, regex = '(log_)|(sqrt_)', replacement = '') %>% 
           map(unique))
   
   ## common variables selected by at least two penalized modeling methods
@@ -93,6 +105,7 @@
     './analysis scripts/ipq_sub1_factors.R',
     './analysis scripts/ipq_sub2_factors.R', 
     './analysis scripts/ipq_factors.R', 
+    './analysis scripts/interaction.R', 
     './analysis scripts/ipq_clust_development.R', 
     './analysis scripts/ipq_clustering.R', 
     './analysis scripts/clust_characteristic.R') %>% 

### 2023-01-09 exec.R

@@ -9,5 +9,3 @@
                      'paper.R'), 
                    crash = FALSE, 
                    message = TRUE))
-  
-  save.image()

### 2023-01-09 exploration scripts/bierbauer2022.R

@@ -1,0 +1,329 @@
+# Dimensionality analysis (factor analysis) an reliability of the IPQ tool 
+# in Long COVID individuals from the study of Bierbauer et al. 
+# (https://osf.io/bwvxg)
+
+  insert_head()
+  
+# container -------
+  
+  lc_ipq <- list()
+  
+# Analysis globals ------
+  
+  insert_msg('Analysis globals')
+  
+  ## analysis table with the 'larger' IPQ dimensions
+  ## inverting the personal control, treatment control and coherence dimensions
+  ## (lacking personal control, lacking treatment control, lacking coherence)
+  ## scaling the identity (37, the maximal symptom number is 5, 0 is 0)
+  
+  lc_ipq$variables <- 
+    c('IPQ consequences', 
+      'IPQ duration', 
+      'IPQ cyclical', 
+      'IPQ lacking personal control', 
+      'IPQ lacking treatment control', 
+      'IPQ identity/# symptoms', 
+      'IPQ lacking coherence', 
+      'IPQ emotional representation')
+  
+  lc_ipq$analysis_tbl <- 
+    read_csv('./data/data_illnessperception_bierbauer2022.csv') %>% 
+    mutate(participant_id = paste0('id_', 1:nrow(.)), 
+           lack_personalcontrol = 5 - personalcontrol, 
+           lack_treatmentcontrol = 5 - treatmentcontrol, 
+           lack_coherence = 5 - coherence) %>% 
+    select(participant_id, 
+           consequence, 
+           duration, 
+           cyclical,
+           lack_personalcontrol, 
+           lack_treatmentcontrol, 
+           identity, 
+           lack_coherence, 
+           emotionalrep) %>% 
+    filter(complete.cases(.)) %>% 
+    mutate(identity = identity * 5/37) %>% 
+    set_names(c('participant_id', 
+                lc_ipq$variables)) %>% 
+    column_to_rownames('participant_id')
+  
+# Median-centered PCA to identify the factor number ------
+  
+  insert_msg('PCA')
+  
+  ## PCA object
+  
+  lc_ipq$pca_obj <- reduce_data(center_data(lc_ipq$analysis_tbl), 
+                                kdim = 6, 
+                                red_fun = 'pca')
+  
+  ## scree plot: 3 - 4 factors seem to be appropriate
+  
+  lc_ipq$pca_scree_plot <- plot(lc_ipq$pca_obj, 
+                                type = 'scree', 
+                                cust_theme = globals$common_theme) +  
[…]

### 2023-01-09 exploration scripts/cohort_characteristic.R

@@ -1,5 +1,11 @@
 # Baseline clinical characteristic of the study cohort and recovery status 
-# at the one-year follow-up
+# at the one-year follow-up.
+#
+# Descriptive statistics for the entire cohort and COVID-19 severity strata
+# are presented. Differences between the cohorts are investigated with 
+# non parametric Chi-squared and Kruskal-Wallis test (multiple numeric 
+# variables are non-normally distributed).
+
 
   insert_head()
   
@@ -53,28 +59,91 @@
     map(mutate, 
         plot_cap = paste(eff_size, significance, sep = ', '))
   
-# Plots of the psych variables in the severity groups ------
+# Plots of variables in the severity groups ------
   
-  insert_msg('Plots of the psych variable values in the severity groups')
+  insert_msg('Plots of variable values in the severity groups')
   
-  cohort$psy_plots <- list(variable = expl$variables$psych, 
-                           plot_title = expl$variables$psych %>% 
-                             translate_var(out_value = 'label_long'), 
-                           plot_subtitle = cohort$test_results$psych$plot_cap, 
-                           y_lab = expl$variables$psych %>% 
-                             translate_var(out_value = 'axis_lab')) %>% 
-    pmap(plot_variable, 
-         cov_data$clear_data, 
-         split_factor = 'cat_WHO', 
-         type = 'violin', 
-         x_lab = 'COVID-19 severity', 
-         cust_theme = globals$common_theme) %>% 
-    map(~.x + 
-          labs(tag = .x$labels$tag %>% 
-                 stri_replace_all(fixed = '\n', replacement = ', ') %>% 
-                 paste0('\n', .)) + 
-          scale_fill_manual(values = globals$sev_colors[1:3])) %>% 
-    set_names(expl$variables$psych)
+  cohort[paste0(names(expl$variables), '_plots')] <- names(expl$variables) %>% 
+    map(function(var_class) list(variable = expl$variables[[var_class]], 
+                                 plot_title = expl$variables[[var_class]] %>% 
+                                   translate_var(out_value = 'label_long'), 
+                                 plot_subtitle = cohort$test_results[[var_class]]$plot_cap, 
+                                 y_lab = expl$variables[[var_class]] %>% 
+                                   translate_var(out_value = 'axis_lab'), 
+                                 type = expl$plot_type[[var_class]]) %>% 
+          pmap(plot_variable, 
+               cov_data$clear_data, 
[…]

### 2023-01-09 exploration scripts/distribution.R

@@ -1,4 +1,6 @@
-# Distribution tests and plots for numeric variables ------
+# Distribution tests and plots for numeric variables
+# Normality (entire cohort) is investigated by Shapiro-Wilk test
+# Variance homogeneity in the CoV severity strata is assessed by Levene test
 
   insert_head()
 
@@ -26,7 +28,7 @@
   distr_tests$norm_test <- explore(cov_data$clear_data, 
                                    variables = distr_tests$num_vars, 
                                    what = 'normality', 
-                                   pub_styled = TRUE)
+                                   pub_styled = FALSE)
   
   ## QQ plots
   
@@ -54,7 +56,7 @@
               map_dfc(function(x) if(is.numeric(x)) log(x + 1) else x), 
             variables = distr_tests$num_vars, 
             what = 'normality', 
-            pub_styled = TRUE)
+            pub_styled = FALSE)
   
   ## QQ plots
   
@@ -84,7 +86,7 @@
               map_dfc(function(x) if(is.numeric(x)) sqrt(x) else x), 
             variables = distr_tests$num_vars, 
             what = 'normality', 
-            pub_styled = TRUE)
+            pub_styled = FALSE)
   
   ## QQ plots
   
@@ -133,7 +135,8 @@
     map(~cov_data$clear_data[.x, ])
   
   distr_tests$mvr_boot <- distr_tests$boot_tbl %>% 
-    future_map(~get_mvr(.x, variables = distr_tests$num_vars)[c('variable', 'mvr')]) %>% 
+    future_map(~get_mvr(.x, 
+                        variables = distr_tests$num_vars)[c('variable', 'mvr')]) %>% 
     map(column_to_rownames, 'variable') %>% 
     reduce(cbind) %>% 
     as.matrix
@@ -173,6 +176,30 @@
     labs(title = 'Variable mean and variance', 
          subtitle = '95%CI: bootstrap, n = 1000, percentile method', 
          x = 'mean MVR \u00B1 95%CI')
+  
+# A table with the optimal transformations of numeric variables ------
+  
+  insert_msg('Optimal transformation table')
+  
+  distr_tests$optimal_trans <- distr_tests[c('norm_test', 
+                                             'norm_test_log', 
+                                             'norm_test_sqrt')] %>% 
+    map2_dfr(., c('identity', 'log', 'sqrt'), 
+             ~mutate(.x, 
+                     transformation = .y, 
+                     base_variable = variable, 
+                     variable = ifelse(transformation == 'identity', 
+                                       base_variable, 
[…]

### 2023-01-09 exploration scripts/gender.R

@@ -1,0 +1,93 @@
+# Comparison of baseline characteristics, recovery and mental health feattures 
+# between females and males.
+#
+# Participants meeting the inclusion criteria are analyzed. Non-parameteric 
+# testing since multiple numeric variables are non-normally distributed
+
+  insert_head()
+  
+# container ------
+  
+  gender <- list()
+  
+# analysis globals -----
+  
+  insert_msg('Analysis globals')
+  
+  ## test and plot types
+  
+  gender$eff_type <- expl$variables %>% 
+    map(~map_chr(.x, 
+                 ~ifelse(is.numeric(cov_data$clear_data[[.x]]), 
+                         'wilcoxon_r', 'cramer_v')))
+  
+  gender$plot_type <- expl$variables %>% 
+    map(~map_chr(.x, 
+                 ~ifelse(is.numeric(cov_data$clear_data[[.x]]), 
+                         'violin', 'stack')))
+  
+  ## parallel backend
+  
+  plan('multisession')
+  
+# descriptive stats -------
+  
+  insert_msg('Descriptive statistics')
+  
+  gender$desc_stats <- expl$variables %>% 
+    map(~explore(cov_data$clear_data, 
+                 split_factor = 'sex', 
+                 variables = .x, 
+                 what = 'table', 
+                 pub_styled = TRUE)) %>% 
+    map(reduce, 
+        left_join, 
+        by = 'variable') %>% 
+    map(set_names, 
+        c('variable', levels(cov_data$clear_data$sex)))
+  
+# Testing for differences between the genders -------
+  
+  insert_msg('Testing for differences between the genders')
+  
+  gender$test_results <- list(variables = expl$variables, 
+                              types = gender$eff_type) %>% 
+    future_pmap(compare_variables,
+                cov_data$clear_data, 
+                split_factor = 'sex', 
+                what = 'eff_size', 
+                ci = FALSE, 
+                exact = FALSE, 
+                pub_styled = TRUE, 
+                adj_method = 'none', 
+                .options = furrr_options(seed = TRUE)) %>% 
+    map(mutate, 
+        plot_cap = paste(eff_size, significance, sep = ', '))
+  
+# Plotting -------
+  
+  insert_msg('Plotting')
+  
+  gender[paste0(names(expl$variables), '_plots')] <- names(expl$variables) %>% 
+    map(function(var_class) list(variable = expl$variables[[var_class]], 
+                                 plot_title = expl$variables[[var_class]] %>% 
[…]

### 2023-01-09 exploration scripts/included_excluded.R

@@ -1,0 +1,109 @@
+# Comparison of the CovILD subset included in the analysis with the participants
+# excluded from the analysis because of variable or BIPQ missingness or lacking
+# COVID-19 sequelae at the one-year follow-up.
+#
+# Done with non-parametric tests since multiple numeric features are 
+# non-normally distributed.
+
+  insert_head()
+  
+# container ------
+  
+  inc_bias <- list()
+  
+# analysis globals -------
+  
+  insert_msg('Analysis globals')
+  
+  ## analysis table
+  
+  inc_bias$analysis_tbl <- cov_data$raw_data %>% 
+    mutate(analysis_status = ifelse(ID %in% cov_data$clear_data$ID, 
+                                    'included', 'excluded'), 
+           analysis_status = factor(analysis_status, 
+                                    c('included', 'excluded')))
+  
+  ## analysis variables
+  
+  inc_bias$variables <- expl$variables[c('baseline', 'fup', 'lab')] %>% 
+    map(~.x[.x %in% names(inc_bias$analysis_tbl)])
+  
+  inc_bias$eff_type <- inc_bias$variables %>% 
+    map(~map_chr(.x, 
+             ~ifelse(is.numeric(inc_bias$analysis_tbl[[.x]]), 
+                'wilcoxon_r', 'cramer_v')))
+  
+  inc_bias$plot_type <- inc_bias$variables %>% 
+    map(~map_chr(.x, 
+                 ~ifelse(is.numeric(inc_bias$analysis_tbl[[.x]]), 
+                         'violin', 'stack')))
+  
+  ## parallel backend
+  
+  plan('multisession')
+  
+# Descriptive stats ------
+  
+  insert_msg('Descriptive stats')
+  
+  inc_bias$desc_stats <- inc_bias$variables %>% 
+    map(~explore(inc_bias$analysis_tbl, 
+                 split_factor = 'analysis_status', 
+                 variables = .x, 
+                 what = 'table', 
+                 pub_styled = TRUE)) %>% 
+    map(reduce, 
+        left_join, 
+        by = 'variable') %>% 
+    map(set_names, 
+        c('variable', levels(inc_bias$analysis_tbl$analysis_status)))
+  
+# Testing for differences between the analysis inclusion strata ------
+  
+  insert_msg('Testing for differences')
+  
+  inc_bias$test_results <- list(variables = inc_bias$variables, 
+                                types = inc_bias$eff_type) %>% 
[…]

### 2023-01-09 exploration scripts/omega_excluded.R

@@ -1,0 +1,93 @@
+# Comparison of the internal consistency of BIPQ in the entire available 
+# dataset and in the patients included in the analysis
+
+  insert_head()
+  
+# container ------
+  
+  inc_omega <- list()
+  
+# analysis globals ------
+  
+  insert_msg('Analysis globals')
+  
+  ## reading all available data
+  ## inverting the negative terms
+  
+  inc_omega$raw_tbl <- read_excel('./data/ipq_sub.xlsx') %>% 
+    select(- Folder) %>% 
+    set_names(c('ID', paste0('ipq_q', 1:8)))
+  
+  for(i in paste0('ipq_q', 1:8)) {
+    
+    inc_omega$raw_tbl <- inc_omega$raw_tbl %>% 
+      mutate(!!i := as.numeric(.data[[i]]))
+    
+  }
+  
+  for(i in c('ipq_q3', 'ipq_q4', 'ipq_q7')) {
+    
+    inc_omega$raw_tbl <- inc_omega$raw_tbl %>% 
+      mutate(!!i := abs(10 - .data[[i]]))
+    
+  }
+  
+  inc_omega$analysis_tbl[c('all', 'included')] <- 
+    list(inc_omega$raw_tbl, 
+         filter(inc_omega$raw_tbl, ID %in% cov_data$clear_data$ID)) %>% 
+    map(~filter(.x, complete.cases(.x)))
+  
+# calculation of omegas ------
+  
+  insert_msg('Omega calculation')
+  
+  inc_omega$omega_obj <- inc_omega$analysis_tbl %>% 
+    map(select, starts_with('ipq_')) %>% 
+    map(psych::omega, 
+        nfactors = 2, 
+        fm = 'ml', 
+        title = 'BIPQ', 
+        option = 'equal', 
+        niter = 1000)
+  
+# Plotting the omegas -------
+  
+  insert_msg('Plotting the omega values')
+  
+  ## plotting data
+  
+  inc_omega$omega_summary <- inc_omega$omega_obj %>% 
+    map(~.x$omega.group) %>% 
+    map(mutate, 
+        factor = c('total IP score', 
+                   'emotion/concern/consequences', 
+                   'control/coherence')) %>% 
+    map2_dfr(., names(.), 
+             ~mutate(.x, subset = .y)) %>% 
+    as_tibble
+  
+  ## plot
+  
+  inc_omega$omega_plot <- inc_omega$omega_summary %>% 
+    ggplot(aes(x = total, 
[…]

### 2023-01-09 exploration scripts/overlap.R

@@ -1,0 +1,115 @@
+# Overlap between the cardiopulmonary findings and persistent symptoms
+# at the one-year follow-up
+
+  insert_head()
+  
+# container -------
+  
+  fup_overlap <- list()
+  
+# globals ---- 
+  
+  insert_msg('Analysis globals')
+  
+  fup_overlap$analysis_sets <- 
+    cov_data$clear_data[c('sympt_present', 
+                          'ct_severity_any', 
+                          'lufo_red', 
+                          'diastolic_dysf')] %>% 
+    map_dfc(~.x == 'yes') %>% 
+    set_names(c('PSS', 
+                'CT findings', 
+                'LFT abnormality', 
+                'TTE abnormality'))
+  
+  fup_overlap$pairs <- combn(c('PSS', 
+                               'CT findings', 
+                               'LFT abnormality', 
+                               'TTE abnormality'), 
+                             m = 2, 
+                             simplify = FALSE)
+  
+# Calculating the intersection sizes ------
+  
+  insert_msg('Intersection sizes')
+  
+  ## pairwise intersections
+  
+  fup_overlap$intersections <- fup_overlap$pairs %>% 
+    set_names(map_chr(fup_overlap$pairs, paste, collapse = '_')) %>% 
+    map(~fup_overlap$analysis_sets[.x]) %>% 
+    map(reduce, `*`) %>% 
+    map_dbl(sum) %>% 
+    map2_dfr(., names(.), 
+             ~tibble(pair = .y, 
+                     count = .x)) %>% 
+    mutate(variable1 = stri_split_fixed(pair, 
+                                        pattern = '_', 
+                                        simplify = TRUE)[, 1], 
+           variable2 = stri_split_fixed(pair, 
+                                        pattern = '_', 
+                                        simplify = TRUE)[, 2])
+  
+  ## the intersection between all sequelae
+  
+  fup_overlap$all_intersection <-  fup_overlap$analysis_sets %>% 
+    reduce(`*`) %>% 
+    sum
+  
+  ## common table
+  
+  fup_overlap$intersections <- 
+    rbind(fup_overlap$intersections, 
+          tibble(pair = 'all', 
+                 count = fup_overlap$all_intersection, 
+                 variable1 = NA, 
+                 variable2 = NA)) %>% 
+    mutate(percent = count/nrow(fup_overlap$analysis_sets) * 100)
+
+# Venn plot -------
+  
+  insert_msg('Venn plot')
+  
[…]

### 2023-01-09 exploration scripts/psy_correlation.R

@@ -1,4 +1,8 @@
-# Correlation of the IPQ items, sum and subscores with other psy variables
+# Correlation of the IPQ items, sum and subscores with 
+# other mental health readouts variables
+#
+# Due to non-normal distribution of multiple mental health variables
+# Spearman correlation is employed
 
   insert_head()
   
@@ -11,7 +15,7 @@
   insert_msg('Globals: variables')
   
   psy_cor$var1 <- 
-    expl$variables$ipq[!expl$variables$ipq %in% c('ipq_sub1', 'ipq_sub2')]
+    expl$variables$ipq
   
   psy_cor$var2 <- globals$psy_variables
   

### 2023-01-09 exploration scripts/respiratory.R

@@ -1,0 +1,89 @@
+# Characteristic of the patients with and without respiratory comorbidity
+
+  insert_head()
+
+# container ------
+
+  respi <- list()
+
+# analysis globals -----
+
+  insert_msg('Analysis globals')
+
+  ## test and plot types
+
+  respi$eff_type <- expl$variables %>% 
+    map(~map_chr(.x, 
+                 ~ifelse(is.numeric(cov_data$clear_data[[.x]]), 
+                         'wilcoxon_r', 'cramer_v')))
+
+  respi$plot_type <- expl$variables %>% 
+    map(~map_chr(.x, 
+                 ~ifelse(is.numeric(cov_data$clear_data[[.x]]), 
+                         'violin', 'stack')))
+
+  ## parallel backend
+
+  plan('multisession')
+  
+# Descriptive stats ------
+  
+  insert_msg('Descriptive stats')
+  
+  respi$desc_stats <- expl$variables %>% 
+    map(~explore(cov_data$clear_data, 
+                 split_factor = 'respi_comorb', 
+                 variables = .x, 
+                 what = 'table', 
+                 pub_styled = TRUE)) %>% 
+    map(reduce, 
+        left_join, 
+        by = 'variable') %>% 
+    map(set_names, 
+        c('variable', levels(cov_data$clear_data$respi_comorb)))
+  
+# Checking for significance of the differences ------
+  
+  insert_msg('Checking for differences between the strata')
+  
+  respi$test_results <- list(variables = expl$variables, 
+                             types = respi$eff_type) %>% 
+    future_pmap(compare_variables,
+                cov_data$clear_data, 
+                split_factor = 'respi_comorb', 
+                what = 'eff_size', 
+                ci = FALSE, 
+                exact = FALSE, 
+                pub_styled = TRUE, 
+                adj_method = 'none', 
+                .options = furrr_options(seed = TRUE)) %>% 
+    map(mutate, 
+        plot_cap = paste(eff_size, significance, sep = ', '))
+  
+# Plotting -------
+  
+  insert_msg('Plotting')
+  
+  respi[paste0(names(expl$variables), '_plots')] <- names(expl$variables) %>% 
+    map(function(var_class) list(variable = expl$variables[[var_class]], 
+                                 plot_title = expl$variables[[var_class]] %>% 
+                                   translate_var(out_value = 'label_long'), 
+                                 plot_subtitle = respi$test_results[[var_class]]$plot_cap, 
[…]

### 2023-01-09 exploration scripts/score_coherence.R

@@ -1,4 +1,7 @@
-# IPQ score coherence with Cronbach's alpha
+# IPQ score coherence with MdDonald's omega.
+#
+# The choice of the consistency statistic was motivated by multi-dimensionality 
+# of the BIPQ tool with two independent components identified by factor analysis
 
   insert_head()
   
@@ -69,7 +72,7 @@
                                                           geom_smooth(method = 'gam')))
   
   
-# calculation of the alphas -----
+# calculation of the Cronbach's alpha -----
   
   insert_msg('Alpha calculation')
   
@@ -91,6 +94,18 @@
     map2_dfr(., names(.), ~mutate(.x, response = .y)) %>% 
     as_tibble
   
+# Calculation of the omegas -------
+  
+  insert_msg('Calculation of the omegas')
+  
+  ipq_co$omega_obj <- 
+    psych::omega(ipq_co$analysis_tbl$ipq_total, 
+                 nfactors = 2, 
+                 fm = 'ml', 
+                 title = 'BIPQ', 
+                 option = 'equal', 
+                 niter = 1000)
+
 # visualization of the correlation matrix ------
   
   insert_msg('Plotting the correlation matrix')
@@ -151,17 +166,44 @@
                                             vjust = 0.5, 
                                             angle = -90)))
   
-  ## appending the plots with titles n numbers and Cronbach's alpha
+  ## appending the plots with titles n numbers and omegas
   
   ipq_co$corr_mtx_plots <- 
     list(x = ipq_co$corr_mtx_plots, 
          y = translate_var(names(ipq_co$corr_mtx_plots)), 
-         z = paste0(ipq_co$alpha_stats$plot_cap, 
+         z = paste0('\u03C9 = ', 
+                    signif(ipq_co$omega_obj$omega.group$total, 2), 
                     ', n = ', 
                     map_dbl(ipq_co$analysis_tbl, nrow))) %>% 
     pmap(function(x, y, z) x + 
            labs(title = y, 
                 subtitle = z))
+  
+# A common summary plot with the faceting by the BIPQ components ------
+  
+  insert_msg('A common bubble plot')
+  
+  ipq_co$summ_omega_lab <- 
+    map2_chr(c('total (Q1 - Q8)', 
+               'emotion/concern/consequences (Q1/2/3/6/8)', 
+               'lacking control/coherence (Q3/Q4/Q7)'), 
[…]

### 2023-01-09 exploration scripts/score_fa.R

@@ -1,0 +1,63 @@
+# Factor analysis of the BIPQ items. No pre-processing.
+# The number of factors (k = 3) was chosen based on the PCA results 
+# (see the scree plot)
+
+  insert_head()
+  
+# container -----
+  
+  ipq_fa <- list()
+  
+# globals -----
+  
+  insert_msg('Globals: analysis table')
+  
+  ipq_fa$analysis_tbl <- cov_data$clear_data %>% 
+    select(ID, starts_with('ipq_q')) %>% 
+    column_to_rownames('ID') %>% 
+    set_names(globals$ipq_sublabs[names(.)])
+  
+# FA: 3 factor ------
+  
+  insert_msg('3D factor analysis')
+  
+  ## FA object
+  
+  ipq_fa$fa_obj <- reduce_data(ipq_fa$analysis_tbl, 
+                               kdim = 3, 
+                               red_fun = 'fa', 
+                               scores = 'regression')
+  
+# Score and loadings plots --------
+  
+  insert_msg('Plotting the score and loadings plots')
+  
+  ## FA regression scores
+  
+  ipq_fa$score_plot <- plot(ipq_fa$fa_obj, 
+                            type = 'scores', 
+                            cust_theme = globals$common_theme) + 
+    geom_hline(yintercept = 0, 
+               linetype = 'dashed') + 
+    geom_vline(xintercept = 0, 
+               linetype = 'dashed') + 
+    labs(title = 'FA regression scores', 
+         subtitle = '3D factor analysis', 
+         tag = paste('\nn =', nrow(ipq_fa$analysis_tbl)))
+  
+  ## loadings
+  
+  ipq_fa$loadings_plot <- plot(ipq_fa$fa_obj, 
+                                type = 'loadings', 
+                                cust_theme = globals$common_theme) + 
+    geom_hline(yintercept = 0, 
+               linetype = 'dashed') + 
+    geom_vline(xintercept = 0, 
+               linetype = 'dashed') + 
+    labs(title = 'FA loadings', 
+         subtitle = '3D factor analysis', 
+         tag = paste('\nn =', nrow(ipq_fa$analysis_tbl)))
+  
+# END ------
+  
+  insert_tail()

### 2023-01-09 exploration scripts/score_pca.R

@@ -1,4 +1,4 @@
-# Principal component analysis of the IPQ sub-scores. Median centering
+# Principal component analysis of the BIPQ items. Median centering
 
   insert_head()
   
@@ -95,7 +95,3 @@
 # END ------
   
   insert_tail()
-
-  
-  
-  

### 2023-01-09 exploration scripts/symptom_kinetic.R

@@ -1,0 +1,193 @@
+# Kinetic of the persistent symptoms in the study collective
+
+  insert_head()
+  
+# container -----
+  
+  sympt_kinet <- list()
+  
+# Analysis globals -------
+  
+  insert_msg('Analysis globals')
+  
+  ## variables
+  
+  sympt_kinet$variables <- 
+    names(cov_data$sympt_data)[stri_detect(names(cov_data$sympt_data), 
+                                           fixed = 'sympt')]
+  
+  sympt_kinet$variables <- set_names(sympt_kinet$variables, 
+                                     sympt_kinet$variables)
+
+  ## analysis table with longitudinal symptom data for 
+  ## all and the symptomatic individuals 
+  ## specific symptom present at the one-year FUP)
+  
+  sympt_kinet$all_tbl <- cov_data$sympt_data %>% 
+    mutate(time = car::recode(time, 
+                              "1 = '2'; 2 = '3'; 3 = '6'; 4 = '12'"), 
+           time = as.numeric(time))
+  
+  sympt_kinet$all_tbl <- sympt_kinet$variables %>% 
+    map(~sympt_kinet$all_tbl[c('ID', 'time', .x)])
+  
+  sympt_kinet$symptomatic_id <- sympt_kinet$all_tbl %>% 
+    map2(., names(.), 
+         ~filter(.x, 
+                 time == 12, 
+                 .data[[.y]] == 'yes')) %>% 
+    map(~.x$ID)
+  
+  sympt_kinet$symp_tbl <- 
+    map2(sympt_kinet$all_tbl, 
+         sympt_kinet$symptomatic_id, 
+         ~filter(.x, ID %in% .y))
+  
+# Descriptive stats --------
+  
+  insert_msg('Descriptive stats')
+  
+  ## frequencies of any symptoms and specific symptoms 
+  ## at the consecutive visits
+  
+  sympt_kinet$desc_stats$all <- sympt_kinet$all_tbl %>% 
+    map2(., names(.), 
+         ~count(group_by(.x, time), 
+               .data[[.y]])) %>% 
+    map(mutate, 
+        percent = n/sum(n) * 100, 
+        n_total = sum(n)) %>% 
+    map(ungroup) %>% 
+    map(set_names, c('time', 'strata', 'n', 'percent', 'n_total')) %>% 
+    map2_dfr(., names(.), 
+             ~mutate(.x, symptom = .y)) %>% 
[…]

### 2023-01-09 exploration.R

@@ -1,5 +1,41 @@
 # Exploratory data analysis
-# Remova of participants with > 25% missing answers if any detected
+# Removal of participants with > 25% missing answers if any detected 
+# (done during data import, see: 'import.R')
+#
+# 1) Characteristic of the study collective, i.e. participants with long-term 
+# sequelae of COVID-19 
+# (symptoms or cardiopulmonary findings at the one-year FUP).
+#
+# 2) Comparison of the individuals included in the analysis and excluded due to
+# variable missingness, missing BIPQ and lack of long term-sequelae.
+#
+# 3) Comparison of baseline, recovery and mental health features between 
+# the genders done for the participants fulfilling the inclusion criteria
+#
+# 4) Analysis of distribution (normality/homogeneity) of the modeling responses
+# and explanatory variables
+#
+# 5) Correlation of illness perception score, subscores and particular items
+# with other readouts of mental health (stress, anxiety, depression, 
+# somatization and others)
+#
+# 6) Dimensionality of the BIPQ tool investigated by median centered PCA 
+# (to get some idea on number of dimensions) and factor analysis
+#
+# 7) Analysis of the internal consistency of the BIPQ tool with McDonald omega
+#
+# 8) analysis of overlap between the cardiopulmonary findings and symptoms at 
+# the one year follow-up
+#
+# 9) Characteristic of patients with and without respiratory comorbidity 
+#
+# 10) Comparison of the BIPQ consistency in all patients and the patients 
+# included in the analysis with McDonald omega, as describe above
+#
+# 11) Kinetics of symptoms in participants included in the analysis
+#
+# 12) Dimensionality and McDonald's omega for the illness perception dataset
+# by Bierbauer et al. 2022 (DOI: 10.1080/23311908.2022.2105007)
 
 # tools -------
 
@@ -11,6 +47,9 @@
   library(furrr)
   library(psych)
   library(clustTools)
+  library(ggvenn)
+  library(kinet)
+  library(rstatix)
 
   insert_head()
   
@@ -101,13 +140,21 @@
   
   insert_msg('Analysis scripts')
   
-  c('./exploration scripts/distribution.R', 
-    './exploration scripts/score_pca.R', 
-    './exploration scripts/score_coherence.R', 
+  c('./exploration scripts/cohort_characteristic.R', 
+    './exploration scripts/included_excluded.R', 
+    './exploration scripts/gender.R', 
[…]

### 2023-01-09 import.R

@@ -34,7 +34,7 @@
                          'weight_class', 
                          'comorb_present', 
                          'no_comorb', 
-                         'no_comorb_sq', 
+                         'no_comorb_class', 
                          'cardiovascular_comorb', 
                          'hypertension_comorb', 
                          'endometabolic_comorb', 
@@ -47,10 +47,9 @@
                          'immdef_comorb', 
                          'cat_WHO', 
                          'WHO', 
-                         'WHO_sq', 
                          'sympt_present', 
                          'sympt_number', 
-                         'sympt_number_sq', 
+                         'sympt_number_class', 
                          'fatigue_sympt', 
                          'dyspnoe_sympt', 
                          'cough_sympt', 
@@ -70,19 +69,31 @@
                          'lufo_red', 
                          'ct_severity_any', 
                          'ct_severity_score',
-                         'ctss_sq', 
+                         'ctss_class', 
                          'diastolic_dysf', 
                          'rehabilitation', 
                          'anemia', 
                          'Hb', 'Hb_sq', 
-                         'ferritin', 'ferritin_sq', 
-                         'TSAT', 'TSAT_sq', 
-                         'sTFR', 'sTFR_sq', 
-                         'Hepcidin', 'Hepcidin_sq', 
-                         'NTproBNP', 'NTproBNP_sq', 
-                         'DDimer', 'DDimer_sq', 
-                         'CRP', 'CRP_sq', 
-                         'HbA1c', 'HbA1c_sq')
+                         'ferritin', 
+                         'log_ferritin', 
+                         'log_ferritin_sq', 
+                         'TSAT', 
+                         'sqrt_TSAT', 
+                         'sqrt_TSAT_sq', 
+                         'sTFR', 
+                         'log_sTFR', 
+                         'log_sTFR_sq', 
+                         'Hepcidin', 
+                         'sqrt_Hepcidin', 
+                         'sqrt_Hepcidin_sq', 
+                         'NTproBNP', 
+                         'NTproBNP_class', 
[…]

### 2023-01-09 issues.R

@@ -1,0 +1,84 @@
+# IPQ score issues
+
+  library(tidyverse)
+  library(writexl)
+  library(cowplot)
+  library(figur)
+  library(ggplotify)
+  
+# container ------
+  
+  issues <- list()
+
+# Issue 1: data discrepancy -------
+
+  insert_msg('Data discrepancy')
+
+  issues$score_discrepancy <- cov_data$data %>% 
+    select(starts_with('ipq_q'), 
+           'ipq_sum', 
+           'KW_IPQ') %>% 
+    set_names(c(paste('IPQ item', 1:8), 
+                'IPQ item sum (done now)', 
+                'IPQ score (from the CovILD pipeline)')) 
+  
+  issues$score_discrepancy %>% 
+    writexl::write_xlsx('./issues/IPQ_discrepancy.xlsx')
+  
+# Issue 2: item correlations ------
+  
+  insert_msg('Item correlation')
+  
+  issues$correlogram <- as.ggplot(~pairs.panels(ipq_co$analysis_table %>% 
+                                                  set_names(paste0('Q', 1:8)), 
+                                                method = 'spearman', 
+                                                stars = TRUE)) %>% 
+    as_figure('item_correlation', 
+              ref_name = 'correlogram', 
+              w = 180, 
+              h = 180)
+  
+# Issue 3: is the inversion really correct? ------
+  
+  insert_msg('Inversion')
+  
+  issues$items_12568 <- cohort$psy_plots[c('ipq_q1', 
+                                           'ipq_q2', 
+                                           'ipq_q5', 
+                                           'ipq_q6', 
+                                           'ipq_q8')] %>% 
+    map(~.x + theme(legend.position = 'none')) %>% 
+    plot_grid(plotlist = ., 
+              ncol = 2, 
+              align = 'hv') %>% 
+    as_figure('items12568_severity', 
+              ref_name = 'items_12568', 
+              w = 180, 
+              h = 210)
+  
+  issues$items_347 <- cohort$psy_plots[c('ipq_q3', 
+                                         'ipq_q4', 
+                                         'ipq_q7')] %>% 
+    map(~.x + theme(legend.position = 'none')) %>% 
+    plot_grid(plotlist = ., 
+              ncol = 2, 
+              align = 'hv') %>% 
+    as_figure('items347_severity', 
+              ref_name = 'items_12568', 
+              w = 180, 
+              h = 140)
+  
+# saving the plots ------
[…]

### 2023-01-09 modeling scripts/elanet.R

@@ -15,7 +15,7 @@
   eln_mod$response <- mod$responses %>% 
     map(~mod$multi_tbl[[.x]])
 
-  eln_mod$variables <- model.matrix(~., data = mod$multi_tbl[globals$variables])
+  eln_mod$variables <- model.matrix(~., data = mod$multi_tbl[mod$variables])
   
   ## alpha and n repetitions
   
@@ -57,8 +57,9 @@
   
   eln_mod$lambda_tbl <- eln_mod$lambda_tune %>% 
     map(~map(.x, 
-             ~as_tibble(.x[c('lambda', 'cvm', 'cvup', 'cvlo', 'lambda.min')]))) %>% 
-    map(~map_dfr(.x, filter, lambda == lambda.min))
+             ~as_tibble(.x[c('lambda', 'cvm', 'cvup', 'cvlo', 
+                             'lambda.min', 'lambda.1se')]))) %>% 
+    map(~map_dfr(.x, filter, lambda == lambda.1se))
   
   eln_mod$opt_lambda <- eln_mod$lambda_tbl %>% 
     map(filter, cvm == min(cvm)) %>% 

### 2023-01-09 modeling scripts/elimination.R

@@ -1,0 +1,208 @@
+# Multi-variate modeling with backward elimination of the terms
+
+  insert_head()
+
+# container -----
+  
+  eli_mod <- list()
+  
+  registerDoParallel()
+  
+# Full model construction -----
+  
+  insert_msg('Building full models')
+  
+  eli_mod$full_models <- list(formula = mod$multi_form) %>% 
+    pmap(make_lm, 
+         data = mod$multi_tbl, 
+         mod_fun = lm, 
+         family = NULL)
+  
+# Model optimization -------
+  
+  insert_msg('Model optimization')
+  
+  eli_mod$opt_models <- eli_mod$full_models %>% 
+    map(~safely(step)(.x, 
+                      step_fun = MASS::stepAIC, 
+                      direction = 'both', 
+                      trace = FALSE, 
+                      k = log(lmqc::nobs(.x)))) %>% 
+    map(~.x$result) %>% 
+    compact
+
+# Model assumptions ------
+  
+  insert_msg('Model assumptions: normality testing and plots')
+  
+  ## tabular results
+  
+  eli_mod$resid_tbl <- eli_mod$opt_models %>% 
+    map(resid, type.predict = 'response')
+  
+  ## Shapiro - Wilk test on residuals
+  
+  eli_mod$resid_norm <- eli_mod$resid_tbl %>% 
+    map(explore, 
+        what = 'normality', 
+        variables = '.resid') %>% 
+    map2_dfr(., names(.), ~mutate(.x, response = .y))
+  
+  ## diagnostic plots
+  
+  eli_mod$resid_plots <- eli_mod$opt_models %>% 
+    map(plot, 
+        type = 'residuals', 
+        cust_theme = globals$common_theme, 
+        type.predict = 'response')
+  
+# Model fit stats -------
+  
+  insert_msg('Model fit stats')
+  
+  eli_mod$fit_stats <- eli_mod$opt_models %>% 
+    map_dfr(summary, type = 'fit') %>% 
+    mutate(plot_cap = paste0('R\u00B2 = ', signif(raw_rsq, 2), 
+                             ', RMSE = ', signif(rmse, 2), 
+                             ', total n = ', n_complete))
+  
+# Model inference -----
+  
+  insert_msg('Model inference')
+  
+  eli_mod$inference <- eli_mod$opt_models %>% 
+    map(summary, type = 'inference')
+  
+# Forest plots ------
+  
+  insert_msg('Forest plots')
+  
[…]

### 2023-01-09 modeling scripts/lasso.R

@@ -15,7 +15,7 @@
   lasso_mod$response <- mod$responses %>% 
     map(~mod$multi_tbl[[.x]])
 
-  lasso_mod$variables <- model.matrix(~., data = mod$multi_tbl[globals$variables])
+  lasso_mod$variables <- model.matrix(~., data = mod$multi_tbl[mod$variables])
   
   ## alpha and n repetitions
   
@@ -57,8 +57,9 @@
   
   lasso_mod$lambda_tbl <- lasso_mod$lambda_tune %>% 
     map(~map(.x, 
-             ~as_tibble(.x[c('lambda', 'cvm', 'cvup', 'cvlo', 'lambda.min')]))) %>% 
-    map(~map_dfr(.x, filter, lambda == lambda.min))
+             ~as_tibble(.x[c('lambda', 'cvm', 'cvup', 'cvlo', 
+                             'lambda.min', 'lambda.1se')]))) %>% 
+    map(~map_dfr(.x, filter, lambda == lambda.1se))
   
   lasso_mod$opt_lambda <- lasso_mod$lambda_tbl %>% 
     map(filter, cvm == min(cvm)) %>% 

### 2023-01-09 modeling scripts/uni_variate.R

@@ -13,8 +13,8 @@
   ## without normalization
   
   uni_mod$models <- mod$responses  %>% 
-    map(function(resp) set_names(globals$variables, 
-                                 globals$variables) %>% 
+    map(function(resp) set_names(mod$variables, 
+                                 mod$variables) %>% 
           map(~make_lm(data = mod$multi_tbl, 
                        response = resp, 
                        indep_variable = .x, 

### 2023-01-09 modeling.R

@@ -1,14 +1,35 @@
 # Analysis of differences in mental health status as a function
 # of baseline factors, symptoms and cardiopulmonary features
-# responses: the concern, identity, time (ipq_sub1) and 
+# responses: the emotion. concern, consequences (ipq_sub1) and 
 # the control, coherence (ipq_sub2) components of the IPQ questionnaire
 # and the total IPQ item sum.
 # Square roots of them used because of improved normality
 # methods: 
-# (1) uni-variate Gaussian-family modeling, 
-# (2) multi-variate backward elimination Gaussian-family modeling
-# (3) multi-variate elastic net modeling
-# multi-variate models are validated by 10-fold CV
+# (1) uni-variable Gaussian-family modeling, 
+#
+# (2) multi-parameter linear modeling with backward elimination 
+# (just to show that it is badly over-parameterized)
+#
+# (3) multi-parameter Elastic Net modeling
+#
+# (4) multi-parameter LASSO
+#
+# (5) multi-parameter Bayesian LASSO
+#
+# (4) multi-variate elastic net modeling
+#
+# Numeric explanatory variables are normalized and mean centered and their 
+# second order terms are included to account for curvature of the 
+# relationship with the reponse
+#
+# Tuning of lambda is accomplished by 10-fold cross validation 
+# repeated 200 times (Elastic Net and LASSO). The optimal lambda associated 
+# with the lowest cross-validation error is selected from the set of lambdas 
+# of maximal regularization models (parameter: 'lambda.1se'). Note: using 
+# 'lambda.min' (minimal model deviance) results in an expansion of 
+# cross-validation error suggestive of over-parameterization of the model.
+#
+# Multi-variate models are validated by 10-fold CV using caret.
 
 # tools -----
 
@@ -36,13 +57,77 @@
                      ipq_sub2 = 'sqrt_ipq_sub2', 
                      ipq_total = 'sqrt_ipq_total')
   
+  ## explanatory variables. As defined in the globals list
+  ## WHO ordinal score is removed since this variable is not really quantitative
+  ## including variables with transformations improving normality and 
+  ## second order terms
+  ##
+  ## Note: variables with strong non-normality are stratified with usual cutoffs
+  
+  mod$variables <- c('sex', 
+                     'age', 
+                     'age_sq', 
+                     'smoking_history', 
+                     'weight_class', 
+                     'comorb_present', 
[…]

### 2023-01-09 paper scripts/figures.R

@@ -1,12 +1,11 @@
-# Paper and supplement figures
+# Main figures of the manuscript
 
   insert_head()
   
 # containers -----
   
   paper_fig <- list()
-  suppl_fig <- list()
-  
+
 # Figure 1: study CONSORT diagram ------
   
   insert_msg('Figure 1: study consort')
@@ -24,62 +23,134 @@
   
   insert_msg('Figure 2: key factors impacting on disease perception')
   
-  ## top panel: Venn plot
+  ## Venn plots are present and the key factors for the total illness
+  ## perception score, the emotion/concern/consequences 
+  ## and the control/coherence sub-scores (no factors identified, actually) 
+  ## are shown
   
-  paper_fig$key_factors$upper_panel <- 
+  paper_fig$key_factors$left_panel <- 
     plot_grid(mod_plots$venn_plots$ipq_total$venn, 
-              plot_grid(mod_plots$venn_plots$ipq_total$legend_grob, 
-                        ggdraw() + 
-                          draw_text(paste(an$cmm_variables$ipq_total %>% 
-                                            translate_var(out_value = 'label_long'), 
-                                          collapse = '\n'), 
-                                    hjust = 0, 
-                                    x = 0.1, 
-                                    size = 8), 
-                        ncol = 2, 
-                        rel_widths = c(0.4, 0.6)), 
+              ggdraw() + 
+                draw_text(paste(an$cmm_variables$ipq_total %>% 
+                                  translate_var(out_value = 'label'), 
+                                collapse = '\n'), 
+                          hjust = 0, 
+                          x = 0.1, 
+                          size = 8), 
               ncol = 2, 
-              rel_widths = c(0.5, 0.5))
+              rel_widths = c(0.57, 0.43))
   
-  ## middle panel: violin plots
+  paper_fig$key_factors$right_panel <- 
+    plot_grid(mod_plots$venn_plots$ipq_sub1$venn, 
+              ggdraw() + 
+                draw_text(paste(an$cmm_variables$ipq_sub1 %>% 
+                                  translate_var(out_value = 'label'), 
+                                collapse = '\n'), 
+                          hjust = 0, 
+                          x = 0.1, 
+                          size = 8), 
+              ncol = 2, 
+              rel_widths = c(0.57, 0.43))
   
-  paper_fig$key_factors$middle_panel <- 
-    ipq_factors$violin_plots[c('fatigue_sympt', 
-                               'respi_comorb', 
[…]

### 2023-01-09 paper scripts/render.R

@@ -11,6 +11,15 @@
                                         reference_docx = 'ms_template.docx'), 
          output_dir = './paper')
   
+# Reviewer material ------
+  
+  insert_msg('Rendering the supplements')
+  
+  render('./paper/markdown/reviewer_material.Rmd', 
+         output_format = word_document2(number_sections = FALSE, 
+                                        reference_docx = 'ms_template.docx'), 
+         output_dir = './paper')
+  
 # paper -----
   
   insert_msg('Rendering the paper')

### 2023-01-09 paper scripts/supplement.R

@@ -1,0 +1,497 @@
+# Supplementary figures for the manuscript and figures for the Reviewers
+
+  insert_head()
+
+# container ------
+
+  suppl_fig <- list()
+  rev_fig <- list()
+  
+# Figure S1: frequency of symptoms in the severity strata ------
+  
+  insert_msg('Figure S1: frequency of symptoms in the severity strata')
+  
+  ## left panel
+  
+  suppl_fig$sev_symptoms$left <- 
+    plot_grid(sympt_kinet$plots$all + 
+                theme(legend.position = 'none'), 
+              get_legend(sympt_kinet$plots$all), 
+              nrow = 3)
+  
+  ## right panel: symptom frequencies in the severity strata
+  
+  suppl_fig$sev_symptoms$right <-
+    plot_grid(cohort$symptom_plot + 
+                theme(legend.position = 'none', 
+                      strip.text = element_text(size = 7)), 
+              plot_grid(get_legend(cohort$symptom_plot), 
+                        ggdraw() + 
+                          draw_text(paste0('ambulatory: n = ', 
+                                           table(cov_data$clear_data$cat_WHO)[1], 
+                                           '\nmoderate: n = ', 
+                                           table(cov_data$clear_data$cat_WHO)[2], 
+                                           '\nsevere: n = ', 
+                                           table(cov_data$clear_data$cat_WHO)[3]), 
+                                    size = 8, 
+                                    hjust = 0,
+                                    vjust = 0.7, 
+                                    x = 0.1), 
+                        ncol = 2), 
+              nrow = 2, 
+              rel_heights = c(0.92, 0.08))
+  
+  ## the entire figure
+  
+  suppl_fig$sev_symptoms <- 
+    plot_grid(suppl_fig$sev_symptoms$left, 
+              suppl_fig$sev_symptoms$right, 
+              ncol = 2, 
+              labels = LETTERS, 
+              label_size = 10)
+  
+  suppl_fig$sev_symptoms <- suppl_fig$sev_symptoms %>% 
+    as_figure(label = 'figure_s1_severity_symptoms', 
+              ref_name = 'sev_symptoms', 
+              caption = paste('Kinetic of persistent symptoms and', 
+                              'frequency of persistent symptoms', 
+                              'in the COVID-19 severity strata at the', 
+                              'one-year follow-up.'), 
+              w = 180, 
+              h = 220)
+  
+# Figure S2: overlap between sequelae -------
+  
+  insert_msg('Figure S2: sequelae overlap')
+
[…]

### 2023-01-09 paper scripts/tables.R

@@ -6,6 +6,8 @@
   
   paper_tbl <- list()
   suppl_tbl <- list()
+  
+  rev_tables <- list() ## tables for the reviewer
 
 # Table 1: characteristic of the study cohort -----
   
@@ -43,7 +45,12 @@
     as_mdtable(paper_tbl$baseline , 
                label = 'cohort_baseline', 
                ref_name = 'cohort_baseline', 
-               caption = 'Baseline demographic and clinical characteristic of the study cohort.')
+               caption = paste('Baseline demographic and clinical', 
+                               'characteristic of the study cohort.', 
+                               'Numeric variables are presented as medians', 
+                               'with interquartile ranges (IQR).',  
+                               'Categorical variables are shown as percentages', 
+                               'and counts within the strata.'))
   
 # Table 2: IPQ components and sum in the cohort and severity groups -----
   
@@ -54,7 +61,6 @@
                                                          'significance', 
                                                          'eff_size')], 
                                by = 'variable') %>% 
-    filter(variable %in% c(paste0('ipq_q', 1:8), 'ipq_total')) %>% 
     format_summ_tbl(rm_n = TRUE) %>% 
     rbind(tibble(variable = 'n, participants', 
                  cohort = pap$n_baseline$n[1], 
@@ -79,7 +85,16 @@
     as_mdtable(paper_tbl$psych , 
                label = 'cohort_ipq', 
                ref_name = 'psych', 
-               caption = 'Illness perception score and item values of the Brief Illness Perception Questionnaire (BIPQ, Q1 - Q8) at one year after COVID-19.')
+               caption = paste('Illness perception score', 
+                               '(sum of all BIPQ items),', 
+                               'emotion/concern/consequences subscore', 
+                               '(sum of items 1, 2, 5, 6 and 8)', 
+                               'control/coherence subscore', 
+                               '(sum of items 3, 4 and 7) and values of sigle', 
+                               'Brief Illness Perception Questionnaire items', 
+                               '(BIPQ, Q1 - Q8) at one year after COVID-19.', 
+                               'Values are presented as medians with', 
+                               'interquartile ranges (IQR).'))
   
 # Supplementary Table S1: variable dictionary -------
   
@@ -95,15 +110,92 @@
                 'Unit', 
                 'Description'))
   
+  ## adding the levels
+  
+  suppl_tbl$var_lexicon <- suppl_tbl$var_lexicon$`R name` %>% 
[…]

### 2023-01-09 paper scripts/values.R

@@ -67,7 +67,7 @@
   insert_msg('N numbers and names of sgnificant explanatory variables')
   
   paper_val$n_indep_vars <- 
-    mdexpr(length(globals$variables), 
+    mdexpr(length(mod$variables), 
            ref_name = 'n_indep_vars', 
            caption = 'Number of independent variables.')
 
@@ -159,6 +159,32 @@
            ref_name = 'key_factors_and', 
            caption = 'The key factors for IPQ.')
   
+# Numbers and names of the key factors for the emotion/concern/consequences -------
+  
+  insert_msg('Key factors for the emotion/concern/consequences')
+  
+  paper_val$key_factor_number_sub1 <-  
+    mdexpr(length(an$cmm_variables$ipq_sub1), 
+           ref_name = 'key_factor_number_sub1', 
+           caption = paste('Number of the key factors for', 
+                           'emotion/concern/consequences.'))
+  
+  paper_val$key_factors_sub1 <- 
+    mdexpr(an$cmm_variables$ipq_sub1 %>% 
+             translate_var(out_value = 'label_long') %>% 
+             de_capitalize %>% 
+             paste(collapse = ', '), 
+           ref_name = 'key_factors_sub1', 
+           caption = 'The key factors for emotion/concern/consequences.')
+  
+  paper_val$key_factors_and_sub1 <- 
+    mdexpr(an$cmm_variables$ipq_sub1 %>% 
+             translate_var(out_value = 'label_long') %>% 
+             de_capitalize %>% 
+             collapse_and, 
+           ref_name = 'key_factors_and_sub1', 
+           caption = 'The key factors for emotion/concern/consequences.')
+  
 # Cohort characteristic ------
   
   insert_msg('Cohort characteristic')
@@ -226,9 +252,9 @@
            ref_name = 'diastolic_dysf', 
            caption = 'Diastolic dysfunction, cohort')
   
-# IPQ ------
+# IPQ in the cohort ------
   
-  insert_msg('IPQ')
+  insert_msg('IPQ in the cohort')
   
   paper_val$ipq_cohort <- 
     mdexpr(get_median_iqr(cov_data$clear_data, 'ipq_total'), 
@@ -322,9 +348,9 @@
            ref_name = 'clusters_percent', 
            caption = 'Clusters, percents of the cohort.')
   
-# Optimal lambda values -----
[…]

### 2023-01-09 paper.R

@@ -35,7 +35,7 @@
   
   ## common scales for the estimate plots
   
-  pap$est_limits <- c(-0.35, 0.49)
+  pap$est_limits <- c(-0.5, 0.5)
   
   pap$est_fills <- scale_fill_gradient2(low = 'steelblue', 
                                         high = 'firebrick', 
@@ -50,6 +50,7 @@
   
   c('./paper scripts/tables.R', 
     './paper scripts/figures.R', 
+    './paper scripts/supplement.R', 
     './paper scripts/values.R', 
     './paper scripts/render.R') %>% 
     source_all(message = TRUE, crash = TRUE)

### 2023-01-09 tools/project_tools.R

@@ -261,6 +261,7 @@
                           labelRegions = FALSE, 
                           fontScale = 2,
                           plot_leg_ratio = c(0.9, 0.1), 
+                          comp_heights = c(0.08, 0.06, 0.8, 0.06), 
                           panel = TRUE, ...) {
     
     ## plots a Venn diagram employing nVennR
@@ -269,8 +270,9 @@
       
       subset_names <- stri_replace(subset_names, 
                                    fixed = '\n', 
-                                   replacement = ' ') %>% 
-        set_names(data, subset_names)
+                                   replacement = ' ')
+      
+      data <- set_names(data, subset_names)
       
       
     } 
@@ -318,7 +320,8 @@
     
     ## the legend will be obtained from a 'fake plot'
     
-    gg_legend <- tibble(subset = factor(subset_names, levels = rev(subset_names)), 
+    gg_legend <- tibble(subset = factor(subset_names, 
+                                        levels = subset_names), 
                         n = 1) %>% 
       ggplot(aes(x = subset_names, 
                  y = n, 
@@ -326,7 +329,8 @@
       geom_bar(stat = 'identity', 
                alpha = if(2.5 * opacity > 1) 1 else 2.5 * opacity, 
                size = borderWidth) + 
-      scale_fill_manual(values = rev(tolower(fill_color)), 
+      scale_fill_manual(values = set_names(fill_color, 
+                                           subset_names), 
                         name = '') + 
       globals$common_theme + 
       theme(legend.position = legend_position)
@@ -353,6 +357,126 @@
            legend_grob = gg_legend)
       
     }
+    
+  }
+  
+# Plotting two-way interactions ------
+  
+  two_way_corr <- function(data, 
+                           response, 
+                           num_variable, 
+                           split_factor = 'ct_severity_any', 
+                           plot_title = translate_var(num_variable), 
+                           plot_subtitle = NULL, 
+                           x_lab = translate_var(num_variable, 
+                                                 out_value = 'axis_lab'), 
+                           y_lab = translate_var(response, 
+                                                 out_value = 'axis_lab'), 
+                           labeller = 'label_value',  
+                           plot_tag = NULL, 
+                           fill_colors = c('steelblue', 'coral3'), ...) {
+    
+    ## correlations with a illness perception score 
+    ## in the strata defined by a split factor
+    
+    data %>% 
[…]

### 2023-02-09 paper/markdown/manuscript.Rmd
Second revision
Minor Reviewer Issues, duplicated references removed

@@
 
-A subset of coronavirus disease 2019 (COVID-19) patients is affected by protracted somatic symptoms, cardiopulmonary pathology and mental health disorders [@Sonnweber2020; @Sahanic2021; @Luger2022; @Sonnweber2022; @Hufner2022; @Al-Aly2021; @Davis2021; @Soriano2022; @Evans2021a; @Sudre2021; @Chen2022].
+A subset of coronavirus disease 2019 (COVID-19) patients is affected by protracted somatic symptoms, cardiopulmonary pathology and mental health disorders [@Sonnweber2020; @Sahanic2021; @Luger2022; @Sonnweber2022; @Hufner2022; @Al-Aly2021; @Davis2021; @Soriano2022; @Evans2021; @Sudre2021; @Chen2022].
 Persistent COVID-19-related symptoms have been initially described by patient initiatives [@McCorkell2021] and subsequently recognized as 'post-COVID-19 condition' by the clinical community [@Sonnweber2020; @Sahanic2021; @Davis2021; @Soriano2022; @Al-Aly2021; @Sudre2021].
@@
 The clearance rate is particularly fast for non-specific infection complaints [@Sudre2021; @Sahanic2021; @Rass2022]. 
-In a recent metaanalysis of 1.7 million patients, the pooled prevalence estimate of post-COVID-19 condition, i.e. symptoms present for 3 months or longer, was 43% [@Soriano2022; @Chen2022]. 
+In a recent</span> <span custom-style = 'revision2'>meta-analysis</span> <span custom-style = 'revision'>of 1.7 million patients, the pooled prevalence estimate of post-COVID-19 condition, i.e.
+symptoms present for 3 months or longer, was 43% [@Soriano2022; @Chen2022].
 Yet, in a rigorously controlled study, approximately 12.7% of SARS-CoV-2-positive individuals suffered from persistent somatic symptoms directly associated with COVID-19 [@Ballering2022].
-Fatigue, concentration and memory problems, sleep disorders, physical performance deficits, smell/taste disorders and hair loss, dyspnea and pain are the most common, frequently overlapping and relapsing manifestations [@Chen2022; @Sonnweber2022; @Sahanic2021; @Davis2021; @Rass2022; @Evans2021a; @Huang2021; @Ballering2022].</span>
+Fatigue, concentration and memory problems, sleep disorders, physical performance deficits, smell/taste disorders and hair loss, dyspnea and pain are the most common, frequently overlapping and relapsing manifestations [@Chen2022; @Sonnweber2022; @Sahanic2021; @Davis2021; @Rass2022; @Evans2021; @Huang2021; @Ballering2022].</span>
 
@@
 <span custom-style = 'revision'>We aimed to describe IP and IP components, identify demographic and clinical variables affecting IP and explore IP patterns in convalescents suffering from persistent somatic symptoms (PSS), lung or heart abnormalities one year after COVID-19.
-To this end, we gauged IP in a long-term observatory cross-sectional COVID-19 collective [@Sonnweber2020; @Sahanic2021; @Luger2022; @Sonnweber2022] with the Brief Illness Perception Questionnaire (BIPQ) [@Broadbent2006] and investigated components, influential factors and patterns of IP by factor analysis, regularized multi-parameter modeling and unsupervised PAM clustering.</span>
+To this end, we gauged IP in a long-term observatory cross-sectional COVID-19 collective [@Sonnweber2020; @Sahanic2021; @Luger2022; @Sonnweber2022] with the Brief Illness Perception Questionnaire (BIPQ) [@Broadbent2006] and investigated components, influential factors and patterns of IP by factor analysis, regularized multi-parameter modeling and unsupervised clustering.</span>
 
@@
 Out of 145 enrolled individuals, `r nrow(cov_data$raw_data)` completed the one-year follow-up.</span>
-A subset of participants (n = `r nrow(cov_data$clear_data)`) was analyzed who displayed (1) PSS or (2) any abnormality in chest computed tomography (CT) or (3) any lung function testing (LFT) deficits or (4) any heart abnormality in trans-thoracic echocardiography (TTE) at the one-year follow-up <span custom-style = 'revision'>and had a complete set of BIPQ, demographic and clinical analysis variables (__Figure \@ref(fig:fig-consort)__, __Supplementary Tables S1 - S2__).</span>
+A subset of participants (n = `r nrow(cov_data$clear_data)`) was analyzed who displayed (1) PSS or (2) any abnormality in chest computed tomography or (3) any lung function testing deficits or (4) any heart abnormality in trans-thoracic echocardiography at the one-year follow-up <span custom-style = 'revision'>and had a complete set of BIPQ, demographic and clinical analysis variables (__Figure \@ref(fig:fig-consort)__, __Supplementary Tables S1 - S2__).</span>
 <span custom-style = 'revision'>We decided to restrict the analysis to incomplete COVID-19 recovery, because BIPQ was initially designed for and validated in somatic conditions [@Broadbent2006].</span>
@@
 
-Physical performance was rated with the Eastern Cooperative Oncology Group scale (ECOG). 
-Dyspnea was scored with the Modified Medical British Research Council scale (mMRC).
-Fatigue was rated with likert and bimodal Chalder's Fatigue Scales (CFS) [@Chalder1993; @Morriss1998]. 
-Exertional capacity was assessed by six-minute walking distance (SMWD) [@Crapo2012].
+Physical performance was rated with the Eastern Cooperative Oncology Group scale. 
+Dyspnea was scored with the Modified Medical British Research Council scale.
+Fatigue was rated with likert and bimodal Chalder's Fatigue Scales [@Chalder1993; @Morriss1998]. 
+Exertional capacity was assessed by six-minute walking distance [@Crapo2012].
 
-The following, <span custom-style = 'revision'>PSS,</span> were analyzed at the one-year follow-up: 
-reduced physical performance (ECOG $\geq$ 1), 
-dyspnea (mMRC $\geq$ 1), 
+The following <span custom-style = 'revision'>PSS</span> were analyzed at the one-year follow-up: 
+reduced physical performance (Eastern Cooperative Oncology Group scale $\geq$ 1), 
+dyspnea (Modified Medical British Research Council scale $\geq$ 1), 
 self-reported cough (yes/no item), 
@@
 self-reported hair loss (yes/no), 
-significant fatigue (bimodal CFS $\geq$ 4). 
+significant fatigue (bimodal Chalder's Fatigue Scale $\geq$ 4). 
 <span custom-style = 'revision'>Records of reduced physical performance, dyspnea, cough, sleep problems, night sweating, hyposmia/anosmia and gastrointestinal symptoms were available for acute COVID-19 and all follow-ups [@Sonnweber2022].</span>
 
-LFT abnormality was defined as at least one parameter < 80% (forced vital capacity [FVC], forced expiratory volume in 1 second [FEV1], total lung capacity, diffusion lung capacity for carbon monoxide) or < 70% (FEV1:FVC ratio) of the reference value [@Sonnweber2022].
-CT images were evaluated with the Fleischner Society glossary terms [@Hansell2008] and the CT severity score [@Sonnweber2020; @Luger2022; @Sonnweber2022]. 
+Lung function testing abnormality was defined as at least one parameter < 80% (forced vital capacity, forced expiratory volume in 1 second, total lung capacity, diffusion lung capacity for carbon monoxide) or < 70% (ratio of forced expiratory volume in 1 second to forced vital capacity) of the reference value [@Sonnweber2022].
+Computed tomography images were evaluated with the Fleischner Society glossary terms [@Hansell2008] and the computed tomography severity score [@Sonnweber2020; @Luger2022; @Sonnweber2022]. 
 Blood biomarkers encompassed hemoglobin and parameters of iron turnover, inflammation and coagulation.
@@
 
-Clustering by the BIPQ items was accomplished by PAM (partitioning around medoids) <span custom-style = 'revision'>with Euclidean distance [@Schubert2019; @Drost2018].</span> 
+Clustering by the BIPQ items was accomplished by the partitioning around medoids algorithm <span custom-style = 'revision'>with Euclidean distance [@Schubert2019; @Drost2018].</span> 
 <span custom-style = 'revision'>The clustering algorithm choice was motivated by its optimal stability in 10-fold cross-validation [@Lange2004] and good explanatory performance measured by the fraction of clustering variance as compared with hierarchical and KMEANS clustering.</span>
@@
 The CovILD cohort was recruited <span custom-style = 'revision'>during the first outbreak of SARS-CoV-2</span> between April and June 2020 [@Sonnweber2020]. 
-Out of 145 participants enrolled, `r nrow(cov_data$clear_data)` individuals with PSS, lung CT, LFT or cardiological abnormalities, <span custom-style = 'revision'>complete BIPQ and a complete set of study variables set at the one-year follow-up</span> were analyzed.
+Out of 145 participants enrolled, `r nrow(cov_data$clear_data)` individuals with PSS or abnormalities in lung computed tomography, lung function testing or cardiological abnormalities, <span custom-style = 'revision'>complete BIPQ and a complete set of study variables set at the one-year follow-up</span> were analyzed.
 The major reasons of patient dropout were missing follow-ups and incomplete BIPQ (__Figure \@ref(fig:fig-consort)__).
 <span custom-style = 'revision'>The excluded participants tended towards less severe acute COVID-19, more dermatological and gastrointestinal symptoms at the one-year follow-up, and lower blood hemoglobin and ferritin as compared with the analysis collective; those effects were, however, not significant.
-The excluded participants had also significantly lower SMWD (__Supplementary Figure S3__).</span>
+The excluded participants had also significantly lower six-minute walking distance (__Supplementary Figure S3__).</span>
 
-The <span custom-style = 'revision'>analyzed</span> participants were predominantly male (`r get_percent(cov_data$clear_data, "sex")['male']`%), the median age at COVID-19 diagnosis was `r get_median_iqr(cov_data$clear_data, "age")['median']` years (IQR: `r get_median_iqr(cov_data$clear_data, "age")['iqr']`), over one-third were active or ex-smokers (`r get_percent(cov_data$clear_data, "smoking_history")['yes']`%).
+The <span custom-style = 'revision'>analyzed</span> participants were predominantly male (`r get_percent(cov_data$clear_data, "sex")['male']`%), the median age at COVID-19 diagnosis was `r get_median_iqr(cov_data$clear_data, "age")['median']` years (interquartile range: `r get_median_iqr(cov_data$clear_data, "age")['iqr']`), over one-third were active or ex-smokers (`r get_percent(cov_data$clear_data, "smoking_history")['yes']`%).
 Most participants suffered from comorbidities (`r get_percent(cov_data$clear_data, "comorb_present")['yes']`%), with cardiovascular disease, metabolic and respiratory disorders as leading conditions. 
@@
 
-LFT abnormalities affected `r get_percent(cov_data$clear_data, "lufo_red")['yes']`% of participants and were substantially enriched in severe COVID-19 survivors. 
-Lung CT abnormalities were found in `r get_percent(cov_data$clear_data, "ct_severity_any")['yes']`% individuals and their frequency and scoring was significantly higher in moderate-to-severe COVID-19 than in the ambulatory subset.
+Lung function testing abnormalities affected `r get_percent(cov_data$clear_data, "lufo_red")['yes']`% of participants and were substantially enriched in severe COVID-19 survivors. 
+Abnormalities in lung computed tomography were found in `r get_percent(cov_data$clear_data, "ct_severity_any")['yes']`% individuals and their frequency and scoring was significantly higher in moderate-to-severe COVID-19 than in the ambulatory subset.
 The leading cardiological finding was low grade diastolic dysfunction (`r get_percent(cov_data$clear_data, "diastolic_dysf")['yes']`%), which was significantly enriched in moderate and severe COVID-19 survivors.
-<span custom-style = 'revision'>Concomitant PSS, abnormal lung CT, LFT and TTE one year after COVID-19 were observed in `r signif(fup_overlap$all_intersection/nrow(cov_data$clear_data), 3) * 100`% of participants.
+<span custom-style = 'revision'>Concomitant PSS, abnormal lung computed tomography, lung function testing and diastolic dysfunction in echocardiography one year after COVID-19 were observed in `r signif(fup_overlap$all_intersection/nrow(cov_data$clear_data), 3) * 100`% of participants.
 In `r signif(14/nrow(cov_data$clear_data), 3) * 100`% of participants, PSS were not accompanied by cardiopulmonary abnormalities (__Supplementary Figure S2__).</span>
@@
 Most laboratory parameters at the one-year follow-up were within their normal values. 
-Mild anemia and improper glycemia control (HbA1c) were evident primarily in moderate and severe COVID-19 survivors (__Supplementary Table S5__).
+Mild anemia and improper glycemia control assessed by glycated hemoglobin levels were evident primarily in moderate and severe COVID-19 survivors (__Supplementary Table S5__).
 
-<span custom-style = 'revision'>Females displayed significantly less severe COVID-19 and less chest CT abnormalities but higher PSS rates and counts than males at the one-year follow-up. 
+<span custom-style = 'revision'>Females displayed significantly less severe COVID-19 and less chest computed tomography abnormalities but higher PSS rates and counts than males at the one-year follow-up.
 Differences between the genders in frequencies and scoring of specific PSS were not significant (__Supplementary Table S6__ and not shown).</span>
@@
 
-The median <span custom-style = 'revision'>total</span> IP score representing the sum of all BIPQ items was `r get_median_iqr(cov_data$clear_data, "ipq_total")['median']` (IQR: `r get_median_iqr(cov_data$clear_data, "ipq_total")['iqr']`) in the study collective, the differences between ambulatory, moderate and severe COVID-19 were not significant.
+The median <span custom-style = 'revision'>total</span> IP score representing the sum of all BIPQ items was `r get_median_iqr(cov_data$clear_data, "ipq_total")['median']` (interquartile range: `r get_median_iqr(cov_data$clear_data, "ipq_total")['iqr']`) in the study collective, the differences between ambulatory, moderate and severe COVID-19 were not significant.
 <span custom-style = 'revision'>The emotion/concern/consequences IP component rated by the Q1, Q2, Q5, Q6 and Q8 item sum was the lowest in moderate and the highest in severe COVID-19. 
@@
 The key explanatory factors for the emotion/concern/consequences IP component identified by all three modeling algorithms were 
-PSS count, reduced physical performance, fatigue scoring, significant fatigue, hair loss, cough, diastolic dysfunction, elevated C-reactive protein (CRP) at the one-year follow-up as well as rehabilitation during COVID-19 recovery, respiratory comorbidity and body weight class.
+PSS count, reduced physical performance, fatigue scoring, significant fatigue, hair loss, cough, diastolic dysfunction, elevated C-reactive protein at the one-year follow-up as well as rehabilitation during COVID-19 recovery, respiratory comorbidity and body weight class.
 In uni-variable analysis, PSS number, fatigue scoring and significant fatigue, reduced performance, hair loss, rehabilitation and respiratory comorbidity were significantly associated with increased emotion/concern/consequences scores (__Figure \@ref(fig:fig-sub1-bipq)__).</span>
 
-<span custom-style = 'revision'>Age, gender, CT, LFT or TTE abnormalities alone had no significant effects on total IP or its components in uni-variable analysis (__Supplementary Figures S12 - S13__).
-However, lung CT abnormalities at the one-year follow-up amplified the effects of diastolic dysfunction, fatigue and fatigue rating on total IP and the emotion/concern/consequences component, as revealed by significant interactions in two-way ANOVA.
-In particular, individuals with concomitant diastolic dysfunction plus CT abnormalities or with high fatigue scoring plus CT abnormalities displayed significantly higher emotion/concern/consequences scores than participants without lung CT findings.
-Similar interaction effects of LFT deficits, fatigue and symptom numbers and of diastolic dysfunction and fatigue were discerned as well (__Supplementary Figures S14__ - __S16__).
-Participants with respiratory comorbidities had significantly elevated fatigue scores, LFT abnormality rates, total IP and emotion/concern/consequences ratings as compared with the remaining collective.
-This may suggest additional interaction effects of respiratory conditions, LFT deficits and persistent fatigue on IP (__Supplementary Table S7__).</span>
+<span custom-style = 'revision'>Age, gender, or abnormalities in chest computed tomography or lung function testing, or diastolic dysfunction alone had no significant effects on total IP or its components in uni-variable analysis (__Supplementary Figures S12 - S13__).
+However, lung abnormalities in computed tomography at the one-year follow-up amplified the effects of diastolic dysfunction, fatigue and fatigue rating on total IP and the emotion/concern/consequences component, as revealed by significant interactions in two-way ANOVA.
+In particular, individuals with concomitant diastolic dysfunction plus lung computed tomography abnormalities or with high fatigue scoring plus lung computed tomography abnormalities displayed significantly higher emotion/concern/consequences scores than participants without lung computed tomography findings.
+Similar interaction effects of lung function testing deficits, fatigue and symptom numbers and of diastolic dysfunction and fatigue were discerned as well (__Supplementary Figures S14__ - __S16__).
+Participants with respiratory comorbidities had significantly elevated fatigue scores, lung function testing abnormality rates, total IP and emotion/concern/consequences ratings as compared with the remaining collective.
+This may suggest additional interaction effects of respiratory conditions, lung function testing deficits and persistent fatigue on IP (__Supplementary Table S7__).</span>
 
@@
 <span custom-style = 'revision'>IP patterns were investigated by unsupervised clustering of the study participants in respect to the BIPQ items. 
-Among several algorithms tested, PAM with Euclidean distance [@Drost2018; @Schubert2019] demonstrated the best reproducibility [@Lange2004] and good explanatory performance and was hence used for definition of three IP clusters (__Supplementary Figure S17__).</span>
+Among several algorithms tested, partition aroud medoids with Euclidean distance [@Drost2018; @Schubert2019] demonstrated the best reproducibility [@Lange2004] and good explanatory performance and was hence used for definition of three IP clusters (__Supplementary Figure S17__).</span>
 Roughly half of participants <span custom-style = 'revision'>assigned to the cluster #1 displayed low scores for each BIPQ item.</span> 
-This translated to low total IP severity, <span custom-style = 'revision'>, and low scoring of the emotion/concern/consequences and lacking control/coherence IP components.</span>
+This translated to low total IP severity, <span custom-style = 'revision'>and low scoring of the emotion/concern/consequences and lacking control/coherence IP components.</span>
 Another `r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[2]`% of participants assigned to the cluster #2 were characterized by low self-perceived personal or treatment control but also by low rating of the concern, identity, timeline, consequences and emotional representation items.
@@
 <span custom-style = 'revision'>The cluster #1 individuals tended towards the lowest rates of smoking, metabolic and respiratory comorbidities (__Supplementary Table S8__). 
-In the cluster #2 the lowest acute COVID-19 severity, rates of rehabilitation and lung CT abnormality grades were observed. 
+In the cluster #2 the lowest acute COVID-19 severity, rates of rehabilitation and lung computed tomography abnormality grades were observed. 
 The cluster #2 tended also towards the lowest frequency of diastolic dysfunction.
-By contrast, in the cluster #3 affected by disease-related concerns, the frequency of severe COVID-19, COVID-19-specific rehabilitation rate, chest CT finding severity and PSS frequency at the one-year follow-up were the highest (__Figure \@ref(fig:fig-clust-cov)__).
+By contrast, in the cluster #3 affected by disease-related concerns, the frequency of severe COVID-19, COVID-19-specific rehabilitation rate, chest computed tomography finding severity and PSS frequency at the one-year follow-up were the highest (__Figure \@ref(fig:fig-clust-cov)__).
 In particular, cluster #3 individuals suffered from protracted fatigue, reduced physical performance, dyspnea and sleep problems to the highest extend, and hair loss and gastrointestinal complaints were observed solely in the cluster #3.
@@
 Effects of differences in 
-lung CT abnormality (`r clust_chara$test_results$fup %>% filter(variable == 'ct_severity_any') %>% .$plot_cap`), 
-chest CT abnormality scoring (`r clust_chara$test_results$fup %>% filter(variable == 'ct_severity_score') %>% .$plot_cap`), 
-LFT deficits (`r clust_chara$test_results$fup %>% filter(variable == 'lufo_red') %>% .$plot_cap`) or 
+lung computed tomography abnormality (`r clust_chara$test_results$fup %>% filter(variable == 'ct_severity_any') %>% .$plot_cap`), 
+chest computed tomography abnormality scoring (`r clust_chara$test_results$fup %>% filter(variable == 'ct_severity_score') %>% .$plot_cap`), 
+lung function testing deficits (`r clust_chara$test_results$fup %>% filter(variable == 'lufo_red') %>% .$plot_cap`) or 
 diastolic dysfunction (`r clust_chara$test_results$fup %>% filter(variable == 'diastolic_dysf') %>% .$plot_cap`) were substantially weaker (__Supplementary Table S8__).</span>
@@
 
-<span custom-style = 'revision'>In our observational collective affected by PSS or abnormalities in cardiopulmonary tests one year after COVID, two IP facets were identified: an emotion, concern and consequences component, and a lacking control and coherence component.
+<span custom-style = 'revision'>In our observational collective affected by PSS or abnormalities in cardiopulmonary tests one year after COVID-19, two IP facets were identified: an emotion, concern and consequences component, and a lacking control and coherence component.
 By multi-parameter modeling, we established signatures of total IP and the emotion/concern/consequences IP component. 
 They included reduced physical performance, fatigue intensity and PSS number. 
-The  emotions, concerns and consequences IP component correlated additionally with significant fatigue, hair loss, COVID-19-specific rehabilitation, respiratory comorbidity, and, to a lesser degree, with diastolic dysfunction, elevated levels of the inflammatory marker CRP and obesity.
+The  emotions, concerns and consequences IP component correlated additionally with significant fatigue, hair loss, COVID-19-specific rehabilitation, respiratory comorbidity, and, to a lesser degree, with diastolic dysfunction, elevated levels of the inflammatory marker C-reactive protein and obesity.
 Three IP clusters of study participants were identified based on the BIPQ items. 
-Differences between the IP clusters in PPS readouts such as PSS number, fatigue rating or sleep problems were much more pronounced than differing rates of llung and heart abnormalities in CT, LFT and TTE.</span>
+Differences between the IP clusters in PPS readouts such as PSS number, fatigue rating or sleep problems were much more pronounced than differing rates of lung and heart abnormalities in computed tomography, lung function testing or echocardiography.</span>
 
-We identified <span custom-style = 'revision'>fatigue, a leading persistent COVID-19 manifestation [@Chen2022; @Sudre2021; @Davis2021; @Sahanic2021; @Bierbauer2022]</span> as a strong covariate <span custom-style = 'revision'>of total IP severity and the emotion/concern/consequences IP component.</span>.
+We identified <span custom-style = 'revision'>fatigue, a leading persistent COVID-19 manifestation [@Chen2022; @Sudre2021; @Davis2021; @Sahanic2021; @Bierbauer2022]</span> as a strong covariate <span custom-style = 'revision'>of total IP severity and the emotion/concern/consequences IP component</span>.
 Similar effects were described in arthritis [@Schwartz2020], hematological malignancy [@Schoormans2020], <span custom-style = 'revision'>and recently in post-COVID-19 condition [@Bierbauer2022].</span>
@@
 <span custom-style = 'revision'>Hair loss belongs to the most frequent post-COVID-19 condition manifestations [@Chen2022; @Sahanic2021; @Jimeno-Almazan2022] and may result from psychological stress and/or inflammation [@Sattur2021].
-Hair loss, along with CRP, was linked by us with the emotion/concern/consequences IP component and may reflect an interplay between protracted inflammation and deteriorated mental health.</span>
+Hair loss, along with C-reactive protein, was linked by us with the emotion/concern/consequences IP component and may reflect an interplay between protracted inflammation and deteriorated mental health.</span>
 Post-COVID-19 condition encompass various respiratory symptoms [@Sahanic2021; @Sonnweber2020; @Sonnweber2022; @Davis2021; @Sudre2021; @Chen2022; @Bierbauer2022]. 
 In addition, COVID-19 was found to exacerbate symptoms and worsen disease control in asthma [@Philip2022]. 
-<span custom-style = 'revision'>In our collective, respiratory comorbidity was associated with the emotion/concern/consequences IP component, higher fatigue rating and LFT abnormalities. 
-Such superimposed subjective fatigue feeling and objective LFT deficits may explain more severe IP in individuals with respiratory comorbidity. 
+<span custom-style = 'revision'>In our collective, respiratory comorbidity was associated with the emotion/concern/consequences IP component, higher fatigue rating and lung function testing abnormalities.
+Such superimposed subjective fatigue feeling and objective lung function deficits may explain more severe IP in individuals with respiratory comorbidity. 
 We proposed PSS count as a strong covariate of IP severity and the emotion/concern/consequences IP component. 
@@
 However, IP rating was comparable between the genders. 
-This may be explained by the less severe COVID-19 and less lung CT abnormalities in females. 
-Interestingly, we observed significant super-additive effects of fatigue, reduced physical performance plus lung CT abnormalities on IP. 
-Since pulmonary CT findings affected nearly 75% of male participants, these interactions may influence IP in a male-specific manner and compensate for the higher PSS burden in females.</span>
+This may be explained by the less severe COVID-19 and less lung computed tomography abnormalities in females. 
+Interestingly, we observed significant super-additive effects of fatigue, reduced physical performance plus lung computed tomography abnormalities on IP. 
+Since pulmonary computed tomography findings affected nearly 75% of male participants, these interactions may influence IP in a male-specific manner and compensate for the higher PSS burden in females.</span>
 
@@
 This may reflect two non-exclusive phenomena. 
-First, the internal consistency of the lacking control/coherence component was remarkably lower than consistency of the BIPQ tool or the emotion/concern/consequences component. 
-Second, the lacking control/coherence component covering cognitive IP aspects [@Broadbent2006] may be influenced primarily by COVID-19-independent factors such as socioeconomic background, education level or follow-up care.
+First, the internal consistency of the lacking control/coherence component was remarkably lower than consistency of the BIPQ tool or the emotion/concern/consequences component.</span>
+<span custom-style = 'revision2'>This low reliability may attenuate observed associations [@Spearman1904] and hence obscure true effects of e.g.
+persistent somatic symptoms or cardiopulmonary findings on the lacking control/coherence component of IP.</span>
+<span custom-style = 'revision'>Second, the lacking control/coherence component covering cognitive IP aspects [@Broadbent2006] may be influenced primarily by COVID-19-independent factors such as socioeconomic background, education level or follow-up care.
 Low socioeconomic status was proposed as a risk factor of post-COVID-19 condition [@Dumont2022] and can affect IP measured by BIPQ [@Mafla2018].</span>
@@
 <span custom-style = 'revision'>The cluster #1 encompassing roughly</span> half of the participants displayed low total IP severity, good self-perceived coherence, personal and treatment control paralleled by low PSS rates.
-Another `r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[2]`% of participants <span custom-style = 'revision'>assigned to the cluster #2 enriched in moderate COVID-19 survivors with the lowest grade of lung CT abnormalities</span> showed a similarly low levels of PSS or fatigue.
+Another `r signif(ngroups(ipq_clust$clust_obj)$n/sum(ngroups(ipq_clust$clust_obj)$n) * 100, 2)[2]`% of participants <span custom-style = 'revision'>assigned to the cluster #2 enriched in moderate COVID-19 survivors with the lowest grade of lung computed tomography abnormalities</span> showed a similarly low levels of PSS or fatigue.
 Yet their IP was hallmarked by poor disease understanding and disbelief in personal or treatment control. 
@@
   footnote(1, 6:7, 
-           value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'),
+           value = as_paragraph('COVID-19 severity strata comparison; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic'),
            ref_symbols = 'a', 
@@
   width(7, width = 2.2, unit = 'cm') %>% 
-  footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>% 
+  footnote(1, 6:7, value = as_paragraph('COVID-19 severity strata comparison; Kruskal-Wallis test with \u03B7\u00B2 effect size statistic'), ref_symbols = 'a', part = 'header') %>% 
   theme_vanilla
@@
 
-_Three subsets of study participants (illness perception [IP] clusters) were identified by clustering in respect to the Brief Illness Perception Questionnaire items (BIPQ, Q1 - Q8) with PAM (partitioning around medoids) and Euclidean distance metric.
+_Three subsets of study participants (illness perception [IP] clusters) were identified by clustering in respect to the Brief Illness Perception Questionnaire items (BIPQ, Q1 - Q8) with the partitioning around medoids algorithm and Euclidean distance metric.
 Numbers of observations assigned to the clusters are displayed next to the plots or in the plot X axes._

### 2023-02-09 paper/markdown/supplementary_material.Rmd
Second revision
Minor Reviewer Issues, duplicated references removed

@@
 
-set_flextable_defaults(font.family = 'Cambria', font.size = 10)
+set_flextable_defaults(font.family = 'Cambria', font.size = 9)
 
@@
 The follow-up visits were scheduled at two, three, six months and one year after COVID-19 diagnosis. 
-In the current report, a subset of the original CovILD cohort was investigated, which displayed (1) persistent somatic symptoms (PSS) or (2) any abnormality in chest computed tomography (CT) or (4) any deficits in lung function testing (LFT) or (5) any cardiological abnormality in trans-thoracic echocardiography (TTE) at the one-year follow-up.
-The final analysis inclusion criterion was availability of complete Brief Illness Perception Questionnaire (IPQ) [@Broadbent2006] and the complete set of variables used in explorative data analysis, modeling and clustering (__Supplementary Table S\@ref(tab:tab-var-dict)__ and __S\@ref(tab:tab-modeling-vars)__).
+In the current report, a subset of the original CovILD cohort was investigated, which displayed (1) persistent somatic symptoms (PSS) or (2) any abnormality in chest computed tomography or (4) any deficits in lung function testing or (5) any cardiological abnormality in trans-thoracic echocardiography at the one-year follow-up.
+The final analysis inclusion criterion was availability of complete Brief Illness Perception Questionnaire (BIPQ) [@Broadbent2006] and the complete set of variables used in explorative data analysis, modeling and clustering (__Supplementary Table S\@ref(tab:tab-var-dict)__ and __S\@ref(tab:tab-modeling-vars)__).
 
-The decision to analyze the study participant subset with COVID-19 sequlae was motivated by the design of the illness perception-measuring tool used in our report, the Brief Illness Perception Questionnaire (BIPQ), which was initially developed for and validated in somatic conditions [@Broadbent2006].
+The decision to analyze the study participant subset with COVID-19 sequlae was motivated by the design of the illness perception-measuring tool used in our report, the BIPQ, which was initially developed for and validated in somatic conditions [@Broadbent2006].
 
@@
 
-The protocol of each follow-up visit included a survey of PSS, blood laboratory parameter assessment, LFT, TTE and chest CT as described in more detail before [@Sonnweber2020; @Sonnweber2022; @Luger2022].
+The protocol of each follow-up visit included a survey of PSS, blood laboratory parameter assessment, lung function testing, trans-thoracic echocardiography and chest computed tomography as described in more detail before [@Sonnweber2020; @Sonnweber2022; @Luger2022].
 Comorbidities and data on acute COVID-19 course (symptoms, laboratory parameters, functional tests and imaging) were recorded retrospectively at the two-month follow-up visit based on information provided by the participant and electronic patient records.
@@
 The following pre-existing comorbidities were surveyed: obesity, metabolic disease, cardiovascular disease, hypertension, type II diabetes, hypercholesterolemia, gastrointestinal disease, malignancy, chronic kidney disease, immune deficiency, pulmonary and respiratory conditions.
-Among the surveyed comorbidities, asthma, COPD (chronic obstructive pulmonary disease), interstitial lung disease and other chronic pulmonary disorders were subsumed under the 'respiratory disease' variable included in the independent modeling variable set.
+Among the surveyed comorbidities, asthma, chronic obstructive pulmonary disease, interstitial lung disease and other chronic pulmonary disorders were subsumed under the 'respiratory disease' variable included in the independent modeling variable set.
 Study participants were stratified according to the acute COVID-19 severity as ambulatory (outpatient, WHO ordinal scale for clinical improvement 1 - 2), moderate (hospitalized at normal infection ward, no oxygen therapy, WHO 3 - 4) and severe COVID-19 survivors (hospitalized with oxygen therapy or mechanical ventilation or intensive care, WHO 5 - 7).
@@
 
-LFT abnormality was defined as FVC (forced vital capacity) < 80% or FEV1 (forced expiratory volume in 1 second) < 80% or TLC (total lung capacity) < 80% or DLCO (diffusion lung capacity for carbon monoxide) < 80% of the predicted reference value or FEV1:FVC ratio < 70% of the predicted reference value [@Sonnweber2022].
-CT images were screened for ground glass opacities (GGO), consolidations, bronchiectasis and reticulations according to the Fleischner Society glossary [@Hansell2008]. 
-Chest CT abnormalities were scored separately for each lobe with the CT severity score (CTSS) [@Sonnweber2020; @Sonnweber2022; @Luger2022]: 0 - no abnormality, 1 - minimal (subtle GGO), 2 - mild (several GGO, subtle reticulation), 3 - moderate (multiple GGO, reticulation, small consolidation), 4 - severe (extensive GGO, consolidation, reticulation with distortion), 5 - massive (multiple findings, parenchymal destruction).
-The sum CTSS for all five lobes was used in the analysis. 
-Any chest CT abnormality is defined as CTSS $\geq$ 1. 
-The most frequent cardiological abnormality in TTE was low grade diastolic dysfunction. 
+Lung function testing abnormality was defined as 
+forced vital capacity < 80% or 
+forced expiratory volume in 1 second < 80% or 
+total lung capacity < 80% or 
+diffusion lung capacity for carbon monoxide < 80% of the predicted reference value or 
+the ratio of forced expiratory volume in 1 second to forced vital capacity < 70% of the predicted reference value [@Sonnweber2022]. 
+Computed tomography images were screened for ground glass opacities, consolidations, bronchiectasis and reticulations according to the Fleischner Society glossary [@Hansell2008]. 
+Chest computed tomography abnormalities were scored separately for each lobe with the computed tomography severity score [@Sonnweber2020; @Sonnweber2022; @Luger2022]: 0 - no abnormality, 1 - minimal (subtle ground glass opacities), 2 - mild (several ground glass opacities, subtle reticulation), 3 - moderate (multiple ground glass opacities, reticulation, small consolidation), 4 - severe (extensive ground glass opacities, consolidation, reticulation with distortion), 5 - massive (multiple findings, parenchymal destruction).
+The sum computed tomography severity score for all five lobes was used in the analysis. 
+Any chest computed tomography abnormality is defined as computed tomography severity score $\geq$ 1. 
+The most frequent cardiological abnormality in echocardiography was low grade diastolic dysfunction. 
 No reduced left ventricular ejection fraction was observed in the participants included in the current analysis. 
@@
 
-- __reduced physical performance__ assessed with the ECOG scale (Eastern Cooperative Oncology Group). Reduced physical performance was defined as ECOG $\geq$ 1
-- __dyspnea__ rated with the mMRC scale (Modified Medical British Research Council). Dyspnea was defined as mMRC $\geq$ 1
+- __reduced physical performance__ defined as Eastern Cooperative Oncology Group (ECOG) scale $\geq$ 1
+- __dyspnea__ defined as Modified Medical British Research Council (mMRC) scale $\geq$ 1
 - __cough__, self-reported, surveyed as a yes/no item
@@
 - __hair loss__, self-reported, surveyed as a yes/no item
-- __significant fatigue__, measured with the bimodal Chalder's Fatigue Scale (CFS). Significant fatigue was defined as bimodal CFS $\geq$ 4 [@Chalder1993; @Morriss1998]
+- __significant fatigue__, defined as bimodal Chalder's Fatigue Scale $\geq$ 4 [@Chalder1993; @Morriss1998]
 
@@
 
-Since only few participants displayed values of ECOG $\geq$ 2 or mMRC $\geq$ 2, the dichotomous 'reduced performance (ECOG $\geq$ 1)' and 'dyspnea (mMRC $\geq$ 1)' independent variables were used in modeling instead of the full numeric scales.
-Fatigue intensity was rated with the likert Chalder's Fatigue Scale (CFS) [@Chalder1993; @Morriss1998]. 
-Exertional capacity was assessed by six-minute walking distance (SMWD) test performed according to the American Thoracic Society guidelines [@Crapo2012]. 
-The reference SMWD values were obtained as described by Crapo and colleagues [@Crapo2012] and differences between the actual and reference values were calculated.
+Since only few participants displayed values of Eastern Cooperative Oncology Group scale $\geq$ 2 or Modified Medical British Research Council scale $\geq$ 2, the dichotomous 'reduced performance (ECOG $\geq$ 1)' and 'dyspnea (mMRC $\geq$ 1)' independent variables were used in modeling instead of the full numeric scales.
+Fatigue intensity was rated with the likert Chalder's Fatigue Scale [@Chalder1993; @Morriss1998]. 
+Exertional capacity was assessed by six-minute walking test performed according to the American Thoracic Society guidelines [@Crapo2012]. 
+The reference six-minute walking distance values were obtained as described by Crapo and colleagues [@Crapo2012] and differences between the actual and reference values were calculated.
 Frequencies of symptoms, ratings of fatigue and exertional capacity at the one-year follow-up are shown in __Supplementary Table S\@ref(tab:tab-cohort-fup)__, __Supplementary Figures S\@ref(fig:fig-sev-symptoms)__ and  __S\@ref(fig:fig-overlap)__.
@@
 
-Laboratory parameters encompassing hemoglobin (total and glycated hemoglobin [HbA1c]), biomarkers of iron turnover (ferritin [FT], transferrin saturation [TF-Sat], hepcidin and soluble transferrin receptor [sTFR]), inflammation and coagulation (C-reactive protein [CRP] and D-dimer) and cardiovascular pathology (N-terminal pro-brain natriuretic peptide [NT-proBNP]) were determined routinely at the certified laboratory of the Central Institute for Medical and Chemical Diagnostic at the University Hospital in Innsbruck.
-Other markers of systemic inflammation, interleukin 6 (IL6) and procalcitonin (PCT), were within the normal ranges in the analysis collective. 
+Laboratory parameters encompassing hemoglobin (total and glycated hemoglobin), biomarkers of iron turnover (ferritin, transferrin saturation, hepcidin and soluble transferrin receptor), inflammation and coagulation (C-reactive protein and D-dimer) and cardiovascular pathology (N-terminal pro-brain natriuretic peptide) were determined routinely at the certified laboratory of the Central Institute for Medical and Chemical Diagnostic at the University Hospital in Innsbruck.
+Other markers of systemic inflammation, interleukin 6 and procalcitonin, were within the normal ranges in the analysis collective. 
 Values of laboratory parameters at the one-year follow-up are listed in __Supplementary Table S\@ref(tab:tab-cohort-biochemistry)__.
@@
 
-Illness perception (IP) was scored with a German translation of the Brief Illness Perception Questionnaire (BIPQ) [@Broadbent2006]. 
+Illness perception (IP) was scored with a German translation of BIPQ [@Broadbent2006]. 
 Each of the 8 BIPQ items (questions Q1 - Q8) was scored in with an 11-point likert scale, the items Q3, Q4 and Q7 were conceptualized as negative items are were inverted prior to the analysis:
@@
 As shown by the results of factor analysis [@Bartlett1937] of the BIPQ tool items (__Supplementary Figure S\@ref(fig:fig-factor-analysis)__), two separate factors could be identified: the emotion/concern/consequences (items Q1, Q2, Q5, Q6, Q8) and the lacking control/coherence component (items Q3, Q4, Q7).
-For this reason, two IP component scores were calculated and analysed as well: the emotion/concern/consequences score defined as the Q1, Q2, Q5, Q6, Q8 item sum and the lacking control/coherence score defined as the Q3, Q4, Q7 item sum.
+For this reason, two IP component scores were calculated and analyzed: the emotion/concern/consequences score defined as the Q1, Q2, Q5, Q6, Q8 item sum and the lacking control/coherence score defined as the Q3, Q4, Q7 item sum.
 Values of the total IP score, the emotion/concern/consequences and lacking control/coherence IP component scores as well as values of the BIPQ items recorded at the one-year follow-up are presented in __Table 2__.
@@
 
-For descriptive statistic, correlation and statistical testing for differences between analysis groups the package _rstatix_ [@Kassambara2021] and the development package _ExDA_ were used (https://github.com/PiotrTymoszuk/ExDA) were employed.
-Factor analysis was accomplished with the _stats_ package of base R and the development package _clustTools_ (https://github.com/PiotrTymoszuk/clustTools). 
+For descriptive statistic, correlation and statistical testing for differences between analysis groups the package _rstatix_ [@Kassambara2021] and the development package [_ExDA_](https://github.com/PiotrTymoszuk/ExDA) were employed.
+Factor analysis was accomplished with the _stats_ package of base R and the development package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools). 
 The BIPQ score consistency was investigated with the _psych_ package [@Revelle2022].
 
-For modeling and ANOVA, the development package _lmqc_ (https://github.com/PiotrTymoszuk/lmqc, backward elimination and two-way ANOVA) and the packages _glmnet_ (Elastic Net and LASSO) [@Friedman2010], _monomvn_ [@Gramacy2022] (Bayesian LASSO) and _caret_ [@Kuhn2008] were employed.
-Model quality control, cross-validation and fit statistics were computed with _caret_ and the development package _caretExtra_ (https://github.com/PiotrTymoszuk/caretExtra).
+For modeling and ANOVA, the development package [_lmqc_](https://github.com/PiotrTymoszuk/lmqc, backward elimination and two-way ANOVA) and the packages _glmnet_ (Elastic Net and LASSO) [@Friedman2010], _monomvn_ [@Gramacy2022] (Bayesian LASSO) and _caret_ [@Kuhn2008] were employed.
+Model quality control, cross-validation and fit statistics were computed with _caret_ and the development package [_caretExtra_](https://github.com/PiotrTymoszuk/caretExtra).
 
-Clustering was accomplished with tools provided by the packages _cluster_ [@Schubert2019], _philentropy_ [@Drost2018], _factoextra_ [@Kassambara2020] and the development package _clustTools_ (https://github.com/PiotrTymoszuk/clustTools).
+Clustering was accomplished with tools provided by the packages _cluster_ [@Schubert2019], _philentropy_ [@Drost2018], _factoextra_ [@Kassambara2020] and the development package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools).
 
 The _ggplot2_ [@Wickham2016] package were used for data visualization. 
-Bar and violin plots were generated with _ExDA_. 
+Bar and violin plots were generated with [_ExDA_](https://github.com/PiotrTymoszuk/ExDA). 
 Quasi-proportional Venn diagrams were generated with the _nVennR_ package [@Perez-Silva2018]. 
@@
 Tables were generated with _flextable_ [@Gohel2022]. 
-The manuscript and supplementary material files were created with the _rmarkdown_ [@Allaire2022], _knitr_ [@Xie2022], _bookdown_ [@Xie2016] packages and _Affiliations_ Pandoc filter set (https://github.com/drdanholmes/Affiliations).
-For management of figures, tables and R code in the RMarkdown documents, the development package _figur_ (https://github.com/PiotrTymoszuk/figur) was used.
+The manuscript and supplementary material files were created with the _rmarkdown_ [@Allaire2022], _knitr_ [@Xie2022], _bookdown_ [@Xie2016] packages and [_Affiliations_](https://github.com/drdanholmes/Affiliations) Pandoc filter set.
+For management of figures, tables and R code in the RMarkdown documents, the development package [_figur_](https://github.com/PiotrTymoszuk/figur) was used.
 
@@
 
-Descriptive statistic: median with interquartile ranges for numeric variables, and percentages and counts for categorical variables were calculated with the `explore()` function (_ExDA_ package). 
-Statistical significance of differences in frequency of categorical variable levels was assessed by $\chi^2$ test with Cramer V effect size statistic (`compare_variables()`, the _ExDA_ package). 
-Normality was tested with Shapiro-Wilk test and quantile-quantile plots (`explore()`, _ExDA_ package).
+Descriptive statistic: median with interquartile ranges for numeric variables, and percentages and counts for categorical variables were calculated with the `explore()` function ([_ExDA_](https://github.com/PiotrTymoszuk/ExDA) package).
+Statistical significance of differences in frequency of categorical variable levels was assessed by $\chi^2$ test with Cramer V effect size statistic (`compare_variables()`, the [_ExDA_](https://github.com/PiotrTymoszuk/ExDA) package).
+Normality was tested with Shapiro-Wilk test and quantile-quantile plots (`explore()`, [_ExDA_](https://github.com/PiotrTymoszuk/ExDA) package).
 
 Since multiple study variables were non-normally distributed, differences in numeric variables were assessed by non-parametric Mann-Whitney (effect size: r) or Kruskal-Wallis test (effect size: $\eta^2$), as appropriate.
-Group comparisons were done with the `compare_variables()` function (package _ExDA_).
-Correlations were analyzed with Spearman test (`correlate_variables()` package _ExDA_) and visualized by fitting a second order trend (`geom_smooth()`, package _ggplot_, formula: $y \sim x + I(x^2)$).
+Group comparisons were done with the `compare_variables()` function (package [_ExDA_](https://github.com/PiotrTymoszuk/ExDA)).
+Correlations were analyzed with Spearman test (`correlate_variables()` package [_ExDA_](https://github.com/PiotrTymoszuk/ExDA)) and visualized by fitting a second order trend (`geom_smooth()`, package _ggplot_, formula: $y \sim x + I(x^2)$).
 Statistical significance for changes in frequency of PSS in time was investigated with Cochran Q test (function `cochran_qtest()`, package _rstatix_) [@Kassambara2021].
@@
 
-BIPQ dimensionality was assessed by maximum  likelihood factor analysis [@Bartlett1937] (`reduce_data(red_fun = 'fa')`, package _clustTools_, a wrapper around `factanal()` from _stats_) and visual investigation of the BIPQ item loadings (method `plot(type = 'loadings')` called for the factor analysis object).
+BIPQ dimensionality was assessed by maximum  likelihood factor analysis [@Bartlett1937] (`reduce_data(red_fun = 'fa')`, package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools), a wrapper around `factanal()` from _stats_) and visual investigation of the BIPQ item loadings (method `plot(type = 'loadings')` called for the factor analysis object).
 The factor analysis revealed presence of two components of the BIPQ tool: the emotion/concern/consequences and the lacking control/coherence factor (__Figure \@ref(fig:fig-factor-analysis)__).
-Spearman's pair-wise correlations between the BIPQ items were calculated (function 'correlate_variables()', package _ExDA_) and visualized in a bubble plot. 
+Spearman's pair-wise correlations between the BIPQ items were calculated (function `correlate_variables()`, package [_ExDA_](https://github.com/PiotrTymoszuk/ExDA)) and visualized in a bubble plot. 
 Due to multi-dimensionality of the BIPQ tool McDonald's $\omega$ was calculated with the function `omega()` from the package _psych_ [@Revelle2022; @McDonald1999] was used as a reliability measure for the global BIPQ and the emotion/concern/consequences and lacking control/coherence factors (__Supplementary Figure S\@ref(fig:fig-coherence)__).
@@
 In multi-parameter modeling, a set of `r length(mod$variables)` independent variables was used (__Supplementary Table S\@ref(tab:tab-modeling-vars)__). 
-To improve normality, FT and sTFR were transformed with the logarithm function, and TF-Sat and hepcidin were transformed with square root. 
+To improve normality, ferritin and soluble transferrin receptor concentrations were transformed with the logarithm function, and transferrin saturation and hepcidin were transformed with square root. 
 Strongly non-normally or bimodally distributed numeric explanatory variables: 
@@
 number of symptoms (`r levels(cov_data$clear_data$sympt_number_class) %>% paste(collapse = ', ')`), 
-CTSS ('none': CTSS 0, 'mild': 1 - 5, 'moderate': 5 - 10, 'severe': 11+) [@Luger2022], 
-NT-proBNP (`r levels(cov_data$clear_data$NTproBNP_class) %>% paste(collapse = ', ')`), 
+computed tomography severity score ('none': 0, 'mild': 1 - 5, 'moderate': 5 - 10, 'severe': 11+) [@Luger2022], 
+N-terminal pro-brain natriuretic peptide (`r levels(cov_data$clear_data$NTproBNP_class) %>% paste(collapse = ', ')`), 
 D-Dimer (`r levels(cov_data$clear_data$DDimer_class) %>% paste(collapse = ', ')`), 
-CRP (`r levels(cov_data$clear_data$DDimer_class) %>% paste(collapse = ', ')`) and 
-HbA1c (`r levels(cov_data$clear_data$DDimer_class) %>% paste(collapse = ', ')`) were stratified with clinically applicable cutoffs or cutoffs determined by visual analysis of the plots of modeling variable and the explanatory factor with fitted LOESS trends.
+C-reactive protein (`r levels(cov_data$clear_data$DDimer_class) %>% paste(collapse = ', ')`) and 
+glycated hemoglobin (`r levels(cov_data$clear_data$DDimer_class) %>% paste(collapse = ', ')`) were stratified with clinically applicable cutoffs or cutoffs determined by visual analysis of scatter plots of modeling variable and the explanatory factor with fitted LOESS trends.
 For numeric explanatory variables, both first- and second-order terms were included in the model to account for possible non-linear relationship between the independent variable and response. 
@@
 The choice of penalized regression over canonical backward elimination or forward selection was motivated by highly multi-dimensional nature of the data set [@Kirpich2018]. 
-As expected, modeling of the IP responses with canonical linear regression with Akaike information criterion-driven backward elimination (function `stepAIC()`, package MASS) [@Ripley2022] led to highly over-parameterized models as evident from huge expansion of the fit errors in cross-validation (CV) (not shown).
+As expected, modeling of the IP responses with canonical linear regression with Akaike information criterion-driven backward elimination (function `stepAIC()`, package MASS) [@Ripley2022] led to highly over-parameterized models as evident from huge expansion of the fit errors in cross-validation (not shown).
 
@@
 `r filter(lasso_mod$opt_lambda, response == "ipq_sub2")$lambda %>% signif(2)` for the lacking control/coherence IP component score. 
-Elastic Net and LASSO models with the optimal $\lambda$ values were constructed with _caret_ package (function `train(method = 'glmnet')`) [@Kuhn2008] and their performance in the training data set and 10-repeats 10-fold CV was tested.
-As shown in __Supplementary Figure S\@ref(fig:fig-performance)__, for the total IP score and the emotion/concern/consequences IP component score, meaningful, reproducible and acceptably parameterized models could be built with either of the algorithms as suggested by comparison of the model performance in the training and CV data sets.
+Elastic Net and LASSO models with the optimal $\lambda$ values were constructed with _caret_ package (function `train(method = 'glmnet')`) [@Kuhn2008] and their performance in the training data set and 10-repeats 10-fold cross-validation was tested.
+As shown in __Supplementary Figure S\@ref(fig:fig-performance)__, for the total IP score and the emotion/concern/consequences IP component score, meaningful, reproducible and acceptably parameterized models could be built with either of the algorithms as suggested by comparison of the model performance in the training and the cross-validation data sets.
 In turn, poorly performing, estimate-only models could be obtained for the lacking control/coherence IP component. 
-Estimates of non-zero coefficients were extracted from the final models with the _coef_ method (the `s` parameter set to the optimal $\lambda$ value) (__Supplementary Figures S\@ref(fig:fig-eln-est-total)__ - __S\@ref(fig:fig-blasso-est-sub1)__).
+Estimates of non-zero coefficients were extracted from the final models with the `coef()` method (the `s` parameter set to the optimal $\lambda$ value) (__Supplementary Figures S\@ref(fig:fig-eln-est-total)__ - __S\@ref(fig:fig-blasso-est-sub1)__).
 
@@
 The Bayesian LASSO model was constructed with the _caret_ package (function `train(method = 'blasso')`, iteration number: T = 1000). 
-The optimal value of the sparsity parameter, which controls the fraction of non-zero coefficients in the final model based on the posterior distributions, was found by 10-repeats 10-fold CV with the RMSE (root mean squared error) statistic as a selection criterion.
+The optimal value of the sparsity parameter, which controls the fraction of non-zero coefficients in the final model based on the posterior distributions, was found by 10-repeats 10-fold cross-validation with the root mean squared error as a selection criterion.
 The optimal sparsity values selected from the 0.1 - 0.7 range were `r blass_mod$models$ipq_total$bestTune[1, "sparsity"] %>% signif(2)`, `r blass_mod$models$ipq_sub1$bestTune[1, "sparsity"] %>% signif(2)` and `r blass_mod$models$ipq_sub2$bestTune[1, "sparsity"] %>% signif(2)` for the total IP score, the emotion/concern/consequences and the lacking control/coherence IP component scores, respectively.
 Similar to the Elastic Net and LASSO algorithms, meaningful models could be obtained for the total IP score and the emotion/concern/consequences IP component score but not for the lacking control/coherence IP component score (__Supplementary Figure S\@ref(fig:fig-performance)__).
-The model coefficient matrix X was extracted from the 'finalModel' slot of the caret model. 
+The model coefficient matrix X was extracted from the `finalModel` slot of the caret model. 
 The coefficient values (__Supplementary Figures S\@ref(fig:fig-blasso-est-total)__ and __S\@ref(fig:fig-blasso-est-sub1)__) were calculated as matrix column medians over all algorithm iterations [@Park2012].
 
-Extraction of caret model performance measures: RMSE and $R^2$ defined as $1 - MSE/Var(y)$ (__Supplementary Figure S\@ref(fig:fig-performance)__) was accomplished by the in-house-developed package _caretExtra_ (modeling objects created with `as_caretex()`).
-Model assumption testing (normality of residuals, Shapiro-Wilk test, method `residuals()`) and visual quality control based on plots of residuals versus fitted and residuals quantile-quantile plots was done with the _caretExtra_ package (method `plot()` called for the `caretex` object).
+Extraction of caret model performance measures: root mean squared error and $R^2$ defined as $1 - mean\;squared\;error/variance(y)$ (__Supplementary Figure S\@ref(fig:fig-performance)__) was accomplished by the in-house-developed package [_caretExtra_](https://github.com/PiotrTymoszuk/caretExtra) (modeling objects created with `as_caretex()`).
+Model assumption testing (normality of residuals, Shapiro-Wilk test, method `residuals()`) and visual quality control based on plots of residuals versus fitted and residuals quantile-quantile plots was done with the [_caretExtra_](https://github.com/PiotrTymoszuk/caretExtra) package (method `plot()` called for the `caretex` object).
 
@@
 
-Interaction between the key factors affecting the total IP score and the emotion/concern/consequences IP component score and presence of cardiopulmonary abnormalities (CT, LFT findings or diastolic dysfunction in TTE) was analyzed by two-way ANOVA.
-Briefly, the interaction ANOVA models were constructed with the function `make_lm()` (employing `lm()` from base R, package _lmqc_). 
+Interaction between the key factors affecting the total IP score and the emotion/concern/consequences IP component score and presence of cardiopulmonary abnormalities (computed tomography, lung function findings or diastolic dysfunction in echocardiography) was analyzed by two-way ANOVA.
+Briefly, the interaction ANOVA models were constructed with the function `make_lm()` (employing `lm()` from base R, package [_lmqc_](https://github.com/PiotrTymoszuk/lmqc)). 
 The dependent variables were square root-transformed to improve normality. 
 The assumptions of normality and homogeneity of distribution of the model residuals were checked by Shapiro-Wilk and Levene tests, respectively (method `summary(type = 'assumptions')`) and standard diagnostic plots of model residuals (method `plot()`).
-ANOVA results with the $\eta^2$ effect size statistic (fraction of variance explained by the specific ANOVA term) were extracted from the models with the `anova()` method (package _lmqc_). 
+ANOVA results with the $\eta^2$ effect size statistic (fraction of variance explained by the specific ANOVA term) were extracted from the models with the `anova()` method (package [_lmqc_](https://github.com/PiotrTymoszuk/lmqc)).
 The significant and near significant (p < 0.1) interactions are presented in __Supplementary Figures S\@ref(fig:fig-interaction-total)__ - __S\@ref(fig:fig-interaction-sub1-lufo-dia)__.
@@
 
-Study participants were clustered in respect to the BIPQ items by PAM (partition around medoids, originally implemented by the package _cluster_) [@Schubert2019] with Euclidean distance between the observations (originally implemented by the package _philentropy_) [@Drost2018].
+Study participants were clustered in respect to the BIPQ items by the partition around medoids algorithm (originally implemented by the package _cluster_) [@Schubert2019] with Euclidean distance between the observations (originally implemented by the package _philentropy_) [@Drost2018].
 The BIPQ item scores were not pre-processed, since they were measured with the same likert scale. 
-Technically, the PAM clustering object was constructed with the `kcluster(distance_method = 'euclidean', clust_fun = 'pam')` function from the development package _clustTools_.
+Technically, the clustering object was constructed with the `kcluster(distance_method = 'euclidean', clust_fun = 'pam')` function from the development package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools).
 
-The PAM/Euclidean distance procedure was chosen based on the high explained clustering variance (ratio of the between-cluster sum of squares to the total sum of squares, method `var()`, package _clustTools_) and stability (cluster assignment accuracy in 10-fold CV, cluster assignment by 7-nearest neighbors classifier, method `cv()`, package _clustTools_) [@Lange2004; @Leng2014] as compared with several other clustering algorithms (hierarchical clustering and KMEANS) (__Supplementary Figure S\@ref(fig:fig-clust-dev)A__).
-The choice of cluster number (k = `r ipq_clust$clust_obj$clust_assignment$clust_id %>% levels %>% length`) was based on the bend of the curve of within-cluster sum of squares (method `plot()`, package _clustTools_, a wrapper around `fviz_nbclust()`, package _factoextra_) [@Kassambara2020] and visual analysis of the distance heat map (__Supplementary Figure S\@ref(fig:fig-clust-dev)BC__).
+The partition around medoids/Euclidean distance procedure was chosen based on the high explained clustering variance (ratio of the between-cluster sum of squares to the total sum of squares, method `var()`, package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools)) and stability (cluster assignment accuracy in 10-fold cross-validation, cluster assignment by 7-nearest neighbors classifier, method `cv()`, package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools)) [@Lange2004; @Leng2014] as compared with several other clustering algorithms (hierarchical clustering and KMEANS) (__Supplementary Figure S\@ref(fig:fig-clust-dev)A__).
+The choice of cluster number (k = `r ipq_clust$clust_obj$clust_assignment$clust_id %>% levels %>% length`) was based on the bend of the curve of within-cluster sum of squares (method `plot()`, package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools), a wrapper around `fviz_nbclust()`, package _factoextra_) [@Kassambara2020] and visual analysis of the distance heat map (__Supplementary Figure S\@ref(fig:fig-clust-dev)BC__).
 
@@
   footnote(1, 6:7, 
-           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic'), 
            ref_symbols = 'a', 
@@
   footnote(1, 5:6, 
-           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03C7\u00B2 effect size statistic'), 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic'), 
            ref_symbols = 'a', 
@@
 The data points are labeled with their $\beta$ values.
-CFS: Chalder's fatigue score; reduced performance: ECOG (Eastern Cooperative Oncology Group) $\geq$ 1; # PSS: number of persistent somatic symptoms at the one-year follow-up; CT lesion severity: chest computed tomography abnormality severity at the one-year follow-up; GID: gastrointestinal disease; CRP: serum C-reactive protein; COVID-19 severity: HS: hospitalized/severe; NT-proBNP; N-terminal pro-brain natriuretic peptide._
+CFS: Chalder's fatigue score; reduced performance: ECOG (Eastern Cooperative Oncology Group) $\geq$ 1; # PSS: number of persistent somatic symptoms at the one-year follow-up; CT lesion severity: chest computed tomography abnormality severity at the one-year follow-up; GID: gastrointestinal disease; CRP: serum C-reactive protein; COVID-19 severity: HS: hospitalized/severe; NT-proBNP: N-terminal pro-brain natriuretic peptide._
 

### 2023-02-09 analysis.R

@@ -11,7 +11,11 @@
 # hypothesis testing and analysis of their interaction with cardiopulmonary
 # abnormalities (two-way ANOVA)
 #
-# 3) Unsupervised clustering of the study participants in respect to 
+# 3) Unsupervised clustering of the study participants in respect 
+# to the BIPQ items
+#
+# 4) Analysis of effects of the interaction between persistent somatic 
+# symptoms and cardiopulmonary findings on IP; two-way ANOVA
 
 
 # tools ------

### 2023-02-22 paper/markdown/supplementary_material.Rmd
AMA format of p values

@@
 ANOVA results with the $\eta^2$ effect size statistic (fraction of variance explained by the specific ANOVA term) were extracted from the models with the `anova()` method (package [_lmqc_](https://github.com/PiotrTymoszuk/lmqc)).
-The significant and near significant (p < 0.1) interactions are presented in __Supplementary Figures S\@ref(fig:fig-interaction-total)__ - __S\@ref(fig:fig-interaction-sub1-lufo-dia)__.
+The significant and near significant (p < .1) interactions are presented in __Supplementary Figures S\@ref(fig:fig-interaction-total)__ - __S\@ref(fig:fig-interaction-sub1-lufo-dia)__.
 
@@
 
-```{r tab-excluded, tab.cap = 'Significant and near-significant (p < 0.1) factors differentiating between the CovILD cohort participants included in the analysis and excluded due to variable missingness or lacking long-term COVID-19 sequelae.
+```{r tab-excluded, tab.cap = 'Significant and near-significant (p < .1) factors differentiating between the CovILD cohort participants included in the analysis and excluded due to variable missingness or lacking long-term COVID-19 sequelae.
 Numeric variables are presented as medians with interquartile ranges (IQR).
@@
 
-```{r tab-gender, tab.cap = 'Significant and near-significant (p < 0.1) factors differentiating between female and male participants and illness perception in the genders (Brief Illness Perception Questionnaire, BIPQ).
+```{r tab-gender, tab.cap = 'Significant and near-significant (p < .1) factors differentiating between female and male participants and illness perception in the genders (Brief Illness Perception Questionnaire, BIPQ).
 Numeric variables are presented as medians with interquartile ranges (IQR).
@@
 
-```{r tab-respi, tab.cap = 'Significant and near-significant (p < 0.1) factors differentiating between participants with and without respiratory comorbidity.
+```{r tab-respi, tab.cap = 'Significant and near-significant (p < .1) factors differentiating between participants with and without respiratory comorbidity.
 Numeric variables are presented as medians with interquartile ranges (IQR).
@@
 
-```{r tab-ipq-clusters, tab.cap = 'Significant and near-significant (p < 0.1) differences between the illness perception clusters.
+```{r tab-ipq-clusters, tab.cap = 'Significant and near-significant (p < .1) differences between the illness perception clusters.
 Numeric variables are presented as medians with interquartile ranges (IQR).
@@
 _Interaction effects between the key factors affecting the total illness perception (IP) score and cardiopulmonary abnormalities (A: chest computed tomography [CT] abnormality, B: lung function testing [LFT] abnormality, C: diastolic dysfunction) at the one-year follow-up were investigated by two-way ANOVA with $\eta^2$ effect size statistic.
-The key factors with significant and near-significant interaction effects (p < 0.1) are presented.
+The key factors with significant and near-significant interaction effects (p < .1) are presented.
 The associations with fatigue scoring are presented in a point plots; the black line represents the fitted linear trend and the tinted ribbon depicts the 95% confidence intervals.
@@
 _Interaction effects between the key factors affecting the emotion/concern/consequences IP component score and chest computed tomography (CT) abnormalities at the one-year follow-up were investigated by two-way ANOVA with $\eta^2$ effect size statistic.
-The key factors with significant and near-significant interaction effects (p < 0.1) are presented.
+The key factors with significant and near-significant interaction effects (p < .1) are presented.
 The associations with fatigue scoring are presented in a point plots; the black line represents the fitted linear trend and the tinted ribbon depicts the 95% confidence intervals.
@@
 _Interaction effects between the key factors affecting the emotion/concern/consequences IP component score and cardiopulmonary abnormalities (A: lung function testing [LFT] abnormality, B: diastolic dysfunction) at the one-year follow-up were investigated by two-way ANOVA with $\eta^2$ effect size statistic.
-The key factors with significant and near-significant interaction effects (p < 0.1) are presented.
+The key factors with significant and near-significant interaction effects (p < .1) are presented.
 The associations with fatigue scoring are presented in a point plots; the black line represents the fitted linear trend and the tinted ribbon depicts the 95% confidence intervals.

### 2023-02-22 analysis scripts/clust_characteristic.R

@@ -59,6 +59,7 @@
                 pub_styled = TRUE, 
                 adj_method = 'none', 
                 .options = furrr_options(seed = TRUE)) %>% 
+    map(format_test_jps) %>% 
     map(mutate, 
         plot_cap = paste(eff_size, significance, sep = ', '))
   

### 2023-02-22 analysis scripts/interaction.R

@@ -295,14 +295,8 @@
                                        translate_var(variable)), 
                       eff_size = paste('\u03B7\u00B2 = ', 
                                        signif(frac_explained, 2)), 
-                      significance = ifelse(`Pr(>F)` > 0.1, 
-                                            'ns', 
-                                            ifelse(`Pr(>F)` >= 0.05, 
-                                                   paste0('ns (p = ', 
-                                                          signif(`Pr(>F)`, 2), 
-                                                          ')'), 
-                                                   paste('p =', 
-                                                         signif(`Pr(>F)`, 2)))), 
+                      eff_size = eff_size_jps(eff_size), 
+                      significance = p_value_jps(`Pr(>F)`), 
                       plot_cap = paste(eff_size, significance))))
   
   inter[c('plot_tags_ct', 

### 2023-02-22 analysis scripts/ipq_clustering.R

@@ -121,6 +121,7 @@
                       types = ipq_clust$clust_lexicon$test_type, 
                       ci = FALSE, 
                       pub_styled = TRUE) %>% 
+    format_test_jps %>% 
     mutate(plot_cap = paste(eff_size, significance, sep = ', '))
 
   ## plot labels for the ribbon plot

### 2023-02-22 analysis scripts/ipq_factors.R

@@ -54,6 +54,7 @@
              ci = FALSE, 
              exact = FALSE, 
              pub_styled = TRUE) %>% 
+    format_test_jps %>% 
     mutate(split_factor = ipq_factors$variables$eff_size$variable, 
            plot_cap = paste(eff_size, significance, sep = ', '))
   
@@ -67,12 +68,9 @@
                                  type = 'spearman', 
                                  ci = FALSE, 
                                  pub_styled = TRUE)) %>% 
-    safely(mutate)(plot_cap = paste(eff_size, significance, sep = ', '), 
-                   plot_cap = stri_replace(plot_cap, 
-                                           fixed = 'rho', 
-                                           replacement = '\u03C1')) %>% 
-    .$result
-  
+    format_test_jps(correlation = TRUE) %>% 
+    mutate(plot_cap = paste(eff_size, significance, sep = ', '))
+
 # Plotting the comparisons ------
   
   insert_msg('Violin plots')

### 2023-02-22 analysis scripts/ipq_sub1_factors.R

@@ -55,6 +55,7 @@
              ci = FALSE, 
              exact = FALSE, 
              pub_styled = TRUE) %>% 
+    format_test_jps %>% 
     mutate(split_factor = sub1_factors$variables$eff_size$variable, 
            plot_cap = paste(eff_size, significance, sep = ', '))
   
@@ -68,12 +69,9 @@
                                  type = 'spearman', 
                                  ci = FALSE, 
                                  pub_styled = TRUE)) %>% 
-    safely(mutate)(plot_cap = paste(eff_size, significance, sep = ', '), 
-                   plot_cap = stri_replace(plot_cap, 
-                                           fixed = 'rho', 
-                                           replacement = '\u03C1')) %>% 
-    .$result
-  
+    format_test_jps(correlation = TRUE) %>% 
+    mutate(plot_cap = paste(eff_size, significance, sep = ', '))
+
 # Plotting the comparisons ------
   
   insert_msg('Violin plots')

### 2023-02-22 analysis scripts/ipq_sub2_factors.R

@@ -57,6 +57,7 @@
              ci = FALSE, 
              exact = FALSE, 
              pub_styled = TRUE) %>% 
+    format_test_jps %>% 
     mutate(split_factor = sub2_factors$variables$eff_size$variable, 
            plot_cap = paste(eff_size, significance, sep = ', '))
   
@@ -70,11 +71,8 @@
                                  type = 'spearman', 
                                  ci = FALSE, 
                                  pub_styled = TRUE)) %>% 
-    safely(mutate)(plot_cap = paste(eff_size, significance, sep = ', '), 
-                   plot_cap = stri_replace(plot_cap, 
-                                           fixed = 'rho', 
-                                           replacement = '\u03C1')) %>% 
-    .$result
+    format_test_jps(correlation = TRUE) %>% 
+    mutate(plot_cap = paste(eff_size, significance, sep = ', '))
   
 # Plotting the comparisons ------
   

### 2023-02-22 exploration scripts/bierbauer2022.R

@@ -230,7 +230,8 @@
                       types = 'kruskal_test', 
                       exact = FALSE, 
                       ci = FALSE, 
-                      pub_styled = TRUE)
+                      pub_styled = TRUE) %>% 
+    format_test_jps
   
   lc_ipq$ax_labs <- lc_ipq$test_results %>% 
     mutate(ax_lab = paste(variable, significance, 
@@ -321,7 +322,7 @@
           axis.text.x = element_text(hjust = 0, 
                                      vjust = 0.5, 
                                      angle = -90)) + 
-    labs(title = 'Correlation analysis, IPQ, bierbauer et al. 2022', 
+    labs(title = 'Correlation analysis, IPQ, Bierbauer et al. 2022', 
          subtitle = 'Spearman correlation')
   
 # END ------

### 2023-02-22 exploration scripts/cohort_characteristic.R

@@ -56,6 +56,7 @@
                 pub_styled = TRUE, 
                 adj_method = 'none', 
                 .options = furrr_options(seed = TRUE)) %>% 
+    map(format_test_jps) %>% 
     map(mutate, 
         plot_cap = paste(eff_size, significance, sep = ', '))
   

### 2023-02-22 exploration scripts/gender.R

@@ -61,6 +61,7 @@
                 pub_styled = TRUE, 
                 adj_method = 'none', 
                 .options = furrr_options(seed = TRUE)) %>% 
+    map(format_test_jps) %>% 
     map(mutate, 
         plot_cap = paste(eff_size, significance, sep = ', '))
   

### 2023-02-22 exploration scripts/included_excluded.R

@@ -73,6 +73,7 @@
                 pub_styled = TRUE, 
                 adj_method = 'none', 
                 .options = furrr_options(seed = TRUE)) %>% 
+    map(format_test_jps) %>% 
     map(mutate, 
         plot_cap = paste(eff_size, significance, sep = ', '))
   

### 2023-02-22 exploration scripts/respiratory.R

@@ -57,6 +57,7 @@
                 pub_styled = TRUE, 
                 adj_method = 'none', 
                 .options = furrr_options(seed = TRUE)) %>% 
+    map(format_test_jps) %>% 
     map(mutate, 
         plot_cap = paste(eff_size, significance, sep = ', '))
   

### 2023-02-22 exploration scripts/symptom_kinetic.R

@@ -149,9 +149,7 @@
   sympt_kinet$plot_legend <- sympt_kinet$cohran_test %>% 
     map(mutate, 
         var_lab = translate_var(.y.), 
-        significance = ifelse(p >= 0.05, 
-                              paste0('ns (p = ', signif(p, 2), ')'), 
-                              paste('p =', signif(p, 2))), 
+        significance = p_value_jps(p), 
         var_lab = paste(var_lab, significance, sep = ', '))
   
   sympt_kinet$plot_legend <- sympt_kinet$plot_legend %>% 

### 2023-02-22 paper scripts/supplement.R

@@ -94,7 +94,9 @@
   
   insert_msg('Figure S4: BIPQ consistency')
   
-  suppl_fig$coherence <- ipq_co$corr_summ_plot %>% 
+  suppl_fig$coherence <- plot_grid(ipq_co$corr_summ_plot + 
+                                     guides(alpha = 'none', 
+                                            color = 'none')) %>% 
     as_figure(label = 'figure_s4_score_coherence',
               ref_name = 'coherence', 
               caption = 'Coherence of the illness perception score.', 

### 2023-02-22 tools/project_tools.R

@@ -672,5 +672,134 @@
       c(complete = n_complete))
   
   }
+  
+  p_value_jps <- function(x) {
+    
+    ## formats p values according 
+    ## to the style of J. Psychosomatic Research
+    
+    if(length(x) > 1) {
+      
+      return(map_chr(x, p_value_jps))
+      
+    }
+    
+    stopifnot(is.numeric(x))
+    
+    if(x < 0.001) {
+      
+      return('p < .001')
+      
+    }
+    
+    if(x > 0.99) {
+      
+      return('ns (p > .99)')
+      
+    }
+    
+    if(x >= 0.05) {
+      
+      new_x <- as.character(round(x, 2))
+      
+      new_x <- stri_replace(new_x, 
+                            fixed = '0.', 
+                            replacement = '.')
+      
+      if(stri_length(new_x) == 2) new_x <- paste0(new_x, '0')
+      
+      return(paste0('ns (p = ', 
+                    new_x, ')'))
+      
+    }
+    
+    if(x < 0.01) {
+      
+      new_x <- round(x, 3)
+      
+      if(new_x == 0.01) new_x <- paste0(as.character(new_x), '0')
+      
+      new_x <- as.character(new_x)
+      
+    } else if(x < 0.05) {
+      
+      new_x <- round(x, 2)
+      
+      if(new_x == 0.05) new_x <- round(x, 3)
+      
+    }
+    
+    new_x <- as.character(new_x)
+    
+    new_x <- stri_replace(new_x, 
+                          fixed = '0.', 
+                          replacement = '.')
+    
+    return(paste('p =', new_x))
+    
+    
+  }
+  
+  eff_size_jps <- function(x) {
+    
+    ## formats the effect size value according to the JPS style
+    ## takes a text entry and formats its numeric part
+    
+    stopifnot(is.character(x))
+    
+    effect_txt <- stri_extract(x, regex = '(V|r|d|(\u03B7\u00B2))\\s{1}=')
+    
+    effect_digits <- stri_extract(x, regex = '\\d{1}\\.\\d+')
+    
+    effect_digits <- as.numeric(effect_digits)
+    
+    effect_digits <- ifelse(effect_digits < 0, 0, effect_digits)
+    
+    effect_digits <- as.character(effect_digits)
+    
+    effect_digits <- stri_replace(effect_digits, 
+                                  fixed = '0.', 
+                                  replacement = '.')
+    
+    return(paste(effect_txt, effect_digits))
+    
+  }
+  
+  corr_jps <- function(x) {
[…]

### 2023-03-06 paper/markdown/manuscript.Rmd
Change of the manuscript title

@@
 ---
-title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19 in a cross-sectional prospective cohort study"
+title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19 in a cross-sectional analysis of a prospective cohort study"
 subtitle: "Illness perception one year after COVID-19"

### 2023-03-06 paper/markdown/supplementary_material.Rmd
Change of the manuscript title

@@
 ---
-title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19 in a cross-sectional prospective cohort study"
+title: "Persistent somatic symptoms are key to individual illness perception at one year after COVID-19 in a cross-sectional analysis of a prospective cohort study"
 subtitle: "Supplementary Material"

### 2023-03-17 analysis scripts/ipq_clust_development.R
R code update
To accommodate R version 4.3.0 and newest stable versions of 'trafo' and figure

@@ -67,8 +67,8 @@
         nfolds = 10, 
         kNN = 7, 
         resolve_ties = TRUE) %>% 
-    map2_dfr(., names(.), 
-             ~mutate(.x$summary, method = .y))
+    map(~.x$summary) %>% 
+    compress(names_to = 'method')
   
 # common comparison results, plotting ----
   

### 2023-03-17 analysis scripts/ipq_sub1_factors.R

@@ -15,16 +15,7 @@
     filter(variable %in% unique(c(an$cmm_variables$ipq_sub1, 
                                   an$conf_variables))) %>% 
     dlply(.(what))
-  
-  ## n numbers to be presented in the X axis
-  
-  sub1_factors$n_numbers <- sub1_factors$variables$eff_size$variable %>% 
-    map(~count(mod$multi_tbl, .data[[.x]])) %>% 
-    set_names(sub1_factors$variables$eff_size$variable)
-  
-  sub1_factors$x_axes <- sub1_factors$n_numbers %>% 
-    map(~map2_chr(.x[[1]], .x[[2]], paste, sep = '\nn = '))
-  
+
 # Serial comparisons and correlations -----
   
   insert_msg('Serial comparisons')
@@ -84,16 +75,18 @@
                         split_factor = .x, 
                         variable = 'ipq_sub1', 
                         type = 'violin', 
-                        plot_title = translate_var(.x, 
-                                                   out_value = 'label_long'), 
+                        plot_title = exchange(.x,
+                                              dict = globals$var_lexicon, 
+                                              value = 'label_long'), 
                         plot_subtitle = .y, 
-                        y_lab = translate_var('ipq_sub1'), 
-                        x_lab = translate_var(.x), 
+                        y_lab = exchange('ipq_sub1', 
+                                         dict = globals$var_lexicon), 
+                        x_lab = exchange(.x, 
+                                         dict = globals$var_lexicon), 
                         point_hjitter = 0, 
-                        cust_theme = globals$common_theme)) %>% 
-    map2(sub1_factors$x_axes, 
-         ~.x + 
-          scale_x_discrete(labels = .y) + 
+                        cust_theme = globals$common_theme, 
+                        x_n_labs = TRUE)) %>% 
+    map(~.x + 
           scale_fill_brewer() + 
           theme(plot.tag = element_blank())) %>% 
     set_names(sub1_factors$factor_test$split_factor)
@@ -109,10 +102,16 @@
                              filter(complete.cases(.)), 
                            variables = c(.x, 'ipq_sub1'), 
                            type = 'correlation',
-                           plot_title = translate_var(.x, out_value = 'label_long'), 
+                           plot_title = exchange(.x, 
+                                                 dict = globals$var_lexicon, 
+                                                 value = 'label_long'), 
                            plot_subtitle = .y, 
[…]

### 2023-03-17 analysis scripts/ipq_sub2_factors.R

@@ -17,16 +17,7 @@
     filter(variable %in% unique(c(an$cmm_variables$ipq_sub2, 
                                   an$conf_variables))) %>% 
     dlply(.(what))
-  
-  ## n numbers to be presented in the X axis
-  
-  sub2_factors$n_numbers <- sub2_factors$variables$eff_size$variable %>% 
-    map(~count(mod$multi_tbl, .data[[.x]])) %>% 
-    set_names(sub2_factors$variables$eff_size$variable)
-  
-  sub2_factors$x_axes <- sub2_factors$n_numbers %>% 
-    map(~map2_chr(.x[[1]], .x[[2]], paste, sep = '\nn = '))
-  
+
 # Serial comparisons and correlations -----
   
   insert_msg('Serial comparisons')
@@ -86,15 +77,18 @@
                         split_factor = .x, 
                         variable = 'ipq_sub2', 
                         type = 'violin', 
-                        plot_title = translate_var(.x, out_value = 'label_long'), 
+                        plot_title = exchange(.x, 
+                                              dict = globals$var_lexicon, 
+                                              value = 'label_long'), 
                         plot_subtitle = .y, 
-                        y_lab = translate_var('ipq_sub2'), 
-                        x_lab = translate_var(.x), 
+                        y_lab = exchange('ipq_sub2', 
+                                         dict = globals$var_lexicon), 
+                        x_lab = exchange(.x, 
+                                         dict = globals$var_lexicon), 
                         point_hjitter = 0, 
-                        cust_theme = globals$common_theme)) %>% 
-    map2(sub2_factors$x_axes, 
-        ~.x + 
-          scale_x_discrete(labels = .y) + 
+                        cust_theme = globals$common_theme, 
+                        x_n_labs = TRUE)) %>% 
+    map(~.x + 
           scale_fill_brewer() + 
           theme(plot.tag = element_blank())) %>% 
     set_names(sub2_factors$factor_test$split_factor)
@@ -110,10 +104,16 @@
                              filter(complete.cases(.)), 
                            variables = c(.x, 'ipq_sub2'), 
                            type = 'correlation',
-                           plot_title = translate_var(.x, out_value = 'label_long'), 
+                           plot_title = exchange(.x, 
+                                                 dict = globals$var_lexicon, 
+                                                 value = 'label_long'), 
                            plot_subtitle = .y, 
-                           x_lab = translate_var(.x, out_value = 'axis_lab'), 
-                           y_lab = translate_var('ipq_sub2', out_value = 'axis_lab'), 
[…]

### 2023-03-17 analysis scripts/plots.R

@@ -29,7 +29,7 @@
   mod_plots$perf_plots <- list(cv_stats =  mod_plots$perf_data %>% 
                                  dlply(.(response)), 
                                plot_subtitle = an$responses %>% 
-                                 translate_var %>% 
+                                 exchange(dict = globals$var_lexicon) %>% 
                                  paste('1-year follow-up', sep = ', ')) %>% 
     pmap(plot_rsq_rmse,
          y_factor = 'method') %>% 
@@ -48,7 +48,7 @@
     list(data = map(an$variables, 
                     ~set_names(.x, an$methods[names(.x)])), 
          plot_title = an$responses %>% 
-           translate_var) %>% 
+           exchange(dict = globals$var_lexicon)) %>% 
     pmap(safely(plot_n_venn),  
          fill_color = set_names(an$method_colors, 
                                 an$methods), 

### 2023-03-17 analysis.R

@@ -27,8 +27,11 @@
   library(soucer)
   library(clustTools)
   library(furrr)
+  library(trafo)
 
   insert_head()
+  
+  explore <- exda::explore
   
   source_all('./tools/project_tools.R')
 

### 2023-03-17 exec.R

@@ -7,5 +7,5 @@
                      'modeling.R', 
                      'analysis.R', 
                      'paper.R'), 
-                   crash = FALSE, 
+                   crash = TRUE, 
                    message = TRUE))

### 2023-03-17 exploration scripts/missingness.R

@@ -16,7 +16,8 @@
            n_missing = .) %>% 
     mutate(frac_missing = n_missing/nrow(cov_data$clear_data), 
            perc_missing = frac_missing * 100) %>% 
-    mutate(var_lab = translate_var(variable))
+    mutate(var_lab = exchange(variable, 
+                              dict = globals$var_lexicon))
   
 # Plotting variable missingness -------
   

### 2023-03-17 exploration scripts/psy_correlation.R

@@ -51,7 +51,8 @@
   psy_cor$corr_mtx_plot <- psy_cor$corr_results %>% 
     mutate(variable1 = ifelse(variable1 %in% names(globals$ipq_sublabs), 
                               globals$ipq_sublabs[variable1], 
-                              translate_var(variable1))) %>% 
+                              exchange(variable1, 
+                                       dict = globals$var_lexicon))) %>% 
     ggplot(aes(x = variable1, 
                y = variable2, 
                size = abs(estimate), 
@@ -64,7 +65,8 @@
               size = 2.75, 
               vjust = -1.5) + 
     scale_y_discrete(limits = psy_cor$var2, 
-                     labels = translate_var(psy_cor$var2)) + 
+                     labels = exchange(psy_cor$var2, 
+                                       dict = globals$var_lexicon)) + 
     scale_size_area(max_size = 5, 
                     limits = c(0, 1), 
                     name = expression('abs(' * rho * ')')) + 

### 2023-03-17 exploration scripts/respiratory.R

@@ -68,21 +68,20 @@
   respi[paste0(names(expl$variables), '_plots')] <- names(expl$variables) %>% 
     map(function(var_class) list(variable = expl$variables[[var_class]], 
                                  plot_title = expl$variables[[var_class]] %>% 
-                                   translate_var(out_value = 'label_long'), 
+                                   exchange(dict = globals$var_lexicon,
+                                            value = 'label_long'), 
                                  plot_subtitle = respi$test_results[[var_class]]$plot_cap, 
                                  y_lab = expl$variables[[var_class]] %>% 
-                                   translate_var(out_value = 'axis_lab'), 
+                                   exchange(dict = globals$var_lexicon, 
+                                            out_value = 'axis_lab'), 
                                  type = respi$plot_type[[var_class]]) %>% 
           pmap(plot_variable, 
                cov_data$clear_data, 
                split_factor = 'respi_comorb', 
                scale = 'percent', 
                x_lab = 'Respiratory comorbidity', 
-               cust_theme = globals$common_theme) %>% 
-          map(~.x + 
-                labs(tag = .x$labels$tag %>% 
-                       stri_replace_all(fixed = '\n', replacement = ', ') %>% 
-                       paste0('\n', .))) %>% 
+               cust_theme = globals$common_theme, 
+               x_n_labs = TRUE) %>% 
           set_names(expl$variables[[var_class]]))
   
 # END ------

### 2023-03-17 exploration scripts/score_coherence.R

@@ -170,7 +170,8 @@
   
   ipq_co$corr_mtx_plots <- 
     list(x = ipq_co$corr_mtx_plots, 
-         y = translate_var(names(ipq_co$corr_mtx_plots)), 
+         y = exchange(names(ipq_co$corr_mtx_plots),
+                      dict = globals$var_lexicon), 
          z = paste0('\u03C9 = ', 
                     signif(ipq_co$omega_obj$omega.group$total, 2), 
                     ', n = ', 

### 2023-03-17 exploration scripts/symptom_kinetic.R

@@ -148,7 +148,8 @@
   
   sympt_kinet$plot_legend <- sympt_kinet$cohran_test %>% 
     map(mutate, 
-        var_lab = translate_var(.y.), 
+        var_lab = exchange(.y., 
+                           dict = globals$var_lexicon), 
         significance = p_value_jps(p), 
         var_lab = paste(var_lab, significance, sep = ', '))
   

### 2023-03-17 import.R

@@ -11,6 +11,7 @@
   library(stringi)
   library(soucer)
   library(rlang)
+  library(trafo)
 
   insert_head()
   

### 2023-03-17 modeling scripts/bayes_lasso.R

@@ -43,14 +43,14 @@
              explore, 
              what = 'normality', 
              variables = '.resid') %>% 
-          map2_dfr(., names(.), ~mutate(.x, mod_type = .y))) %>% 
-    map2_dfr(., names(.), ~mutate(.x, response = .y))
+          compress(names_to = 'mod_type')) %>% 
+    compress(names_to = 'response')
   
   ## residual plots
   
   blass_mod$resid_plots <- list(x = blass_mod$models, 
                                 plot_title = names(blass_mod$models) %>% 
-                                  translate_var %>% 
+                                  exchange(dict = globals$var_lexicon) %>% 
                                   map(paste, c('training', 'CV'), sep = ', ')) %>% 
     pmap(plot, 
          type = 'diagnostic', 
@@ -83,7 +83,8 @@
     map2(., c(0.7, 1.1), 
          ~.x + 
            expand_limits(x = .y) + 
-           scale_y_discrete(labels = translate_var(names(blass_mod$models))))
+           scale_y_discrete(labels = exchange(names(blass_mod$models), 
+                                              dict = globals$var_lexicon)))
   
 # Calibration plots ------
   
@@ -91,7 +92,7 @@
   
   blass_mod$fit_plots <- list(x = blass_mod$models, 
                               plot_title = names(blass_mod$models) %>% 
-                                translate_var %>% 
+                                exchange(dict = globals$var_lexicon) %>% 
                                 map(paste, c('training', 'CV'), sep = ', ')) %>% 
     pmap(plot, 
          type = 'fit', 
@@ -135,7 +136,8 @@
   
   blass_mod$estimate_plots <- 
     list(est_data = blass_mod$inference, 
-         plot_title = translate_var(names(blass_mod$inference)) %>% 
+         plot_title = exchange(names(blass_mod$inference), 
+                               dict = globals$var_lexicon) %>% 
            paste('1-year follow-up', sep = ', '), 
          plot_subtitle = blass_mod$plot_cap[names(blass_mod$inference)]) %>% 
     pmap(est_bubble, 

### 2023-03-17 modeling scripts/elanet.R

@@ -63,7 +63,7 @@
   
   eln_mod$opt_lambda <- eln_mod$lambda_tbl %>% 
     map(filter, cvm == min(cvm)) %>% 
-    map2_dfr(., names(.), ~mutate(.x, response = .y))
+    compress(names_to = 'response')
   
   eln_mod$tune_grids <- eln_mod$opt_lambda$lambda %>% 
     map(~data.frame(alpha = eln_mod$alpha, 
@@ -103,14 +103,14 @@
              explore, 
              what = 'normality', 
              variables = '.resid') %>% 
-          map2_dfr(., names(.), ~mutate(.x, mod_type = .y))) %>% 
-    map2_dfr(., names(.), ~mutate(.x, response = .y))
+          compress(names_to = 'mod_type')) %>% 
+    compress(names_to = 'response')
 
   ## residual plots
   
   eln_mod$resid_plots <- list(x = eln_mod$models, 
                               plot_title = names(eln_mod$models) %>% 
-                                translate_var %>% 
+                                exchange(dict = globals$var_lexicon) %>% 
                                 map(paste, c('training', 'CV'), sep = ', ')) %>% 
     pmap(plot, 
          type = 'diagnostic', 
@@ -143,7 +143,8 @@
     map2(., c(0.7, 1.1), 
          ~.x + 
            expand_limits(x = .y) + 
-           scale_y_discrete(labels = translate_var(names(eln_mod$models))))
+           scale_y_discrete(labels = exchange(names(eln_mod$models), 
+                                              dict = globals$var_lexicon)))
 
 # Model calibration -----  
   
@@ -151,7 +152,7 @@
   
   eln_mod$fit_plots <- list(x = eln_mod$models, 
                               plot_title = names(eln_mod$models) %>% 
-                                translate_var %>% 
+                                exchange(dict = globals$var_lexicon) %>% 
                                 map(paste, c('training', 'CV'), sep = ', ')) %>% 
     pmap(plot, 
          type = 'fit', 
@@ -182,7 +183,8 @@
   
   eln_mod$estimate_plots <- 
     list(est_data = eln_mod$inference, 
-         plot_title = translate_var(names(eln_mod$inference)) %>% 
+         plot_title = exchange(names(eln_mod$inference), 
+                               dict = globals$var_lexicon) %>% 
            paste('1-year follow-up', sep = ', '), 
          plot_subtitle = eln_mod$plot_cap[names(eln_mod$inference)]) %>% 
     pmap(est_bubble, 

### 2023-03-17 modeling scripts/elimination.R

@@ -46,7 +46,7 @@
     map(explore, 
         what = 'normality', 
         variables = '.resid') %>% 
-    map2_dfr(., names(.), ~mutate(.x, response = .y))
+    compress(names_to = 'response')
   
   ## diagnostic plots
   
@@ -79,9 +79,10 @@
   
   eli_mod$forest_plots <- list(x = map(eli_mod$inference, 
                                        mutate, 
-                                       variable = translate_var(variable)), 
+                                       variable = exchange(variable, 
+                                                           dict = globals$var_lexicon)), 
                                plot_title = names(eli_mod$inference) %>% 
-                                 translate_var %>% 
+                                 exchange(dict = globals$var_lexicon) %>% 
                                  paste('1-year follow-up', sep = ', '), 
                                plot_subtitle = eli_mod$fit_stats$plot_cap) %>% 
     pmap(plot_forest, 
@@ -118,10 +119,11 @@
   
   eli_mod$dev_plots <- map2(eli_mod$deviances %>% 
                               map(mutate, 
-                                variable = translate_var(variable)) %>% 
+                                variable = exchange(variable, 
+                                                    dict = globals$var_lexicon)) %>% 
                               map(filter, variable != 'Residuals'), 
                             names(eli_mod$inference) %>% 
-                              translate_var %>% 
+                              exchange(dict = globals$var_lexicon) %>% 
                               paste('1-year follow-up', sep = ', '), 
                             ~ggplot(.x, 
                                     aes(x = frac_explained, 
@@ -182,7 +184,8 @@
     map2(., c(0.65, 1.1), 
          ~.x + 
            expand_limits(x = .y) + 
-           scale_y_discrete(labels = translate_var(names(eli_mod$caret_models))))
+           scale_y_discrete(labels = exchange(names(eli_mod$caret_models), 
+                                              dict = globals$var_lexicon)))
 
 # Calibration plots: response vs fitted -----
   
@@ -190,7 +193,7 @@
   
   eli_mod$fit_plots <- list(x = eli_mod$caret_models, 
                             plot_title = names(eli_mod$caret_models) %>% 
-                              translate_var %>% 
+                              exchange(dict = globals$var_lexicon) %>% 
                               map(paste, c('training', 'CV'), sep = ', ')) %>% 
     pmap(plot, 
          type = 'fit', 

### 2023-03-17 modeling scripts/lasso.R

@@ -63,7 +63,7 @@
   
   lasso_mod$opt_lambda <- lasso_mod$lambda_tbl %>% 
     map(filter, cvm == min(cvm)) %>% 
-    map2_dfr(., names(.), ~mutate(.x, response = .y))
+    compress(names_to = 'response')
   
   lasso_mod$tune_grids <- lasso_mod$opt_lambda$lambda %>% 
     map(~data.frame(alpha = lasso_mod$alpha, 
@@ -103,14 +103,14 @@
              explore, 
              what = 'normality', 
              variables = '.resid') %>% 
-          map2_dfr(., names(.), ~mutate(.x, mod_type = .y))) %>% 
-    map2_dfr(., names(.), ~mutate(.x, response = .y))
+          compress(names_to = 'mod_type')) %>% 
+    compress(names_to = 'response')
 
   ## residual plots
   
   lasso_mod$resid_plots <- list(x = lasso_mod$models, 
                               plot_title = names(lasso_mod$models) %>% 
-                                translate_var %>% 
+                                exchange(dict = globals$var_lexicon) %>% 
                                 map(paste, c('training', 'CV'), sep = ', ')) %>% 
     pmap(plot, 
          type = 'diagnostic', 
@@ -143,7 +143,8 @@
     map2(., c(0.7, 1.1), 
          ~.x + 
            expand_limits(x = .y) + 
-           scale_y_discrete(labels = translate_var(names(lasso_mod$models))))
+           scale_y_discrete(labels = exchange(names(lasso_mod$models), 
+                                              dict = globals$var_lexicon)))
 
 # Model calibration -----  
   
@@ -151,7 +152,7 @@
   
   lasso_mod$fit_plots <- list(x = lasso_mod$models, 
                               plot_title = names(lasso_mod$models) %>% 
-                                translate_var %>% 
+                                exchange(dict = globals$var_lexicon) %>% 
                                 map(paste, c('training', 'CV'), sep = ', ')) %>% 
     pmap(plot, 
          type = 'fit', 
@@ -182,7 +183,8 @@
   
   lasso_mod$estimate_plots <- 
     list(est_data = lasso_mod$inference, 
-         plot_title = translate_var(names(lasso_mod$inference)) %>% 
+         plot_title = exchange(names(lasso_mod$inference), 
+                               dict = globals$var_lexicon) %>% 
            paste('1-year follow-up', sep = ', '), 
[…]

### 2023-03-17 modeling scripts/uni_variate.R

@@ -30,7 +30,7 @@
              summary, 
              type = 'assumptions', 
              type.predict = 'response') %>% 
-          map2_dfr(., names(.), ~mutate(.x, variable = .y))) %>% 
+          compress(names_to = 'variable')) %>% 
     map2(., names(.), ~mutate(.x, response = .y))
 
 # model fit statistics -----
@@ -41,7 +41,7 @@
     map(~map(.x, 
              summary, 
              type = 'fit') %>% 
-          map2_dfr(., names(.), ~mutate(.x, variable = .y))) %>% 
+          compress(names_to = 'variable')) %>% 
     map(mutate, 
         rsq_eff_size = cut(raw_rsq, 
                            c(-Inf, 0.01, 0.09, 0.25, Inf), 
@@ -71,7 +71,8 @@
   insert_msg('R square value plots')
   
   uni_mod$rsq_plots <- uni_mod$fit_stats %>% 
-    map2(., translate_var(names(.)), 
+    map2(., exchange(names(.), 
+                     dict = globals$var_lexicon), 
          ~ggplot(.x, 
                  aes(x = raw_rsq, 
                      y = reorder(variable, raw_rsq), 
@@ -84,7 +85,8 @@
                       linetype = 'dashed') + 
            geom_vline(xintercept = 0.25, 
                       linetype = 'dashed') + 
-           scale_y_discrete(labels = translate_var(globals$variables)) + 
+           scale_y_discrete(labels = exchange(globals$variables,
+                                              dict = globals$var_lexicon)) + 
            scale_x_continuous(limits = c(0, 0.45)) + 
            scale_fill_manual(values = mod$size_colors, 
                              name = 'Effect size') + 
@@ -114,14 +116,15 @@
                                          variable %in% .y, 
                                          parameter != '(Intercept)')) %>% 
     map(mutate, 
-        variable = translate_var(variable))
+        variable = exchange(variable,
+                            dict = globals$var_lexicon))
   
   ## Forest plots
   
   uni_mod$forest_plots <- 
     list(x =  uni_mod$forest_results,
          plot_title = names(uni_mod$forest_results) %>% 
-           translate_var %>% 
+           exchange(globals$var_lexicon) %>% 
            paste('1-year follow-up', sep = ', ')) %>% 
     pmap(plot_forest, 
          p_value = 'p_value', 

### 2023-03-17 modeling.R

@@ -42,8 +42,12 @@
   library(soucer)
   library(glmnet)
   library(doParallel)
+  library(trafo)
 
   insert_head()
+  
+  explore <- exda::explore
+  train <- caret::train
   
   source_all('./tools/project_tools.R')
   
@@ -168,7 +172,7 @@
   
 # modeling scripts ------
   
-  insert_msg('Aanlysis scripts')
+  insert_msg('Analysis scripts')
   
   ## working with cache (time costly lambda turning)
 

### 2023-03-17 paper scripts/tables.R

@@ -130,7 +130,7 @@
   insert_msg('Table S2: modeling, explanatory variables')
   
   suppl_tbl$modeling_vars <- mod$variables %>% 
-    translate_var %>% 
+    exchange(dict = globals$var_lexicon) %>% 
     paste(collapse = ', ') %>% 
     tibble(`Explanatory variable` = .) %>% 
     as_mdtable(label = 'modeling_vars', 

### 2023-03-17 paper scripts/values.R

@@ -46,7 +46,8 @@
   
   paper_val$clust_factors <- 
     mdexpr(colnames(model.frame(ipq_clust$clust_obj)) %>% 
-             translate_var(out_value = 'label_long') %>% 
+             exchange(dict = globals$var_lexicon, 
+                      value = 'label_long') %>% 
              stri_replace(fixed = 'Illness perception, ', 
                           replacement = '') %>% 
              paste(collapse = ', '), 
@@ -55,7 +56,8 @@
   
   paper_val$clust_factors_and <- 
     mdexpr(colnames(model.frame(ipq_clust$clust_obj)) %>% 
-             translate_var(out_value = 'label_long') %>% 
+             exchange(dict = globals$var_lexicon, 
+                      value = 'label_long') %>% 
              stri_replace(fixed = 'Illness perception, ', 
                           replacement = '') %>% 
              collapse_and, 
@@ -88,7 +90,8 @@
   
   paper_val$factor_elnet <- 
     mdexpr(an$variables$ipq_total$eln_mod %>% 
-             translate_var(out_value = 'label_long') %>% 
+             exchange(dict = globals$var_lexicon,
+                      value = 'label_long') %>% 
              de_capitalize %>% 
              paste(collapse = ', '), 
            ref_name = 'factor_elnet', 
@@ -96,7 +99,8 @@
   
   paper_val$factor_elnet_and <- 
     mdexpr(an$variables$ipq_total$eln_mod %>% 
-             translate_var(out_value = 'label_long') %>% 
+             exchange(dict = globals$var_lexicon, 
+                      value = 'label_long') %>% 
              de_capitalize %>% 
              collapse_and, 
            ref_name = 'factor_elnet_and', 
@@ -104,7 +108,8 @@
   
   paper_val$factor_lasso <- 
     mdexpr(an$variables$ipq_total$lasso_mod %>% 
-             translate_var(out_value = 'label_long') %>% 
+             exchange(dict = globals$var_lexicon, 
+                      value = 'label_long') %>% 
              de_capitalize %>% 
              paste(collapse = ', '), 
            ref_name = 'factor_lasso', 
@@ -112,7 +117,8 @@
   
   paper_val$factor_lasso_and <- 
     mdexpr(an$variables$ipq_total$lasso_mod %>% 
-             translate_var(out_value = 'label_long') %>% 
+             exchange(dict = globals$var_lexicon, 
+                      value = 'label_long') %>% 
[…]

### 2023-03-17 paper.R

@@ -14,8 +14,12 @@
   library(flextable)
   library(writexl)
   library(clustTools)
+  library(trafo)
 
   insert_head()
+  
+  explore <- exda::explore
+  train <- caret::train
   
   source_all('./tools/project_tools.R')
   

### 2023-03-17 paper/markdown/supplementary_material.Rmd
Manuscript/Supplement update
Update of the R code in Markdown documents to accommodate the recent versions of development packages figur and trafo. No effects on results

@@
 The number of non-zero model coefficients was found to differ substantially between the regularized regression models (__Supplementary Figures S\@ref(fig:fig-eln-est-total)__ - __S\@ref(fig:fig-blasso-est-sub1)__), which was reported for data sets of similar size [@Morozova2015].
-Hence, the key factors impacting on the total IP score and the emotion/concern/consequences IP component at one year after COVID-19 (`r an$cmm_variables$ipq_total %>% translate_var(out_value = "label_long") %>% de_capitalize %>% collapse_and`) were defined as variables with non-zero coefficients in all multi-parameter models.
+Hence, the key factors impacting on the total IP score and the emotion/concern/consequences IP component at one year after COVID-19 (`r an$cmm_variables$ipq_total %>% exchange(dict = globals$var_lexicon, value = "label_long") %>% de_capitalize %>% collapse_and`) were defined as variables with non-zero coefficients in all multi-parameter models.
 Since no meaningful models for the lacking control/coherence could be obtained, no influential factors for the lacking control/coherence component could be identified (__Figure 2__).

### 2023-03-20 paper/markdown/manuscript.Rmd
Compacting footnotes of Suppl. Tables

@@
 fatigue rating (`r clust_chara$test_results$fup %>% filter(variable == 'Chalder_FS') %>% .$plot_cap`), 
-significant fatigue ($\chi^2$ test: `r clust_chara$test_results$fup %>% filter(variable == 'Chalder_FS-bimodal') %>% .$plot_cap`), 
+significant fatigue ($\chi^2$ test: `r clust_chara$test_results$fup %>% filter(variable == 'Chalder_FS_bimodal') %>% .$plot_cap`), 
 sleep problems (`r clust_chara$test_results$fup %>% filter(variable == 'sleep_sympt') %>% .$plot_cap`) or 

### 2023-03-20 paper/markdown/supplementary_material.Rmd
Compacting footnotes of Suppl. Tables

@@
   width(1, width = 14, unit = 'cm') %>% 
+  footnote(1, 1, 
+           value = as_paragraph("CVD: cardiovascular disease; #: number of; CKD: chronic kidney disease; GID: gastrointestinal disease; PSS: persistent somatic symptoms; CFS: Chalder's fatigue score; SMWD: six-minute walking distance; SMWD vs ref.: Difference of SMWD and the expected reference SMWD value; LFT: lung function testing; CT: computed tomography of the chest; HB: blood hemoglobin; FT: serum ferritin; TF-Sat: transferrin saturation; sTFR: serum soluble transferrin receptor; NT-proBNP: serum N-terminal pro-brain natriuretic peptide; CRP: serum C-reactive protein; HbA1c: glycated blood Hb"),
+           part = 'header', 
+           ref_symbols = 'a') %>% 
   theme_vanilla
@@
   width(5, width = 2.1, unit = 'cm') %>% 
+  footnote(1, 1, 
+           value = as_paragraph('WHO COVID-19 severity: WHO ordinal scale for clinical improvement; SMWD: Six minute walking distance; Hb: Blood hemoglobin; FT: serum ferritin'), 
+           ref_symbols = 'a', 
+           part = 'header') %>% 
   footnote(1, 4:5, 
            value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Mann-Whitney test with r effect size statistic'), 
-           ref_symbols = 'a', 
+           ref_symbols = 'b', 
            part = 'header') %>% 
-  footnote(2, 1, 
-           value = as_paragraph('WHO ordinal scale for clinical improvement'), 
-           ref_symbols = 'b', 
-           part = 'body') %>% 
-  footnote(5, 1, 
-           value = as_paragraph('Six minute walking distance'), 
-           ref_symbols = 'b', 
-           part = 'body') %>% 
-  footnote(6, 1, 
-           value = as_paragraph('Blood hemoglobin'), 
-           ref_symbols = 'b', 
-           part = 'body') %>% 
-  footnote(7, 1, 
-           value = as_paragraph('Serum ferritin'), 
-           ref_symbols = 'b', 
-           part = 'body') %>% 
   theme_vanilla()
@@
   width(7, width = 2.1, unit = 'cm') %>%
+  footnote(1, 1, 
+           value = as_paragraph("PSS: persistent somatic symptoms; ECOG: Eastern Cooperative Oncology Group performance score; mMRC: Modified Medical Research Council dyspnea score; CFS: Chalder's fatigure score; LFT: lung function testing; CT: computed tomography of the chest; SMWD: Six minute walking distance; SMWD vs reference: SMWD, difference between the actual and reference value"),
+           part = 'header', 
+           ref_symbols = 'a') %>% 
   footnote(1, 6:7, 
            value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic'), 
-           ref_symbols = 'a', 
+           ref_symbols = 'b', 
            part = 'header') %>% 
-  footnote(2:3, 1, 
-           value = as_paragraph('PSS: persistent somatic symptoms'), 
-           ref_symbols = 'c', 
-           part = 'body') %>% 
-  footnote(4, 1, 
-           value = as_paragraph('ECOG: Eastern Cooperative Oncology Group performance score'), 
-           ref_symbols = 'c', 
-           part = 'body') %>% 
-  footnote(5, 1, 
-           value = as_paragraph('mMRC: Modified Medical Research Council dyspnea score'), 
-           ref_symbols = 'd', 
-           part = 'body') %>% 
-  footnote(13:14, 1, 
-           value = as_paragraph("CFS: Chalder's Fatigure Score"), 
-           ref_symbols = 'e', 
-           part = 'body') %>% 
   footnote(15, 1, 
            value = as_paragraph('Lung function testing abnormality: FVC < 80% or FEV1 < 80% or TLC < 80% or DLCO < 80% predicted or FEV1:FVC < 70% predicted reference value'), 
-           ref_symbols = 'f', 
-           part = 'body') %>% 
-  footnote(17, 1, 
-           value = as_paragraph('Chest computer tomography severity score'), 
-           ref_symbols = 'g', 
-           part = 'body') %>% 
-  footnote(19, 1, 
-           value = as_paragraph('SMWD: Six minute walking distance'), 
-           ref_symbols = 'h', 
-           part = 'body') %>% 
-  footnote(20, 1, 
-           value = as_paragraph('SMWD, difference between the actual and reference value'), 
-           ref_symbols = 'i', 
+           ref_symbols = 'c', 
            part = 'body') %>% 
@@
   width(7, width = 2.1, unit = 'cm') %>%
+  footnote(1, 1, 
+           value = as_paragraph('Hb: blood hemoglobin; FT: serum ferritin; TF-Sat: serum transferrin saturation; sTFR: serum soluble transferrin receptor; NT-pro-BNP: serum N-terminal pro-brain natriuretic peptide; CRP: serum C-reactive protein; HbA1c: glycated blood Hb'),
+           part = 'header', 
+           ref_symbols = 'a') %>% 
   footnote(1, 6:7, 
            value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic'), 
-           ref_symbols = 'a', 
+           ref_symbols = 'b', 
            part = 'header') %>% 
@@
            value = as_paragraph('Females: hemoglobin (Hb) < 120 g/dL, males: Hb < 140 g/dL'), 
-           ref_symbols = 'b', 
-           part = 'body') %>% 
-  footnote(3, 1, 
-           as_paragraph('Blood hemoglobin'), 
            ref_symbols = 'c', 
-           part = 'body') %>% 
-  footnote(4, 1, 
-           as_paragraph('Serum ferritin'), 
-           ref_symbols = 'd', 
-           part = 'body') %>%
-  footnote(5, 1, 
-           as_paragraph('Serum transferrin saturation'), 
-           ref_symbols = 'e', 
-           part = 'body') %>% 
-  footnote(6, 1, 
-           as_paragraph('Serum soluble transferrin receptor'), 
-           ref_symbols = 'f', 
-           part = 'body') %>% 
-  footnote(8, 1, 
-           as_paragraph('Serum N-terminal pro-brain natriuretic peptide'), 
-           ref_symbols = 'g', part = 'body') %>% 
-  footnote(10, 1, as_paragraph('Serum C-reactive protein'), 
-           ref_symbols = 'h', 
-           part = 'body') %>% 
-  footnote(11, 1, 
-           as_paragraph('Blood glycated hamoglobin'), 
-           ref_symbols = 'i', 
            part = 'body') %>% 
@@
 flextable::flextable(suppl_tbl$gender) %>% 
-  width(1, width = 5.2, unit = 'cm') %>% 
+  width(1, width = 6, unit = 'cm') %>% 
   width(2:3, width = 3.5, unit = 'cm') %>% 
@@
   width(5, width = 2.1, unit = 'cm') %>% 
+  footnote(1, 1, 
+           value = as_paragraph('WHO COVID-19 severity: WHO ordinal scale for clinical improvement; PSS: persistent somatic symptoms; mMRC: Modified Medical Research Council dyspnea score; CT: computed tomography of the chest; Hb: blood hemoglobin; FT: serum ferritin; BIPQ: brief illness perception questionnaire'),
+           part = 'header', 
+           ref_symbols = 'a') %>% 
   footnote(1, 4:5, 
            value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Mann-Whitney test with r effect size statistic'), 
-           ref_symbols = 'a', 
+           ref_symbols = 'b', 
            part = 'header') %>% 
@@
            value = as_paragraph('Former or active smoker'), 
-           ref_symbols = 'b', 
+           ref_symbols = 'c', 
            part = 'body') %>% 
@@
            value = as_paragraph('Overweight: body mass index 25 - 30 kg/m\u00B2, obesity: > 30 kg/m\u00B2'), 
-           ref_symbols = 'c', 
-           part = 'body') %>% 
-  footnote(5, 1, 
-           value = as_paragraph('WHO ordinal scale for clinical improvement'), 
            ref_symbols = 'd', 
-           part = 'body') %>% 
-  footnote(7:8, 1, 
-           value = as_paragraph('PSS: persistent somatic symptoms'), 
-           ref_symbols = 'e', 
-           part = 'body') %>% 
-  footnote(9, 1, 
-           value = as_paragraph('mMRC: Modified Medical Research Council dyspnea score'), 
-           ref_symbols = 'f', 
-           part = 'body') %>% 
-  footnote(11, 1, 
-           value = as_paragraph('Chest computer tomography severity score'), 
-           ref_symbols = 'g', 
-           part = 'body') %>% 
-  footnote(12, 1, 
-           value = as_paragraph('Blood hemoglobin'), 
-           ref_symbols = 'h', 
-           part = 'body') %>% 
-  footnote(13, 1, 
-           value = as_paragraph('Serum ferritin'), 
-           ref_symbols = 'i', 
            part = 'body') %>% 
@@
 flextable::flextable(suppl_tbl$respi) %>% 
-  width(1, width = 4.8, unit = 'cm') %>% 
+  width(1, width = 6, unit = 'cm') %>% 
   width(2:3, width = 3.3, unit = 'cm') %>% 
@@
   width(5, width = 2.1, unit = 'cm') %>% 
+  footnote(1, 1, 
+           value = as_paragraph("CFS: Chalder's fatigue score; BIPQ: brief illness perception questionnaire"), 
+           ref_symbols = 'a', 
+           part = 'header') %>% 
   footnote(1, 4:5, 
            value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Mann-Whitney test with r effect size statistic'), 
-           ref_symbols = 'a', 
+           ref_symbols = 'b', 
            part = 'header') %>% 
-  footnote(5:6, 1, 
-           value = as_paragraph("CFS: Chalder's fatigue score"), 
-           ref_symbols = 'b', 
-           part = 'body') %>% 
   footnote(7, 1, 
@@
            ref_symbols = 'c', 
-           part = 'body') %>% 
-  footnote(8:12, 1, 
-           as_paragraph('BIPQ: brief illness perception questionnaire'), 
-           ref_symbols = 'd', 
            part = 'body') %>% 
@@
   width(6, width = 2.1, unit = 'cm') %>%
+  footnote(1, 1, 
+           value = as_paragraph("WHO COVID-19 severity: WHO ordinal scale for clinical improvement; PSS: persistent somatic symptoms; ECOG: Eastern Cooperative Oncology Group performance score; mMRC: Modified Medical Research Council dyspnea score; CFS: Chalder's Fatigure Score"),
+           part = 'header', 
+           ref_symbols = 'a') %>% 
   footnote(1, 5:6, 
            value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic, numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic'), 
-           ref_symbols = 'a', 
+           ref_symbols = 'b', 
            part = 'header') %>% 
@@
            value = as_paragraph('Former or active smoker'), 
-           ref_symbols = 'b', 
-           part = 'body') %>% 
-  footnote(5, 1, 
-           value = as_paragraph('WHO ordinal scale for clinical improvement'), 
            ref_symbols = 'c', 
-           part = 'body') %>% 
-  footnote(6:7, 1, 
-           value = as_paragraph('PSS: persistent somatic symptoms'), 
-           ref_symbols = 'd', 
-           part = 'body') %>% 
-  footnote(8, 1, 
-           value = as_paragraph('ECOG: Eastern Cooperative Oncology Group performance score'), 
-           ref_symbols = 'e', 
-           part = 'body') %>% 
-  footnote(9, 1, 
-           value = as_paragraph('mMRC: Modified Medical Research Council dyspnea score'), 
-           ref_symbols = 'f', 
-           part = 'body') %>% 
-  footnote(13:14, 1, value = as_paragraph("Chalder's Fatigure Score"), 
-           ref_symbols = 'g', 
            part = 'body') %>% 
@@
            value = as_paragraph('Lung function testing abnormality: FVC < 80% or FEV1 < 80% or TLC < 80% or DLCO < 80% predicted or FEV1:FVC < 70% predicted reference value'), 
-           ref_symbols = 'h', 
+           ref_symbols = 'd', 
            part = 'body') %>% 
